### 2012-10-03 Recomb/conclusion_RECOMB.tex
started recomb

%!TEX root = main_RECOMB.tex
\section{Discussion}
\label{sec:conclusion}

### 2012-10-03 Recomb/introduction_RECOMB.tex
started recomb

%!TEX root = main_RECOMB.tex
\section{Introduction}
\label{sec:introduction}

### 2012-10-03 Recomb/main_RECOMB.tex
started recomb

\documentclass[11pt]{article}
\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage[numbers]{natbib}

\usepackage{subfig} %Subfigures

\newcommand{\ourprog}{\texttt{RNA-MoIP}\xspace}

\usepackage[applemac]{inputenc} %for the encoding

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\farna}{\texttt{FARNA}\xspace}
\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
\newcommand{\nast}{\texttt{NAST}\xspace}
\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
\newcommand{\contrafold}{\texttt{contrafold}\xspace}
\newcommand{\unafold}{\texttt{unafold}\xspace}
\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
\newcommand{\assemble}{\texttt{assemble}\xspace}
\newcommand{\fred}{\texttt{FR3D}\xspace}
\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
\newcommand{\contextfold}{\texttt{contextfold}\xspace}
%\newcommand{\citep}{\cite}

\title{RNApyro}
\author{}
\date{}

%Last thing before BEGIN DOC for links to appear in pdf

%  \usepackage{hyperref}
%  \hypersetup{colorlinks}
%              citecolor=black,
%              filecolor=black%
%              linkcolor=black%
%              urlcolor=black%
%              pdftex}

\begin{document}
\maketitle
%\section{}
%\subsection{}
\begin{abstract}
Some stuff~\cite{Waldispuhl2008}

\noindent
\textbf{Key words:} RNA
\end{abstract}

\newpage

\input{introduction_RECOMB.tex}

\input{methods_RECOMB.tex}

\input{results_RECOMB.tex}

\input{conclusion_RECOMB.tex}

\input{acknowledgments_RECOMB.tex}

\newpage

\bibliographystyle{plainnat}
\bibliography{RNApyro}

\end{document}

### 2012-10-03 src/RNAPyroEx.py

@@ -330,12 +330,12 @@ class memoize(dict):
   def __init__(self, f):
       self.fun = f
   
-  def __call__(self,seq,struct,*args):
+  def __call__(self,seq,ref_seq,struct,*args):
       nargs = (args)
       if nargs in self:
           return self[nargs]
       else:
-          val = mpf(self.fun(seq,struct,*args))
+          val = mpf(self.fun(seq,ref_seq,struct,*args))
           self[nargs] = val
           return val
   def resetCache(self):
@@ -354,13 +354,14 @@ def energy((a,b),(a2,b2),alpha=1.0):
   E = STACKING_ENERGY[a,a2,b2,b]
   return  math.exp(-(alpha*E)/(BOLTZMANN*T))
 
-def isostericity(seq,(i,j),(a,b), alpha=1.0):
+def isostericity(seq,ref_seq,(i,j),(a,b), alpha=1.0):
   #isostericity of going from original base pair to (a,b)
-  iso = ISO[(seq[i],seq[j]),(a,b)]
-  return  math.exp(-((1-alpha)*iso)/(BOLTZMANN*T))
+  iso1 = ISO[(ref_seq[i],ref_seq[j]),(a,b)]
+  iso2 = ISO[(ref_seq[i],ref_seq[j]),(seq[i],seq[j])]
+  return  math.exp(-((1-alpha)*(iso1-iso2))/(BOLTZMANN*T))
 
 @memoize
-def forward(seq,struct,(i,j),(a,b),m, alpha=1.0):
+def forward(seq,ref_seq,struct,(i,j),(a,b),m, alpha=1.0):
   #alpha gives the weight energy vs isostericity
   result = 0.
   if m<0: return 0
@@ -375,7 +376,7 @@ def forward(seq,struct,(i,j),(a,b),m, alpha=1.0):
       for a2 in BASES:
         d = delta(seq,i,a2)
         if d <= m:
-          result += forward(seq,struct,
+          result += forward(seq,ref_seq,struct,
                             (i+1,j),
                             (a2,b),
                             m-d)
@@ -386,36 +387,36 @@ def forward(seq,struct,(i,j),(a,b),m, alpha=1.0):
           #if not stacked outside (or border, then no stack possible)
           if i==0 or j==len(seq)-1 or not (j==k and struct[i-1]==j+1):
             for m2 in range(m-d+1):
-              result += forward(seq,struct,
+              result += forward(seq,ref_seq,struct,
                                   (i+1,k-1),
[…]

### 2012-10-03 Recomb/methods_RECOMB.tex
started recomb
added raw equations to recomb

--- 2012-10-03 Recomb/introduction_RECOMB.tex
@@
 %!TEX root = main_RECOMB.tex
-\section{Introduction}
-\label{sec:introduction}
+\section{Methods}
+\label{sec:methods}
+
+Given:
+\begin{itemize}
+	\item an RNA sequence $s$ of length $n$
+	\item $s[i]:=$ nucleotide at position $i$ in $s$
+	\item $\displaystyle
+		S[i]:=\left\{
+		\begin{array}{rl}
+			-1 & \text{If position }i \text{ unpaired}\\
+	 		k  & \text{If position }i \text{ paired with position }k
+		\end{array}\right.
+		$\\
+		S[-1] = -1, special value allowing to compute $S[i-1]$ when $i\equiv 0$
+	\item $B:=\{A, U, G, C\}$
+	\item $ \Kron_{a,a'}:=\left\{
+		\begin{array}{ll}
+			1 &\text{If }a \equiv a'\\
+			0 &\text{Else}
+		\end{array}\right.$
+	\item $\Z{i,j}{m}{\alpha,\beta}:=$ Partition function of the subsequence of $s$ from nucleotides 
+	$i$ to $j$ with $m$ mutations when nucleotides at positions $i-1=\alpha$ and $j+1=\beta$
+	\item $\Y{i,j}{m}{\alpha,\beta}:=$ Partition function of $s$ excluding subsequence  from $i+1$ to 
+	$j-1$ with $m$ mutations when nucleotides at positions $i+1=\alpha$ and $j-1=\beta$ 	
+	\item Some function $ES_{ab \to a'b'}$ in function of turner starking energy, isostericity and
+	  	some $\alpha$. If we choose $\alpha\times\text{StackEnergy} + (1-\alpha)\text{SumIsostericity}$ 
+		we 	have the problem that on any invalid base pair (e.g $GG$) the infinite energy given to the 
+		stacking 	will remove all information of the isostericity. 
+	\item $EI_{ab,a'b'}$ which take into account the isostericity of base pairs not stacked $ab$ when 
+	change to $a'b'$.
+\end{itemize}
+
+\subsection{Inside}
+\subsubsection{Initial conditions}
+$$
+	\forall i \in (0,\cdots,n-1):\, \Z{i+1,i}{m}{\alpha,\beta}=\left\{
+	\begin{array}{ll}
+		1 &\text{If } m = 0\\
+		0 &\text{Else }
+	\end{array}\right.
+$$
+\subsubsection{Recursion}
+$$
+	\Z{i,j}{m}{\alpha,\beta}:=\left\{
+  \begin{array}{ll}
+  		\displaystyle
+      \sum_{\substack{a'\in \B,\\ \Kron_{a',s[i]}\le m}}  
+      \Z{i+1,j}{m-\Kron_{a',s[i]}}{a',\beta} & \text{If }S{[i]}=-1\\
+      \displaystyle
+      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s[i]s[j]}\le m}}  
+			 e^{\frac{-ES_{\alpha \beta \to a' b'}}{RT}}
+			 \Z{i+1,j-1}{m-\Kron_{a'b',s[i]s[j]}}{a',b'}&
+			 \text{Elif }S{[i]}=j \land S{[i-1]}=j+1\\
+			 \displaystyle
+      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s[i]s[k]}\le m}}
+      \sum_{m'=0}^{m-\Kron_{a'b',s[i]s[k]}}
+   		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
+      \Z{i+1,k-1}{m-\Kron_{a'b',s[i]s[k]}-m'}{a',b'}
+      \Z{k+1,j}{m'}{b',\beta} & \text{Elif }S[i]=k \land i < k \leq j\\
+      0 &\text{Else}
+	\end{array}\right.
+$$
+
+\subsection{Outside}	
+\subsubsection{Initial conditions}
+$$
+	\Y{-1,j}{m}{X,X}:=
+		\displaystyle
+	  \Z{j,n-1}{m}{X,X}
+$$
+\subsubsection{Recursion}
+$$
+	\Y{i,j}{m}{\alpha,\beta} = \left\{
+  \begin{array}{ll}
+		\displaystyle
+    \sum_{\substack{a'\in \B,\\ \Kron_{a',s[i]}\le m}}
+    \Y{i-1,j}{m- \Kron_{a',s[i]}}{a',\beta} &
+    \text{Elif }S[i]=-1 \\
+    \displaystyle
+    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[i]s[j]}\le m}}
+		 e^{\frac{-ES_{ a' b' \to \alpha \beta }}{RT}}
+    \Y{i-1,j+1}{m- \Kron_{a'b',s[i]s[j]}}{a',b'} &
+   	 \text{Elif }S{[i]}=j \land S{[i+1]}=j-1\\
+		 \displaystyle
+		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[i]s[k]}\le m}}
+		 \sum_{m'=0}^{m-\Kron_{a'b',s[i]s[k]}}
+  		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
+		 \Y{i-1,k+1}{m- \Kron_{a'b',s[i]s[k]} - m'}{a',b'}
+     \Z{j,k-1}{m'}{\beta,b'} &
+		 \text{Elif }S{[i]}=k \geq j\\
+		 \displaystyle
+		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[k]s[i]}\le m}}
+		 \sum_{m'=0}^{m-\Kron_{a'b',s[k]s[i]}}
+  		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
+		 \Y{k-1,j}{m- \Kron_{a'b',s[k]s[i]} - m'}{a',\beta}
+     \Z{k+1,i-1}{m'}{a',b'} &
+		 \text{Elif }-1 < S{[i]}=k < i\\
+		 0 & \text{Else}
+  \end{array}\right.
+$$
+\subsection*{}
+We must have (if $S[k] = -1)$:
+$$
+	\Z{0,n-1}{m}{X,X} \equiv     
+	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
+	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a};\qquad
+	\forall k \in \{0,\cdots,n-1\}
+$$
+
+\subsection{Probability of a position being a given nucleotide}
+Given $x\in B$,
+$$
+	\mathbb{P}(s[i] = x\mid s, S,m):=\left\{
+	\begin{array}{ll}
+		\displaystyle
+		\frac{
+			\Y{i-1,i+1}{m-\Kron_{x,s[k]}}{x,x}
+		}{
+			\Z{0,n-1}{m}{X,X}
+		}
+		&\text{If }S[i] = -1\\
+		\displaystyle
+ 		\frac{
+			\displaystyle
+			\sum_{\substack{b\in Bases\\\Kron_{bx,s[k]s[i]\leq m}}}
+			\sum_{m'=0}^{m-\Kron_{bx,s[k]s[i]}}
+			\Y{k-1,i+1}{m-\Kron_{bx,s[k]s[i]-m'}}{b,x}
+			\Z{k+1,i-1}{m'}{b,x}
+		}{
+			\Z{0,n-1}{m}{X,X}
+		}
+		&\text{If }S[i]=k<i\\
+		\displaystyle
+		\frac{
+			\displaystyle
+			\sum_{\substack{b\in Bases\\\Kron_{xb,s[i]s[k]\leq m}}}
+			\sum_{m'=0}^{m-\Kron_{xb,s[i]s[k]}}
+			\Y{i-1,k+1}{m-\Kron_{xb,s[i]s[k]-m'}}{x,b}
+			\Z{i+1,k-1}{m'}{x,b}
+		}{
+			\Z{0,n-1}{m}{X,X}
+		}
+		&\text{If }S[i]=k>i
+	\end{array}\right.
+$$

### 2012-10-03 src/RNAPyroEx.py
if only one seq in fasta, use it for starting AND ref

@@ -588,7 +588,10 @@ def parse_fasta(file_name):
         seq.append(line)
       if all(x in '(.)' for x in line):
         struct = line 
-  return seq[0],seq[1], parseStruct(struct)
+  if len(seq) == 1:
+    return seq[0], seq[0], parseStruct(struct)
+  else:
+    return seq[0],seq[1], parseStruct(struct)
 
 def all_probabilities(seq,ref_seq, stuct, m, alpha):
   n = len(seq)

### 2012-10-03 src/RNAPyroEx.py
input must be alignment + new seq

@@ -356,9 +356,10 @@ def energy((a,b),(a2,b2),alpha=1.0):
 
 def isostericity(seq,ref_seq,(i,j),(a,b), alpha=1.0):
   #isostericity of going from original base pair to (a,b)
-  iso1 = ISO[(ref_seq[i],ref_seq[j]),(a,b)]
-  iso2 = ISO[(ref_seq[i],ref_seq[j]),(seq[i],seq[j])]
-  return  math.exp(-((1-alpha)*(iso1-iso2))/(BOLTZMANN*T))
+  iso_mut = sum(ISO[(ref[i],ref[j]),(a,b)] for ref in ref_seq)
+  iso_start = sum(ISO[(ref[i],ref[j]),(seq[i],seq[j])] for ref in ref_seq)
+  iso = (iso_mut-iso_start)/len(ref_seq)
+  return  math.exp(-((1-alpha)*iso)/(BOLTZMANN*T))
 
 @memoize
 def forward(seq,ref_seq,struct,(i,j),(a,b),m, alpha=1.0):
@@ -588,10 +589,7 @@ def parse_fasta(file_name):
         seq.append(line)
       if all(x in '(.)' for x in line):
         struct = line 
-  if len(seq) == 1:
-    return seq[0], seq[0], parseStruct(struct)
-  else:
-    return seq[0],seq[1], parseStruct(struct)
+  return seq[:-1],seq[-1], parseStruct(struct)
 
 def all_probabilities(seq,ref_seq, stuct, m, alpha):
   n = len(seq)
@@ -612,8 +610,6 @@ def help():
     the nb of mutants allowed and the value of alpha, between 0 and 1"""
 
 if __name__ == "__main__":
-  test()
-  sys.exit(0)
   opts = sys.argv
   if len(opts) < 4:
     help()
@@ -638,7 +634,7 @@ if __name__ == "__main__":
 
 
 
-  seq,ref_seq,struct = parse_fasta(file_name)
+  ref_seq,seq,struct = parse_fasta(file_name)
   results = all_probabilities(seq,ref_seq,struct,m,alpha)
   display_all_probabilities(results)
 

### 2012-10-03 src/RNAPyroEx.py
added overrid maxint

@@ -607,9 +607,13 @@ def display_all_probabilities(results):
 
 def help():
     print """You need at least three arguments, the file path,
-    the nb of mutants allowed and the value of alpha, between 0 and 1"""
+    the nb of mutants allowed and the value of alpha, between 0 and 1
+    An optional 4th argument can be given if you want to change the max penality for an invalid 
+    base pair
+    """
 
 if __name__ == "__main__":
+  global STACKING_ENERGY
   opts = sys.argv
   if len(opts) < 4:
     help()
@@ -631,6 +635,14 @@ if __name__ == "__main__":
   if not 0 <= alpha <= 1:
     help()
     sys.exit(1)
+  if len(opts) == 4:
+    try:
+      z = float(opts[4])
+      for x in STACKING_ENERGY:
+        if STACKING_ENERGY[x] == sys.maxint:
+          STACKING_ENERGY[x] = z
+    except ValueError:
+      pass
 
 
 

### 2012-10-03 Recomb/methods_RECOMB.tex
started method formally

--- 2012-10-03 Recomb/introduction_RECOMB.tex
@@
 %!TEX root = main_RECOMB.tex
-\section{Introduction}
-\label{sec:introduction}
+\section{Methods}
+\label{sec:methods}
+Let $\Omega$ be an un-gapped RNA alignment, $S$ its associate secondary structure, $s$ an RNA 
+sequence and $m\geq 0$. $S$ is considered as  one derivation of the SCFG generating  
+all possible secondary structures of the same length as $s$. We define a variant of the
+ \emph{Inside-Outside algorithm}~\cite{Lari1990} with the two following functions,
+$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former, ass the 
+equivalent of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
+knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing $m$ mutations. The latter computes the \emph{outside}, in particular, knowing 
+that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing, outside 
+including $i$ and $j$, $m$ mutations.
+The Boltzmann weights are a combination of the base pairs stacking energy, 
+using as values those of the NNDB~\cite{Turner2010}, and
+the average isostericity with $\Omega$, as defined in~\cite{Stombaugh2009}.
+
+This allows us to compute the probability of any position being a specific nucleotides,
+over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$.
+
+\subsection{Definitions}
+Let be $B:=\left\{\text{A},\text{C},\text{G},\text{U}\right\}$, the set of nucleotides.
+Given $s\in B^n$ an RNA sequence, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
+length $n$, and $S$ a secondary structure without pseudoknots. 
+Formally, if $(i,j)$ and $(k,l)$ are base pairs in $S$, there is no overlapping extremities
+ $\{i,j\}\cap \{k,l\}=\varnothing$ and either the intersection is empty 
+ ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
+ $[i,j]\subset[k,l])$. Let $R$ be the Boltzmann constant, $T$ the temperature in Kelvin and
+  the function $\delta$ such that: 
+ $\forall a,a' \in B, \delta_{a,a'}:=\left\{\begin{array}{ll}
+															1 & \text{If } a\equiv a'\\
+															0 & \text{Else}
+														\end{array}\right.$
+
+\subsection{Energy Model}
+The energy we will be composed of two function, $ES^{\beta}_{ab\to a'b'}$ and 
+$EI_{ab\to a'b'}$. $ES^\beta_{ab\to a'b'}$ is equal to the 
+stacking energy of the base pair with nucleotides $ab$ on top of the base pair with nucleotides 
+$a'b'$, as set in the NNDB~\cite{Turner2010}. If one of the base pair is not valid (i.e. not in 
+$\{\text{GU},\text{UG},\text{CG},\text{GC}, \text{AU or UA}\}$, the value is a parameter 
+$\beta \in [1,\infty]$. This allows
+to completely forbid a sequence where a base pair is non valid, when $\beta = \infty$ or only 
+penalize it.
+$EI_{ab\to a'b'}$ is the isostericity between
+ the base pair $ab$ and $a'b'$ as defined in~\cite{Stombaugh2009}. Let be $\alpha\in[0,1]$, it 
+ will be used to balance the weight given to the stacking energy or the isostericity.
+
+\subsection{Inside}
+We define the \emph{Inside} function $\Z{i,j}{m}{a,b}$ as follows.
+\subsubsection{Initial conditions}
+$$
+	\forall i \in (0,\cdots,n-1):\, \Z{i+1,i}{m}{a,b}=\left\{
+	\begin{array}{ll}
+		1 &\text{If } m = 0\\
+		0 &\text{Else }
+	\end{array}\right.
+$$
+\subsubsection{Recursion}
+$$
+	\Z{i,j}{m}{a,b}:=\left\{
+  \begin{array}{ll}
+  		\displaystyle
+      \sum_{\substack{a'\in \B,\\ \Kron_{a',s[i]}\le m}}  
+      \Z{i+1,j}{m-\Kron_{a',s[i]}}{a',b} & \text{If }S{[i]}=-1\\
+      \displaystyle
+      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s[i]s[j]}\le m}}  
+			 e^{\frac{-ES_{a b \to a' b'}}{RT}}
+			 \Z{i+1,j-1}{m-\Kron_{a'b',s[i]s[j]}}{a',b'}&
+			 \text{Elif }S{[i]}=j \land S{[i-1]}=j+1\\
+			 \displaystyle
+      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s[i]s[k]}\le m}}
+      \sum_{m'=0}^{m-\Kron_{a'b',s[i]s[k]}}
+   		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
+      \Z{i+1,k-1}{m-\Kron_{a'b',s[i]s[k]}-m'}{a',b'}
+      \Z{k+1,j}{m'}{b',b} & \text{Elif }S[i]=k \land i < k \leq j\\
+      0 &\text{Else}
+	\end{array}\right.
+$$
+
+\subsection{Outside}	
+\subsubsection{Initial conditions}
+$$
+	\Y{-1,j}{m}{X,X}:=
+		\displaystyle
+	  \Z{j,n-1}{m}{X,X}
+$$
+\subsubsection{Recursion}
+$$
+	\Y{i,j}{m}{a,b} = \left\{
+  \begin{array}{ll}
+		\displaystyle
+    \sum_{\substack{a'\in \B,\\ \Kron_{a',s[i]}\le m}}
+    \Y{i-1,j}{m- \Kron_{a',s[i]}}{a',b} &
+    \text{Elif }S[i]=-1 \\
+    \displaystyle
+    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[i]s[j]}\le m}}
+		 e^{\frac{-ES_{ a' b' \to a b }}{RT}}
+    \Y{i-1,j+1}{m- \Kron_{a'b',s[i]s[j]}}{a',b'} &
+   	 \text{Elif }S{[i]}=j \land S{[i+1]}=j-1\\
+		 \displaystyle
+		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[i]s[k]}\le m}}
+		 \sum_{m'=0}^{m-\Kron_{a'b',s[i]s[k]}}
+  		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
+		 \Y{i-1,k+1}{m- \Kron_{a'b',s[i]s[k]} - m'}{a',b'}
+     \Z{j,k-1}{m'}{b,b'} &
+		 \text{Elif }S{[i]}=k \geq j\\
+		 \displaystyle
+		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[k]s[i]}\le m}}
+		 \sum_{m'=0}^{m-\Kron_{a'b',s[k]s[i]}}
+  		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
+		 \Y{k-1,j}{m- \Kron_{a'b',s[k]s[i]} - m'}{a',b}
+     \Z{k+1,i-1}{m'}{a',b'} &
+		 \text{Elif }-1 < S{[i]}=k < i\\
+		 0 & \text{Else}
+  \end{array}\right.
+$$
+\subsection*{}
+We must have (if $S[k] = -1)$:
+$$
+	\Z{0,n-1}{m}{X,X} \equiv     
+	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
+	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a};\qquad
+	\forall k \in \{0,\cdots,n-1\}
+$$
+
+\subsection{Probability of a position being a given nucleotide}
+Given $x\in B$,
+$$
+	\mathbb{P}(s[i] = x\mid s, S,m):=\left\{
+	\begin{array}{ll}
+		\displaystyle
+		\frac{
+			\Y{i-1,i+1}{m-\Kron_{x,s[k]}}{x,x}
+		}{
+			\Z{0,n-1}{m}{X,X}
+		}
+		&\text{If }S[i] = -1\\
+		\displaystyle
+ 		\frac{
+			\displaystyle
+			\sum_{\substack{b\in Bases\\\Kron_{bx,s[k]s[i]\leq m}}}
+			\sum_{m'=0}^{m-\Kron_{bx,s[k]s[i]}}
+			\Y{k-1,i+1}{m-\Kron_{bx,s[k]s[i]-m'}}{b,x}
+			\Z{k+1,i-1}{m'}{b,x}
+		}{
+			\Z{0,n-1}{m}{X,X}
+		}
+		&\text{If }S[i]=k<i\\
+		\displaystyle
+		\frac{
+			\displaystyle
+			\sum_{\substack{b\in Bases\\\Kron_{xb,s[i]s[k]\leq m}}}
+			\sum_{m'=0}^{m-\Kron_{xb,s[i]s[k]}}
+			\Y{i-1,k+1}{m-\Kron_{xb,s[i]s[k]-m'}}{x,b}
+			\Z{i+1,k-1}{m'}{x,b}
+		}{
+			\Z{0,n-1}{m}{X,X}
+		}
+		&\text{If }S[i]=k>i
+	\end{array}\right.
+$$

### 2012-10-03 Recomb/methods_RECOMB.tex
first draft inside

@@
  \emph{Inside-Outside algorithm}~\cite{Lari1990} with the two following functions,
-$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former, ass the 
+$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former, as the 
 equivalent of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
 knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing $m$ mutations. The latter computes the \emph{outside}, in particular, knowing 
-that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing, outside 
+that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing outside,
 including $i$ and $j$, $m$ mutations.
@@
 The energy we will be composed of two function, $ES^{\beta}_{ab\to a'b'}$ and 
-$EI_{ab\to a'b'}$. $ES^\beta_{ab\to a'b'}$ is equal to the 
+$EI_{(i,j),ab}$. The former is equal to the 
 stacking energy of the base pair with nucleotides $ab$ on top of the base pair with nucleotides 
@@
 penalize it.
-$EI_{ab\to a'b'}$ is the isostericity between
- the base pair $ab$ and $a'b'$ as defined in~\cite{Stombaugh2009}. Let be $\alpha\in[0,1]$, it 
+$EI_{(i,j), a'b'}$ is the average of the sum of differences between the isostericity
+of base pairs at positions $(i,j)$ in $\Omega$ and $s_is_j$, and the isostericity of base pairs
+ at positions $(i,j)$ in $\Omega$ and $ab$. If gives us an indication 
+ if the base pair $ab$ is more isosteric to the set $\Omega$ than the one on the sequence 
+ $\omega$. Where the isostericity values 
+ between any pair of canonical base pairs is  as defined in~\cite{Stombaugh2009}. 
+ Let be $\alpha\in[0,1]$, it 
  will be used to balance the weight given to the stacking energy or the isostericity.
@@
 \subsection{Inside}
-We define the \emph{Inside} function $\Z{i,j}{m}{a,b}$ as follows.
-\subsubsection{Initial conditions}
-$$
+To define the \emph{Inside} function $\Z{i,j}{m}{a,b}$ as a recurrence, we will use as initial conditions:
+
+\[
 	\forall i \in (0,\cdots,n-1):\, \Z{i+1,i}{m}{a,b}=\left\{
@@
 	\end{array}\right.
-$$
-\subsubsection{Recursion}
+\]
+In other words, when we evaluate the function $\mathcal Z$  over only one nucleotide, there 
+is only one possible solution if there is $0$ mutations left, and none else. Since the 
+energetics term only consider base pairs, they are not involved in the initial conditions.
+
+The recurrence is composed of four terms:
 $$
@@
   		\displaystyle
-      \sum_{\substack{a'\in \B,\\ \Kron_{a',s[i]}\le m}}  
-      \Z{i+1,j}{m-\Kron_{a',s[i]}}{a',b} & \text{If }S{[i]}=-1\\
+      \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}  
+      \Z{i+1,j}{m-\Kron_{a',s_i}}{a',b} & \text{If }S_{i}=-1\\
       \displaystyle
-      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s[i]s[j]}\le m}}  
-			 e^{\frac{-ES_{a b \to a' b'}}{RT}}
-			 \Z{i+1,j-1}{m-\Kron_{a'b',s[i]s[j]}}{a',b'}&
-			 \text{Elif }S{[i]}=j \land S{[i-1]}=j+1\\
+      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
+			 e^{\frac{-(\alpha ES_{a b \to a' b'}+(1-\alpha)EI_{(i,j),a'b'})}{RT}}
+			 \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
+			 \text{Elif }S_i=j \land S_{i-1}=j+1\\
 			 \displaystyle
-      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s[i]s[k]}\le m}}
-      \sum_{m'=0}^{m-\Kron_{a'b',s[i]s[k]}}
-   		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
-      \Z{i+1,k-1}{m-\Kron_{a'b',s[i]s[k]}-m'}{a',b'}
-      \Z{k+1,j}{m'}{b',b} & \text{Elif }S[i]=k \land i < k \leq j\\
+      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
+      \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
+   		 e^{\frac{-(1-\alpha)EI_{(i,k),a'b'}}{RT}}
+      \Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
+      \Z{k+1,j}{m'}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
       0 &\text{Else}
@@
 $$
+The cases can be breaked down as follows.
+\begin{description}
+\item[$S_{i}=-1$:] If the nucleotide at position $i$ is not paired, then the value is the same
+as if we increase the lower interval bound by $1$ (i.e. $i+1$), and consider all possible
+ nucleotides $a'$ at position $i$, taking into account the 
+\item[$S_i=j$ and $S_{i-1}=j+1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i-1$ is
+paired with $j+1$, we are in the only case were stacked base pairs can occur. We thus add
+the energy of the stacking and of the isostericity of the base pair $(i,j)$. What is left
+to compute is the \emph{inside} value of the interval $[i+1,j-1]$ over all possible nucleotides 
+$a',b'\in B^2$ at positions $i$ and $j$ respectively.
+\item[$S_i=k$ and $i<k \leq j$:] If nucleotide $i$ is paired with position $k$ but is not stacked outside, the 
+only term contributing to the energy is the isostericity of the base pair $(i,k)$. This creates
+two different intervals for which we must compute the values, $[i+1,k-1]$ and $[k+1,j-1]$, for 
+all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $j$ respectively.
+\item[Else:] In all other cases, we are in a derivation of the SCFG that does not correspond to the 
+secondary structure $S$, and we return $0$.
+
+\end{description}
 
 \subsection{Outside}	
-\subsubsection{Initial conditions}
+The \emph{Outside} function, $\Y{i,j}{m}{a,b}$, will also be defined as a recurrence and have 
+as initial conditions the following.
 $$

### 2012-10-04 Recomb/methods_RECOMB.tex

@@
 \subsection{Energy Model}
-The energy we will be composed of two function, $ES^{\beta}_{ab\to a'b'}$ and 
-$EI_{(i,j),ab}$. The former is equal to the 
+The energy we will be composed of two function, $\text{ES}^{\beta}_{ab\to a'b'}$ and 
+$\text{EI}_{(i,j),ab}$. The former is equal to the 
 stacking energy of the base pair with nucleotides $ab$ on top of the base pair with nucleotides 
@@
 penalize it.
-$EI_{(i,j), a'b'}$ is the average of the sum of differences between the isostericity
+$\text{EI}_{(i,j), a'b'}$ is the average of the sum of differences between the isostericity
 of base pairs at positions $(i,j)$ in $\Omega$ and $s_is_j$, and the isostericity of base pairs
@@
 \subsection{Inside}
-To define the \emph{Inside} function $\Z{i,j}{m}{a,b}$ as a recurrence, we will use as initial conditions:
+The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function considering only the 
+energy in subsequence $[i,j]$ over mutants of $s$ having exactly $m$ mutations between $[i,j]$ and who nucleotide at position $i-1$ is $a$ (resp. in position $j+1$ it is $b$).
+We define function $\Z{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
 
@@
 \]
-In other words, when we evaluate the function $\mathcal Z$  over only one nucleotide, there 
+In other words, when we evaluate the function $\mathcal Z$, after exhausting all positions, there 
 is only one possible solution if there is $0$ mutations left, and none else. Since the 
-energetics term only consider base pairs, they are not involved in the initial conditions.
+energetic terms only depend on base pairs, they are not involved in the initial conditions.
 
-The recurrence is composed of four terms:
+The recursion itself is composed of four terms:
 $$
@@
       \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
-			 e^{\frac{-(\alpha ES_{a b \to a' b'}+(1-\alpha)EI_{(i,j),a'b'})}{RT}}
+			 e^{\frac{-(\alpha \text{ES}^\beta_{a b \to a' b'}+(1-\alpha)\text{EI}_{(i,j),a'b'})}{RT}}
 			 \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
@@
       \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-   		 e^{\frac{-(1-\alpha)EI_{(i,k),a'b'}}{RT}}
+   		 e^{\frac{-(1-\alpha)\text{EI}_{(i,k),a'b'}}{RT}}
       \Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
@@
 as if we increase the lower interval bound by $1$ (i.e. $i+1$), and consider all possible
- nucleotides $a'$ at position $i$, taking into account the 
+ nucleotides $a'$ at position $i$. 
 \item[$S_i=j$ and $S_{i-1}=j+1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i-1$ is
@@
 $a',b'\in B^2$ at positions $i$ and $j$ respectively.
-\item[$S_i=k$ and $i<k \leq j$:] If nucleotide $i$ is paired with position $k$ but is not stacked outside, the 
-only term contributing to the energy is the isostericity of the base pair $(i,k)$. This creates
+\item[$S_i=k$ and $i<k \leq j$:] If nucleotide $i$ is paired with position $k$ 
+but is not stacked outside, the 
+only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. This 
+creates
 two different intervals for which we must compute the values, $[i+1,k-1]$ and $[k+1,j-1]$, for 
@@
 \subsection{Outside}	
-The \emph{Outside} function, $\Y{i,j}{m}{a,b}$, will also be defined as a recurrence and have 
-as initial conditions the following.
+The \emph{Outside} function, $\mathcal Y$, is the partition function considering only the 
+energy in subsequences $[0,i]\cup[j,n-1]$ over the mutants of $s$ having exactly $m$ mutations between $[0,i]\cup[j,n-1]$ and whose nucleotide at position $i+1$ is $a$ 
+(resp. in position $j-1$ it is $b$).
+We define function $\Y{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
+
 $$
@@
 $$
-\subsubsection{Recursion}
+The recurrence, as shown below, will increase the interval $[i,j]$ by decreasing $i$ when
+it is not base paired. If it is with a position $k>j$, we increase $j$ to include it.
+ Thus, when we need
+to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
+structure between $(j,n-1)$ must be a set of independent stems. Thus, all the outside energy is
+equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself is the following:
 $$
@@
 		\displaystyle
-    \sum_{\substack{a'\in \B,\\ \Kron_{a',s[i]}\le m}}
-    \Y{i-1,j}{m- \Kron_{a',s[i]}}{a',b} &
-    \text{Elif }S[i]=-1 \\
+    \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}
+    \Y{i-1,j}{m- \Kron_{a',s_i}}{a',b} &
+    \text{Elif }S_i=-1 \\
     \displaystyle
-    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[i]s[j]}\le m}}
-		 e^{\frac{-ES_{ a' b' \to a b }}{RT}}
-    \Y{i-1,j+1}{m- \Kron_{a'b',s[i]s[j]}}{a',b'} &
-   	 \text{Elif }S{[i]}=j \land S{[i+1]}=j-1\\
+    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
+		 e^{\frac{-(\alpha \text{ES}^\beta_{a b \to a' b'}+(1-\alpha)\text{EI}_{(i,j),a'b'})}{RT}}
+    \Y{i-1,j+1}{m- \Kron_{a'b',s_is_j}}{a',b'} &
+   	 \text{Elif }S_{i}=j \land S_{i+1}=j-1\\
 		 \displaystyle
-		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[i]s[k]}\le m}}
-		 \sum_{m'=0}^{m-\Kron_{a'b',s[i]s[k]}}
-  		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
-		 \Y{i-1,k+1}{m- \Kron_{a'b',s[i]s[k]} - m'}{a',b'}
+		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
+		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
+  		 e^{\frac{-(1-\alpha)\text{EI}_{(i,k),a'b'}}{RT}}
+		 \Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
      \Z{j,k-1}{m'}{b,b'} &
-		 \text{Elif }S{[i]}=k \geq j\\
+		 \text{Elif }S_{i}=k \geq j\\
 		 \displaystyle
-		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[k]s[i]}\le m}}
-		 \sum_{m'=0}^{m-\Kron_{a'b',s[k]s[i]}}
-  		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
-		 \Y{k-1,j}{m- \Kron_{a'b',s[k]s[i]} - m'}{a',b}
+		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_ks_i}\le m}}
+		 \sum_{m'=0}^{m-\Kron_{a'b',s_ks_i}}
+   	 e^{\frac{-(1-\alpha)\text{EI}_{(k,i),a'b'}}{RT}}
+		 \Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
      \Z{k+1,i-1}{m'}{a',b'} &
-		 \text{Elif }-1 < S{[i]}=k < i\\
+		 \text{Elif }-1 < S_{i}=k < i\\
 		 0 & \text{Else}
@@
 $$
-\subsection*{}
-We must have (if $S[k] = -1)$:
+The five cases can be break down as follows.
+\begin{description}
+\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, then the value is the same
+as if we decrease the lower interval bound by $1$ (i.e. $i-1$), and consider all possible
+nucleotides $a'$ at position $i$.
+\item[$S_{i}=j$ and $S_{i+1}=j-1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i+1$ is
+paired with $j-11$, we are in the only case were stacked base pairs can occur. We thus add
+the energy of the stacking and of the isostericity of the base pair $(i,j)$. What is left
+to compute is the \emph{outside} value for the interval $[i-1,j+1]$ over all possible nucleotides 
+$a',b'\in B^2$ at positions $i$ and $j$ respectively.
+\item[$S_{i}=k \geq j$:]If nucleotide $i$ is paired with position $k\geq j$, 
+and is not stacked inside, the 
+only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
+We can then consider the outside interval $[i-1,k+1]$ by multiplying it by the the \emph{forward}
+value of the newly included interval (i.e. $[j,k-1]$), for 
+all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $k$ respectively.
+\item[$-1<S_{i}<i$:]As above but if the pairing is to the left.
+\item[Else:] In all other cases, we are in a derivation of the SCFG that does not correspond to the 
+secondary structure $S$, and we return $0$.
+
+\end{description}
+
+\section{Inside-Outside}
+By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
+be described in function of the \emph{forward} term as $\Z{0,n-1}{m}{X,X}$,
+ for any nucleotide $X\in B$ or
+in function of the \emph{backward} term, for any position $k$ such that $S_k=-1$:
 $$
-	\Z{0,n-1}{m}{X,X} \equiv     
+	\Z{0,n-1}{m}{X,X}
+	\equiv
 	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
-	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a};\qquad
-	\forall k \in \{0,\cdots,n-1\}
+	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a}
 $$
 
-\subsection{Probability of a position being a given nucleotide}
-Given $x\in B$,
+We are now interested in knowing, under our model, 
+ the probability that a given position is a given nucleotide.
+We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
+Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of mutations allowed. 
 $$
-	\mathbb{P}(s[i] = x\mid s, S,m):=\left\{
+	\mathbb{P}(s_i = x\mid s,\Omega, S,M):=\left\{
 	\begin{array}{ll}
@@
 		\frac{
-			\Y{i-1,i+1}{m-\Kron_{x,s[k]}}{x,x}
+			\displaystyle
+			\sum_{m=0}^{M}
+			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
 		}{
+			\displaystyle
+			\sum_{m=0}^{M}
 			\Z{0,n-1}{m}{X,X}
 		}
-		&\text{If }S[i] = -1\\
+		&\text{If }S_i = -1\\
 		\displaystyle
@@
 			\displaystyle
-			\sum_{\substack{b\in Bases\\\Kron_{bx,s[k]s[i]\leq m}}}
-			\sum_{m'=0}^{m-\Kron_{bx,s[k]s[i]}}
-			\Y{k-1,i+1}{m-\Kron_{bx,s[k]s[i]-m'}}{b,x}
+			\sum_{m=0}^{M}
+			\sum_{\substack{b\in Bases\\\Kron_{bx,s_ks_i\leq m}}}
+			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
+			\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
 			\Z{k+1,i-1}{m'}{b,x}
 		}{
+			\displaystyle
+			\sum_{m=0}^{M}
 			\Z{0,n-1}{m}{X,X}
 		}
-		&\text{If }S[i]=k<i\\
+		&\text{If }S_i=k<i\\
 		\displaystyle
@@
 			\displaystyle
-			\sum_{\substack{b\in Bases\\\Kron_{xb,s[i]s[k]\leq m}}}
-			\sum_{m'=0}^{m-\Kron_{xb,s[i]s[k]}}
-			\Y{i-1,k+1}{m-\Kron_{xb,s[i]s[k]-m'}}{x,b}
+			\sum_{m=0}^{M}
+			\sum_{\substack{b\in Bases\\\Kron_{xb,s_is_k\leq m}}}
+			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
+			\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
 			\Z{i+1,k-1}{m'}{x,b}
 		}{
+			\displaystyle
+			\sum_{m=0}^{M}
 			\Z{0,n-1}{m}{X,X}
@@
 $$
+
+In every case, the denominator is the sum of the partition function of exactly $m$ mutations, for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
+\begin{description}
+\item[$S_i=-1$:]
+\item[$S_i=k>i$:]
+\item[$S_i=k<i$:]
+\end{description}

### 2012-10-04 Recomb/methods_RECOMB.tex
added probs explanation

@@
 
-In every case, the denominator is the sum of the partition function of exactly $m$ mutations, for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
+In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
 \begin{description}
-\item[$S_i=-1$:]
-\item[$S_i=k>i$:]
-\item[$S_i=k<i$:]
+\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, we are concerned by the weights
+over all sequences which have at position $i$ nucleotide $x$, which is exactly the sum
+of the values of $\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}$, for all $m$ between $0$ and $M$.
+\item[$S_i=k>i$:] Since we need to respect the derivation of the secondary structure $S$, if 
+position $i$ is paired, we must consider the two partition functions. The \emph{outside} of the 
+base pair, and the \emph{inside}, for all possible values for the nucleotide at position $k$, and
+all possible distribution of the mutant positions between the inside and outside of the base pair.
+\item[$S_i=k<i$:] Same as above, but with $i$ pairing with a lower position.
 \end{description}

### 2012-10-04 Recomb/methods_RECOMB.tex
removed typo

@@
 The energy we will be composed of two function, $\text{ES}^{\beta}_{ab\to a'b'}$ and 
-$\text{EI}_{(i,j),ab}$. The former is equal to the 
+$\text{EI}^{\Omega}_{(i,j),ab}$. The former is equal to the 
 stacking energy of the base pair with nucleotides $ab$ on top of the base pair with nucleotides 
@@
 penalize it.
-$\text{EI}_{(i,j), a'b'}$ is the average of the sum of differences between the isostericity
+$\text{EI}^{\Omega}_{(i,j), a'b'}$ is the average of the sum of differences between the isostericity
 of base pairs at positions $(i,j)$ in $\Omega$ and $s_is_j$, and the isostericity of base pairs
@@
 The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function considering only the 
-energy in subsequence $[i,j]$ over mutants of $s$ having exactly $m$ mutations between $[i,j]$ and who nucleotide at position $i-1$ is $a$ (resp. in position $j+1$ it is $b$).
+energy in subsequence $[i,j]$ over mutants of $s$ having exactly $m$ mutations between $[i,j]$ and whose nucleotide at position $i-1$ is $a$ (resp. in position $j+1$ it is $b$).
 We define function $\Z{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
@@
       \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
-			 e^{\frac{-(\alpha \text{ES}^\beta_{a b \to a' b'}+(1-\alpha)\text{EI}_{(i,j),a'b'})}{RT}}
+			 e^{\frac{-(\alpha \text{ES}^\beta_{a b \to a' b'}+(1-\alpha)\text{EI}^{\Omega}_{(i,j),a'b'})}{RT}}
 			 \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
@@
       \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-   		 e^{\frac{-(1-\alpha)\text{EI}_{(i,k),a'b'}}{RT}}
+   		 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(i,k),a'b'}}{RT}}
       \Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
@@
 $$
-The cases can be breaked down as follows.
+The cases can be broked down as follows.
 \begin{description}
@@
     \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
-		 e^{\frac{-(\alpha \text{ES}^\beta_{a b \to a' b'}+(1-\alpha)\text{EI}_{(i,j),a'b'})}{RT}}
+		 e^{\frac{-(\alpha \text{ES}^\beta_{a b \to a' b'}+(1-\alpha)\text{EI}^{\Omega}_{(i,j),a'b'})}{RT}}
     \Y{i-1,j+1}{m- \Kron_{a'b',s_is_j}}{a',b'} &
@@
 		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-  		 e^{\frac{-(1-\alpha)\text{EI}_{(i,k),a'b'}}{RT}}
+  		 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(i,k),a'b'}}{RT}}
 		 \Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
@@
 		 \sum_{m'=0}^{m-\Kron_{a'b',s_ks_i}}
-   	 e^{\frac{-(1-\alpha)\text{EI}_{(k,i),a'b'}}{RT}}
+   	 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(k,i),a'b'}}{RT}}
 		 \Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
@@
 $$
-The five cases can be break down as follows.
+The five cases can be broked down as follows.
 \begin{description}
@@
 			\sum_{m=0}^{M}
+			\sum_{\substack{b\in Bases\\\Kron_{xb,s_is_k\leq m}}}
+			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
+			\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
+			\Z{i+1,k-1}{m'}{x,b}
+		}{
+			\displaystyle
+			\sum_{m=0}^{M}
+			\Z{0,n-1}{m}{X,X}
+		}
+		&\text{If }S_i=k>i\\
+		\displaystyle
+ 		\frac{
+			\displaystyle
+			\sum_{m=0}^{M}
 			\sum_{\substack{b\in Bases\\\Kron_{bx,s_ks_i\leq m}}}
@@
 		}
-		&\text{If }S_i=k<i\\
-		\displaystyle
-		\frac{
-			\displaystyle
-			\sum_{m=0}^{M}
-			\sum_{\substack{b\in Bases\\\Kron_{xb,s_is_k\leq m}}}
-			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
-			\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
-			\Z{i+1,k-1}{m'}{x,b}
-		}{
-			\displaystyle
-			\sum_{m=0}^{M}
-			\Z{0,n-1}{m}{X,X}
-		}
-		&\text{If }S[i]=k>i
+		&\text{If }S_i=k<i
 	\end{array}\right.

### 2012-10-04 src/RNAPyroEx.py
fixed bug

@@ -635,7 +635,7 @@ if __name__ == "__main__":
   if not 0 <= alpha <= 1:
     help()
     sys.exit(1)
-  if len(opts) == 4:
+  if len(opts) == 5:
     try:
       z = float(opts[4])
       for x in STACKING_ENERGY:

### 2012-10-04 src/RNAPyroEx.py
fixed bug

@@ -23,6 +23,7 @@ BOLTZMANN = 0.0019872041
 T = 310.15
 
 #Just generate all possible combination with maxint
+global STACKING_ENERGY
 STACKING_ENERGY = {k:sys.maxint for k in itertools.product(
                   BASES, repeat=4)}
 #Adjust with turner04
@@ -613,7 +614,6 @@ def help():
     """
 
 if __name__ == "__main__":
-  global STACKING_ENERGY
   opts = sys.argv
   if len(opts) < 4:
     help()

### 2012-10-04 benchmark/RfamParser.py
add passer features

@@ -182,7 +182,7 @@ def clusterUnGapped(data):
 
 #################################################################################
 
-def main(filename):
+def main(filename,nmut,idc,listflag):
 
     seqline_re = re.compile("(\S+)(\s+)(\S+)");
 
@@ -235,13 +235,12 @@ def main(filename):
 
     cleandata,idMap = clusterUnGapped(info)
 
-    if False:
+    if listflag:
         for hcode in cleandata.keys():
             if len(cleandata[hcode])>1:
                 print '>>>',idMap['code2num'][hcode],', size=' + str(len(cleandata[hcode]))
+        sys.exit(1)
             
-    idc=60
-    nmut=10
     hcode=idMap['num2code'][idc]
     listout = []
     listout.append(random.randint(0,len(cleandata[hcode])-1))
@@ -267,7 +266,7 @@ def main(filename):
         if not ssamatch:
             print 'WARNING: Structure do not match'
 
-        print '>',idout,'(mutant)'
+        print '>',idout,'(' + str(nmut) + '-mutant)'
         cleanseq = cleandata[hcode][idout].replace('.','')
         mutant = mutate(cleanseq,nmut)
         print mutant
@@ -280,11 +279,14 @@ def main(filename):
 if __name__ == '__main__':
 
   try:
-      opts, args = getopt.getopt(sys.argv[1:], "hf:", ["help", "file="]);
+      opts, args = getopt.getopt(sys.argv[1:], "hf:m:n:l", ["help", "file=", "mut=", "id=", "list"]);
   except getopt.GetoptError:
       usage(sys.argv[0]);
 
   filename='';
+  nmut=0;
+  id=0;
+  listflag=False;
 
   argStart=len(sys.argv);
   for o,a in opts:
@@ -293,8 +295,17 @@ if __name__ == '__main__':
     if o in ("-f", "--file"):
       filename = a;
       argStart-=2;  
+    if o in ("-m", "--file"):
+      nmut = int(a);
+      argStart -=2;  
+    if o in ("-n", "--file"):
+      id = int(a);
+      argStart-=2;  
+    if o in ("-l", "--list"):
+      listflag = True;
+      argStart-=1;  
 
-  main(filename);
+  main(filename,nmut,id,listflag);
 
 
 

### 2012-10-04 src/RNAPyroEx.py
fixed bug

@@ -583,14 +583,19 @@ def test():
 def parse_fasta(file_name):
   #first seq is reference, second seq in starting point, one struct only
   seq = []
+  struct = []
   with open(file_name) as f:
     for line in f:
       line = line.strip()
+      if not line:
+        continue
       if all(x in 'AUGC' for x in line):
         seq.append(line)
+        continue
       if all(x in '(.)' for x in line):
-        struct = line 
-  return seq[:-1],seq[-1], parseStruct(struct)
+        struct.append(line)
+        continue
+  return seq[:-1],seq[-1], parseStruct(struct[0])
 
 def all_probabilities(seq,ref_seq, stuct, m, alpha):
   n = len(seq)
@@ -647,6 +652,7 @@ if __name__ == "__main__":
 
 
   ref_seq,seq,struct = parse_fasta(file_name)
+  print ref_seq,seq,struct
   results = all_probabilities(seq,ref_seq,struct,m,alpha)
   display_all_probabilities(results)
 

### 2012-10-04 src/RNAPyroEx.py
fix iso error, added in prob count

@@ -350,12 +350,12 @@ def delta(seq,i,c):
   else:
     return 1
 
-def energy((a,b),(a2,b2),alpha=1.0):
+def energy((a,b),(a2,b2),alpha):
   #stacking energy of base pair (a,b) around base pair (a2,b2)
   E = STACKING_ENERGY[a,a2,b2,b]
   return  math.exp(-(alpha*E)/(BOLTZMANN*T))
 
-def isostericity(seq,ref_seq,(i,j),(a,b), alpha=1.0):
+def isostericity(seq,ref_seq,(i,j),(a,b), alpha):
   #isostericity of going from original base pair to (a,b)
   iso_mut = sum(ISO[(ref[i],ref[j]),(a,b)] for ref in ref_seq)
   iso_start = sum(ISO[(ref[i],ref[j]),(seq[i],seq[j])] for ref in ref_seq)
@@ -363,7 +363,7 @@ def isostericity(seq,ref_seq,(i,j),(a,b), alpha=1.0):
   return  math.exp(-((1-alpha)*iso)/(BOLTZMANN*T))
 
 @memoize
-def forward(seq,ref_seq,struct,(i,j),(a,b),m, alpha=1.0):
+def forward(seq,ref_seq,struct,(i,j),(a,b),m, alpha):
   #alpha gives the weight energy vs isostericity
   result = 0.
   if m<0: return 0
@@ -381,7 +381,7 @@ def forward(seq,ref_seq,struct,(i,j),(a,b),m, alpha=1.0):
           result += forward(seq,ref_seq,struct,
                             (i+1,j),
                             (a2,b),
-                            m-d)
+                            m-d,alpha)
     elif i < k <= j: #If k > j we return 0
       for a2 in BASES:
         for b2 in BASES:
@@ -392,10 +392,10 @@ def forward(seq,ref_seq,struct,(i,j),(a,b),m, alpha=1.0):
               result += forward(seq,ref_seq,struct,
                                   (i+1,k-1),
                                   (a2,b2),
-                                  m2)*forward(seq,ref_seq,struct,
+                                  m2,alpha)*forward(seq,ref_seq,struct,
                                   (k+1,j),
                                   (b2,b),
-                                  m-m2-d)*isostericity(seq,ref_seq,
+                                  m-m2-d,alpha)*isostericity(seq,ref_seq,
                                                        (i,k),
                                                        (a2,b2),
[…]

### 2012-10-04 src/RNAPyroEx.py
added mpf for iso output

@@ -359,7 +359,7 @@ def isostericity(seq,ref_seq,(i,j),(a,b), alpha):
   #isostericity of going from original base pair to (a,b)
   iso_mut = sum(ISO[(ref[i],ref[j]),(a,b)] for ref in ref_seq)
   iso_start = sum(ISO[(ref[i],ref[j]),(seq[i],seq[j])] for ref in ref_seq)
-  iso = (iso_mut-iso_start)/len(ref_seq)
+  iso = mpf(iso_mut-iso_start)/len(ref_seq)
   return  math.exp(-((1-alpha)*iso)/(BOLTZMANN*T))
 
 @memoize

### 2012-10-04 Recomb/methods_RECOMB.tex
fixed probs in tex
formalized EI

@@
  if the base pair $ab$ is more isosteric to the set $\Omega$ than the one on the sequence 
- $\omega$. Where the isostericity values 
- between any pair of canonical base pairs is  as defined in~\cite{Stombaugh2009}. 
+ $\omega$. Formally, given $s$, $\Omega$, two positions $(i,S_i)$ and two nucleotides $a,b\in B$:
+ $$
+ 	\text{EI}^{\Omega}_{(i,j),ab}:=
+	\frac{
+		\displaystyle
+		\sum_{s'\in\Omega}
+		\left(
+			\text{ISO}((s_i',s_j'),(a,b))
+		\right)-
+		\left(
+			\text{ISO}((s_i',s_j'),(s_i,s_j))		
+		\right)
+	}{
+		\displaystyle
+		|\Omega|
+	}
+ $$
+  Where $\text{ISO}((a',b'),(a,b))$ is the isostericity values 
+ between a the canonical base pairs $(a',b')$ and $(a,b)$  as defined in~\cite{Stombaugh2009}. 
  Let be $\alpha\in[0,1]$, it 
@@
 			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
+     	 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(i,k),xb}}{RT}}
 			\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
@@
 			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
+     	 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(k,i),bx}}{RT}}
 			\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
@@
 base pair, and the \emph{inside}, for all possible values for the nucleotide at position $k$, and
-all possible distribution of the mutant positions between the inside and outside of the base pair.
+all possible distribution of the mutant positions between the inside and outside of the base pair. We also add the term of isostericity for this specific base pair.
 \item[$S_i=k<i$:] Same as above, but with $i$ pairing with a lower position.

### 2012-10-04 Recomb/methods_RECOMB.tex
advanced method

@@
  $$
-  Where $\text{ISO}((a',b'),(a,b))$ is the isostericity values 
- between a the canonical base pairs $(a',b')$ and $(a,b)$  as defined in~\cite{Stombaugh2009}. 
+  Where $\text{ISO}((a',b'),(a,b))$ is the isostericity value 
+ between  the canonical base pairs $(a',b')$ and $(a,b)$  as defined in~\cite{Stombaugh2009}. 
  Let be $\alpha\in[0,1]$, it 
- will be used to balance the weight given to the stacking energy or the isostericity.
+ will be used to balance the weight given to the stacking energy and the isostericity.
 

### 2012-10-05 Recomb/methods_RECOMB.tex
method

@@
 \label{sec:methods}
-Let $\Omega$ be an un-gapped RNA alignment, $S$ its associate secondary structure, $s$ an RNA 
+Let $\Omega$ be an un-gapped RNA alignment, $S$ its associated secondary structure, $s$ an RNA 
 sequence and $m\geq 0$. $S$ is considered as  one derivation of the SCFG generating  
-all possible secondary structures of the same length as $s$. We define a variant of the
- \emph{Inside-Outside algorithm}~\cite{Lari1990} with the two following functions,
+all possible secondary structures of length  $s$. We are interested in  
+the probability of a given position being a specific nucleotides,
+over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$ (i.e.
+$\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
+We define a variant of the
+ \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
+ desired probability, with the two following functions,
 $\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former, as the 
@@
 the average isostericity with $\Omega$, as defined in~\cite{Stombaugh2009}.
-
-This allows us to compute the probability of any position being a specific nucleotides,
-over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$.
 
@@
 secondary structure $S$, and we return $0$.
-
 \end{description}
@@
 We define function $\Y{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
-
 $$

### 2012-10-05 Recomb/methods_RECOMB.tex
methods

@@
 sequence and $m\geq 0$. $S$ is considered as  one derivation of the SCFG generating  
-all possible secondary structures of length  $s$. We are interested in  
+all  secondary structures of length  $|s|$. We are interested in  
 the probability of a given position being a specific nucleotides,
-over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$ (i.e.
+over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$ 
+(formally 
 $\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
-We define a variant of the
+We define as a variant of the
  \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
- desired probability, with the two following functions,
+ desired probability,  the two  functions
 $\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former, as the 
 equivalent of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
-knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing $m$ mutations. The latter computes the \emph{outside}, in particular, knowing 
-that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing outside,
-including $i$ and $j$, $m$ mutations.
+knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing $m$ mutations. The latter computes the \emph{outside}, in particular,  
+the partition function considering only between $[0,i]\cup[j,n-1]$
+knowing  that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing 
+ $m$ mutations outside.
+
 The Boltzmann weights are a combination of the base pairs stacking energy, 
 using as values those of the NNDB~\cite{Turner2010}, and
-the average isostericity with $\Omega$, as defined in~\cite{Stombaugh2009}.
+the average isostericity difference
+between the mutant and $\Omega$, and $s$ with $\Omega$, using the isostericity values
+ as defined in~\cite{Stombaugh2009}. The value of $10$ was used for the isostericity
+ of any base pair 
+ compared with $GG$ given that the latter base pair was not
+  found in the~\cite{Stombaugh2009} tables,
+ and the range of values is $[0,9.7]$, $0$ for a perfect isostericity.
 

### 2012-10-05 Recomb/methods_RECOMB.tex
cleaned energy model

@@
 \subsection{Energy Model}
-The energy we will be composed of two function, $\text{ES}^{\beta}_{ab\to a'b'}$ and 
+The energy used is composed of two function, $\text{ES}^{\beta}_{ab\to a'b'}$ and 
 $\text{EI}^{\Omega}_{(i,j),ab}$. The former is equal to the 
@@
 penalize it.
-$\text{EI}^{\Omega}_{(i,j), a'b'}$ is the average of the sum of differences between the isostericity
-of base pairs at positions $(i,j)$ in $\Omega$ and $s_is_j$, and the isostericity of base pairs
- at positions $(i,j)$ in $\Omega$ and $ab$. If gives us an indication 
- if the base pair $ab$ is more isosteric to the set $\Omega$ than the one on the sequence 
- $\omega$. Formally, given $s$, $\Omega$, two positions $(i,S_i)$ and two nucleotides $a,b\in B$:
+$\text{EI}^{\Omega}_{(i,j), a'b'}$ is the average of the sum of differences between 
+the isostericity of base pairs
+ at positions $(i,j)$ in $\Omega$ and $ab$, and the isostericity
+of base pairs at positions $(i,j)$ in $\Omega$ and $s_is_j$. 
+It gives us an estimation of which  base pair, between $ab$ in the mutant sequence
+ and $s_is_j$, is in average more isosteric to $\Omega$.
+Formally, given $s$, $\Omega$, two positions $(i,S_i)$ and two nucleotides $a,b\in B$:
  $$
@@
  Let be $\alpha\in[0,1]$, it 
- will be used to balance the weight given to the stacking energy and the isostericity.
+ will be used to balance the weight given to the stacking energy and the isostericity, 
+ by considering from now on $\alpha\text{ES}^{\beta}_{ab\to a'b'}$ and 
+$(1-\alpha)\text{EI}^{\Omega}_{(i,j),ab}$.
 

### 2012-10-05 Recomb/methods_RECOMB.tex
method inside/outside

@@
 The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function considering only the 
-energy in subsequence $[i,j]$ over mutants of $s$ having exactly $m$ mutations between $[i,j]$ and whose nucleotide at position $i-1$ is $a$ (resp. in position $j+1$ it is $b$).
+energy in subsequence $[i,j]$ over mutants of $s$ having exactly $m$ mutations between $[i,j]$ and whose nucleotide at position $i-1$ is $a$ (resp. $b$ in position $j+1$).
 We define function $\Z{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
@@
 as if we increase the lower interval bound by $1$ (i.e. $i+1$), and consider all possible
- nucleotides $a'$ at position $i$. 
+ nucleotides $a'$ at position $i$, updating the number of mutants in function of $\Kron_{a',s_i}$. 
 \item[$S_i=j$ and $S_{i-1}=j+1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i-1$ is
@@
 to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
-structure between $(j,n-1)$ must be a set of independent stems. Thus, all the outside energy is
-equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself is the following:
+structure between $(j,n-1)$ must be a set of independent stems. Therefore,
+ all the outside energy between $[j,n-1]$ is
+equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself is as follows.
 $$
@@
 as if we decrease the lower interval bound by $1$ (i.e. $i-1$), and consider all possible
-nucleotides $a'$ at position $i$.
+nucleotides $a'$ at position $i$, correcting the number of mutants
+in function of $\Kron_{a',s_i}$.
 \item[$S_{i}=j$ and $S_{i+1}=j-1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i+1$ is
@@
 all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $k$ respectively.
-\item[$-1<S_{i}<i$:]As above but if the pairing is to the left.
+\item[$-1<S_{i}<i$:]As above but if position $i$ is to pairing with a lower value.
 \item[Else:] In all other cases, we are in a derivation of the SCFG that does not correspond to the 
@@
 all possible distribution of the mutant positions between the inside and outside of the base pair. We also add the term of isostericity for this specific base pair.
-\item[$S_i=k<i$:] Same as above, but with $i$ pairing with a lower position.
+\item[$S_i=k<i$:] Same as above, but with position $i$ pairing with a lower position.
 \end{description}

### 2012-10-05 Recomb/methods_RECOMB.tex
method

@@
  desired probability,  the two  functions
-$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former, as the 
-equivalent of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
-knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing $m$ mutations. The latter computes the \emph{outside}, in particular,  
+$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former presented in 
+Eq.~\ref{eq:Z_in} and Eq.~\ref{eq:Z_rec}, our 
+version of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
+knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing 
+$m$ mutations. The latter in Eq.~\ref{eq:Y_in} and Eq.\ref{eq:Y_rec},
+ computes the \emph{outside}, in particular,  
 the partition function considering only between $[0,i]\cup[j,n-1]$
@@
 
-\[
+\begin{equation}
 	\forall i \in (0,\cdots,n-1):\, \Z{i+1,i}{m}{a,b}=\left\{
@@
 	\end{array}\right.
-\]
+\label{eq:Z_in}
+\end{equation}
 In other words, when we evaluate the function $\mathcal Z$, after exhausting all positions, there 
@@
 The recursion itself is composed of four terms:
-$$
+\begin{equation}
 	\Z{i,j}{m}{a,b}:=\left\{
@@
 	\end{array}\right.
-$$
+\label{eq:Z_rec}
+\end{equation}
 The cases can be broked down as follows.
@@
 We define function $\Y{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
-$$
+\begin{equation}
 	\Y{-1,j}{m}{X,X}:=
@@
 	  \Z{j,n-1}{m}{X,X}
-$$
+\label{eq:Y_in}
+\end{equation}
 The recurrence, as shown below, will increase the interval $[i,j]$ by decreasing $i$ when
@@
 equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself is as follows.
-$$
+\begin{equation}
 	\Y{i,j}{m}{a,b} = \left\{
@@
   \end{array}\right.
-$$
+\label{eq:Y_rec}
+\end{equation}
 The five cases can be broked down as follows.
@@
 Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of mutations allowed. 
-$$
+\begin{equation}
 	\mathbb{P}(s_i = x\mid s,\Omega, S,M):=\left\{
@@
 	\end{array}\right.
-$$
+\end{equation}
 

### 2012-10-05 benchmark/runbenchmark.sh
fixes bug test

@@ -1,0 +1,9 @@
+python ../scripts/benchmark.py -i RF00001_0.ref -o RF00001_0_alpha_50.out
+python ../scripts/benchmark.py -i RF00001_1.ref -o RF00001_1_alpha_50.out
+python ../scripts/benchmark.py -i RF00001_2.ref -o RF00001_2_alpha_50.out
+python ../scripts/benchmark.py -i RF00001_3.ref -o RF00001_3_alpha_50.out
+python ../scripts/benchmark.py -i RF00001_4.ref -o RF00001_4_alpha_50.out
+python ../scripts/benchmark.py -i RF00001_5.ref -o RF00001_5_alpha_50.out
+python ../scripts/benchmark.py -i RF00001_6.ref -o RF00001_6_alpha_50.out
+python ../scripts/benchmark.py -i RF00001_7.ref -o RF00001_7_alpha_50.out
+python ../scripts/benchmark.py -i RF00001_8.ref -o RF00001_8_alpha_50.out

### 2012-10-05 benchmark/runtest.sh

@@ -1,0 +1,8 @@
+python ../src/RNAPyroEx.py RF00001_1.in 10 0.5 > RF00001_1_alpha_50.out
+python ../src/RNAPyroEx.py RF00001_2.in 10 0.5 > RF00001_2_alpha_50.out
+python ../src/RNAPyroEx.py RF00001_3.in 10 0.5 > RF00001_3_alpha_50.out
+python ../src/RNAPyroEx.py RF00001_4.in 10 0.5 > RF00001_4_alpha_50.out
+python ../src/RNAPyroEx.py RF00001_5.in 10 0.5 > RF00001_5_alpha_50.out
+python ../src/RNAPyroEx.py RF00001_6.in 10 0.5 > RF00001_6_alpha_50.out
+python ../src/RNAPyroEx.py RF00001_7.in 10 0.5 > RF00001_7_alpha_50.out
+python ../src/RNAPyroEx.py RF00001_8.in 10 0.5 > RF00001_8_alpha_50.out

### 2012-10-05 scripts/RfamParser.py

@@ -182,7 +182,7 @@ def clusterUnGapped(data):
 
 #################################################################################
 
-def main(filename,nmut,idc,listflag):
+def main(filename,nmut,idc,listflag,out):
 
     seqline_re = re.compile("(\S+)(\s+)(\S+)");
 
@@ -242,37 +242,53 @@ def main(filename,nmut,idc,listflag):
         sys.exit(1)
             
     hcode=idMap['num2code'][idc]
-    listout = []
-    listout.append(random.randint(0,len(cleandata[hcode])-1))
+    if out=="":
+        listout = []
+        listout.append(random.randint(0,len(cleandata[hcode])-1))
+    else:
+        listout = range(len(cleandata[hcode]))
+
     for iseqout in listout:
         ssamatch=True
         prevssa=''
         idout=''
         cmpt=0
-        #print '>>> cluster',idc,'size =',len(cleandata[hcode])
+
+        if out=="":
+            fout_in = sys.stdout
+            fout_ref = sys.stdout
+        else:
+            fout_in = open(out + "_" + str(iseqout) + ".in","w")
+            fout_ref = open(out + "_" + str(iseqout) + ".ref","w")
+        
         for myid,myseq in cleandata[hcode].iteritems():
             cleanseq,cleanssa = fitstruct2seq(myseq,info['consensus'],False)
             if cmpt!=iseqout:
-                print '>',myid
-                print cleanseq
+                print >>fout_in, '> ' + str(myid)
+                print >>fout_in, cleanseq
             else:
                 idout=myid
             prevssa=cleanssa
             cmpt+=1
             if prevssa != '' and prevssa != cleanssa:
                 ssamatch=False
-        print '> structure'
-        print prevssa
+        print >>fout_in, '> structure'
+        print >>fout_in, prevssa
+        if out!="":
+            print >>fout_ref, '> structure'
+            print >>fout_ref, prevssa
         if not ssamatch:
             print 'WARNING: Structure do not match'
 
-        print '>',idout,'(' + str(nmut) + '-mutant) read'
         cleanseq = cleandata[hcode][idout].replace('.','')
         mutant = mutate(cleanseq,nmut)
-        print mutant
-
-        print '>',idout,'target'
-        print cleanseq
[…]

### 2012-10-05 scripts/benchmark.py

@@ -253,7 +253,7 @@ def basiccorrelation(mutlist,data):
 
 ## read infile
 
-def fullcorrelation(mutlist,indata,outdata):
+def fullcorrelation(indata,outdata,withread):
     
     read=indata['read'].replace('.','').replace('-','').upper()
     original=indata['target'].replace('.','').replace('-','').upper()
@@ -266,19 +266,32 @@ def fullcorrelation(mutlist,indata,outdata):
     badscore=0.
     cmptgood=0
     cmptbad=0
+    scores={'TP':0.0,'TN':0.0,'FP':0.0,'FN':0.0}
+    counter={'TP':0,'TN':0,'FP':0,'FN':0}
     for i in range(len(read)):
+        #print i,(outdata[i]['A']+outdata[i]['C']+outdata[i]['G']+outdata[i]['U'])
         for nt in ['A','C','G','U']:
-            if nt!=original[i]:
-                if nt!=read[i]:
-                    badscore+=outdata[i][nt]
-                    cmptbad+=1
+            if nt!=read[i] or withread:
+                if nt==original[i]:
+                    # nt is the good one (T)
+                    if original[i]!=read[i]: # this is a mutation site
+                        scores['TP']+=outdata[i][nt]
+                        counter['TP']+=1
+                    else: # nt is NOT a mutation site (N)
+                        scores['TN']+=outdata[i][nt]
+                        counter['TN']+=1
                 else:
-                    goodscore+=outdata[i][nt]
-                    cmptgood+=1
+                    # nt is not correct (F)
+                    if original[i]!=read[i]: # nt is a mutation site(P)
+                        scores['FP']+=outdata[i][nt]
+                        counter['FP']+=1
+                    else:
+                        # nt is NOT a mutation site (N)
+                        scores['FN']+=outdata[i][nt]
+                        counter['FN']+=1
 
-    print '# sum scores',goodscore+badscore
-    return goodscore/cmptgood,badscore/cmptgood
-
+    return scores,counter
+    
 
 ## read infile
 
@@ -304,7 +317,6 @@ def checkconsistency(mutlist,indata,outdata):
                 else:
                     goodscore+=outdata[i][nt]
                     cmptgood+=1
-    
     print goodscore,badscore,goodscore+badscore
 
 
@@ -343,19 +355,20 @@ def fullprediction(mutlist,indata,outdata,onlybpregion,verbose,customsteps):
         FN=0
[…]

### 2012-10-05 benchmark/runtest.py
improved scripts

@@ -1,0 +1,43 @@
+import sys,os,os.path
+
+def main(nmut,alpha,prefix):
+
+    for numex in range(10):
+        infile = "RF00001_%d.in" % numex
+        if os.path.exists(infile):
+            alphanorm = int (alpha * 100)
+            commandline1="python ../src/RNAPyroEx.py RF00001_%d.in %d %.1f > RF00001_%d_alpha_%d.out" % (numex,nmut,alpha,numex,alphanorm)
+            os.system(commandline1)
+            commandline1="python ../scripts/benchmark.py -i RF00001_%d.ref -o RF00001_%d_alpha_%d.out" % (numex,numex,alphanorm)
+            os.system(commandline1)
+
+
+
+###############################################################################
+
+if __name__ == '__main__':
+    
+    nmut=10
+    alpha=0.5
+    prefix="RF00001"
+    
+    if len(sys.argv)==1:
+        usage(sys.argv[0])
+    
+    for i in range(len(sys.argv)):
+        myarg=sys.argv[i]
+        if myarg=='-h' or myarg=='--help':
+            usage(sys.argv[0])
+        if myarg=='-n':
+            nmut=int(sys.argv[i+1])
+        if myarg=='-a':
+            alpha=float(sys.argv[i+1])
+        if myarg=='-p':
+            prefix=sys.argv[i+1]
+    
+    
+    main(nmut,alpha,prefix)
+
+
+
+

### 2012-10-05 scripts/benchmark.py

@@ -9,7 +9,7 @@ from pylab import *;
 ## usage
 
 def usage(softname):
-    print "%s [-cvs] -i <RNApyro infile> -o <RNApyro outfile>" % (softname)
+    print "%s [-cvsw] -i <RNApyro infile> -o <RNApyro outfile>" % (softname)
     sys.exit(1)
 
 ## compute bp set
@@ -322,7 +322,7 @@ def checkconsistency(mutlist,indata,outdata):
 
 ## read infile
 
-def fullprediction(mutlist,indata,outdata,onlybpregion,verbose,customsteps):
+def fullprediction(mutlist,indata,outdata,onlybpregion,verbose,customsteps,withread):
     
     read=indata['read'].replace('.','').replace('-','').upper()
     original=indata['target'].replace('.','').replace('-','').upper()
@@ -357,16 +357,17 @@ def fullprediction(mutlist,indata,outdata,onlybpregion,verbose,customsteps):
         for i in range(len(read)):
             if myssa[i]!='.' or onlybpregion: # mutation in a base pair
                 for nt in ['A','C','G','U']:
-                    if outdata[i][nt]>=threshold: # predicted (P)
-                        if nt==original[i]: # nt is correct (T)
-                            TP+=1
-                        else: # nt is wrong (F)
-                            FP+=1
-                    else: # not predicted (N)
-                        if nt!=original[i]: # nt is NOT the good nt (T)
-                            TN+=1
-                        else: # nt is the good one (F)
-                            FN+=1
+                    if withread or nt!=read[i]:
+                        if outdata[i][nt]>=threshold: # predicted (P)
+                            if nt==original[i]: # nt is correct (T)
+                                TP+=1
+                            else: # nt is wrong (F)
+                                FP+=1
+                        else: # not predicted (N)
+                            if nt!=original[i]: # nt is NOT the good nt (T)
+                                TN+=1
+                            else: # nt is the good one (F)
+                                FN+=1
                     
                     
         #print TP,TN,FP,FN
@@ -429,7 +430,7 @@ def fullprediction(mutlist,indata,outdata,onlybpregion,verbose,customsteps):
 
 ## main
 
-def main(infilename,outfilename,onlybpregion,verbose,mysteps):
[…]

### 2012-10-05 Recomb/methods_RECOMB.tex
Initial commit

@@
 \label{sec:methods}
+
+\subsection{Model}
+Our probabilistic model aims at capturing both the stability of the folded RNA and its ability to adopt a predefined 3D conformation.
+To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function which combines free-energy and isostericity contributions.
+
+The free-energy contribution corresponds to the stacking base-pairs free-erngies of the Turner XXXXX model, as retrieved from the NNDB~\cite{Turner2010}.
+This simplification of the Turner energy .
+The imprecision introduced by our simplification of the Turner model not only
+greatly eases the design of dynamic-programming equations but  also causes 
+a limited imprecision since the targeted secondary structure is fixed 
+(e.g. multiloops do not account for base-specific contributions).
+
+The isostericity term captures the superimposability of corresponding base-pairs in the initial and mutated structures,
+%$\Omega$, and $s$ with $\Omega$, 
+using isostericity values defined in~\cite{Stombaugh2009}. 
+The range of values is $[0,9.7]$, with $0$ standing for a perfect isostericity, 
+and $10$ being used for missing values in Stombaugh~\emph{et al}~\cite{Stombaugh2009} tables.
+The relative contribution of these terms in the pseudo-energy function is then adjusted using a parameter $\alpha\in[0,1]$.
+
 Let $\Omega$ be an un-gapped RNA alignment, $S$ its associated secondary structure, $s$ an RNA 
@@
  $m$ mutations outside.
-
-The Boltzmann weights are a combination of the base pairs stacking energy, 
-using as values those of the NNDB~\cite{Turner2010}, and
-the average isostericity difference
-between the mutant and $\Omega$, and $s$ with $\Omega$, using the isostericity values
- as defined in~\cite{Stombaugh2009}. The value of $10$ was used for the isostericity
- of any base pair 
- compared with $GG$ given that the latter base pair was not
-  found in the~\cite{Stombaugh2009} tables,
- and the range of values is $[0,9.7]$, $0$ for a perfect isostericity.
 

### 2012-10-05 Recomb/introduction_RECOMB.tex
intro start

@@
 \label{sec:introduction}
+
+A particular class of molecules,
+ribonucleic acids (RNAs) are  now  an ubiquitous class of molecules, being
+found in every living organisms and having a broad range of functions, from catalyzing
+chemical reactions as the RNase P or the group II introns,
+hybridizing  messenger RNA to regulate gene expression,
+to ribosomal RNA (rRNA) synthesizing protein.
+Those functions  require specific structures, 
+encoded in their nucleotide sequence. Although the functions
+need to be preserved through various organisms, and therefore
+their structure must be similar,  the sequences
+can greatly differ from one another.
+For half a century, biological molecules have been studied as a proxy to understand
+evolution~\cite{Zuckerkandl1965}, and with all their characteristics, rRNAs have
+become a prime candidate to study phylogeny~\cite{Olsen1986, Olsen1993}.

### 2012-10-05 Recomb/introduction_RECOMB.tex
intro

@@
 
-A particular class of molecules,
-ribonucleic acids (RNAs) are  now  an ubiquitous class of molecules, being
+Ribonucleic acids (RNAs) are  now  an ubiquitous class of molecules, being
 found in every living organisms and having a broad range of functions, from catalyzing
@@
 hybridizing  messenger RNA to regulate gene expression,
-to ribosomal RNA (rRNA) synthesizing protein.
+to ribosomal RNA (rRNA) synthesizing proteins.
 Those functions  require specific structures, 
@@
 become a prime candidate to study phylogeny~\cite{Olsen1986, Olsen1993}.
+
+In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007}, 
+leveraging the NGS techniques to sequence as many new organisms 
+as possible, are producing a wealth of new information. Although
+those techniques have a huge throughput, they yield a sequencing error rate of around
+$4\%$~\cite{Huse2007}. This error can be highly reduced in genome projects when highly 
+redundant consensus 
+assemblies are available, but in studies of new or not well known organisms, there is not
+ enough  similarity to differentiate between the sequencing errors and the natural 
+ polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
+  In rRNAs, we have as additional 
+ information the conserved secondary structure, and we want to use it to identify
+highly probable sequencing errors.

### 2012-10-05 Recomb/introduction_RECOMB.tex
intro

@@
 their structure must be similar,  the sequences
-can greatly differ from one another.
+can greatly differ from one organism to another.
 For half a century, biological molecules have been studied as a proxy to understand
 evolution~\cite{Zuckerkandl1965}, and with all their characteristics, rRNAs have
-become a prime candidate to study phylogeny~\cite{Olsen1986, Olsen1993}.
+become a prime candidate for phylogenetic studies~\cite{Olsen1986, Olsen1993}.
 
@@
 highly probable sequencing errors.
+
+The first challenge is to efficiently explore the mutant space, which grows exponentially. 
+Leveraging the techniques  in \texttt{RNAmutants}~\cite{Waldispuhl2008}, we develop 
+here

### 2012-10-05 benchmark/runtest.sh
improved and clean dataset construction (clustering requires cd-hit)

@@ -1,8 +1,0 @@
-python ../src/RNAPyroEx.py RF00001_1.in 10 0.5 > RF00001_1_alpha_50.out
-python ../src/RNAPyroEx.py RF00001_2.in 10 0.5 > RF00001_2_alpha_50.out
-python ../src/RNAPyroEx.py RF00001_3.in 10 0.5 > RF00001_3_alpha_50.out
-python ../src/RNAPyroEx.py RF00001_4.in 10 0.5 > RF00001_4_alpha_50.out
-python ../src/RNAPyroEx.py RF00001_5.in 10 0.5 > RF00001_5_alpha_50.out
-python ../src/RNAPyroEx.py RF00001_6.in 10 0.5 > RF00001_6_alpha_50.out
-python ../src/RNAPyroEx.py RF00001_7.in 10 0.5 > RF00001_7_alpha_50.out
-python ../src/RNAPyroEx.py RF00001_8.in 10 0.5 > RF00001_8_alpha_50.out

### 2012-10-05 Recomb/methods_RECOMB.tex
Reorganized method section

--- 2012-10-05 Recomb/methods_RECOMB.tex
@@
 \label{sec:methods}
-Let $\Omega$ be an un-gapped RNA alignment, $S$ its associated secondary structure, $s$ an RNA 
-sequence and $m\geq 0$. $S$ is considered as  one derivation of the SCFG generating  
+
+\newcommand{\PE}[1]{E(#1)}
+\newcommand{\EI}{\text{EI}}
+\newcommand{\ES}{\text{ES}}
+\newcommand{\ISO}{\text{ISO}}
+
+\subsection{Probabilistic model}
+Our probabilistic model aims at capturing both the stability of the folded RNA and its ability to adopt a predefined 3D conformation.
+To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both 
+the free-energy and the isostericity. 
+More precisely, let $\Omega$ be an gapless RNA alignment, $S$ its associated secondary structure, 
+then each sequence $s$ has probability proportional to its Boltzmann factor
+\begin{align*}
+  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT} &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
+\end{align*}
+where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ are the free-energy and isostericity contributiuons, further described below, and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
+
+\subsubsection{Energy contribution}
+The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner \todo[fancyline]{which version of the Turner model?}XXXXX model retrieved from the NNDB~\cite{Turner2010}.
+Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
+\begin{align*}
+  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
+\end{align*}
+where $\ES^{\beta}_{ab\to a'b'}$ is the tabulated free-energy of a stacking pair in the Turner model if available, $0$ if $ab=\varnothing$ (no base-pair to stack onto), and using default value $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
+not in $\{\text{GU},\text{UG},\text{CG},\text{GC}, \text{AU or UA}\}$).
+This latter parameter allows to choose whether to completely forbid a sequence inducing invalid base pairs ($\beta = \infty$), or to simply penalize it.
+The imprecision introduced by this simplification of the Turner model remains reasonable, since the targeted secondary structure is fixed 
+(e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
+\subsection{isostericity contribution}
+The concept of isostericity captures the geometric superimposability of two base-pairs, using individual scores given by Stombaugh~\emph{et al}~\cite{Stombaugh2009} based on a probabilistic model.
+Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gapless RNA alignment $\Omega$,  the contribution of the isostericity to the pseudo-energy is given by
+\begin{align*}
+  \ES(s,S,\Omega) &= \sum_{\substack{(i,j)\in S\\ \text{pairs}}}\EI^{\Omega}_{(i,j),s_i s_j}, & \text{where}&& 	\EI^{\Omega}_{(i,j),ab}:=
+	\frac{
+		\sum_{s'\in\Omega}
+			\text{\ISO}((s_i',s_j'),(a,b))}
+%-		\left(			\text{\ISO}((s_i',s_j'),(s_i,s_j))				\right)	
+{		
+		|\Omega|
+	}
+\end{align*}
+is the average isostericity of a base-pair in the candidate sequence in comparison with the reference alignment.
+The $\ISO$ function uses the \todo[fancyline, inline]{\relsize{-1}Is this true?
+If so, we might want to try taking the most likely edge/orientation combination for each couple instead.}{Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
+Isostericity scores range between $0$ and $9.7$, with $0$ being assigned to a perfect isostericity, and a penalty of $10$ is used for missing entries.
+The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the alignment sequences.
+
+\subsection{Mutational profiles of sequences}
+
+Let $s$ an RNA 
+sequence and $m\geq 0$.
+$S$ is considered as  one derivation of the \todo[fancyline]{A bit tough: SCFG's were never mentioned before.
+Maybe introduce this with a sentence relating problem to parsing}{SCFG} generating
 all  secondary structures of length  $|s|$. We are interested in  
-the probability of a given position being a specific nucleotides,
+the probability of a given position being a specific nucleotide,
 over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$ 
@@
 $\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former presented in 
-Eq.~\ref{eq:Z_in} and Eq.~\ref{eq:Z_rec}, our 
+Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec}, our 
 version of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
 knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing 
-$m$ mutations. The latter in Eq.~\ref{eq:Y_in} and Eq.\ref{eq:Y_rec},
+$m$ mutations. The latter in Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
  computes the \emph{outside}, in particular,  
@@
 
-The Boltzmann weights are a combination of the base pairs stacking energy, 
-using as values those of the NNDB~\cite{Turner2010}, and
-the average isostericity difference
-between the mutant and $\Omega$, and $s$ with $\Omega$, using the isostericity values
- as defined in~\cite{Stombaugh2009}. The value of $10$ was used for the isostericity
- of any base pair 
- compared with $GG$ given that the latter base pair was not
-  found in the~\cite{Stombaugh2009} tables,
- and the range of values is $[0,9.7]$, $0$ for a perfect isostericity.
-
-\subsection{Definitions}
+\subsubsection{Definitions}
 Let be $B:=\left\{\text{A},\text{C},\text{G},\text{U}\right\}$, the set of nucleotides.
@@
 															1 & \text{If } a\equiv a'\\
-															0 & \text{Else}
+															0 & \text{Otherwise.}
 														\end{array}\right.$
+ Let us finally denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
+\begin{equation}
+  E_{(i,j),ab \to a'b'}^{\Omega,\beta}  = \alpha \cdot\ES^\beta_{a b \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
+\end{equation}
 
-\subsection{Energy Model}
-The energy used is composed of two function, $\text{ES}^{\beta}_{ab\to a'b'}$ and 
-$\text{EI}^{\Omega}_{(i,j),ab}$. The former is equal to the 
-stacking energy of the base pair with nucleotides $ab$ on top of the base pair with nucleotides 
-$a'b'$, as set in the NNDB~\cite{Turner2010}. If one of the base pair is not valid (i.e. not in 
-$\{\text{GU},\text{UG},\text{CG},\text{GC}, \text{AU or UA}\}$, the value is a parameter 
-$\beta \in [1,\infty]$. This allows
-to completely forbid a sequence where a base pair is non valid, when $\beta = \infty$ or only 
-penalize it.
-$\text{EI}^{\Omega}_{(i,j), a'b'}$ is the average of the sum of differences between 
-the isostericity of base pairs
- at positions $(i,j)$ in $\Omega$ and $ab$, and the isostericity
-of base pairs at positions $(i,j)$ in $\Omega$ and $s_is_j$. 
-It gives us an estimation of which  base pair, between $ab$ in the mutant sequence
- and $s_is_j$, is in average more isosteric to $\Omega$.
-Formally, given $s$, $\Omega$, two positions $(i,S_i)$ and two nucleotides $a,b\in B$:
- $$
- 	\text{EI}^{\Omega}_{(i,j),ab}:=
-	\frac{
-		\displaystyle
-		\sum_{s'\in\Omega}
-		\left(
-			\text{ISO}((s_i',s_j'),(a,b))
-		\right)-
-		\left(
-			\text{ISO}((s_i',s_j'),(s_i,s_j))		
-		\right)
-	}{
-		\displaystyle
-		|\Omega|
-	}
- $$
-  Where $\text{ISO}((a',b'),(a,b))$ is the isostericity value 
- between  the canonical base pairs $(a',b')$ and $(a,b)$  as defined in~\cite{Stombaugh2009}. 
- Let be $\alpha\in[0,1]$, it 
- will be used to balance the weight given to the stacking energy and the isostericity, 
- by considering from now on $\alpha\text{ES}^{\beta}_{ab\to a'b'}$ and 
-$(1-\alpha)\text{EI}^{\Omega}_{(i,j),ab}$.
-
-\subsection{Inside}
+\subsubsection{Inside}
 The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function considering only the 
@@
 \begin{equation}
-	\forall i \in (0,\cdots,n-1):\, \Z{i+1,i}{m}{a,b}=\left\{
+	\forall i \in [0,n-1]:\, \Z{i+1,i}{m}{a,b}=\left\{
 	\begin{array}{ll}
 		1 &\text{If } m = 0\\
-		0 &\text{Else }
+		0 &\text{Otherwise.}
 	\end{array}\right.
@@
       \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
-			 e^{\frac{-(\alpha \text{ES}^\beta_{a b \to a' b'}+(1-\alpha)\text{EI}^{\Omega}_{(i,j),a'b'})}{RT}}
-			 \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
+			 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}
+			 \cdot \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
 			 \text{Elif }S_i=j \land S_{i-1}=j+1\\
@@
       \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-   		 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(i,k),a'b'}}{RT}}
-      \Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
-      \Z{k+1,j}{m'}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
-      0 &\text{Else}
+   		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
+      \cdot\Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
+      \cdot\Z{k+1,j}{m'}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
+      0 &\text{Otherwise}
 	\end{array}\right.
@@
 \end{equation}
-The cases can be broked down as follows.
+The cases can be broken down as follows:
 \begin{description}
@@
 all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $j$ respectively.
-\item[Else:] In all other cases, we are in a derivation of the SCFG that does not correspond to the 
+\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the 
 secondary structure $S$, and we return $0$.
@@
 
-\subsection{Outside}	
+\subsubsection{Outside}	
 The \emph{Outside} function, $\mathcal Y$, is the partition function considering only the 
@@
     \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
-		 e^{\frac{-(\alpha \text{ES}^\beta_{a b \to a' b'}+(1-\alpha)\text{EI}^{\Omega}_{(i,j),a'b'})}{RT}}
+		 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}\cdot
     \Y{i-1,j+1}{m- \Kron_{a'b',s_is_j}}{a',b'} &
@@
 		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-  		 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(i,k),a'b'}}{RT}}
-		 \Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
-     \Z{j,k-1}{m'}{b,b'} &
+  		 e^{\frac{E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
+		 \cdot\Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
+     \cdot\Z{j,k-1}{m'}{b,b'} &
 		 \text{Elif }S_{i}=k \geq j\\
@@
 		 \sum_{m'=0}^{m-\Kron_{a'b',s_ks_i}}
-   	 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(k,i),a'b'}}{RT}}
-		 \Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
-     \Z{k+1,i-1}{m'}{a',b'} &
+   	 e^{\frac{E_{(k,i),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
+		 \cdot\Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
+     \cdot\Z{k+1,i-1}{m'}{a',b'} &
 		 \text{Elif }-1 < S_{i}=k < i\\
@@
 
-\section{Inside-Outside}
+\subsubsection{Inside-Outside}
 By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
@@
  for any nucleotide $X\in B$ or
-in function of the \emph{backward} term, for any position $k$ such that $S_k=-1$:
+in function of the \emph{backward} term, for any unpaired position $k$:
 $$
@@
 
-We are now interested in knowing, under our model, 
- the probability that a given position is a given nucleotide.
+We are now left to compute the probability that a \todo[fancyline]{Maybe worth doing it for base-pairs? maybe not\ldots}{given position is a given nucleotide}.
 We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
-Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of mutations allowed. 
-\begin{equation}
-	\mathbb{P}(s_i = x\mid s,\Omega, S,M):=\left\{
+Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations. 
+\begin{align*}
+	\mathbb{P}(s_i = x\mid s,\Omega, S,M) &:= \frac{\mathcal{B}(i,x,s,\Omega,S,M)}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\\ 
+\mathcal{B}(i,x,s,\Omega,S,M)&=
+ \left\{
 	\begin{array}{ll}
-		\displaystyle
-		\frac{
-			\displaystyle
 			\sum_{m=0}^{M}
 			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
-		}{
-			\displaystyle
+		&\text{If }S_i = -1\\
 			\sum_{m=0}^{M}
-			\Z{0,n-1}{m}{X,X}
-		}
-		&\text{If }S_i = -1\\
-		\displaystyle
-		\frac{
-			\displaystyle
-			\sum_{m=0}^{M}
-			\sum_{\substack{b\in Bases\\\Kron_{xb,s_is_k\leq m}}}
+			\sum_{\substack{b\in B\\\Kron_{xb,s_is_k\leq m}}}
 			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
-     	 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(i,k),xb}}{RT}}
+     	 e^{\frac{E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
 			\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
 			\Z{i+1,k-1}{m'}{x,b}
-		}{
-			\displaystyle
+		&\text{If }S_i=k>i\\
 			\sum_{m=0}^{M}
-			\Z{0,n-1}{m}{X,X}
-		}
-		&\text{If }S_i=k>i\\
-		\displaystyle
- 		\frac{
-			\displaystyle
-			\sum_{m=0}^{M}
-			\sum_{\substack{b\in Bases\\\Kron_{bx,s_ks_i\leq m}}}
+			\sum_{\substack{b\in B\\\Kron_{bx,s_ks_i\leq m}}}
 			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
-     	 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(k,i),bx}}{RT}}
+     	 e^{E_{(k,i),\varnothing\to bx}{RT}}
 			\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
 			\Z{k+1,i-1}{m'}{b,x}
-		}{
-			\displaystyle
-			\sum_{m=0}^{M}
-			\Z{0,n-1}{m}{X,X}
-		}
 		&\text{If }S_i=k<i
 	\end{array}\right.
-\end{equation}
+\end{align*}
 
-In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
+In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, \todo[fancyline,inline]{\relsize{-1}Seems weird to do this for a given upper bound $M$ (as opposed to exact $m$), and it makes the equations a bit more complicated.
+Anyway, there is roughly $3n$ times more sequences with $m+1$ mutations than with $m$ muts, so I expect the $m+1$ guys are so dominating in the final probabilities that it may not be worth the trouble of combining them with the \emph{little guys} ($\#muts < m$)}{for $m$ smaller or equal to our target $M$}.
+The numerators are divided in the following three cases.
 \begin{description}

### 2012-10-05 Recomb/introduction_RECOMB.tex
RNApyro

@@
 those techniques have a huge throughput, they yield a sequencing error rate of around
-$4\%$~\cite{Huse2007}. This error can be highly reduced in genome projects when highly 
-redundant consensus 
-assemblies are available, but in studies of new or not well known organisms, there is not
+$4\%$~\cite{Huse2007}. This error can be highly reduced  when highly 
+redundant multiple sequence allignments 
+ are available, but in studies of new or not well known organisms, there is not
  enough  similarity to differentiate between the sequencing errors and the natural 
  polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
-  In rRNAs, we have as additional 
+  In rRNAs, on top of multiple sequence alignments, we have as additional 
  information the conserved secondary structure, and we want to use it to identify
 highly probable sequencing errors. 
+Given a new rRNA sequence, a first challenge to find the sequencing errors
+  is to  efficiently explore the mutant space, 
+which grows exponentially, guiding ourselves with the alignment and the consensus structure.
 
-The first challenge is to efficiently explore the mutant space, which grows exponentially. 
-Leveraging the techniques  in \texttt{RNAmutants}~\cite{Waldispuhl2008}, we develop 
-here
+Leveraging the techniques  in \texttt{RNAmutants}~\cite{Waldispuhl2008}, and building on top 
+of the \emph{Inside-Outside algorithm}, we define here a new method called \texttt{RNApyro}
+efficiently computing for large RNAs those probabilities.
+ Classical techniques define a probabilistic model using a Boltzmann distribution 
+whose weights are based on the free energy of the structure, using as energy parameter
+the values of Turner found in the NNDB~\cite{Turner2010}  for stacked, 
+canonical and wobble, base pairs. As shown by Leontis and Westhof~\cite{Leontis2001},
+this  does not encapsulated the large diversity of base pairs that any nucleotide
+can form with any other, although with an energy too small to be yet determined
+by experimental techniques. To quantify geometrical differences, they
+ define an isostericity distance, increasing as two base pairs differ 
+ more from one another in space. We incorporate this second measure in the Boltzmann weights.
+
+To benchmark

### 2012-10-05 Recomb/methods_RECOMB.tex
Quelques typos et suppression des fichiers temporaires
fixed typos
RNApyro

@@
 \subsubsection{Energy contribution}
-The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner \todo[fancyline]{which version of the Turner model?}XXXXX model retrieved from the NNDB~\cite{Turner2010}.
+The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
 Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
@@
 		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-  		 e^{\frac{E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
+  		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
 		 \cdot\Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
@@
 		 \sum_{m'=0}^{m-\Kron_{a'b',s_ks_i}}
-   	 e^{\frac{E_{(k,i),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
+   	 e^{\frac{-E_{(k,i),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
 		 \cdot\Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
@@
 			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
-     	 e^{\frac{E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
+     	 e^{\frac{-E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
 			\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
@@
 			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
-     	 e^{E_{(k,i),\varnothing\to bx}{RT}}
+     	 e^{\frac{-E^{\Omega, \beta}_{(k,i),\varnothing\to bx}}{RT}}
 			\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}

### 2012-10-06 Recomb/introduction_RECOMB.tex
intro

@@
 \label{sec:introduction}
+
+In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007}, 
+leveraging the NGS techniques to sequence as many new organisms 
+as possible, are producing a wealth of new information. Although
+those techniques have a huge throughput, they yield a sequencing error rate of around
+$4\%$~\cite{Huse2007}. This error can be highly reduced  when highly 
+redundant multiple sequence alignments 
+ are available, but in studies of new or not well known organisms, there is not
+ enough  similarity to differentiate between the sequencing errors and the natural 
+ polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
 
@@
 
-In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007}, 
-leveraging the NGS techniques to sequence as many new organisms 
-as possible, are producing a wealth of new information. Although
-those techniques have a huge throughput, they yield a sequencing error rate of around
-$4\%$~\cite{Huse2007}. This error can be highly reduced  when highly 
-redundant multiple sequence allignments 
- are available, but in studies of new or not well known organisms, there is not
- enough  similarity to differentiate between the sequencing errors and the natural 
- polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
-  In rRNAs, on top of multiple sequence alignments, we have as additional 
- information the conserved secondary structure, and we want to use it to identify
-highly probable sequencing errors. 
-Given a new rRNA sequence, a first challenge to find the sequencing errors
-  is to  efficiently explore the mutant space, 
-which grows exponentially, guiding ourselves with the alignment and the consensus structure.
+In this paper, we tackle the problem of exploring an rRNA sequence mutant space, to
+identify positions which are probably sequencing error, given the  rRNA family sequences
+and its consensus structure.
 

### 2012-10-06 Recomb/introduction_RECOMB.tex
intro

@@
 
-In this paper, we tackle the problem of exploring an rRNA sequence mutant space, to
-identify positions which are probably sequencing error, given the  rRNA family sequences
-and its consensus structure.
+In this paper, we tackle the problem of, given an rRNA sequence its family and 
+consensus secondary structure, 
+identifing the positions which are probably sequencing error.
 
@@
 of the \emph{Inside-Outside algorithm}, we define here a new method called \texttt{RNApyro}
-efficiently computing for large RNAs those probabilities.
+efficiently computing for large RNAs those probabilities under a new 
+pseudo-energetic model.
  Classical techniques define a probabilistic model using a Boltzmann distribution 
@@
 
-To benchmark
+To benchmark our technique we use the 5S ribosomal RNA family,
+ which is around $120$ nucleotides longs. We show

### 2012-10-06 Recomb/introduction_RECOMB.tex
intro

@@
 \label{sec:introduction}
-
-In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007}, 
-leveraging the NGS techniques to sequence as many new organisms 
-as possible, are producing a wealth of new information. Although
-those techniques have a huge throughput, they yield a sequencing error rate of around
-$4\%$~\cite{Huse2007}. This error can be highly reduced  when highly 
-redundant multiple sequence alignments 
- are available, but in studies of new or not well known organisms, there is not
- enough  similarity to differentiate between the sequencing errors and the natural 
- polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
 
@@
 evolution~\cite{Zuckerkandl1965}, and with all their characteristics, rRNAs have
-become a prime candidate for phylogenetic studies~\cite{Olsen1986, Olsen1993}.
+always been a prime candidate for phylogenetic studies~\cite{Olsen1986, Olsen1993}.
 
-In this paper, we tackle the problem of, given an rRNA sequence its family and 
-consensus secondary structure, 
-identifing the positions which are probably sequencing error.
+In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007}, 
+leveraging NGS techniques to sequence as many new organisms 
+as possible, are producing a wealth of new information. Although
+those techniques have a huge throughput, they yield a sequencing error rate of around
+$4\%$~\cite{Huse2007}. This error can be highly reduced  when highly 
+redundant multiple sequence alignments 
+ are available, but in studies of new or not well known organisms, there is not
+ enough  similarity to differentiate between the sequencing errors and the natural 
+ polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
+
+In this paper, we hypothesize that the family and consensus secondary structure hold 
+information allowing to identify the positions most probable to be sequencing errors.
 

### 2012-10-06 benchmark/runtest.py
update script & add results section (partial)

@@ -2,13 +2,15 @@ import sys,os,os.path
 
 def main(nmut,alpha,prefix):
 
-    for numex in range(10):
+    for numex in range(100):
         infile = "RF00001_%d.in" % numex
         if os.path.exists(infile):
             alphanorm = int (alpha * 100)
-            commandline1="python ../src/RNAPyroEx.py RF00001_%d.in %d %.1f > RF00001_%d_nmut_%d_alpha_%d.out" % (numex,nmut,alpha,numex,nmut,alphanorm)
+            outfile = "RF00001_%d_nmut_%d_alpha_%d.out" % (numex,nmut,alphanorm)
+            commandline1="python ../src/RNAPyroEx.py %s %d %.1f 1.5 > %s" % (infile,nmut,alpha,outfile)
             print commandline1
-            os.system(commandline1)
+            if not os.path.exists(outfile):
+                os.system(commandline1)
             commandline2="python ../scripts/benchmark.py -i RF00001_%d.ref -o RF00001_%d_nmut_%d_alpha_%d.out" % (numex,numex,nmut,alphanorm)
             print commandline2
             os.system(commandline2)
@@ -22,14 +24,11 @@ if __name__ == '__main__':
     alpha=0.5
     prefix="RF00001"
     
-    if len(sys.argv)==1:
-        usage(sys.argv[0])
-    
     for i in range(len(sys.argv)):
         myarg=sys.argv[i]
         if myarg=='-h' or myarg=='--help':
             usage(sys.argv[0])
-        if myarg=='-n':
+        if myarg=='-m':
             nmut=int(sys.argv[i+1])
         if myarg=='-a':
             alpha=float(sys.argv[i+1])

### 2012-10-06 Recomb/results_RECOMB.tex
started recomb
mpmath citation

--- 2012-10-03 Recomb/introduction_RECOMB.tex
@@
 %!TEX root = main_RECOMB.tex
-\section{Introduction}
-\label{sec:introduction}
+\section{Results}
+\label{sec:results}
+
+\subsection{Implementation}
+The software was implemented in Python2.7 using the \textit{mpmath}~\cite{mpmath} library
+for  arbitrary floating point precision. The source code is freely available at \verb+https://github.com/McGill-CSB/RNApyro+.
+
+\subsection{Error correction in 5s rRNA}
+
+To illustrate the potential of our algorithm, we applied our techniques to identify and correct point-wise errors in RNA sequences
+with conserved secondary structures. More precisely, we used \RNApyro to reconstruct 5s rRNA sequences with randomly distributed
+mutations. This experiment has been designed to suggest further applications to error-corrections in pyrosequencing data.
+
+We build our data set from the 5S rRNA multiple sequence alignment (MSA) available in the Rfam Database 11.0 (Rfam id: \texttt{RF00001}).
+Since our software does not currently implement gaps (mainly because scoring indels is a challenging issue that cannot be fully addressed
+in this work),  we clustered together the sequences with identical gap locations. From the $54$ MSAs without gap produced, we selected the
+biggest MSA  which contains $130$ sequences (out of $712$ in the original Rfam MSA). Then, in order to avoid any bias, we used \texttt{cd-hit}
+\cite{CDHIT} to remove sequences with more than 80\% of sequence similarity. This operation resulted in a data set of $45$ sequences.
+
+We design our benchmark using a leave-one-out strategy. We randomly picked one sequence from our data set and performed $12$ random
+mutations. Our sequences have $119$ nucleotides, thus the number of mutations corresponds to an error-rate of 10\%. We repeated this operation 
+$10$ times.
+
+To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed by
+\RNApyro. We reported in Table \ref{tab:benchmark} the area under the curve (AUC). More specifically, we fix a threshold $\lambda \in [0,1]$ and we
+predict an error at position $i$ in sequence $\omega$ if and only if the probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceed this threshold.
+The set of corrections is thus $\{ n \; | \;  n \in \{ A,C,G,U \} \mbox{ and } P(i,n) > \lambda \mbox{ and }  n \neq \omega[i] \}$, where $\omega[i]$ is the
+nucleotide at position $i$ in the input sequence. Then, we progressively vary $\lambda$ between $0$ and $1$ to calculate the ROC curve and the AUC.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+& 1.0 & 0.5 & 0. \\
+\hline
+6 & 0.69 & 0.72 & 0.74 \\
+12 & 0.80 & 0.84 & 0.84 \\
+24 & 0.76 & & \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Performance of error-correction. The row index indicates the number of mutations performed with \RNApyro, and the column index indicates
+the value of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores. In each cell, we report the average AUC
+values over the 10 experiments.}
+\label{tab:benchmark}
+\end{table}

### 2012-10-06 Recomb/introduction_RECOMB.tex

@@
 
-To benchmark our technique we use the 5S ribosomal RNA family,
- which is around $120$ nucleotides longs. We show
+The 5S ribosomal RNA family, $119$ nucleotides long,  was used to benchmark our method.
+It is a prime example since it has been extensively used for phylogenetic
+reconstructions~\cite{Hori1987} and its sequence has been recovered for over 8000 species
+ (RFAM Id: \texttt{RF00001}).
+ Using a leave one out strategy, we perform random distributed mutations on a sequence. We show that
+\texttt{RNApyro} reconstructs the original sequence with an excellent accuracy.
+
+The pseudo-energetic model and the algorithm is presented in Sec.~\ref{sec:methods}.
+Details of the implementation and benchmarks are in Sec.~\ref{sec:results}. 
+Future applications and a discussion are developed in Sec.~\ref{sec:conclusion}.

### 2012-10-07 Recomb/results_RECOMB.tex
new roc figs
time

@@
 To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed by
-\RNApyro. We reported in Table \ref{tab:benchmark} the area under the curve (AUC). More specifically, we fix a threshold $\lambda \in [0,1]$ and we
+\RNApyro. We report in Table \ref{tab:benchmark} the area under the curve (AUC). More specifically, we fix a threshold $\lambda \in [0,1]$ and we
 predict an error at position $i$ in sequence $\omega$ if and only if the probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceed this threshold.
@@
 12 & 0.80 & 0.84 & 0.84 \\
-24 & 0.76 & & \\
+24 & 0.76 & 0.84 & 0.84 \\
 \hline
@@
 \end{table}
+
+\begin{figure}
+\begin{subfigure}[b]{0.3\textwidth}
+\includegraphics[width=1.2\textwidth]{figures/ROC_6.pdf}
+\caption{6 mutations}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.3\textwidth}
+\includegraphics[width=1.2\textwidth]{figures/ROC_12.pdf}
+\caption{12 mutations}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.3\textwidth}
+\includegraphics[width=1.2\textwidth]{figures/ROC_24.pdf}
+\caption{24 mutations}
+\end{subfigure}
+\end{figure}
+
+\subsection{Time}
+The applications for this algorithms implies a need for efficiency
+and scalability. In Table~\ref{tab:time} we
+present different times needed to compute the probabilities for
+ every nucleotide at every positions for a vast set of parameters.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{llccc}
+$|s|$&$|\Omega|$&\multicolumn{3}{c}{Number of mutations}\\\cline{3-5}
+		 & 				  & 6   &  12  & 24\\\cline{3-5}
+100  & 44				& 35s  & 238s & 1023s\\
+300  & 44				& 135s & 594s &2460s\\
+
+\end{tabular}
+\end{center}
+\caption{Time to compute all probabilities. The first column indicates the length, the second
+ the size of $\Omega$ and  the column indexes indicate the number
+ of mutations. $\alpha$ is
+set at $0.5$, and $\beta=1.5$ and All sequences are generated randomly.}
+\label{tab:time}
+\end{table}

### 2012-10-07 Recomb/methods_RECOMB.tex
Some additional typos

@@
 \subsubsection{Inside}
+\begin{figure}
+\resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
+\caption{Illustration of the principle underlying the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
+can be decomposed as a sequence over $[i+1,j]$ preceded by a, possibly mutated, base at $i$ 
+(Unpaired case), a sequence over $[i+1,j-1]$ surrounded by some base-pair (Stacking-pair case), 
+or as two sequences over $[i+1,k-1]$ and $[k+1,j]$, completed by some base-pair (General base-pairing case). 
+In each case, one has to investigate any possible ways to distribute mutations between the different sequences 
+and locally instanciated bases.}
+\end{figure}
+
 The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function considering only the 
@@
 \subsubsection{Outside}	
+\begin{figure}
+\resizebox{\textwidth}{!}{\input{FigDPOutside.pgf}}
+\caption{Principle of the outside computation.}
+\end{figure}
 The \emph{Outside} function, $\mathcal Y$, is the partition function considering only the 
-energy in subsequences $[0,i]\cup[j,n-1]$ over the mutants of $s$ having exactly $m$ mutations between $[0,i]\cup[j,n-1]$ and whose nucleotide at position $i+1$ is $a$ 
+contributions of subsequences $[0,i]\cup[j,n-1]$ over the mutants of $s$ having exactly $m$ mutations between $[0,i]\cup[j,n-1]$ and whose nucleotide at position $i+1$ is $a$ 
 (resp. in position $j-1$ it is $b$).
@@
 		 \text{Elif }-1 < S_{i}=k < i\\
-		 0 & \text{Else}
+		 0 & \text{Otherwise}
   \end{array}\right.

### 2012-10-07 Recomb/results_RECOMB.tex
time
intro

@@
 present different times needed to compute the probabilities for
- every nucleotide at every positions for a vast set of parameters.
+ every nucleotide at every positions for a vast set of parameters. For those test
+ all the sequences were generated randomly, as the target secondary structure.
 
@@
 \begin{center}
-\begin{tabular}{llccc}
-$|s|$&$|\Omega|$&\multicolumn{3}{c}{Number of mutations}\\\cline{3-5}
-		 & 				  & 6   &  12  & 24\\\cline{3-5}
-100  & 44				& 35s  & 238s & 1023s\\
-300  & 44				& 135s & 594s &2460s\\
-
+\begin{tabular}{lccc}
+$|s|$&\multicolumn{3}{c}{Number of mutations}\\\cline{2-4}
+		 			  & 6   &  12  & 24\\\cline{2-4}
+100  				& 35s  & 238s & 1023s\\
+300  			& 135s & 594s &2460s\\\cline{2-4}
+		 						& 25   & 30   &			\\\cline{2-4}
+300       &      &  3950s&     \\
+500         & 5400s&       &      \\
 \end{tabular}
 \end{center}
-\caption{Time to compute all probabilities. The first column indicates the length, the second
- the size of $\Omega$ and  the column indexes indicate the number
+\caption{Time to compute all probabilities. The first column indicates the length and  the column indexes indicate the number
  of mutations. $\alpha$ is
-set at $0.5$, and $\beta=1.5$ and All sequences are generated randomly.}
+set at $0.5$, and $\beta=1.5$ and $|\Omega|=44$.}
 \label{tab:time}

### 2012-10-07 benchmark/runtest.py
new figs

@@ -1,16 +1,20 @@
 import sys,os,os.path
 
-def main(nmut,alpha,prefix):
+outdir=""
+
+def main(nmut,alpha,prefix,penalty):
 
     for numex in range(100):
         infile = "RF00001_%d.in" % numex
         if os.path.exists(infile):
             alphanorm = int (alpha * 100)
-            outfile = "RF00001_%d_nmut_%d_alpha_%d.out" % (numex,nmut,alphanorm)
-            commandline1="python ../src/RNAPyroEx.py %s %d %.1f 1.5 > %s" % (infile,nmut,alpha,outfile)
+            outfile = "%sRF00001_%d_nmut_%d_alpha_%d.out" % (outdir,numex,nmut,alphanorm)
+            commandline1="python ../src/RNAPyroEx.py %s %d %.1f %.2f > %s" % (infile,nmut,alpha,penalty,outfile)
             print commandline1
             if not os.path.exists(outfile):
                 os.system(commandline1)
+            else:
+                print "WARNING: %s already present" % (outfile)
             commandline2="python ../scripts/benchmark.py -i RF00001_%d.ref -o RF00001_%d_nmut_%d_alpha_%d.out" % (numex,numex,nmut,alphanorm)
             print commandline2
             os.system(commandline2)
@@ -23,6 +27,7 @@ if __name__ == '__main__':
     nmut=10
     alpha=0.5
     prefix="RF00001"
+    penalty=1.5
     
     for i in range(len(sys.argv)):
         myarg=sys.argv[i]
@@ -34,9 +39,11 @@ if __name__ == '__main__':
             alpha=float(sys.argv[i+1])
         if myarg=='-p':
             prefix=sys.argv[i+1]
+        if myarg=='-e':
+            penalty=float(sys.argv[i+1])
     
     
-    main(nmut,alpha,prefix)
+    main(nmut,alpha,prefix,penalty)
 
 
 

### 2012-10-07 scripts/benchmark.py

@@ -232,237 +232,227 @@ def randommutants(data,nmut):
 
     return mutationlist
 
-## read infile
-
-def basiccorrelation(mutlist,data):
-
-    score=0
-    for nt1,index,nt2 in mutlist:
-        score += data[index][nt1]
-
-    if False:
-        sumscore=0
-        for index,probs in data.iteritems():
-            subtotal = probs['A']+probs['C']+probs['G']+probs['U']
-            if False and subtotal!=1:
-                print index,probs
-            sumscore+=subtotal
-        print sumscore
     
-    return score/len(mutlist)
+## correlation 2
 
-## read infile
+def probprofile(indatafull,outdatafull,withread):
+    
+    scores={}
+    counter={}
+    
+    for alpha in indatafull.keys():
 
-def fullcorrelation(indata,outdata,withread):
-    
-    read=indata['read'].replace('.','').replace('-','').upper()
-    original=indata['target'].replace('.','').replace('-','').upper()
-    
-    if len(read)!=len(original):
-        print 'ERROR'
-        sys.exit(1)
-    
-    goodscore=0.
-    badscore=0.
-    cmptgood=0
-    cmptbad=0
-    scores={'TP':0.0,'TN':0.0,'FP':0.0,'FN':0.0}
-    counter={'TP':0,'TN':0,'FP':0,'FN':0}
-    for i in range(len(read)):
-        #print i,(outdata[i]['A']+outdata[i]['C']+outdata[i]['G']+outdata[i]['U'])
-        for nt in ['A','C','G','U']:
-            if nt!=read[i] or withread:
-                if nt==original[i]:
-                    # nt is the good one (T)
-                    if original[i]!=read[i]: # this is a mutation site
-                        scores['TP']+=outdata[i][nt]
-                        counter['TP']+=1
-                    else: # nt is NOT a mutation site (N)
-                        scores['TN']+=outdata[i][nt]
-                        counter['TN']+=1
-                else:
-                    # nt is not correct (F)
-                    if original[i]!=read[i]: # nt is a mutation site(P)
-                        scores['FP']+=outdata[i][nt]
-                        counter['FP']+=1
-                    else:
-                        # nt is NOT a mutation site (N)
-                        scores['FN']+=outdata[i][nt]
-                        counter['FN']+=1
+        indata = indatafull[alpha]
+        outdata = outdatafull[alpha]
+        scores[alpha]={'TP':0.0,'TN':0.0,'FP':0.0,'FN':0.0}
[…]

### 2012-10-07 Recomb/methods_RECOMB.tex
New title + removed todo

@@
 is the average isostericity of a base-pair in the candidate sequence in comparison with the reference alignment.
-The $\ISO$ function uses the \todo[fancyline, inline]{\relsize{-1}Is this true?
-If so, we might want to try taking the most likely edge/orientation combination for each couple instead.}{Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
+The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
 Isostericity scores range between $0$ and $9.7$, with $0$ being assigned to a perfect isostericity, and a penalty of $10$ is used for missing entries.
@@
 Let $s$ an RNA 
-sequence and $m\geq 0$.
-$S$ is considered as  one derivation of the \todo[fancyline]{A bit tough: SCFG's were never mentioned before.
-Maybe introduce this with a sentence relating problem to parsing}{SCFG} generating
+sequence and $m\geq 0$. $S$ is considered as  one derivation of the 
+{SCFG} generating  
 all  secondary structures of length  $|s|$. We are interested in  
@@
 
-We are now left to compute the probability that a \todo[fancyline]{Maybe worth doing it for base-pairs? maybe not\ldots}{given position is a given nucleotide}.
+We are now left to compute the probability that a given position is a given nucleotide.
 We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
@@
 
-In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, \todo[fancyline,inline]{\relsize{-1}Seems weird to do this for a given upper bound $M$ (as opposed to exact $m$), and it makes the equations a bit more complicated.
-Anyway, there is roughly $3n$ times more sequences with $m+1$ mutations than with $m$ muts, so I expect the $m+1$ guys are so dominating in the final probabilities that it may not be worth the trouble of combining them with the \emph{little guys} ($\#muts < m$)}{for $m$ smaller or equal to our target $M$}.
-The numerators are divided in the following three cases.
+In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, 
+for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
 \begin{description}

### 2012-10-07 Recomb/results_RECOMB.tex
new figs
results section completed

@@
 biggest MSA  which contains $130$ sequences (out of $712$ in the original Rfam MSA). Then, in order to avoid any bias, we used \texttt{cd-hit}
-\cite{CDHIT} to remove sequences with more than 80\% of sequence similarity. This operation resulted in a data set of $45$ sequences.
+\cite{Li:2006fk} to remove sequences with more than 80\% of sequence similarity. This operation resulted in a data set of $45$ sequences.
 
@@
 mutations. Our sequences have $119$ nucleotides, thus the number of mutations corresponds to an error-rate of 10\%. We repeated this operation 
-$10$ times.
+$10$ times. The value of $\beta$ has been set to $1.5$ (larger values gave similar results). To estimate the impact of the distribution of the weights
+between the energy term and isostericity  score, we use 4 different values of $\alpha = {0, 0.5, 0.8, 1.0}$. Similarly, we also investigate the impact of 
+an under- and over- estimate of the number of errors, and use values equal to 50\% (6 mutations) and 200\% (24 mutations) of the exact number of
+errors (i.e. 12).
 
-To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed by
-\RNApyro. We report in Table \ref{tab:benchmark} the area under the curve (AUC). More specifically, we fix a threshold $\lambda \in [0,1]$ and we
-predict an error at position $i$ in sequence $\omega$ if and only if the probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceed this threshold.
-The set of corrections is thus $\{ n \; | \;  n \in \{ A,C,G,U \} \mbox{ and } P(i,n) > \lambda \mbox{ and }  n \neq \omega[i] \}$, where $\omega[i]$ is the
-nucleotide at position $i$ in the input sequence. Then, we progressively vary $\lambda$ between $0$ and $1$ to calculate the ROC curve and the AUC.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-& 1.0 & 0.5 & 0. \\
-\hline
-6 & 0.69 & 0.72 & 0.74 \\
-12 & 0.80 & 0.84 & 0.84 \\
-24 & 0.76 & 0.84 & 0.84 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Performance of error-correction. The row index indicates the number of mutations performed with \RNApyro, and the column index indicates
-the value of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores. In each cell, we report the average AUC
-values over the 10 experiments.}
-\label{tab:benchmark}
-\end{table}
+To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed
+by \RNApyro. More specifically, we fix a threshold $\lambda \in [0,1]$ and we predict an error at position $i$ in sequence $\omega$ if and only if the
+probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceed this threshold. To correct the errors we use the set of nucleotides with a probability
+higher than this threshold is  $C(i) = \{ n \; | \;  n \in \{ A,C,G,U \} \mbox{ and } P(i,n) > \lambda \mbox{ and }  n \neq \omega[i] \}$, where $\omega[i]$ is
+the nucleotide at position $i$ in the input sequence. We note that for the lowest thresholds multiple nucleotides can be available in $C(i)$ to correct
+the sequence. Here, we remind that we aim to estimate the potential of error-correction of \RNApyro and not to develop a error-correction pipe-line. 
+The latter will be investigate in further studies. Finally, we progressively vary $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
+under the curve (AUC). We report our results in Figure \ref{fig:ROCall}.
 
@@
 \caption{6 mutations}
+\label{fig:ROC6mut}
 \end{subfigure}
@@
 \caption{12 mutations}
+\label{fig:ROC12mut}
 \end{subfigure}
@@
 \caption{24 mutations}
+\label{fig:ROC24mut}
 \end{subfigure}
+\caption{Performance of error-correction. The row index indicates the number of mutations performed with \RNApyro, and the column index indicates
+the value of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores. In each cell, we report the average AUC
+values over the 10 experiments.}
+\label{fig:ROCall}
 \end{figure}
 
+Our data demonstrate that our algorithm exhibits interesting performance for error-correction applications. First, the AUC values (up to $0.86$) indicate that a
+signal has been successfully extracted. This result has been achieved with errors in loop regions (i.e. without base pairing information) and thus suggests
+that correction rates in structured regions (i.e. base paired regions) could be even higher. Next, the optimal values of $\alpha$ tend to be close to $0.0$. This 
+finding suggests that at this point the information issued from the stacking energy is currently modest. However, specific examples showed improved performance
+using this energy term. Further studies must be conducted to understand how to make the best use it. Finally, our algorithm seems robust to the number of
+mutations considered. Indeed, good AUC values are achieved with low estimates of the number of errors in the sequence (c.f. 50\% of the errors  in
+Fig.~\ref{fig:ROC6mut}) and with large  values (c.f.~200\% of the errors  in Fig.~\ref{fig:ROC24mut}). It is worth noting that scoring scheme with larger weight on
+the isostericity scores (low $\alpha$ values) seem more robust to under- and over-estimate of the number of errors.
+
 \subsection{Time}
-The applications for this algorithms implies a need for efficiency
-and scalability. In Table~\ref{tab:time} we
-present different times needed to compute the probabilities for
- every nucleotide at every positions for a vast set of parameters. For those test
+Further NGS applications of \RNApyro implies a need for efficiency and scalability. In Table~\ref{tab:time} we
+present different times needed to compute the probabilities for  every nucleotide at every positions for a vast set of parameters. For those test
  all the sequences were generated randomly, as the target secondary structure.

### 2012-10-07 Recomb/introduction_RECOMB.tex
intro mod
fixed intro (abstract missing)

@@
 
-Ribonucleic acids (RNAs) are  now  an ubiquitous class of molecules, being
-found in every living organisms and having a broad range of functions, from catalyzing
-chemical reactions as the RNase P or the group II introns,
-hybridizing  messenger RNA to regulate gene expression,
-to ribosomal RNA (rRNA) synthesizing proteins.
-Those functions  require specific structures, 
-encoded in their nucleotide sequence. Although the functions
-need to be preserved through various organisms, and therefore
-their structure must be similar,  the sequences
-can greatly differ from one organism to another.
-For half a century, biological molecules have been studied as a proxy to understand
-evolution~\cite{Zuckerkandl1965}, and with all their characteristics, rRNAs have
-always been a prime candidate for phylogenetic studies~\cite{Olsen1986, Olsen1993}.
+Ribonucleic acids (RNAs) are found in every living organisms and exhibit a broad range of functions, from catalyzing
+chemical reactions as the RNase P or the group II introns, hybridizing  messenger RNA to regulate gene expression,
+to ribosomal RNA (rRNA) synthesizing proteins. Those functions  require specific structures,  encoded in their nucleotide
+sequence. Although the functions, and thus the structures, need to be preserved through various organisms, the sequences
+can greatly differ from one organism to another. This sequence diversity coupled with the structural conservation is a fundamental
+asset for evolutionary studies.
 
-In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007}, 
-leveraging NGS techniques to sequence as many new organisms 
-as possible, are producing a wealth of new information. Although
-those techniques have a huge throughput, they yield a sequencing error rate of around
-$4\%$~\cite{Huse2007}. This error can be highly reduced  when highly 
-redundant multiple sequence alignments 
- are available, but in studies of new or not well known organisms, there is not
- enough  similarity to differentiate between the sequencing errors and the natural 
- polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
+For half a century, biological molecules have been studied as a proxy to understand evolution~\cite{Zuckerkandl1965}, and due
+to their fundamental functions, rRNAs have always been a prime candidate for phylogenetic studies~\cite{Olsen1986, Olsen1993}.
+In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007} benefited of new technologies such as the
+NGS techniques to sequence as many new organisms as possible and extract a unprecedented flow of new information. 
+Nonetheless, these high-throughput techniques typically have high error rates that make their applications to metagenomics (a.k.a. environmental
+genomics) studies challenging. For instance, pyrosequencing as implemented by Roche's 454 produces may have a error rate raising up to 10\%.
+Because there is no cloning step, resequencing to increase accuracy is not possible and it is therefore vital to disentangle noise from true sequence
+diversity in this type of data \cite{Quince:2009uq}. This error can be highly reduced  when large multiple sequence alignments with close homologs
+are available, but in studies of new or not well known organisms, such information is rather sparse. In particular, it is common to is not enough  similarity to 
+differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
+It is therefore essential to develop methods that can exploit any type of signal available to correct errors.
 
-In this paper, we hypothesize that the family and consensus secondary structure hold 
-information allowing to identify the positions most probable to be sequencing errors.
+In this paper, we propose to use the structural information included in physics-based energy models and covariance models to identify and correct errors
+in RNA molecules with conserved secondary structure.
+In particular, we hypothesize that sequence profiles combined with conserved consensus secondary structures provide an information that will allow us to identify and fix sequencing errors.
 
-Leveraging the techniques  in \texttt{RNAmutants}~\cite{Waldispuhl2008}, and building on top 
-of the \emph{Inside-Outside algorithm}, we define here a new method called \texttt{RNApyro}
-efficiently computing for large RNAs those probabilities under a new 
-pseudo-energetic model.
+Here, we introduce \RNApyro and expand the range of algorithmic techniques previously introduced with \RNAmutants~\cite{Waldispuhl2008}.
+Instead of exploring the full conformational landscape and sample mutants, we develop an inside-outside algorithm that enables us
+to explore the complete mutational landscape with a \emph{fixed} secondary structure and calculate exact probability values. In addition to a gain into the
+numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot k^2)$ for the original version of 
+\RNAmutants to $\mathcal{O}(n \cdot k^2)$ for \RNApyro, where $n$ is the size of the sequence and $k$ the number of mutations).
+We design a new scoring scheme combining nearest-neighbor models \cite{Turner2010} to isostericity metrics \cite{Stombaugh2009}.
+
 Classical techniques define a probabilistic model using a Boltzmann distribution 
@@
 
-The 5S ribosomal RNA family, $119$ nucleotides long,  was used to benchmark our method.
-It is a prime example since it has been extensively used for phylogenetic
+We benchmark our method on the 5S ribosomal RNA. It is a prime example since it has been extensively used for phylogenetic
 reconstructions~\cite{Hori1987} and its sequence has been recovered for over 8000 species
@@
  Using a leave one out strategy, we perform random distributed mutations on a sequence. We show that
-\texttt{RNApyro} reconstructs the original sequence with an excellent accuracy.
+\texttt{RNApyro} can reconstruct the original sequence with an excellent accuracy.
 
-The pseudo-energetic model and the algorithm is presented in Sec.~\ref{sec:methods}.
+The scoring scheme and the algorithm are presented in Sec.~\ref{sec:methods}.
 Details of the implementation and benchmarks are in Sec.~\ref{sec:results}. 
-Future applications and a discussion are developed in Sec.~\ref{sec:conclusion}.
+Finally, we discuss future developments and applications in Sec.~\ref{sec:conclusion}.

### 2012-10-07 Recomb/conclusion_RECOMB.tex
conclusion

@@
 %!TEX root = main_RECOMB.tex
-\section{Discussion}
+\section{Conclusion}
 \label{sec:conclusion}
+
+In this article we presented a new 
+ pseudo-energetic model taking into account the energetic contribution 
+of the stacked base pairs and the isostericity of all base pairs. 
+We also showed that a secondary structure allows to efficiently
+explore the mutational landscape and that we can benefit from the
+ conserved consensus secondary structures  information
+to identify and fix sequencing errors.
+By combining the pseudo-energetic model and the 
+mutational landscape exploration we are able  to efficiently 
+correct randomly mutated 5s rRNA sequences. 
+As presented in Sec.~\ref{sec:results},
+we observe that the models
+with larger weights on the
+isostericity seems to hold a higher accuracy on the estimation of errors.
+Importantly, the implementation is fast enough for practical applications.

### 2012-10-07 Recomb/introduction_RECOMB.tex
update intro & figure caption

@@
 can greatly differ from one organism to another. This sequence diversity coupled with the structural conservation is a fundamental
-asset for evolutionary studies.
+asset for evolutionary studies. To this end, algorithms to analyze the relationship between RNA mutants and structures are required.
 
 For half a century, biological molecules have been studied as a proxy to understand evolution~\cite{Zuckerkandl1965}, and due
-to their fundamental functions, rRNAs have always been a prime candidate for phylogenetic studies~\cite{Olsen1986, Olsen1993}.
-In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007} benefited of new technologies such as the
-NGS techniques to sequence as many new organisms as possible and extract a unprecedented flow of new information. 
+to their fundamental functions and remarkably conserved structures, rRNAs have always been a prime candidate for phylogenetic
+studies~\cite{Olsen1986, Olsen1993}. In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007} benefited
+of new technologies such as the NGS techniques to sequence as many new organisms as possible and extract a unprecedented flow of new information. 
 Nonetheless, these high-throughput techniques typically have high error rates that make their applications to metagenomics (a.k.a. environmental
@@
 Because there is no cloning step, resequencing to increase accuracy is not possible and it is therefore vital to disentangle noise from true sequence
-diversity in this type of data \cite{Quince:2009uq}. This error can be highly reduced  when large multiple sequence alignments with close homologs
+diversity in this type of data \cite{Quince:2009uq}. Errors can be significantly reduced  when large multiple sequence alignments with close homologs
 are available, but in studies of new or not well known organisms, such information is rather sparse. In particular, it is common to is not enough  similarity to 
@@
 
-In this paper, we propose to use the structural information included in physics-based energy models and covariance models to identify and correct errors
-in RNA molecules with conserved secondary structure.
-In particular, we hypothesize that sequence profiles combined with conserved consensus secondary structures provide an information that will allow us to identify and fix sequencing errors.
+In this paper, we introduce \RNApyro, a novel algorithm to that enable us to calculate precisely mutational probabilities in RNA sequences with a
+conserved consensus secondary structure. We propose to use the structural information included in physics-based energy models and covariance models to 
+identify and correct errors in RNA molecules with conserved secondary structure. In particular, we hypothesize that conserved consensus secondary structures 
+combined with sequence profiles provide an information that will allow us to identify and fix sequencing errors.
 
-Here, we introduce \RNApyro and expand the range of algorithmic techniques previously introduced with \RNAmutants~\cite{Waldispuhl2008}.
+Here, we expand the range of algorithmic techniques previously introduced with \RNAmutants~\cite{Waldispuhl2008}.
 Instead of exploring the full conformational landscape and sample mutants, we develop an inside-outside algorithm that enables us
-to explore the complete mutational landscape with a \emph{fixed} secondary structure and calculate exact probability values. In addition to a gain into the
-numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot k^2)$ for the original version of 
-\RNAmutants to $\mathcal{O}(n \cdot k^2)$ for \RNApyro, where $n$ is the size of the sequence and $k$ the number of mutations).
+to explore the complete mutational landscape with a \emph{fixed} secondary structure and to calculate exactly mutational probability values. In addition
+to a gain into the numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot k^2)$ for the
+original version of  \RNAmutants to $\mathcal{O}(n \cdot k^2)$ for \RNApyro, where $n$ is the size of the sequence and $k$ the number of mutations).
 We design a new scoring scheme combining nearest-neighbor models \cite{Turner2010} to isostericity metrics \cite{Stombaugh2009}.

### 2012-10-07 Recomb/conclusion_RECOMB.tex
conclusion

@@
 to identify and fix sequencing errors.
-By combining the pseudo-energetic model and the 
-mutational landscape exploration we are able  to efficiently 
+
+By combining those two approaches, the pseudo-energetic model and the 
+mutational landscape exploration, we are able  to efficiently 
 correct randomly mutated 5s rRNA sequences. 

### 2012-10-07 Recomb/results_RECOMB.tex
conclusion

@@
 for  arbitrary floating point precision. The source code is freely available at \verb+https://github.com/McGill-CSB/RNApyro+.
+
+The time benchmarks were done on a MacBook Pro 2010, 2 processors at 2.66GHz, 8GB of RAM.
+Since applications of \RNApyro implies a need for
+ efficiency and scalability. In Table~\ref{tab:time} we
+present different times needed to compute the probabilities for  every nucleotide at every positions for a vast set of parameters. For those test
+ all the sequences were generated randomly, as the target secondary structure.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{lccc}
+$|s|$&\multicolumn{3}{c}{Number of mutations}\\\cline{2-4}
+		 			  & 6   &  12  & 24\\\cline{2-4}
+100  				& 35s  & 238s & 1023s\\
+300  			& 135s & 594s &2460s\\\cline{2-4}
+		 						& 25   & 30   &	50		\\\cline{2-4}
+300       &      &  3950s&     \\
+500         & 5400s&       &  21003s    \\
+\end{tabular}
+\end{center}
+\caption{Time to compute all probabilities. The first column indicates the length and  the column indexes indicate the number
+ of mutations. $\alpha$ is
+set at $0.5$,  $\beta$ to $1.5$ and $|\Omega|=44$.}
+\label{tab:time}
+\end{table}
 
@@
 \end{subfigure}
-\caption{Performance of error-correction. The row index indicates the number of mutations performed with \RNApyro, and the column index indicates
-the value of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores. In each cell, we report the average AUC
-values over the 10 experiments.}
+\caption{Performance of error-correction. Subfigures accuracy with under-estimated error rate (6 mutations), exact estimates (12 mutations) and over estimates 
+(24 mutations). We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values 
+ranging of $\alpha=\{0,0.5,0.8,1.0\}$. The AUC is indicated in the legend of the figures. Each individual ROC curve represent the average performance aver the 10 experiments.}
 \label{fig:ROCall}
@@
 the isostericity scores (low $\alpha$ values) seem more robust to under- and over-estimate of the number of errors.
-
-\subsection{Time}
-Further NGS applications of \RNApyro implies a need for efficiency and scalability. In Table~\ref{tab:time} we
-present different times needed to compute the probabilities for  every nucleotide at every positions for a vast set of parameters. For those test
- all the sequences were generated randomly, as the target secondary structure.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{lccc}
-$|s|$&\multicolumn{3}{c}{Number of mutations}\\\cline{2-4}
-		 			  & 6   &  12  & 24\\\cline{2-4}
-100  				& 35s  & 238s & 1023s\\
-300  			& 135s & 594s &2460s\\\cline{2-4}
-		 						& 25   & 30   &			\\\cline{2-4}
-300       &      &  3950s&     \\
-500         & 5400s&       &      \\
-\end{tabular}
-\end{center}
-\caption{Time to compute all probabilities. The first column indicates the length and  the column indexes indicate the number
- of mutations. $\alpha$ is
-set at $0.5$, and $\beta=1.5$ and $|\Omega|=44$.}
-\label{tab:time}
-\end{table}

### 2012-10-07 Recomb/conclusion_RECOMB.tex
intro/name

--- 2012-10-07 Recomb/conclusion_RECOMB.tex
@@
 In this article we presented a new 
- pseudo-energetic model taking into account the energetic contribution 
+ pseudo energy model taking into account the energetic contribution 
 of the stacked base pairs and the isostericity of all base pairs. 
@@
 to identify and fix sequencing errors.
-By combining the pseudo-energetic model and the 
-mutational landscape exploration we are able  to efficiently 
+
+By combining those two approaches, the pseudo energy and the 
+mutational landscape exploration, we are able  to efficiently 
 correct randomly mutated 5s rRNA sequences. 

### 2012-10-07 Recomb/introduction_RECOMB.tex
intro/name
ref

@@
 differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
-It is therefore essential to develop methods that can exploit any type of signal available to correct errors.
+A few techniques have been developed to remedy to this problem~\cite{Quinlan2010,Medvedev2011} but they do not take
+into account all the available information. It is therefore essential to develop methods that can exploit any type of signal available to correct errors.
 
 In this paper, we introduce \RNApyro, a novel algorithm to that enable us to calculate precisely mutational probabilities in RNA sequences with a
-conserved consensus secondary structure. We propose to use the structural information included in physics-based energy models and covariance models to 
-identify and correct errors in RNA molecules with conserved secondary structure. In particular, we hypothesize that conserved consensus secondary structures 
-combined with sequence profiles provide an information that will allow us to identify and fix sequencing errors.
+conserved consensus secondary structure. We show how our techniques can exploit the structural information embedded in physics-based energy models, 
+covariance models and isostericity scales to identify and correct errors in RNA molecules with conserved secondary structure. In particular, we hypothesize that 
+conserved consensus secondary structures combined with sequence profiles and provide an information that allow us to identify and fix sequencing errors.
 

### 2012-10-07 Recomb/introduction_RECOMB.tex
fix ref

@@
 differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
-A few techniques have been developed to remedy to this problem~\cite{Quinlan2010,Medvedev2011} but they do not take
+A few techniques have been developed to remedy to this problem~\cite{Quinlan2008,Medvedev2011} but they do not take
 into account all the available information. It is therefore essential to develop methods that can exploit any type of signal available to correct errors.

### 2012-10-07 Recomb/main_RECOMB.tex
added raw equations to recomb
improved and clean dataset construction (clustering requires cd-hit)
Reorganized method section
new title
intro
new roc figs
Dynamic programming Figures
New title + removed todo
fixed intro (abstract missing)
added abstract

@@
 \usepackage{xspace}
+\usepackage{relsize}
 \usepackage[numbers]{natbib}
+\usepackage[colorinlistoftodos]{todonotes}
 
-\usepackage{subfig} %Subfigures
+%\usepackage{subfig} %Subfigures
+\usepackage{subcaption} %Subfigures
+
+\usepackage{tikz}
+\usepackage{amsmath}
+\usepackage{ifthen}
+\usepackage{relsize}
+\usetikzlibrary{positioning}
+\usetikzlibrary{matrix}
+\usetikzlibrary{calc}
+\usetikzlibrary{shapes}
+\usetikzlibrary{arrows}
+\usetikzlibrary{fit}
+\usetikzlibrary{decorations}
+\usetikzlibrary{snakes}
+\pgfdeclarelayer{background}
+\pgfdeclarelayer{backbackground}
+\pgfdeclarelayer{foreground}
+\pgfsetlayers{backbackground,background,main,foreground}
 
@@
 \usepackage[applemac]{inputenc} %for the encoding
+
+\newcommand{\RNAmutants}{\texttt{RNAmutants}\xspace}
+\newcommand{\RNApyro}{\texttt{RNApyro}\xspace}
 
@@
 %\newcommand{\citep}{\cite}
+\newcommand{\Z}[3]{\mathcal{Z}_{\substack{(#1)\\ [#3]}}^{#2}}
+\newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
+\newcommand{\B}{\mathcal{B}}
+\newcommand{\Kron}{\delta}
+\newcommand{\ub}{\bullet}
 
-\title{RNApyro}
-\author{}
+%\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
+\title{A linear inside-outside algorithm for correcting sequencing errors in structured RNA sequences}
+\author{Vladimir Reinharz, Yann Ponty$^{*}$, J\'er\^{o}me Waldisp\"{u}hl$^*$}
 \date{}
@@
 \begin{abstract}
-Some stuff~\cite{Waldispuhl2008}
+Analysis of the sequence-structure relationship in RNA molecules are essential to evolutionary studies but also to 
+concrete applications such as error-correction methodologies in sequencing technologies. The prohibitive sizes of the
+mutational and conformational landscapes combined with the volume of data to proceed require efficient algorithms 
+to compute sequence-structure properties. More specifically, here we aim to calculate which mutations increase the most the 
+likelihood of a sequence to a given structure and RNA family.\\
+In this paper, we introduce \RNApyro, an efficient linear-time and space inside-outside algorithm that computes exact mutational
+probabilities under secondary structure and evolutionary constraints given as a multiple sequence alignment with a consensus structure.
+We develop a scoring scheme combining classical stacking base pair energies to novel isostericity scales, and apply our techniques
+to correct point-wise errors in 5s rRNA sequences. Our results suggest that \RNApyro is a promising algorithm to complement existing
+tools in the NGS error-correction pipeline.
 
 \noindent
-\textbf{Key words:} RNA
+\textbf{Key words:} RNA, mutations, secondary structure
 \end{abstract}

### 2012-10-07 Recomb/conclusion_RECOMB.tex
conclusion
conc

--- 2012-10-07 Recomb/conclusion_RECOMB.tex
@@
 
-In this article we presented a new 
- pseudo-energetic model taking into account the energetic contribution 
+In this article we presented a new and efficient way of
+exploring the mutational landscape of an RNA, benefit from the
+ conserved consensus secondary structures  information,
+to identify and fix sequencing errors. We also present a
+ pseudo energy model taking into account the energetic contribution 
 of the stacked base pairs and the isostericity of all base pairs.
-We also showed that a secondary structure allows to efficiently
-explore the mutational landscape and that we can benefit from the
- conserved consensus secondary structures  information
-to identify and fix sequencing errors.
-By combining the pseudo-energetic model and the 
-mutational landscape exploration we are able  to efficiently 
+
+By combining those two approaches,  the 
+mutational landscape exploration and the pseudo energy model,
+ we are able  to efficiently 
 correct randomly mutated 5s rRNA sequences. 
@@
 Importantly, the implementation is fast enough for practical applications.
+
+We must point out that our approach can only detect point wise 
+sequencing errors, and only if they are located in base pairs. It 
+nonetheless supplements well existing tools by using previously discarded
+information.

### 2012-10-07 Recomb/introduction_RECOMB.tex
intro again

@@
 conserved consensus secondary structure. We show how our techniques can exploit the structural information embedded in physics-based energy models, 
-covariance models and isostericity scales to identify and correct errors in RNA molecules with conserved secondary structure. In particular, we hypothesize that 
-conserved consensus secondary structures combined with sequence profiles and provide an information that allow us to identify and fix sequencing errors.
+covariance models and isostericity scales to identify and correct point-wise errors in RNA molecules with conserved secondary structure. In particular, we 
+hypothesize that  conserved consensus secondary structures combined with sequence profiles and provide an information that allow us to identify and fix sequencing errors.
 
@@
 original version of  \RNAmutants to $\mathcal{O}(n \cdot k^2)$ for \RNApyro, where $n$ is the size of the sequence and $k$ the number of mutations).
+
 We design a new scoring scheme combining nearest-neighbor models \cite{Turner2010} to isostericity metrics \cite{Stombaugh2009}.
+Classical approaches use a Boltzmann distribution whose weights are estimated using a nearest-neighbour energy model~\cite{Turner2010}. However, the
+latter only accounts for  canonical and wobble, base pairs. As was shown by Leontis and Westhof~\cite{Leontis2001},
+the diversity of base pairs observed in tertiary structures is much larger, albeit their energetic contribution remains unknown. To quantify geometrical differences, 
+an isostericity distance has been designed \cite{Stombaugh2009}, increasing as two base pairs geometrically differ from each other in space. Therefore, we 
+incorporate these scores in the Boltzmann weights used by \RNApyro.
 
-Classical techniques define a probabilistic model using a Boltzmann distribution 
-whose weights are based on the free energy of the structure, using as energy parameter
-the values of Turner found in the NNDB~\cite{Turner2010}  for stacked, 
-canonical and wobble, base pairs. As shown by Leontis and Westhof~\cite{Leontis2001},
-this  does not encapsulated the large diversity of base pairs that any nucleotide
-can form with any other, although with an energy too small to be yet determined
-by experimental techniques. To quantify geometrical differences, they
- define an isostericity distance, increasing as two base pairs differ 
- more from one another in space. We incorporate this second measure in the Boltzmann weights.
+We illustrate and benchmark our techniques for point-wise error corrections on the 5S ribosomal RNA. We choose the latter since it has been extensively
+used for phylogenetic reconstructions~\cite{Hori1987} and its sequence has been recovered for over 712 species (in the Rfam seed alignment with id
+\texttt{RF00001}). Using a leave one out strategy, we perform random distributed mutations on a sequence. While our methodology is restricted to the correction of 
+point-wise error in structured regions (i.e. with base pairs), we show that \texttt{RNApyro} can successfully extract a signal that can be used to reconstruct the 
+original sequence with an excellent accuracy. This suggests that \RNApyro is a promising algorithm to complement existing tools in the NGS error-correction 
+pipeline.
 
-We benchmark our method on the 5S ribosomal RNA. It is a prime example since it has been extensively used for phylogenetic
-reconstructions~\cite{Hori1987} and its sequence has been recovered for over 8000 species
- (RFAM Id: \texttt{RF00001}).
- Using a leave one out strategy, we perform random distributed mutations on a sequence. We show that
-\texttt{RNApyro} can reconstruct the original sequence with an excellent accuracy.
-
-The scoring scheme and the algorithm are presented in Sec.~\ref{sec:methods}.
-Details of the implementation and benchmarks are in Sec.~\ref{sec:results}. 
+The algorithm and the scoring scheme are presented in Sec.~\ref{sec:methods}. Details of the implementation and benchmarks are in Sec.~\ref{sec:results}. 
 Finally, we discuss future developments and applications in Sec.~\ref{sec:conclusion}.

### 2012-10-07 Recomb/main_RECOMB.tex
affiliations

@@
 \title{A linear inside-outside algorithm for correcting sequencing errors in structured RNA sequences}
-\author{Vladimir Reinharz, Yann Ponty$^{*}$, J\'er\^{o}me Waldisp\"{u}hl$^*$}
-\date{}
+\author{Vladimir Reinharz$^{2}$, Yann Ponty$^{1}\footnote{To whom correspondence should be addressed}\;$, J\'er\^{o}me Waldisp\"{u}hl$^{2*}$}
+\date{$^1$Laboratoire d'Informatique, \'Ecole Polytechnique, Palaiseau, 91120 France $^2$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7}
 

### 2012-10-07 Recomb/introduction_RECOMB.tex
typos

@@
 studies~\cite{Olsen1986, Olsen1993}. In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007} benefited
-of new technologies such as the NGS techniques to sequence as many new organisms as possible and extract a unprecedented flow of new information. 
+of new technologies such as the NGS techniques to sequence as many new organisms as possible and extract an unprecedented flow of new information. 
 Nonetheless, these high-throughput techniques typically have high error rates that make their applications to metagenomics (a.k.a. environmental
@@
 covariance models and isostericity scales to identify and correct point-wise errors in RNA molecules with conserved secondary structure. In particular, we 
-hypothesize that  conserved consensus secondary structures combined with sequence profiles and provide an information that allow us to identify and fix sequencing errors.
+hypothesize that  conserved consensus secondary structures combined with sequence profiles  provide an information that allow us to identify and fix sequencing errors.
 

### 2012-10-07 Recomb/main_RECOMB.tex
fixed affiliations

@@
 \author{Vladimir Reinharz$^{2}$, Yann Ponty$^{1}\footnote{To whom correspondence should be addressed}\;$, J\'er\^{o}me Waldisp\"{u}hl$^{2*}$}
-\date{$^1$Laboratoire d'Informatique, \'Ecole Polytechnique, Palaiseau, 91120 France $^2$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7}
+\date{$^1$CNRS/\'Ecole Polytechnique/INRIA AMIB, LIX UMR 7161 X-CNRS, Palaiseau, France $^2$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7}
 

### 2012-10-07 Recomb/introduction_RECOMB.tex
Quelques typos, révision de méthods

--- 2012-10-07 Recomb/introduction_RECOMB.tex
@@
 latter only accounts for  canonical and wobble, base pairs. As was shown by Leontis and Westhof~\cite{Leontis2001},
-the diversity of base pairs observed in tertiary structures is much larger, albeit their energetic contribution remains unknown. To quantify geometrical differences, 
+the diversity of base pairs observed in tertiary structures is much larger, albeit their energetic contribution remains unknown. To quantify geometrical discrepancies, 
 an isostericity distance has been designed \cite{Stombaugh2009}, increasing as two base pairs geometrically differ from each other in space. Therefore, we 

### 2012-10-07 Recomb/methods_RECOMB.tex
Quelques typos, révision de méthods

@@
 \end{align*}
-where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ are the free-energy and isostericity contributiuons, further described below, and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
+where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
+are the free-energy and isostericity contributions respectively, further described below, and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
 
@@
 (e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
-\subsection{isostericity contribution}
+\subsubsection{Isostericity contribution}
 The concept of isostericity captures the geometric superimposability of two base-pairs, using individual scores given by Stombaugh~\emph{et al}~\cite{Stombaugh2009} based on a probabilistic model.
@@
 
-\subsection{Mutational profiles of sequences}
+\subsection{Mutational profile of sequences}
 
@@
 Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec}, our 
-version of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
+version of the \emph{inside} algorithm, computes the partition function between $i$ and $j$ included, 
 knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing 
@@
 														\end{array}\right.$
- Let us finally denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
+
+Finally, we denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
 \begin{equation}
@@
 \resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
-\caption{Illustration of the principle underlying the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
+\caption{Illustration of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
 can be decomposed as a sequence over $[i+1,j]$ preceded by a, possibly mutated, base at $i$ 
@@
 
-\subsubsection{Inside-Outside}
+\subsubsection{Combining Inside and Outside value into base content probabilities}
 By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 

### 2012-10-07 Recomb/conclusion_RECOMB.tex
a

@@
  conserved consensus secondary structures  information,
-to identify and fix sequencing errors. We also present a
- pseudo energy model taking into account the energetic contribution 
-of the stacked base pairs and the isostericity of all base pairs.
+to identify and fix sequencing errors. We also develop a new
+ pseudo energy model, taking into account the nearest-neighbour energy model 
+and the isostericity,  quantifying geometrical differences, of all base pairs.
 

### 2012-10-07 Recomb/conclusion_RECOMB.tex
bib less space/conclusion

--- 2012-10-07 Recomb/conclusion_RECOMB.tex
@@
 In this article we presented a new and efficient way of
-exploring the mutational landscape of an RNA, benefit from the
+exploring the mutational landscape of an RNA, benefiting from the
  conserved consensus secondary structures  information,
-to identify and fix sequencing errors. We also present a
- pseudo energy model taking into account the energetic contribution 
-of the stacked base pairs and the isostericity of all base pairs.
+to identify and fix sequencing errors. We also showed a new
+ pseudo energy model, taking into account the nearest-neighbour energy model 
+and, to quantify geometrical differences,  the isostericity of all base pairs.
 
@@
 mutational landscape exploration and the pseudo energy model,
- we are able  to efficiently 
+ into \texttt{RNApyro}, we show that we can
 correct randomly mutated 5s rRNA sequences. 

### 2012-10-07 Recomb/results_RECOMB.tex
intro/name
bib/time
hardware/conclusion

@@
 
-The time benchmarks were done on a MacBook Pro 2010, 2 processors at 2.66GHz, 8GB of RAM.
+The time benchmarks were done on a MacMini 2010, 2.3GHz dual-core Intel Core i5, 8GB of RAM.
 Since applications of \RNApyro implies a need for
@@
 300  			& 135s & 594s &2460s\\\cline{2-4}
-		 						& 25   & 30   &	50		\\\cline{2-4}
-300       &      &  3950s&     \\
+		 						& 25   &    &	50		\\\cline{2-4}
 500         & 5400s&       &  21003s    \\
@@
 (24 mutations). We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values 
-ranging of $\alpha=\{0,0.5,0.8,1.0\}$. The AUC is indicated in the legend of the figures. Each individual ROC curve represent the average performance aver the 10 experiments.}
+ranging of $\alpha=\{0,0.5,0.8,1.0\}$. The AUC is indicated in the legend of the figures. Each individual ROC curve represent the average performance over the 10 experiments.}
 \label{fig:ROCall}

### 2012-10-07 Recomb/methods_RECOMB.tex
Complexity chapter ... at last!

@@
 
-\subsubsection{Combining Inside and Outside value into base content probabilities}
+\subsubsection{Combining Inside and Outside values into point-wise mutations probabilities}
 By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
@@
 Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations. 
-\begin{align*}
-	\mathbb{P}(s_i = x\mid s,\Omega, S,M) &:= \frac{\mathcal{B}(i,x,s,\Omega,S,M)}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\\ 
-\mathcal{B}(i,x,s,\Omega,S,M)&=
+\begin{align}
+	\mathbb{P}(s_i = x\mid s,\Omega, S,M) &:= \frac{\mathcal{W}(i,x,s,\Omega,S,M)}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}\\ 
+\mathcal{W}(i,x,s,\Omega,S,M)&=
  \left\{
@@
 		&\text{If }S_i = -1\\
+			\displaystyle
 			\sum_{m=0}^{M}
@@
      	 e^{\frac{-E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
-			\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
-			\Z{i+1,k-1}{m'}{x,b}
+			\cdot\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
+			\cdot\Z{i+1,k-1}{m'}{x,b}
 		&\text{If }S_i=k>i\\
+    \displaystyle
 			\sum_{m=0}^{M}
@@
      	 e^{\frac{-E^{\Omega, \beta}_{(k,i),\varnothing\to bx}}{RT}}
-			\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
-			\Z{k+1,i-1}{m'}{b,x}
+			\cdot\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
+			\cdot\Z{k+1,i-1}{m'}{b,x}
 		&\text{If }S_i=k<i
-	\end{array}\right.
-\end{align*}
+	\end{array}\right.\label{eq:combine}
+\end{align}
 
@@
 \end{description}
+
+\subsection{Complexity}
+Equations~\eqref{eq:Z_rec} and~\eqref{eq:Y_rec} can be computed using dynamic programming, i.e.
+$\mathcal{Z}^{*}_{*}$ or $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and interval lengths, memorizing the results such that they become available in constant time during future stages of the computation.
+Furthermore, energy terms $E(\cdot)$ can be accessed in constant time (after a precomputation in $\Theta(n\cdot|\Omega|)$ for the isostericity part).
+Computing any given term therefore requires $\Theta(m)$ operations.
+
+In principle, this action should be performed $\Theta(m\cdot n^2)$ times to investigate any triplet $(m,i,j)$.
+However, a close inspection of the recurrences reveal that the computation can be safely restricted to a subset of all intervals.
+For instance, in the forward computation, such intervals $[i,j]$ do not break any base-pair, and their next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
+Similar constraints hold for the outside computation.
+Such constraints drastically limit the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored.
+Consequently the overall complexity of the algorithm is $\Theta(n\cdot(|\Omega|+m^2))$ aritmetic operations and $\Theta(n\cdot(|\Omega|+m))$ memory.

### 2012-10-07 Recomb/results_RECOMB.tex
Complexity chapter ... at last!

@@
 
-The time benchmarks were done on a MacMini 2010, 2.3GHz dual-core Intel Core i5, 8GB of RAM.
+The time benchmarks were done on a MacBook Pro 2010, 2 processors at 2.66GHz, 8GB of RAM.
 Since applications of \RNApyro implies a need for
@@
 
-We build our data set from the 5S rRNA multiple sequence alignment (MSA) available in the Rfam Database 11.0 (Rfam id: \texttt{RF00001}).
+We built our data set from the 5S rRNA multiple sequence alignment (MSA) available in the Rfam Database 11.0 (Rfam id: \texttt{RF00001}).
 Since our software does not currently implement gaps (mainly because scoring indels is a challenging issue that cannot be fully addressed

### 2012-10-07 Recomb/conclusion_RECOMB.tex

@@
 
-We must point out that our approach can only detect point wise 
-sequencing errors, and only if they are located in base pairs. It 
-nonetheless supplements well existing tools by using previously discarded
-information.
+We must recall that our approach is restricted to
+ the correction of point-wise error in structured regions (i.e. base paired nucleotides).
+ Nonetheless it supplements well existing tools by using previously discarded
+information holding, as shown, a strong signal.

### 2012-10-07 Recomb/methods_RECOMB.tex
Added caption for outside replaced occurrences of forward/backward with inside/outside

@@
 \resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
-\caption{Illustration of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
+\caption{Principle of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
 can be decomposed as a sequence over $[i+1,j]$ preceded by a, possibly mutated, base at $i$ 
@@
 \resizebox{\textwidth}{!}{\input{FigDPOutside.pgf}}
-\caption{Principle of the outside computation.}
+\caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
+therefore its efficient implementation requires an implementation of the inside computation.}
 \end{figure}
@@
 only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
-We can then consider the outside interval $[i-1,k+1]$ by multiplying it by the the \emph{forward}
+We can then consider the outside interval $[i-1,k+1]$ by multiplying it by the the \emph{inside}
 value of the newly included interval (i.e. $[j,k-1]$), for 
@@
 By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
-be described in function of the \emph{forward} term as $\Z{0,n-1}{m}{X,X}$,
+be described in function of the \emph{inside} term as $\Z{0,n-1}{m}{X,X}$,
  for any nucleotide $X\in B$ or
-in function of the \emph{backward} term, for any unpaired position $k$:
+in function of the \emph{outside} term, for any unpaired position $k$:
 $$
@@
 
-In principle, this action should be performed $\Theta(m\cdot n^2)$ times to investigate any triplet $(m,i,j)$.
+In principle, this computation should be performed $\Theta(m\cdot n^2)$ times to investigate any triplet $(m,i,j)$.
 However, a close inspection of the recurrences reveal that the computation can be safely restricted to a subset of all intervals.
-For instance, in the forward computation, such intervals $[i,j]$ do not break any base-pair, and their next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
+For instance, in the inside computation, such intervals $[i,j]$ do not break any base-pair, and their next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
 Similar constraints hold for the outside computation.

### 2012-10-07 Recomb/results_RECOMB.tex
Added caption for outside replaced occurrences of forward/backward with inside/outside

@@
 
-The time benchmarks were done on a MacBook Pro 2010, 2 processors at 2.66GHz, 8GB of RAM.
+The time benchmarks were done on a MacMini 2010, 2.3GHz dual-core Intel Core i5, 8GB of RAM.
 Since applications of \RNApyro implies a need for
- efficiency and scalability. In Table~\ref{tab:time} we
-present different times needed to compute the probabilities for  every nucleotide at every positions for a vast set of parameters. For those test
- all the sequences were generated randomly, as the target secondary structure.
+ efficiency and scalability, we present in Table~\ref{tab:time} typical runtimes required to compute the probabilities for  every nucleotide at every positions for a vast set of parameters.
+For those test
+ both the sequences and the target secondary structure were generated at random.
 

### 2012-10-07 Recomb/introduction_RECOMB.tex
fix typos

@@
 to explore the complete mutational landscape with a \emph{fixed} secondary structure and to calculate exactly mutational probability values. In addition
-to a gain into the numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot k^2)$ for the
-original version of  \RNAmutants to $\mathcal{O}(n \cdot k^2)$ for \RNApyro, where $n$ is the size of the sequence and $k$ the number of mutations).
+to a gain into the numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot m^2)$ for the
+original version of  \RNAmutants to $\mathcal{O}(n \cdot m^2)$ for \RNApyro, where $n$ is the size of the sequence and $m$ the number of mutations).
 

### 2012-10-07 Recomb/conclusion_RECOMB.tex
Typos Concordance des temps dans la partie results

@@
  conserved consensus secondary structures  information,
-to identify and fix sequencing errors. We also showed a new
- pseudo energy model, taking into account the nearest-neighbour energy model 
-and, to quantify geometrical differences,  the isostericity of all base pairs.
+to identify and fix sequencing errors. We introduced a new
+ pseudo energy model, both taking into account the nearest-neighbour energy model 
+and, to account for geometrical discrepancies,  the isostericity of all base pairs.
 
-By combining those two approaches,  the 
+By combining these two approaches,  the 
 mutational landscape exploration and the pseudo energy model,
- into \texttt{RNApyro}, we show that we can
+ into \texttt{RNApyro}, we obtained a tool to predict the position and 
 correct randomly mutated 5s rRNA sequences. 
@@
  the correction of point-wise error in structured regions (i.e. base paired nucleotides).
- Nonetheless it supplements well existing tools by using previously discarded
+ Nonetheless it should supplement well existing tools, by using previously discarded
 information holding, as shown, a strong signal.

### 2012-10-07 Recomb/results_RECOMB.tex
fix typos
Typos Concordance des temps dans la partie results

@@
 
-\begin{table}
+\begin{table}[b]
 \begin{center}
@@
 in this work),  we clustered together the sequences with identical gap locations. From the $54$ MSAs without gap produced, we selected the
-biggest MSA  which contains $130$ sequences (out of $712$ in the original Rfam MSA). Then, in order to avoid any bias, we used \texttt{cd-hit}
+biggest MSA  which contains $130$ sequences (out of $712$ in the original Rfam MSA). Then, in order to avoid overfitting, we used \texttt{cd-hit}
 \cite{Li:2006fk} to remove sequences with more than 80\% of sequence similarity. This operation resulted in a data set of $45$ sequences.
 
-We design our benchmark using a leave-one-out strategy. We randomly picked one sequence from our data set and performed $12$ random
+We designed our benchmark using a leave-one-out strategy. We randomly picked one sequence from our data set and performed $12$ random
 mutations. Our sequences have $119$ nucleotides, thus the number of mutations corresponds to an error-rate of 10\%. We repeated this operation 
 $10$ times. The value of $\beta$ has been set to $1.5$ (larger values gave similar results). To estimate the impact of the distribution of the weights
-between the energy term and isostericity  score, we use 4 different values of $\alpha = {0, 0.5, 0.8, 1.0}$. Similarly, we also investigate the impact of 
+between the energy term and isostericity  score, we used 4 different values of $\alpha = {0, 0.5, 0.8, 1.0}$. Similarly, we also investigated the impact of 
 an under- and over- estimate of the number of errors, and use values equal to 50\% (6 mutations) and 200\% (24 mutations) of the exact number of
@@
 To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed
-by \RNApyro. More specifically, we fix a threshold $\lambda \in [0,1]$ and we predict an error at position $i$ in sequence $\omega$ if and only if the
-probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceed this threshold. To correct the errors we use the set of nucleotides with a probability
+by \RNApyro. More specifically, we fixed a threshold $\lambda \in [0,1]$, and predicted an error at position $i$ in sequence $\omega$ if and only if the
+probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceeds this threshold. To correct the errors we used the set of nucleotides with a probability
 higher than this threshold is  $C(i) = \{ n \; | \;  n \in \{ A,C,G,U \} \mbox{ and } P(i,n) > \lambda \mbox{ and }  n \neq \omega[i] \}$, where $\omega[i]$ is
 the nucleotide at position $i$ in the input sequence. We note that for the lowest thresholds multiple nucleotides can be available in $C(i)$ to correct
-the sequence. Here, we remind that we aim to estimate the potential of error-correction of \RNApyro and not to develop a error-correction pipe-line. 
-The latter will be investigate in further studies. Finally, we progressively vary $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
-under the curve (AUC). We report our results in Figure \ref{fig:ROCall}.
+the sequence. Here, we remind that our aim is to estimate the potential of error-correction of \RNApyro and not to develop an error-correction pipe-line, which  
+will be the subject of further studies. Finally, we progressively varied $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
+under the curve (AUC). We report our results in Figure~\ref{fig:ROCall}.
 

### 2012-10-07 Recomb/main_RECOMB.tex

--- 2012-10-07 Recomb/main_RECOMB.tex
@@
 \usepackage[numbers]{natbib}
+\usepackage{natbibspacing}
+\setlength{\bibspacing}{0em}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \author{Vladimir Reinharz$^{2}$, Yann Ponty$^{1}\footnote{To whom correspondence should be addressed}\;$, J\'er\^{o}me Waldisp\"{u}hl$^{2*}$}
-\date{$^1$Laboratoire d'Informatique, \'Ecole Polytechnique, Palaiseau, 91120 France $^2$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7}
+\date{$^1$ CNRS/\'Ecole Polytechnique/Inria AMIB, LIX, Palaiseau, France \newline $^2$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7}
 

### 2012-10-07 Recomb/methods_RECOMB.tex

@@
 In principle, this computation should be performed $\Theta(m\cdot n^2)$ times to investigate any triplet $(m,i,j)$.
-However, a close inspection of the recurrences reveal that the computation can be safely restricted to a subset of all intervals.
-For instance, in the inside computation, such intervals $[i,j]$ do not break any base-pair, and their next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
+However, a close inspection of the recurrences reveals that the computation can be safely restricted to a subset of all intervals.
+For instance, in the inside computation, such intervals $[i,j]$ are those that do not break any base-pair, and whose next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
 Similar constraints hold for the outside computation.

### 2012-10-07 Recomb/conclusion_RECOMB.tex
conclusion

@@
 mutational landscape exploration and the pseudo energy model,
- into \texttt{RNApyro}, we obtained a tool to predict the position and 
-correct randomly mutated 5s rRNA sequences. 
-As presented in Sec.~\ref{sec:results},
-we observe that the models
+ into \texttt{RNApyro}, we created a tool predicting the positions
+ yielding sequencing error and their correct value.
+We validated our model with the 5s rRNA,
+as presented in Sec.~\ref{sec:results}.
+We observed that the models
 with larger weights on the

### 2012-10-07 Recomb/conclusion_RECOMB.tex
conclusion

@@
  conserved consensus secondary structures  information,
-to identify and fix sequencing errors. We introduced a new
+to identify and fix sequencing errors. The algorithm runs in  $\Theta(n\cdot(|\Omega|+m^2))$ time and $\Theta(n\cdot(|\Omega|+m))$ memory, where $n$ is the length of the RNA,
+$m$ the number of mutations and $\Omega$ the size of the multiple sequence alignment.
+
+We additionally introduce a new
  pseudo energy model, both taking into account the nearest-neighbour energy model 
@@
 
-By combining these two approaches,  the 
+By combining  into \texttt{RNApyro} these two approaches,  the 
 mutational landscape exploration and the pseudo energy model,
- into \texttt{RNApyro}, we created a tool predicting the positions
- yielding sequencing error and their correct value.
+ we created a tool predicting the positions
+ yielding point-wise sequencing error and correcting them.
 We validated our model with the 5s rRNA,

### 2012-10-07 Recomb/main_RECOMB.tex
main

@@
 \title{A linear inside-outside algorithm for correcting sequencing errors in structured RNA sequences}
-\author{Vladimir Reinharz$^{2}$, Yann Ponty$^{1}\footnote{To whom correspondence should be addressed}\;$, J\'er\^{o}me Waldisp\"{u}hl$^{2*}$}
-\date{$^1$ CNRS/\'Ecole Polytechnique/Inria AMIB, LIX, Palaiseau, France \newline $^2$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7}
+\author{Vladimir Reinharz$^{1}$, Yann Ponty$^{2}\footnote{To whom correspondence should be addressed}\;$, J\'er\^{o}me Waldisp\"{u}hl$^{1*}$}
+\date{$^1$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7\newline $^2$ CNRS/\'Ecole Polytechnique/Inria AMIB, LIX, Palaiseau, France }
 

### 2012-10-07 Recomb/conclusion_RECOMB.tex
conclusion
misc

@@
 
-In this article we presented a new and efficient way of
-exploring the mutational landscape of an RNA, benefiting from the
- conserved consensus secondary structures  information,
-to identify and fix sequencing errors. The algorithm runs in  $\Theta(n\cdot(|\Omega|+m^2))$ time and $\Theta(n\cdot(|\Omega|+m))$ memory, where $n$ is the length of the RNA,
+In this article we presented a new and efficient way of exploring the mutational landscape of an RNA under structural constraints,
+and apply our techniques to identify and fix sequencing errors. In adddition, we introduce a new
+scoring scheme combining the nearest-neighbour energy model to new isostericity matrices in order to account for geometrical
+discrepancies occurring during base pair replacements.
+The algorithm runs in  $\Theta(n\cdot(|\Omega|+m^2))$ time and $\Theta(n\cdot(|\Omega|+m))$ memory, where $n$ is the length of the RNA,
 $m$ the number of mutations and $\Omega$ the size of the multiple sequence alignment.
-
-We additionally introduce a new
- pseudo energy model, both taking into account the nearest-neighbour energy model 
-and, to account for geometrical discrepancies,  the isostericity of all base pairs.
 
@@
 isostericity seems to hold a higher accuracy on the estimation of errors.
+This indicates that an extractable signal is contained in the isostericity.
 Importantly, the implementation is fast enough for practical applications.
@@
 information holding, as shown, a strong signal.
+
+Further research, given the potential of error-correction of \texttt{RNApyro}, 
+will evaluate its impact over large datasets with different existing
+  NGS error-correction pipe-line.

### 2012-10-08 Recomb/conclusion_RECOMB.tex
fixed typo

@@
 In this article we presented a new and efficient way of exploring the mutational landscape of an RNA under structural constraints,
-and apply our techniques to identify and fix sequencing errors. In adddition, we introduce a new
+and apply our techniques to identify and fix sequencing errors. In addition, we introduce a new
 scoring scheme combining the nearest-neighbour energy model to new isostericity matrices in order to account for geometrical

### 2012-10-08 Recomb/main_RECOMB.tex
Revamped part of method section One more hunt for typos
title

--- 2012-10-07 Recomb/main_RECOMB.tex
@@
 %\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
-\title{A linear inside-outside algorithm for correcting sequencing errors in structured RNA sequences}
-\author{Vladimir Reinharz$^{2}$, Yann Ponty$^{1}\footnote{To whom correspondence should be addressed}\;$, J\'er\^{o}me Waldisp\"{u}hl$^{2*}$}
-\date{$^1$CNRS/\'Ecole Polytechnique/INRIA AMIB, LIX UMR 7161 X-CNRS, Palaiseau, France $^2$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7}
+\title{A linear inside-outside algorithm for correcting sequencing errors in structured RNA}
+\author{Vladimir Reinharz, Yann Ponty$^{*}$, J\'er\^{o}me Waldisp\"{u}hl$^*$}
+\date{}
 

### 2012-10-08 benchmark/night.sh

@@ -1,0 +1,4 @@
+./runtest.py -m 24 -a 1.0 -e 10
+./runtest.py -m 24 -a 0.8 -e 10
+./runtest.py -m 24 -a 0.5 -e 10
+./runtest.py -m 24 -a 0.0 -e 10

### 2012-12-18 src/RNAPyroProfile.py
cleaned input options

@@ -728,40 +728,82 @@ def help():
 
 if __name__ == "__main__":
   opts = sys.argv
-  if len(opts) < 5:
-    help()
-    sys.exit(1)
-  file_name = opts[1]
-  if not os.path.isfile(file_name):
-    help()
-    sys.exit(1)
-  profile_path = opts[2]
-  if not os.path.isfile(profile_path):
-    help()
-    sys.exit(1)
-  try:
-    alpha = float(opts[3])
-  except ValueError:
-    help()
-    sys.exit(1)
-  if not 0 <= alpha <= 1:
-    help()
-    sys.exit(1)
-  try:
-    z = float(opts[4])
-    if z >= 0:
-      for x in STACKING_ENERGY:
-        if STACKING_ENERGY[x] == sys.maxint:
-          STACKING_ENERGY[x] = z
-  except ValueError:
-    pass
+  l = len(opts)
+  i = 1
+
+  #Action Flags!!
+  f_no_profile = False#Don't do profile
+  f_backtrack = False#Do backtrack
+
+  while i < l:
+    cmd = opts[i]
+    if not cmd.startswith('-'):
+      i += 1
+    else:
+      #Profile
+      if cmd[1:] == 'p':  
+        file_name = opts[i+1]
+        i += 2
+        if not os.path.isfile(file_name):
+          help()
+          sys.exit(1)
+      #Data (MSA + sec struct)
+      elif cmd[1:] == 'd':
+        profile_path = opts[i+1]
+        i += 2
+        if not os.path.isfile(profile_path):
+          help()
+          sys.exit(1)
+      #alpha
+      elif cmd[1:] == 'a':
+        alpha = opts[i+1]
+        i += 2
+        try:
+          alpha = float(alpha)
+        except ValueError:
+          help()
+          sys.exit(1)
+        if not 0 <= alpha <= 1:
+          help()
+          sys.exit(1)
+      #Max BP penality (def inf.)
+      elif cmd[1:] == 'm':
+        i += 2
+        z = opts[i+1]
+        try:
+          z = float(z)
+          if z >= 0:
+            for x in STACKING_ENERGY:
+              if STACKING_ENERGY[x] == sys.maxint:
+                STACKING_ENERGY[x] = z
+        except ValueError:
+          pass
+      #Don't output profile
+      elif cmd[1:] == 'no_profile':
+        f_no_profile = True
+        i += 1
+      elif cmd[1:] == 'b':
+        f_backtrack = True
[…]

### 2012-12-18 src/bench.sh

@@ -1,0 +1,1 @@
+python RNAPyroProfile.py -p test -d profile.txt -a 0.5 -m 20 -no_profile -b 5

### 2012-12-18 src/RNAPyroProfile.py
fixed help()

@@ -713,17 +713,21 @@ def display_all_probabilities(results):
     print x[0],x[1],x[2],x[3]
 
 def help():
-  print """You need at least three arguments:
-    i) A file containing the MSE and Secondary structure
-    ii)A file containing an RNA profile (i.e. every lines contains
+  print """
+  Required:
+    -p <file_path> A file containing the MSE and Secondary structure
+    -d <file_path> A data file with  RNA profile (i.e. every lines contains
        nucleotides probability in order: 'ACGU')
-    iii) The value of alpha, between 0 and 1.
-    iv) The max penality for an invalid base pair, -1 for infinity
-    v) OPTIONAL:  If a fifth option is inputed, ONE RANDOM best sequence will be backtracked
+    -a <float> The value of alpha, between 0 and 1.
 
-    e.g. 
-      profile)    python RNAPyroProfile mseq_sec_strut.txt profile.txt 0.5 20
-      backtrack)  python RNAPyroProfile mseq_sec_strut.txt profile.txt 0.5 20 y
+  Optional:
+    -m <int> The max penality for an invalid base pair, -1 for infinity
+    -b <int> print 'n' optimal sequences
+    -no_profile <> Doesn't output a profile
+
+  e.g.
+    python RNAPyroProfile -p prof.txt -d data.txt -a 0.5 -m 20
+    python RNAPyroProfile -p prof.txt -d data.txt -a 0.5 -m 20 -b -no_profile
     """
 
 if __name__ == "__main__":
@@ -796,8 +800,13 @@ if __name__ == "__main__":
           sys.exit(1)
 
 
-  ref_seq,struct = parse_fasta(file_name)
-  profile = parse_profile(profile_path)
+  try:
+    ref_seq,struct = parse_fasta(file_name)
+    profile = parse_profile(profile_path)
+  except NameError:
+    help()
+    sys.exit(1)
+
   n = len(struct)
 
   if not f_no_profile:

### 2012-12-18 src/RNAPyroProfile.py
cleaned input

@@ -698,6 +698,14 @@ def parse_profile(file_name):
       profile.append(prob)
   return tuple(profile)
 
+def equiprob_profile(n):
+      prob = {'A':0.25,
+              'C':0.25,
+              'G':0.25,
+              'U':0.25}
+      profile = tuple(prob for _ in xrange(n))
+      return profile
+
 def all_probabilities(profile,ref_seq, stuct, alpha):
   n = len(struct)
   results = []
@@ -715,19 +723,29 @@ def display_all_probabilities(results):
 def help():
   print """
   Required:
-    -p <file_path> A file containing the MSE and Secondary structure
-    -d <file_path> A data file with  RNA profile (i.e. every lines contains
+    -p <file_path> 
+      A file containing the MSE and Secondary structure
+    -d <file_path> 
+      A data file with  RNA profile (i.e. every lines contains
        nucleotides probability in order: 'ACGU')
-    -a <float> The value of alpha, between 0 and 1.
+    -a <float> 
+      The value of alpha, between 0 and 1.
 
   Optional:
-    -m <int> The max penality for an invalid base pair, -1 for infinity
-    -b <int> print 'n' optimal sequences
-    -no_profile <> Doesn't output a profile
+    -m <int> 
+      The max penality for an invalid base pair, -1 for infinity
+    -b <int> 
+      print 'n' optimal sequences
+    -no_profile <> 
+      Doesn't output a profile
+    -s_gc <target_gc> <nb_samples> 
+      Sampling sequences with a 0<=target_gc<=1 and a given
+      number of samples 
 
   e.g.
     python RNAPyroProfile -p prof.txt -d data.txt -a 0.5 -m 20
-    python RNAPyroProfile -p prof.txt -d data.txt -a 0.5 -m 20 -b -no_profile
+    python RNAPyroProfile -p prof.txt -d data.txt -a 0.5 -m 20 -b 5 -no_profile
+    python RNAPyroProfile -d data.txt -a 0.5 -m 20 -no_profile -s_gc 0.5 100
     """
 
 if __name__ == "__main__":
@@ -738,6 +756,7 @@ if __name__ == "__main__":
   #Action Flags!!
   f_no_profile = False#Don't do profile
   f_backtrack = False#Do backtrack
+  f_sample_gc = False#Sample with given gc content
 
   while i < l:
     cmd = opts[i]
@@ -745,14 +764,14 @@ if __name__ == "__main__":
       i += 1
     else:
[…]

### 2012-12-18 src/bench.sh

@@ -1,1 +1,3 @@
-python RNAPyroProfile.py -p test -d profile.txt -a 0.5 -m 20 -no_profile -b 5
+#!/bin/bash
+#python RNAPyroProfile.py -d test -p profile.txt -a 0.5 -m 20 -no_profile -b 5
+python RNAPyroProfile.py -d test -a 0.5 -m 20 -no_profile -s_gc 0.5 100

### 2012-12-18 src/RNAPyroProfile.py
gc_content

@@ -720,6 +720,16 @@ def display_all_probabilities(results):
   for x in results:
     print x[0],x[1],x[2],x[3]
 
+def gc_content(sequence,structure=None):
+  if not structure:
+    nb = len([x for x in sequence if x in 'GC'])
+    return float(nb)/len(sequence)
+  else:
+    nb = len([x for i,x in enumerate(sequence) if x in 'GC' and 
+             structure[i] > -1])
+    nb_bp = len([x for x in structure if x > -1])
+    return float(nb) / nb_bp
+
 def help():
   print """
   Required:
@@ -826,10 +836,10 @@ if __name__ == "__main__":
         nb_gc_sample = opts[i+2]
         i += 3
         try:
-          print "Error s_gc"
           nb_gc_sample = int(nb_gc_sample)
           gc_target = float(gc_target)
         except ValueError:
+          print "Error s_gc"
           help()
           sys.exit(1)
         if not (0 <= gc_target <= 1) or not nb_gc_sample > 0:
@@ -861,7 +871,9 @@ if __name__ == "__main__":
 
   if f_backtrack:
     for _ in xrange(nb_backtrack):
-      print backtrack(profile,ref_seq,struct,(0,n-1),('',''),alpha)
+      res = backtrack(profile,ref_seq,struct,(0,n-1),('',''),alpha)
+      print res, gc_content(res,struct)
+
 
   if f_sample_gc:
     print nb_gc_sample, gc_target

### 2012-12-18 src/bench.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 #python RNAPyroProfile.py -d test -p profile.txt -a 0.5 -m 20 -no_profile -b 5
-python RNAPyroProfile.py -d test -a 0.5 -m 20 -no_profile -s_gc 0.5 100
+python RNAPyroProfile.py -d test -a 0.5 -m 15 -no_profile -b 4 -s_gc 0.5 100

### 2012-12-18 src/RNAPyroProfile.py
cleaned

@@ -21,6 +21,7 @@ sys.setrecursionlimit(10000)
 
 BASES = ['A','C','G','U']
 BOLTZMANN = 0.0019872041
+global T
 T = 310.15
 
 #Just generate all possible combination with maxint
@@ -324,23 +325,6 @@ ISO = {(('A', 'C'), ('C', 'A')):4.93,
         (('G', 'G'), ('U', 'G')):10.0,
         (('A', 'U'), ('U', 'C')):3.5}
 
-class memoize_str(dict):
-  """Generically memoizes a function results."""
-  fun = None
-  
-  def __init__(self, f):
-      self.fun = f
-  
-  def __call__(self,profile,ref_seq,struct,*args):
-      nargs = (args)
-      if nargs in self:
-          return self[nargs]
-      else:
-          val = self.fun(profile,ref_seq,struct,*args)
-          self[nargs] = val
-          return val
-  def resetCache(self):
-      self.clear()
 
 class memoize(dict):
   """Generically memoizes a function results."""
@@ -360,11 +344,30 @@ class memoize(dict):
   def resetCache(self):
       self.clear()
 
+class memoize_iso(dict):
+  """Generically memoizes a function results."""
+  fun = None
+  
+  def __init__(self, f):
+      self.fun = f
+  
+  def __call__(self,ref_seq,*args):
+      nargs = (args)
+      if nargs in self:
+          return self[nargs]
+      else:
+          val = mpf(self.fun(ref_seq,*args))
+          self[nargs] = val
+          return val
+  def resetCache(self):
+      self.clear()
+
 def energy((a,b),(a2,b2),alpha):
   #stacking energy of base pair (a,b) around base pair (a2,b2)
   E = STACKING_ENERGY[a,a2,b2,b]
   return  math.exp(-(alpha*E)/(BOLTZMANN*T))
 
+@memoize_iso
 def isostericity(ref_seq,(i,j),(a,b), alpha):
   #isostericity of going from original base pair to (a,b)
   iso = sum(ISO[(ref[i],ref[j]),(a,b)] for ref in ref_seq)/len(ref_seq)
@@ -751,6 +754,8 @@ def help():
     -s_gc <target_gc> <nb_samples> 
       Sampling sequences with a 0<=target_gc<=1 and a given
       number of samples 
+    -t <float>
+      The temperature (default 310.5K)
 
   e.g.
     python RNAPyroProfile -p prof.txt -d data.txt -a 0.5 -m 20
[…]

### 2012-12-18 src/RNAPyroProfile.py
implemented weights

@@ -733,6 +733,62 @@ def gc_content(sequence,structure=None):
     nb_bp = len([x for x in structure if x > -1])
     return float(nb) / nb_bp
 
+def update_profile(profile,max_bound,min_bound,increase=True):
+  """Given a profile, will increase (resp. decrease) by half the probability
+  of GC content at all positions. Will equaly distribute the remaining between
+  AU
+  """
+  new_profile = []
+  for position in profile:
+    if increase:
+      up_G = (max_bound-position['G'])/2
+      up_C = (max_bound-position['C'])/2
+    else:
+      up_G = -(position['G']-min_bound)/2
+      up_C = -(position['C']-min_bound) /2
+    to_remove = up_G+up_C
+    new_profile.append(
+      {'G':position['G']+up_G,
+       'C':position['C']+up_C,
+       'A':position['A']-to_remove/2})
+    #make sure sums to 1
+    new_profile[-1]['U'] = 1.0 - sum(new_profile[-1][x] for x in 'GCA')
+  return new_profile
+     
+def sample_gc_target(profile,ref_seq,struct,alpha,nb_gc_sample,gc_target,
+                     f_gc_only_structure,max_err=0.1,sample_before_update=100):
+  """Will sample the required number of sequences, inside the gc_target
+
+  """
+  lower_gc = gc_target-max_err
+  upper_gc = gc_target+max_err
+  max_bound = 1.0
+  min_bound = 0.0
+  l_all_sample = []
+  l_correct_gc = []
+  while len(l_correct_gc) < nb_gc_sample:
+    l_all_sample.append([backtrack(profile,ref_seq,struct,(0,n-1),('',''),alpha)
+                         for _ in xrange(sample_before_update)])
+    over_under = 0 
+    for sample in l_all_sample[-1]:
+      if f_gc_only_structure:
+        content = gc_content(sample,structure=struct)
+      else:
+        content = gc_content(sample)
+      over_under += (content - gc_target)
+      if lower_gc <= content <= upper_gc: 
+        l_correct_gc.append(sample)
+    if over_under < 0:
+      min_bound = profile[0]['C']
+      profile = update_profile(profile,max_bound,min_bound) 
+      forward.clear()
+    elif over_under > 0:
+      max_bound = profile[0]['C']
+      profile = update_profile(profile,max_bound,min_bound,increase=False)
+      forward.clear()
+
+  return l_correct_gc
+
 def help():
   print """
   Required:
[…]

### 2012-12-18 src/bench.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 #python RNAPyroProfile.py -d test -p profile.txt -a 0.5 -m 20 -no_profile -b 5
-python RNAPyroProfile.py -d test -a 0.5 -m 15 -no_profile -b 4 -s_gc 0.5 100
+python RNAPyroProfile.py -d test -a 0.5 -m 15 -no_profile -s_gc 0.75 100 -gc_sec_struct -t 899

### 2012-12-18 ISMB/conclusion_ISMB.tex
started ISMB folder

--- 2012-10-07 Recomb/conclusion_RECOMB.tex
@@
 \label{sec:conclusion}
-
-In this article we presented a new and efficient way of
-exploring the mutational landscape of an RNA, benefiting from the
- conserved consensus secondary structures  information,
-to identify and fix sequencing errors. The algorithm runs in  $\Theta(n\cdot(|\Omega|+m^2))$ time and $\Theta(n\cdot(|\Omega|+m))$ memory, where $n$ is the length of the RNA,
-$m$ the number of mutations and $\Omega$ the size of the multiple sequence alignment.
-
-We additionally introduce a new
- pseudo energy model, both taking into account the nearest-neighbour energy model 
-and, to account for geometrical discrepancies,  the isostericity of all base pairs.
-
-By combining  into \texttt{RNApyro} these two approaches,  the 
-mutational landscape exploration and the pseudo energy model,
- we created a tool predicting the positions
- yielding point-wise sequencing error and correcting them.
-We validated our model with the 5s rRNA,
-as presented in Sec.~\ref{sec:results}.
-We observed that the models
-with larger weights on the
-isostericity seems to hold a higher accuracy on the estimation of errors.
-Importantly, the implementation is fast enough for practical applications.
-
-We must recall that our approach is restricted to
- the correction of point-wise error in structured regions (i.e. base paired nucleotides).
- Nonetheless it should supplement well existing tools, by using previously discarded
-information holding, as shown, a strong signal.

### 2012-12-18 ISMB/introduction_ISMB.tex
started ISMB folder

--- 2012-10-07 Recomb/introduction_RECOMB.tex
@@
 \label{sec:introduction}
-
-Ribonucleic acids (RNAs) are found in every living organisms and exhibit a broad range of functions, from catalyzing
-chemical reactions as the RNase P or the group II introns, hybridizing  messenger RNA to regulate gene expression,
-to ribosomal RNA (rRNA) synthesizing proteins. Those functions  require specific structures,  encoded in their nucleotide
-sequence. Although the functions, and thus the structures, need to be preserved through various organisms, the sequences
-can greatly differ from one organism to another. This sequence diversity coupled with the structural conservation is a fundamental
-asset for evolutionary studies. To this end, algorithms to analyze the relationship between RNA mutants and structures are required.
-
-For half a century, biological molecules have been studied as a proxy to understand evolution~\cite{Zuckerkandl1965}, and due
-to their fundamental functions and remarkably conserved structures, rRNAs have always been a prime candidate for phylogenetic
-studies~\cite{Olsen1986, Olsen1993}. In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007} benefited
-of new technologies such as the NGS techniques to sequence as many new organisms as possible and extract an unprecedented flow of new information. 
-Nonetheless, these high-throughput techniques typically have high error rates that make their applications to metagenomics (a.k.a. environmental
-genomics) studies challenging. For instance, pyrosequencing as implemented by Roche's 454 produces may have a error rate raising up to 10\%.
-Because there is no cloning step, resequencing to increase accuracy is not possible and it is therefore vital to disentangle noise from true sequence
-diversity in this type of data \cite{Quince:2009uq}. Errors can be significantly reduced  when large multiple sequence alignments with close homologs
-are available, but in studies of new or not well known organisms, such information is rather sparse. In particular, it is common to is not enough  similarity to 
-differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
-A few techniques have been developed to remedy to this problem~\cite{Quinlan2008,Medvedev2011} but they do not take
-into account all the available information. It is therefore essential to develop methods that can exploit any type of signal available to correct errors.
-
-In this paper, we introduce \RNApyro, a novel algorithm to that enable us to calculate precisely mutational probabilities in RNA sequences with a
-conserved consensus secondary structure. We show how our techniques can exploit the structural information embedded in physics-based energy models, 
-covariance models and isostericity scales to identify and correct point-wise errors in RNA molecules with conserved secondary structure. In particular, we 
-hypothesize that  conserved consensus secondary structures combined with sequence profiles  provide an information that allow us to identify and fix sequencing errors.
-
-Here, we expand the range of algorithmic techniques previously introduced with \RNAmutants~\cite{Waldispuhl2008}.
-Instead of exploring the full conformational landscape and sample mutants, we develop an inside-outside algorithm that enables us
-to explore the complete mutational landscape with a \emph{fixed} secondary structure and to calculate exactly mutational probability values. In addition
-to a gain into the numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot m^2)$ for the
-original version of  \RNAmutants to $\mathcal{O}(n \cdot m^2)$ for \RNApyro, where $n$ is the size of the sequence and $m$ the number of mutations).
-
-We design a new scoring scheme combining nearest-neighbor models \cite{Turner2010} to isostericity metrics \cite{Stombaugh2009}.
-Classical approaches use a Boltzmann distribution whose weights are estimated using a nearest-neighbour energy model~\cite{Turner2010}. However, the
-latter only accounts for  canonical and wobble, base pairs. As was shown by Leontis and Westhof~\cite{Leontis2001},
-the diversity of base pairs observed in tertiary structures is much larger, albeit their energetic contribution remains unknown. To quantify geometrical discrepancies, 
-an isostericity distance has been designed \cite{Stombaugh2009}, increasing as two base pairs geometrically differ from each other in space. Therefore, we 
-incorporate these scores in the Boltzmann weights used by \RNApyro.
-
-We illustrate and benchmark our techniques for point-wise error corrections on the 5S ribosomal RNA. We choose the latter since it has been extensively
-used for phylogenetic reconstructions~\cite{Hori1987} and its sequence has been recovered for over 712 species (in the Rfam seed alignment with id
-\texttt{RF00001}). Using a leave one out strategy, we perform random distributed mutations on a sequence. While our methodology is restricted to the correction of 
-point-wise error in structured regions (i.e. with base pairs), we show that \texttt{RNApyro} can successfully extract a signal that can be used to reconstruct the 
-original sequence with an excellent accuracy. This suggests that \RNApyro is a promising algorithm to complement existing tools in the NGS error-correction 
-pipeline.
-
-The algorithm and the scoring scheme are presented in Sec.~\ref{sec:methods}. Details of the implementation and benchmarks are in Sec.~\ref{sec:results}. 
-Finally, we discuss future developments and applications in Sec.~\ref{sec:conclusion}.

### 2012-12-18 ISMB/main_ISMB.tex
started ISMB folder

--- 2012-10-08 Recomb/main_RECOMB.tex
@@
 
-%\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
-\title{A linear inside-outside algorithm for correcting sequencing errors in structured RNA}
-\author{Vladimir Reinharz, Yann Ponty$^{*}$, J\'er\^{o}me Waldisp\"{u}hl$^*$}
+\newcommand{\Ab}{{\sf{A}}}
+\newcommand{\Cb}{{\sf{C}}}
+\newcommand{\Gb}{{\sf{G}}}
+\newcommand{\Ub}{{\sf{U}}}
+
+\title{}
+\author{}
 \date{}
@@
 \begin{abstract}
-Analysis of the sequence-structure relationship in RNA molecules are essential to evolutionary studies but also to 
-concrete applications such as error-correction methodologies in sequencing technologies. The prohibitive sizes of the
-mutational and conformational landscapes combined with the volume of data to proceed require efficient algorithms 
-to compute sequence-structure properties. More specifically, here we aim to calculate which mutations increase the most the 
-likelihood of a sequence to a given structure and RNA family.\\
-In this paper, we introduce \RNApyro, an efficient linear-time and space inside-outside algorithm that computes exact mutational
-probabilities under secondary structure and evolutionary constraints given as a multiple sequence alignment with a consensus structure.
-We develop a scoring scheme combining classical stacking base pair energies to novel isostericity scales, and apply our techniques
-to correct point-wise errors in 5s rRNA sequences. Our results suggest that \RNApyro is a promising algorithm to complement existing
-tools in the NGS error-correction pipeline.
 
@@
 
-\input{introduction_RECOMB.tex}
+\input{introduction_ISMB.tex}
 
-\input{methods_RECOMB.tex}
+\input{methods_ISMB.tex}
 
-\input{results_RECOMB.tex}
+\input{results_ISMB.tex}
 
-\input{conclusion_RECOMB.tex}
+\input{conclusion_ISMB.tex}
 
-\input{acknowledgments_RECOMB.tex}
+\input{acknowledgments_ISMB.tex}
 

### 2012-12-18 ISMB/methods_ISMB.tex
started ISMB folder

--- 2012-10-07 Recomb/methods_RECOMB.tex
@@
 \newcommand{\ISO}{\text{ISO}}
+We introduce a probabilistic model, which aims at capturing both the stability of the folded RNA and its ability to adopt a predefined 3D conformation.
+To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment.
+In this model, the probability that the nucleotide at a given position needs to be mutated (i.e.
+corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990}.
 
 \subsection{Probabilistic model}
-Our probabilistic model aims at capturing both the stability of the folded RNA and its ability to adopt a predefined 3D conformation.
-To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both 
-the free-energy and the isostericity. 
-More precisely, let $\Omega$ be an gapless RNA alignment, $S$ its associated secondary structure, 
-then each sequence $s$ has probability proportional to its Boltzmann factor
+Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
+then any sequence $s$ is assigned a probability proportional to its Boltzmann factor
 \begin{align*}
-  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT} &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
+  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
 \end{align*}
 where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
-are the free-energy and isostericity contributions respectively, further described below, and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
+are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
 
@@
 \end{align*}
-where $\ES^{\beta}_{ab\to a'b'}$ is the tabulated free-energy of a stacking pair in the Turner model if available, $0$ if $ab=\varnothing$ (no base-pair to stack onto), and using default value $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
-not in $\{\text{GU},\text{UG},\text{CG},\text{GC}, \text{AU or UA}\}$).
-This latter parameter allows to choose whether to completely forbid a sequence inducing invalid base pairs ($\beta = \infty$), or to simply penalize it.
-The imprecision introduced by this simplification of the Turner model remains reasonable, since the targeted secondary structure is fixed 
+where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
+not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$).
+This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).
+The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed 
 (e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
 \subsubsection{Isostericity contribution}
-The concept of isostericity captures the geometric superimposability of two base-pairs, using individual scores given by Stombaugh~\emph{et al}~\cite{Stombaugh2009} based on a probabilistic model.
-Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gapless RNA alignment $\Omega$,  the contribution of the isostericity to the pseudo-energy is given by
+The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
+Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the contribution of the isostericity to the pseudo-energy as
 \begin{align*}
@@
 \end{align*}
-is the average isostericity of a base-pair in the candidate sequence in comparison with the reference alignment.
+is the average isostericity of a base-pair in the candidate sequence, compared with the reference alignment.
 The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
 Isostericity scores range between $0$ and $9.7$, with $0$ being assigned to a perfect isostericity, and a penalty of $10$ is used for missing entries.
-The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the alignment sequences.
+The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
 
@@
 
-Let $s$ an RNA 
-sequence and $m\geq 0$. $S$ is considered as  one derivation of the 
-{SCFG} generating  
-all  secondary structures of length  $|s|$. We are interested in  
-the probability of a given position being a specific nucleotide,
-over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$ 
-(formally 
-$\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
-We define as a variant of the
+Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
+We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $m$ mutations from $s$
+%  under the SCFG derivation $S$ 
+(formally $\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
+We define a variant of the
  \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
  desired probability,  the two  functions
-$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former presented in 
-Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec}, our 
-version of the \emph{inside} algorithm, computes the partition function between $i$ and $j$ included, 
-knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing 
-$m$ mutations. The latter in Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
- computes the \emph{outside}, in particular,  
-the partition function considering only between $[0,i]\cup[j,n-1]$
-knowing  that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing 
- $m$ mutations outside.
+$\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.
+
+The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec}, is analogous to the \emph{inside} algorithm.
+It is the partition function, i.e. the sum of Boltzmann factors, over all sequences within $[i,j]$, 
+knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$), within 
+$m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
+ computes the \emph{outside} algorithm, i.e.  
+the partition function over sequences within $m$ mutations of $s$, restricted to two intervals $[0,i]\cup[j,n-1]$, and 
+knowing  that position $i+1$ is composed of $a$ (resp.
+$j-1$ is $b$).
+A suitable combination of these terms, given in Equation~\eqref{eq:combine}, gives the total weight, and in turn the probability, of seeing a specific base at a given position.
+
+%Drawing a parallel with stochastic context-free grammars (SCFG), for which the inside-outside algorithm was introduced, the set of sequences can be seen as the language of words having length $n$, generated from a stochastic context-free grammar.
+%These rules would limit the number of mutations (i.e. )
+% $S$ can be considered as constraining weighing on the shape of parse trees the derivations in a classic 
+% of the {SCFG} generating  all  secondary structures of length  $|s|$.
 
 \subsubsection{Definitions}
-Let be $B:=\left\{\text{A},\text{C},\text{G},\text{U}\right\}$, the set of nucleotides.
+Let $B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
 Given $s\in B^n$ an RNA sequence, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
@@
  ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
- $[i,j]\subset[k,l])$. Let $R$ be the Boltzmann constant, $T$ the temperature in Kelvin and
-  the function $\delta$ such that: 
- $\forall a,a' \in B, \delta_{a,a'}:=\left\{\begin{array}{ll}
-															1 & \text{If } a\equiv a'\\
-															0 & \text{Otherwise.}
-														\end{array}\right.$
+ $[i,j]\subset[k,l])$.
 
-Finally, we denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
+Let us then remind a Hamming distance function $\delta: B^*\times B^* \to \mathbb{N}^+$, which takes two sequences $s'$ and $s''$ as input, $|s'|=|s''|$, and returns the number of differing positions.
+Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
 \begin{equation}
@@
 
-The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function considering only the 
-energy in subsequence $[i,j]$ over mutants of $s$ having exactly $m$ mutations between $[i,j]$ and whose nucleotide at position $i-1$ is $a$ (resp. $b$ in position $j+1$).
-We define function $\Z{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
+The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function, i.e.
+the sum of Boltzmann factors, over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively).
+Such terms can be defined by recurrence, for which the following initial conditions holds:
 
@@
 \end{equation}
-In other words, when we evaluate the function $\mathcal Z$, after exhausting all positions, there 
-is only one possible solution if there is $0$ mutations left, and none else. Since the 
-energetic terms only depend on base pairs, they are not involved in the initial conditions.
+In other words, the set of sequences at distance $m$ of the empty sequence is either empty if $m>0$, or restricted to the empty sequence, having energy $0$, if $m=0$.
+Since the energetic terms only depend on base pairs, they are not involved in the initial conditions.
 
-The recursion itself is composed of four terms:
+The main recursion is composed of four terms:
 \begin{equation}
@@
 \begin{description}
-\item[$S_{i}=-1$:] If the nucleotide at position $i$ is not paired, then the value is the same
-as if we increase the lower interval bound by $1$ (i.e. $i+1$), and consider all possible
- nucleotides $a'$ at position $i$, updating the number of mutants in function of $\Kron_{a',s_i}$. 
-\item[$S_i=j$ and $S_{i-1}=j+1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i-1$ is
-paired with $j+1$, we are in the only case were stacked base pairs can occur. We thus add
-the energy of the stacking and of the isostericity of the base pair $(i,j)$. What is left
-to compute is the \emph{inside} value of the interval $[i+1,j-1]$ over all possible nucleotides 
-$a',b'\in B^2$ at positions $i$ and $j$ respectively.
-\item[$S_i=k$ and $i<k \leq j$:] If nucleotide $i$ is paired with position $k$ 
-but is not stacked outside, the 
-only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. This 
-creates
-two different intervals for which we must compute the values, $[i+1,k-1]$ and $[k+1,j-1]$, for 
-all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $j$ respectively.
-\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the 
-secondary structure $S$, and we return $0$.
+\item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
+any sequence consists in a, possibly mutated, nucleotide $a'$ at position  $i$, 
+followed by a sequence over $[i+1,j]$ having either $m-\Kron_{a',s_i}$, accounting for a possible mutation 
+at position $i$, and having flanking nucleotides $a'$ and $b$.
+\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two, possibly mutated, nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$ having distance $m-\Kron_{a'b',s_is_j}$ (to avoid exceeding the targeted distance $m$).
+Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
+then a stacking energy contribution is added. 
+\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
+only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
+Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$.
+Since the number of mutations sum to $m$ over the whole sequence must , then a parameter $m'$ is introduced to distribute the remaining mutations between the two sequences.
+\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
 \end{description}
@@
 
-\subsection{Complexity}
-Equations~\eqref{eq:Z_rec} and~\eqref{eq:Y_rec} can be computed using dynamic programming, i.e.
-$\mathcal{Z}^{*}_{*}$ or $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and interval lengths, memorizing the results such that they become available in constant time during future stages of the computation.
-Furthermore, energy terms $E(\cdot)$ can be accessed in constant time (after a precomputation in $\Theta(n\cdot|\Omega|)$ for the isostericity part).
+\subsection{Complexity considerations}
+Equations~\eqref{eq:Z_rec} and~\eqref{eq:Y_rec} can be computed using dynamic programming.
+Namely, the $\mathcal{Z}^{*}_{*}$ and $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and interval lengths, memorizing the results as they become available to ensure constant-time access during later stages of the computation.
+Furthermore, energy terms $E(\cdot)$ can be accessed in constant time thanks to a simple precomputation (not described)  of the isostericity contributions in $\Theta(n\cdot|\Omega|)$.
 Computing any given term therefore requires $\Theta(m)$ operations.
 
-In principle, this computation should be performed $\Theta(m\cdot n^2)$ times to investigate any triplet $(m,i,j)$.
-However, a close inspection of the recurrences reveal that the computation can be safely restricted to a subset of all intervals.
-For instance, in the inside computation, such intervals $[i,j]$ do not break any base-pair, and their next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
-Similar constraints hold for the outside computation.
-Such constraints drastically limit the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored.
-Consequently the overall complexity of the algorithm is $\Theta(n\cdot(|\Omega|+m^2))$ aritmetic operations and $\Theta(n\cdot(|\Omega|+m))$ memory.
+In principle, $\Theta(m\cdot n^2)$ terms, identified by $(m,i,j)$ triplets, should be computed.
+However, a close inspection of the recurrences reveals that the computation can be safely restricted to a subset of intervals $(i,j)$.
+For instance, the inside algorithm only requires computing intervals $[i,j]$ that do not break any base-pair, and whose next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
+Similar constraints hold for the outside computation, resulting in a drastic limitation of the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored.
+Consequently the overall complexity of the algorithm is $\Theta(n\cdot(|\Omega|+m^2))$ arithmetic operations and $\Theta(n\cdot(|\Omega|+m))$ memory.

### 2012-12-18 ISMB/results_ISMB.tex
started ISMB folder

--- 2012-10-07 Recomb/results_RECOMB.tex
@@
 \label{sec:results}
-
-\subsection{Implementation}
-The software was implemented in Python2.7 using the \textit{mpmath}~\cite{mpmath} library
-for  arbitrary floating point precision. The source code is freely available at \verb+https://github.com/McGill-CSB/RNApyro+.
-
-The time benchmarks were done on a MacBook Pro 2010, 2 processors at 2.66GHz, 8GB of RAM.
-Since applications of \RNApyro implies a need for
- efficiency and scalability. In Table~\ref{tab:time} we
-present different times needed to compute the probabilities for  every nucleotide at every positions for a vast set of parameters. For those test
- all the sequences were generated randomly, as the target secondary structure.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{lccc}
-$|s|$&\multicolumn{3}{c}{Number of mutations}\\\cline{2-4}
-		 			  & 6   &  12  & 24\\\cline{2-4}
-100  				& 35s  & 238s & 1023s\\
-300  			& 135s & 594s &2460s\\\cline{2-4}
-		 						& 25   &    &	50		\\\cline{2-4}
-500         & 5400s&       &  21003s    \\
-\end{tabular}
-\end{center}
-\caption{Time to compute all probabilities. The first column indicates the length and  the column indexes indicate the number
- of mutations. $\alpha$ is
-set at $0.5$,  $\beta$ to $1.5$ and $|\Omega|=44$.}
-\label{tab:time}
-\end{table}
-
-\subsection{Error correction in 5s rRNA}
-
-To illustrate the potential of our algorithm, we applied our techniques to identify and correct point-wise errors in RNA sequences
-with conserved secondary structures. More precisely, we used \RNApyro to reconstruct 5s rRNA sequences with randomly distributed
-mutations. This experiment has been designed to suggest further applications to error-corrections in pyrosequencing data.
-
-We built our data set from the 5S rRNA multiple sequence alignment (MSA) available in the Rfam Database 11.0 (Rfam id: \texttt{RF00001}).
-Since our software does not currently implement gaps (mainly because scoring indels is a challenging issue that cannot be fully addressed
-in this work),  we clustered together the sequences with identical gap locations. From the $54$ MSAs without gap produced, we selected the
-biggest MSA  which contains $130$ sequences (out of $712$ in the original Rfam MSA). Then, in order to avoid any bias, we used \texttt{cd-hit}
-\cite{Li:2006fk} to remove sequences with more than 80\% of sequence similarity. This operation resulted in a data set of $45$ sequences.
-
-We design our benchmark using a leave-one-out strategy. We randomly picked one sequence from our data set and performed $12$ random
-mutations. Our sequences have $119$ nucleotides, thus the number of mutations corresponds to an error-rate of 10\%. We repeated this operation 
-$10$ times. The value of $\beta$ has been set to $1.5$ (larger values gave similar results). To estimate the impact of the distribution of the weights
-between the energy term and isostericity  score, we use 4 different values of $\alpha = {0, 0.5, 0.8, 1.0}$. Similarly, we also investigate the impact of 
-an under- and over- estimate of the number of errors, and use values equal to 50\% (6 mutations) and 200\% (24 mutations) of the exact number of
-errors (i.e. 12).
-
-To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed
-by \RNApyro. More specifically, we fix a threshold $\lambda \in [0,1]$ and we predict an error at position $i$ in sequence $\omega$ if and only if the
-probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceed this threshold. To correct the errors we use the set of nucleotides with a probability
-higher than this threshold is  $C(i) = \{ n \; | \;  n \in \{ A,C,G,U \} \mbox{ and } P(i,n) > \lambda \mbox{ and }  n \neq \omega[i] \}$, where $\omega[i]$ is
-the nucleotide at position $i$ in the input sequence. We note that for the lowest thresholds multiple nucleotides can be available in $C(i)$ to correct
-the sequence. Here, we remind that we aim to estimate the potential of error-correction of \RNApyro and not to develop a error-correction pipe-line. 
-The latter will be investigate in further studies. Finally, we progressively vary $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
-under the curve (AUC). We report our results in Figure \ref{fig:ROCall}.
-
-\begin{figure}
-\begin{subfigure}[b]{0.3\textwidth}
-\includegraphics[width=1.2\textwidth]{figures/ROC_6.pdf}
-\caption{6 mutations}
-\label{fig:ROC6mut}
-\end{subfigure}
-\hfill
-\begin{subfigure}[b]{0.3\textwidth}
-\includegraphics[width=1.2\textwidth]{figures/ROC_12.pdf}
-\caption{12 mutations}
-\label{fig:ROC12mut}
-\end{subfigure}
-\hfill
-\begin{subfigure}[b]{0.3\textwidth}
-\includegraphics[width=1.2\textwidth]{figures/ROC_24.pdf}
-\caption{24 mutations}
-\label{fig:ROC24mut}
-\end{subfigure}
-\caption{Performance of error-correction. Subfigures accuracy with under-estimated error rate (6 mutations), exact estimates (12 mutations) and over estimates 
-(24 mutations). We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values 
-ranging of $\alpha=\{0,0.5,0.8,1.0\}$. The AUC is indicated in the legend of the figures. Each individual ROC curve represent the average performance over the 10 experiments.}
-\label{fig:ROCall}
-\end{figure}
-
-Our data demonstrate that our algorithm exhibits interesting performance for error-correction applications. First, the AUC values (up to $0.86$) indicate that a
-signal has been successfully extracted. This result has been achieved with errors in loop regions (i.e. without base pairing information) and thus suggests
-that correction rates in structured regions (i.e. base paired regions) could be even higher. Next, the optimal values of $\alpha$ tend to be close to $0.0$. This 
-finding suggests that at this point the information issued from the stacking energy is currently modest. However, specific examples showed improved performance
-using this energy term. Further studies must be conducted to understand how to make the best use it. Finally, our algorithm seems robust to the number of
-mutations considered. Indeed, good AUC values are achieved with low estimates of the number of errors in the sequence (c.f. 50\% of the errors  in
-Fig.~\ref{fig:ROC6mut}) and with large  values (c.f.~200\% of the errors  in Fig.~\ref{fig:ROC24mut}). It is worth noting that scoring scheme with larger weight on
-the isostericity scores (low $\alpha$ values) seem more robust to under- and over-estimate of the number of errors.

### 2012-12-18 src/RNAPyroProfile.py
added -gc_data interface

@@ -756,7 +756,8 @@ def update_profile(profile,max_bound,min_bound,increase=True):
   return new_profile
      
 def sample_gc_target(profile,ref_seq,struct,alpha,nb_gc_sample,gc_target,
-                     f_gc_only_structure,max_err=0.1,sample_before_update=100):
+                     file_gc_data,f_gc_only_structure,
+                     max_err=0.1,sample_before_update=100):
   """Will sample the required number of sequences, inside the gc_target
 
   """
@@ -815,6 +816,9 @@ def help():
       will be considered for the GC content
     -gc_max_err <float>
       Max error from GC target allowed in sample, default 0.1
+    -gc_data <file_name>
+      Create a file with the given name, will print to it a first line
+      the weight of 'C' and on the next the list of sampled GC content
     -t <float>
       The temperature (default 310.5K)
 
@@ -835,6 +839,7 @@ if __name__ == "__main__":
   f_sample_gc = False#Sample with given gc content
   f_gc_only_structure = False#Only take into account GC with interactions
   f_gc_max_error = False#Custom max error for gc content, default 0.1
+  file_gc_data#If a file given, print to it for each iter. weight C, GC content
 
   while i < l:
     cmd = opts[i]
@@ -928,6 +933,7 @@ if __name__ == "__main__":
         except ValueError:
           help()
           sys.exit(1)
+      #Max error from GC_Target allowed
       elif cmd == '-gc_max_err':
         f_gc_max_error = True
         max_error = opts[i+1]
@@ -942,6 +948,10 @@ if __name__ == "__main__":
           print "Error gc_max_err"
           help()
           sys.exit(1)
+      #If want data from GC sampling
+      elif cmd == '-gc_data':
+        file_gc_data = opts[i+1]
+        i += 2
       else:
         print "Unrecognized arguement"
         help()
@@ -976,9 +986,11 @@ if __name__ == "__main__":
   if f_sample_gc:
     if f_gc_max_error:
       res = sample_gc_target(profile,ref_seq,struct,alpha,nb_gc_sample,
-                             gc_target,f_gc_only_structure,max_err=max_error)
+                             gc_target,file_gc_data,f_gc_only_structure,
+                             max_err=max_error)
     else:
       res = sample_gc_target(profile,ref_seq,struct,alpha,
-                             nb_gc_sample,gc_target,f_gc_only_structure)
+                             nb_gc_sample,gc_target,file_gc_data,
+                             f_gc_only_structure)
     for x in res:
       print x

### 2012-12-18 src/RNAPyroProfile.py
can record weights data

@@ -767,18 +767,39 @@ def sample_gc_target(profile,ref_seq,struct,alpha,nb_gc_sample,gc_target,
   min_bound = 0.0
   l_all_sample = []
   l_correct_gc = []
+
+  if file_gc_data:
+    l_contents = []
+    if os.path.isfile(file_gc_data):
+      if  raw_input("""The file, %s, to write the weight/content
+                       data already exist, do you want to
+                       overwrite it? Y/n\n""" % file_gc_data) in ('n','N',
+                                                                  'no','NO'):
+        print 'You decided to exit'
+        sys.exit(0)
+    f = open(file_gc_data,'w')
+
   while len(l_correct_gc) < nb_gc_sample:
     l_all_sample.append([backtrack(profile,ref_seq,struct,(0,n-1),('',''),alpha)
                          for _ in xrange(sample_before_update)])
     over_under = 0 
+    l_contents[:] = []
     for sample in l_all_sample[-1]:
       if f_gc_only_structure:
         content = gc_content(sample,structure=struct)
       else:
         content = gc_content(sample)
+      if file_gc_data:
+        l_contents.append(content)
       over_under += (content - gc_target)
       if lower_gc <= content <= upper_gc: 
         l_correct_gc.append(sample)
+
+    if file_gc_data:
+        f.write('%s\n' % profile[0]['C'])
+        for c in l_contents[:-1]:
+          f.write('%s\t' % c)
+        f.write('%s\n' % l_contents[-1])
     if over_under < 0:
       min_bound = profile[0]['C']
       profile = update_profile(profile,max_bound,min_bound) 
@@ -787,6 +808,9 @@ def sample_gc_target(profile,ref_seq,struct,alpha,nb_gc_sample,gc_target,
       max_bound = profile[0]['C']
       profile = update_profile(profile,max_bound,min_bound,increase=False)
       forward.clear()
+
+  if file_gc_data:
+    f.close()
 
   return l_correct_gc
 
@@ -839,7 +863,7 @@ if __name__ == "__main__":
   f_sample_gc = False#Sample with given gc content
   f_gc_only_structure = False#Only take into account GC with interactions
   f_gc_max_error = False#Custom max error for gc content, default 0.1
-  file_gc_data#If a file given, print to it for each iter. weight C, GC content
+  file_gc_data=None#If a file given, print to it for each iter. weight C, GC content
 
   while i < l:
     cmd = opts[i]

### 2012-12-18 src/bench.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 #python RNAPyroProfile.py -d test -p profile.txt -a 0.5 -m 20 -no_profile -b 5
-python RNAPyroProfile.py -d test -a 0.5 -m 15 -no_profile -s_gc 0.75 100 -gc_sec_struct -t 899
+python RNAPyroProfile.py -d test -a 0.5 -m 15 -no_profile -s_gc 0.75 100 -gc_sec_struct -gc_data weights.dat

### 2012-12-19 benchmark/weight_calibration.sh
removed required MSA

@@ -1,0 +1,4 @@
+#!/bin/bash
+#python RNAPyroProfile.py -d test -p profile.txt -a 0.5 -m 20 -no_profile -b 5
+cd ../src
+python RNAPyroProfile.py -d "../benchmark/RF00001_40.in" -a 0.5 -m 15 -no_profile -s_gc 0.5 100 -gc_sec_struct -gc_data weights.dat 

### 2012-12-19 src/RNAPyroEx.py

@@ -356,6 +356,8 @@ def energy((a,b),(a2,b2),alpha):
   return  math.exp(-(alpha*E)/(BOLTZMANN*T))
 
 def isostericity(seq,ref_seq,(i,j),(a,b), alpha):
+  if not ref_seq:
+    return 1
   #isostericity of going from original base pair to (a,b)
   iso_mut = sum(ISO[(ref[i],ref[j]),(a,b)] for ref in ref_seq)
   iso_start = sum(ISO[(ref[i],ref[j]),(seq[i],seq[j])] for ref in ref_seq)

### 2012-12-19 src/RNAPyroProfile.py

@@ -1,7 +1,8 @@
-import os,sys
 import itertools
 import random
 import math
+import sys
+import os
 
 def MPMATH_MISSING():
   print """The module `mpmath` was not found. This might impeed the
@@ -370,6 +371,8 @@ def energy((a,b),(a2,b2),alpha):
 @memoize_iso
 def isostericity(ref_seq,(i,j),(a,b), alpha):
   #isostericity of going from original base pair to (a,b)
+  if not ref_seq:
+    return 1
   iso = sum(ISO[(ref[i],ref[j]),(a,b)] for ref in ref_seq)/len(ref_seq)
   #iso_start = sum(ISO[(ref[i],ref[j]),(seq[i],seq[j])] for ref in ref_seq)
   #iso = mpf(iso_mut-iso_start)/len(ref_seq)
@@ -571,7 +574,6 @@ def backtrack(profile,ref_seq,struct,(i,j),(a,b),alpha):
       a2 = random.choice(result_list)
       max_seq = a2 + backtrack(profile,ref_seq,struct,(i+1,j),(a2,b),alpha)
       
-
     elif i < k <= j: #If k > j we return 0
       for a2 in BASES:
         for b2 in BASES:
@@ -770,16 +772,10 @@ def sample_gc_target(profile,ref_seq,struct,alpha,nb_gc_sample,gc_target,
 
   if file_gc_data:
     l_contents = []
-    if os.path.isfile(file_gc_data):
-      if  raw_input("""The file, %s, to write the weight/content
-                       data already exist, do you want to
-                       overwrite it? Y/n\n""" % file_gc_data) in ('n','N',
-                                                                  'no','NO'):
-        print 'You decided to exit'
-        sys.exit(0)
     f = open(file_gc_data,'w')
 
   while len(l_correct_gc) < nb_gc_sample:
+    print profile[0]['C']
     l_all_sample.append([backtrack(profile,ref_seq,struct,(0,n-1),('',''),alpha)
                          for _ in xrange(sample_before_update)])
     over_under = 0 
@@ -796,7 +792,6 @@ def sample_gc_target(profile,ref_seq,struct,alpha,nb_gc_sample,gc_target,
         l_correct_gc.append(sample)
 
     if file_gc_data:
-        f.write('%s\n' % profile[0]['C'])
         for c in l_contents[:-1]:
           f.write('%s\t' % c)
         f.write('%s\n' % l_contents[-1])
[…]

### 2012-12-19 src/RNAPyroSeedless.py

@@ -348,6 +348,8 @@ def energy((a,b),(a2,b2),alpha):
   return  math.exp(-(alpha*E)/(BOLTZMANN*T))
 
 def isostericity(ref_seq,(i,j),(a,b), alpha):
+  if not ref_seq:
+    return 1
   #isostericity of going from original base pair to (a,b)
   iso = sum(ISO[(ref[i],ref[j]),(a,b)] for ref in ref_seq)/len(ref_seq)
   #iso_start = sum(ISO[(ref[i],ref[j]),(seq[i],seq[j])] for ref in ref_seq)

### 2012-12-19 src/bench.sh

@@ -1,3 +1,0 @@
-#!/bin/bash
-#python RNAPyroProfile.py -d test -p profile.txt -a 0.5 -m 20 -no_profile -b 5
-python RNAPyroProfile.py -d test -a 0.5 -m 15 -no_profile -s_gc 0.75 100 -gc_sec_struct -gc_data weights.dat

### 2012-12-19 src/RNAPyroProfile.py
stochastic backtrack implemented

@@ -541,6 +541,18 @@ def product_given_i(profile,ref_seq,struct,i,a,alpha):
       result += pro*f*b*iso
   return result
 
+def random_weighted_sampling(l_samples):
+  tot = sum(x[1] for x in l_samples)
+  scaled_weights = [x[1]/tot for x in l_samples]
+  rand_nb = random.random()
+  accumulation = 0
+  for i,x in enumerate(scaled_weights):
+    accumulation += x
+    if accumulation > rand_nb:
+      return l_samples[i][0]
+  return l_samples[-1][0]
+
+
 def backtrack(profile,ref_seq,struct,(i,j),(a,b),alpha):
   #alpha gives the weight energy vs isostericity
   max_result = 0.
@@ -552,6 +564,7 @@ def backtrack(profile,ref_seq,struct,(i,j),(a,b),alpha):
   else:
     k = struct[i]
     if k==-1:
+      l_samples = []
       for a2 in BASES:
         pro = profile[i][a2]
         f = forward(profile,ref_seq,struct,
@@ -559,22 +572,12 @@ def backtrack(profile,ref_seq,struct,(i,j),(a,b),alpha):
                     (a2,b),
                     alpha)
         result =  pro*f
-        if result > max_result:
-          max_result = result
-          result_list = [a2]
-        elif result == max_result:
-          max_result = result
-          #result_list = [a2]
-          result_list.append(a2)
-      """
-      for a2 in result_list:
-        for best in backtrack(profile,ref_seq,struct,(i+1,j),(a2,b),alpha):
-          max_seq_list.append(a2 + best)
-      """
-      a2 = random.choice(result_list)
+        l_samples.append((a2,result))
+      a2 = random_weighted_sampling(l_samples)
       max_seq = a2 + backtrack(profile,ref_seq,struct,(i+1,j),(a2,b),alpha)
       
     elif i < k <= j: #If k > j we return 0
+      l_samples = []
       for a2 in BASES:
         for b2 in BASES:
           pro = profile[i][a2]*profile[k][b2]
@@ -607,19 +610,14 @@ def backtrack(profile,ref_seq,struct,(i,j),(a,b),alpha):
                                (a2,b2),
                                alpha)
             result = pro*f*e*iso
-          if result > max_result:
-            max_result = result
-            result_list = [(a2,b2)]
-          elif result == max_result:
[…]

### 2012-12-19 src/RNAPyroProfile.py
fixed output backtrack

@@ -1019,4 +1019,4 @@ if __name__ == "__main__":
       res = sample_gc_target(profile,ref_seq,struct,alpha,
                              nb_gc_sample,gc_target,file_gc_data,
                              f_gc_only_structure)
-      print diversity_seq(res,struct)
+      print '\n'.join(res)

### 2012-12-19 benchmark/bench_weight.py
added benchmark file/executable (bench_weight.py)

@@ -1,0 +1,50 @@
+import sys
+import os
+import subprocess
+import cPickle
+
+def benchmark(list_sec_struc,weights,nb,output_cPickle=None):
+  """f
+  Will output a dictionary with keys (sec_struct,gc_content_target)
+  and value a tuple (res,data_dump)
+    res = list of sequences with +- 10% of GC target
+    data_dump = list alterning 'C' weight in profile,
+                list of GC contents for sequences in given round
+
+  If a 'output_cPickle' arguement is given, will save the dic in it
+  """
+  d_all_bench = {}
+  tmp_struct_file = '.sec_struct.tmp'
+  content_dump = '.content.tmp'
+  present_loc = os.path.abspath('.')
+  cmd = 'python ../src/RNAPyroProfile.py -d "%s" -a 1 -m 15 -no_profile -s_gc %s %s -gc_sec_struct -gc_data "%s"'
+  for struct in list_sec_struct:
+    results = []
+    with open(tmp_struct_file,'w') as f:
+      f.write(struct)
+    for w in weights:
+      results.append([])
+      full_path_struct = os.path.join(present_loc,tmp_struct_file)
+      full_path_content = os.path.join(present_loc,content_dump)
+      print w,nb,full_path_struct,full_path_content
+      to_do = cmd % (full_path_struct,
+                   w,nb,
+                   full_path_content)
+      res = subprocess.check_output(to_do,shell=True)
+      print struct,w,tac-tic
+      data_dump = [x.strip() for x in open(full_path_content)]
+      d_all_bench[struct,w] = (res,data_dump)
+  if output_cPickle:
+    with open(output_cPickle, 'wb') as f:
+      cPickle.dump(d_all_bench,f,-1)
+    os.remove(tmp_struct_file)
+  return d_all_bench
+
+if __name__ == '__main__':
+  list_sec_struct = [x.strip() for x in open(
+    os.path.join('..','data','rnastrand_dataset_filtered_nodup.txt')
+    )]
+  nb = 1000
+  weights = [0.1,0.3,0.5,0.7,0.9]
+  output_file = 'dump_rnastrand_dataset_filtered_nodup.cPickle'
+  benchmark(list_sec_struct,weights,nb,output_file)

### 2012-12-19 src/RNAPyroProfile.py

@@ -773,7 +773,6 @@ def sample_gc_target(profile,ref_seq,struct,alpha,nb_gc_sample,gc_target,
     f = open(file_gc_data,'w')
 
   while len(l_correct_gc) < nb_gc_sample:
-    print profile[0]['C']
     l_all_sample.append([backtrack(profile,ref_seq,struct,(0,n-1),('',''),alpha)
                          for _ in xrange(sample_before_update)])
     over_under = 0 
@@ -790,9 +789,10 @@ def sample_gc_target(profile,ref_seq,struct,alpha,nb_gc_sample,gc_target,
         l_correct_gc.append(sample)
 
     if file_gc_data:
-        for c in l_contents[:-1]:
-          f.write('%s\t' % c)
-        f.write('%s\n' % l_contents[-1])
+      f.write('%s\n' % profile[0]['C']) 
+      for c in l_contents[:-1]:
+        f.write('%s\t' % c)
+      f.write('%s\n' % l_contents[-1])
     if over_under < 0:
       min_bound = profile[0]['C']
       profile = update_profile(profile,max_bound,min_bound) 

### 2012-12-19 benchmark/bench_weight.py
typo...

@@ -31,7 +31,7 @@ def benchmark(list_sec_struc,weights,nb,output_cPickle=None):
                    w,nb,
                    full_path_content)
       res = subprocess.check_output(to_do,shell=True)
-      print struct,w,tac-tic
+      print struct,w
       data_dump = [x.strip() for x in open(full_path_content)]
       d_all_bench[struct,w] = (res,data_dump)
   if output_cPickle:

### 2013-01-01 src/RNAPyroEx.py
RNAPyroEx backtrack buggy

@@ -1,5 +1,6 @@
 import os,sys
 import itertools
+import random
 import math
 
 def MPMATH_MISSING():
@@ -415,6 +416,79 @@ def forward(seq,ref_seq,struct,(i,j),(a,b),m, alpha):
                                                               alpha)
   return result
 
+def random_weighted_sampling(l_samples):
+  tot = sum(x[1] for x in l_samples)
+  if tot == 0:
+    print 'euurrrr'
+    return random.choice([x[0] for x in l_samples])
+  scaled_weights = [x[1]/tot for x in l_samples]
+  rand_nb = random.random()
+  accumulation = 0
+  for i,x in enumerate(scaled_weights):
+    accumulation += x
+    if accumulation > rand_nb:
+      return l_samples[i][0]
+  return l_samples[-1][0]
+
+def backtrack(seq,ref_seq,struct,(i,j),(a,b),m, alpha):
+  #alpha gives the weight energy vs isostericity
+  result = 0.
+  max_seq = ''
+  if m<0: return seq[i:j+1]
+  if i > j : return ''
+  else:
+    k = struct[i]
+    if k==-1:
+      l_samples = []
+      for a2 in BASES:
+        d = delta(seq,i,a2)
+        if d <= m:
+          result = forward(seq,ref_seq,struct,
+                            (i+1,j),
+                            (a2,b),
+                            m-d,alpha)
+          l_samples.append((a2,result))
+      a2 = random_weighted_sampling(l_samples)
+      d = delta(seq,i,a2)
+      max_seq = a2 + backtrack(seq,ref_seq,struct,(i+1,j),(a2,b),m-d,alpha)
+    elif i < k <= j: #If k > j we return 0
+      l_samples=[]
+      for a2 in BASES:
+        for b2 in BASES:
+          d = delta(seq,i,a2)+delta(seq,k,b2)
+          #if not stacked outside (or border, then no stack possible)
+          if i==0 or j==len(seq)-1 or not (j==k and struct[i-1]==j+1):
+            for m2 in range(m-d+1):
+              result = forward(seq,ref_seq,struct,
+                                  (i+1,k-1),
+                                  (a2,b2),
+                                  m2,alpha)*forward(seq,ref_seq,struct,
+                                  (k+1,j),
+                                  (b2,b),
+                                  m-m2-d,alpha)*isostericity(seq,ref_seq,
+                                                       (i,k),
+                                                       (a2,b2),
[…]

### 2013-01-01 src/RNAPyroProfile.py

@@ -555,7 +555,6 @@ def random_weighted_sampling(l_samples):
 
 def backtrack(profile,ref_seq,struct,(i,j),(a,b),alpha):
   #alpha gives the weight energy vs isostericity
-  max_result = 0.
   result_list = []
   #max_seq_list = []
   max_seq = ''

### 2013-01-04 src/RNAPyroEx.py
bad backtrack RNAPyroEx

@@ -369,7 +369,7 @@ def isostericity(seq,ref_seq,(i,j),(a,b), alpha):
 def forward(seq,ref_seq,struct,(i,j),(a,b),m, alpha):
   #alpha gives the weight energy vs isostericity
   result = 0.
-  if m<0: return 0
+  if m<0 or m>j-i+1: return 0
   if i > j :
     if m==0:
       result=1.
@@ -419,7 +419,7 @@ def forward(seq,ref_seq,struct,(i,j),(a,b),m, alpha):
 def random_weighted_sampling(l_samples):
   tot = sum(x[1] for x in l_samples)
   if tot == 0:
-    print 'euurrrr'
+    print l_samples
     return random.choice([x[0] for x in l_samples])
   scaled_weights = [x[1]/tot for x in l_samples]
   rand_nb = random.random()
@@ -434,7 +434,7 @@ def backtrack(seq,ref_seq,struct,(i,j),(a,b),m, alpha):
   #alpha gives the weight energy vs isostericity
   result = 0.
   max_seq = ''
-  if m<0: return seq[i:j+1]
+  if m==0: return seq[i:j+1]
   if i > j : return ''
   else:
     k = struct[i]
@@ -442,12 +442,11 @@ def backtrack(seq,ref_seq,struct,(i,j),(a,b),m, alpha):
       l_samples = []
       for a2 in BASES:
         d = delta(seq,i,a2)
-        if d <= m:
-          result = forward(seq,ref_seq,struct,
-                            (i+1,j),
-                            (a2,b),
-                            m-d,alpha)
-          l_samples.append((a2,result))
+        result = forward(seq,ref_seq,struct,
+                          (i+1,j),
+                          (a2,b),
+                          m-d,alpha)
+        l_samples.append((a2,result))
       a2 = random_weighted_sampling(l_samples)
       d = delta(seq,i,a2)
       max_seq = a2 + backtrack(seq,ref_seq,struct,(i+1,j),(a2,b),m-d,alpha)
@@ -481,7 +480,7 @@ def backtrack(seq,ref_seq,struct,(i,j),(a,b),m, alpha):
                                                               (i,k),
                                                               (a2,b2),
                                                               alpha)
-            l_samples.append(((a2,b2,0),result))
+            l_samples.append(((a2,b2,m-d),result))
       a2,b2,m2 = random_weighted_sampling(l_samples)
[…]

### 2013-01-04 ISMB/methods_ISMB.tex
RNAPyroEx backtrack buggy
fixed ISMB headers

@@
-%!TEX root = main_RECOMB.tex
+%!TEX root = main_ISMB.tex
 \section{Methods}
@@
 \newcommand{\ISO}{\text{ISO}}
-We introduce a probabilistic model, which aims at capturing both the stability of the folded RNA and its ability to adopt a predefined 3D conformation.
-To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment.
+We introduce a probabilistic model, aiming to design RNA sequences with a 
+specific $\Gb+\Cb$ content, folding into a predefined 2D conformation.
+We use a Boltzmann weighted distribution based on a pseudo-energy function 
+$\PE{\cdot}$ taking into accounts stacked canonical base pairs free-energy.
+
+%To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment.
 In this model, the probability that the nucleotide at a given position needs to be mutated (i.e.

### 2013-01-04 ISMB/methods_ISMB.tex
stuff

@@
 corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990}.
-
-\subsection{Probabilistic model}
-Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
-then any sequence $s$ is assigned a probability proportional to its Boltzmann factor
-\begin{align*}
-  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
-\end{align*}
-where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
-are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
-
-\subsubsection{Energy contribution}
-The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
+%
+%\subsection{Probabilistic model}
+%Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
+%then any sequence $s$ is assigned a probability proportional to its Boltzmann factor
+%\begin{align*}
+%  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
+%\end{align*}
+%where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
+%are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
+%
+%\subsubsection{Energy contribution}
+%The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
 Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
-\begin{align*}
-  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
-\end{align*}
-where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
+%\begin{align*}
+%  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
+%\end{align*}
+%where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
 not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$).
 This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).
-The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed 
-(e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
-\subsubsection{Isostericity contribution}
-The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
+%The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed 
+%(e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
+%\subsubsection{Isostericity contribution}
+%The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
 Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the contribution of the isostericity to the pseudo-energy as
-\begin{align*}
-  \ES(s,S,\Omega) &= \sum_{\substack{(i,j)\in S\\ \text{pairs}}}\EI^{\Omega}_{(i,j),s_i s_j}, & \text{where}&& 	\EI^{\Omega}_{(i,j),ab}:=
-	\frac{
-		\sum_{s'\in\Omega}
-			\text{\ISO}((s_i',s_j'),(a,b))}
-%-		\left(			\text{\ISO}((s_i',s_j'),(s_i,s_j))				\right)	
-{		
-		|\Omega|
-	}
-\end{align*}
-is the average isostericity of a base-pair in the candidate sequence, compared with the reference alignment.
-The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
+%\begin{align*}
+%  \ES(s,S,\Omega) &= \sum_{\substack{(i,j)\in S\\ \text{pairs}}}\EI^{\Omega}_{(i,j),s_i s_j}, & \text{where}&& 	\EI^{\Omega}_{(i,j),ab}:=
+%	\frac{
+%		\sum_{s'\in\Omega}
+%			\text{\ISO}((s_i',s_j'),(a,b))}
+%%-		\left(			\text{\ISO}((s_i',s_j'),(s_i,s_j))				\right)	
+%{		
+%		|\Omega|
+%	}
+%\end{align*}
+%is the average isostericity of a base-pair in the candidate sequence, compared with the reference alignment.
+%The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
 Isostericity scores range between $0$ and $9.7$, with $0$ being assigned to a perfect isostericity, and a penalty of $10$ is used for missing entries.
-The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
-
-\subsection{Mutational profile of sequences}
-
-Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
-We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $m$ mutations from $s$
-%  under the SCFG derivation $S$ 
-(formally $\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
-We define a variant of the
- \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
- desired probability,  the two  functions
-$\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.
-
-The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec}, is analogous to the \emph{inside} algorithm.
-It is the partition function, i.e. the sum of Boltzmann factors, over all sequences within $[i,j]$, 
-knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$), within 
-$m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
- computes the \emph{outside} algorithm, i.e.  
-the partition function over sequences within $m$ mutations of $s$, restricted to two intervals $[0,i]\cup[j,n-1]$, and 
-knowing  that position $i+1$ is composed of $a$ (resp.
+%The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
+%
+%\subsection{Mutational profile of sequences}
+%
+%
+%Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
+%We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $m$ mutations from $s$
+%%  under the SCFG derivation $S$ 
+%(formally $\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
+%We define a variant of the
+% \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
+% desired probability,  the two  functions
+%$\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$. 
+%
+%The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec}, is analogous to the \emph{inside} algorithm.
+%It is the partition function, i.e. the sum of Boltzmann factors, over all sequences within $[i,j]$, 
+%knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$), within 
+%$m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
+% computes the \emph{outside} algorithm, i.e.  
+%the partition function over sequences within $m$ mutations of $s$, restricted to two intervals $[0,i]\cup[j,n-1]$, and 
+%knowing  that position $i+1$ is composed of $a$ (resp.
 $j-1$ is $b$).
 A suitable combination of these terms, given in Equation~\eqref{eq:combine}, gives the total weight, and in turn the probability, of seeing a specific base at a given position.
-
-%Drawing a parallel with stochastic context-free grammars (SCFG), for which the inside-outside algorithm was introduced, the set of sequences can be seen as the language of words having length $n$, generated from a stochastic context-free grammar.
-%These rules would limit the number of mutations (i.e. )
-% $S$ can be considered as constraining weighing on the shape of parse trees the derivations in a classic 
-% of the {SCFG} generating  all  secondary structures of length  $|s|$.
-
-\subsubsection{Definitions}
-Let $B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
-Given $s\in B^n$ an RNA sequence, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
-length $n$, and $S$ a secondary structure without pseudoknots. 
-Formally, if $(i,j)$ and $(k,l)$ are base pairs in $S$, there is no overlapping extremities
- $\{i,j\}\cap \{k,l\}=\varnothing$ and either the intersection is empty 
- ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
- $[i,j]\subset[k,l])$.
-
-Let us then remind a Hamming distance function $\delta: B^*\times B^* \to \mathbb{N}^+$, which takes two sequences $s'$ and $s''$ as input, $|s'|=|s''|$, and returns the number of differing positions.
-Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
-\begin{equation}
-  E_{(i,j),ab \to a'b'}^{\Omega,\beta}  = \alpha \cdot\ES^\beta_{a b \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
-\end{equation}
-
-\subsubsection{Inside}
-\begin{figure}
-\resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
-\caption{Principle of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
-can be decomposed as a sequence over $[i+1,j]$ preceded by a, possibly mutated, base at $i$ 
-(Unpaired case), a sequence over $[i+1,j-1]$ surrounded by some base-pair (Stacking-pair case), 
-or as two sequences over $[i+1,k-1]$ and $[k+1,j]$, completed by some base-pair (General base-pairing case). 
-In each case, one has to investigate any possible ways to distribute mutations between the different sequences 
-and locally instanciated bases.}
-\end{figure}
-
-The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function, i.e.
+%
+%%Drawing a parallel with stochastic context-free grammars (SCFG), for which the inside-outside algorithm was introduced, the set of sequences can be seen as the language of words having length $n$, generated from a stochastic context-free grammar.
+%%These rules would limit the number of mutations (i.e. )
+%% $S$ can be considered as constraining weighing on the shape of parse trees the derivations in a classic 
+%% of the {SCFG} generating  all  secondary structures of length  $|s|$. 
+%
+%
+%
+%
+%
+%\subsubsection{Definitions}
+%Let $B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
+%Given $s\in B^n$ an RNA sequence, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
+%length $n$, and $S$ a secondary structure without pseudoknots. 
+%Formally, if $(i,j)$ and $(k,l)$ are base pairs in $S$, there is no overlapping extremities
+% $\{i,j\}\cap \{k,l\}=\varnothing$ and either the intersection is empty 
+% ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
+% $[i,j]\subset[k,l])$. 
+%
+%Let us then remind a Hamming distance function $\delta: B^*\times B^* \to \mathbb{N}^+$, which takes two sequences $s'$ and $s''$ as input, $|s'|=|s''|$, and returns the number of differing positions.
+%Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
+%\begin{equation}
+%  E_{(i,j),ab \to a'b'}^{\Omega,\beta}  = \alpha \cdot\ES^\beta_{a b \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
+%\end{equation}
+%
+%
+%\subsubsection{Inside}
+%\begin{figure}
+%\resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
+%\caption{Principle of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
+%can be decomposed as a sequence over $[i+1,j]$ preceded by a, possibly mutated, base at $i$ 
+%(Unpaired case), a sequence over $[i+1,j-1]$ surrounded by some base-pair (Stacking-pair case), 
+%or as two sequences over $[i+1,k-1]$ and $[k+1,j]$, completed by some base-pair (General base-pairing case). 
+%In each case, one has to investigate any possible ways to distribute mutations between the different sequences 
+%and locally instanciated bases.}
+%\end{figure}
+%
+%The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function, i.e.
 the sum of Boltzmann factors, over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively).
-Such terms can be defined by recurrence, for which the following initial conditions holds:
-
-\begin{equation}
-	\forall i \in [0,n-1]:\, \Z{i+1,i}{m}{a,b}=\left\{
-	\begin{array}{ll}
-		1 &\text{If } m = 0\\
-		0 &\text{Otherwise.}
-	\end{array}\right.
-\label{eq:Z_in}
-\end{equation}
-In other words, the set of sequences at distance $m$ of the empty sequence is either empty if $m>0$, or restricted to the empty sequence, having energy $0$, if $m=0$.
+%Such terms can be defined by recurrence, for which the following initial conditions holds:
+%
+%\begin{equation}
+%	\forall i \in [0,n-1]:\, \Z{i+1,i}{m}{a,b}=\left\{
+%	\begin{array}{ll}
+%		1 &\text{If } m = 0\\
+%		0 &\text{Otherwise.}
+%	\end{array}\right.
+%\label{eq:Z_in}
+%\end{equation}
+%In other words, the set of sequences at distance $m$ of the empty sequence is either empty if $m>0$, or restricted to the empty sequence, having energy $0$, if $m=0$.
 Since the energetic terms only depend on base pairs, they are not involved in the initial conditions.
@@
 
-\subsubsection{Outside}	
-\begin{figure}
-\resizebox{\textwidth}{!}{\input{FigDPOutside.pgf}}
-\caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
-therefore its efficient implementation requires an implementation of the inside computation.}
-\end{figure}
-The \emph{Outside} function, $\mathcal Y$, is the partition function considering only the 
-contributions of subsequences $[0,i]\cup[j,n-1]$ over the mutants of $s$ having exactly $m$ mutations between $[0,i]\cup[j,n-1]$ and whose nucleotide at position $i+1$ is $a$ 
-(resp. in position $j-1$ it is $b$).
-We define function $\Y{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
-\begin{equation}
-	\Y{-1,j}{m}{X,X}:=
-		\displaystyle
-	  \Z{j,n-1}{m}{X,X}
-\label{eq:Y_in}
-\end{equation}
-The recurrence, as shown below, will increase the interval $[i,j]$ by decreasing $i$ when
-it is not base paired. If it is with a position $k>j$, we increase $j$ to include it.
- Thus, when we need
-to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
-structure between $(j,n-1)$ must be a set of independent stems. Therefore,
- all the outside energy between $[j,n-1]$ is
-equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself is as follows.
-\begin{equation}
-	\Y{i,j}{m}{a,b} = \left\{
-  \begin{array}{ll}
-		\displaystyle
-    \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}
-    \Y{i-1,j}{m- \Kron_{a',s_i}}{a',b} &
-    \text{Elif }S_i=-1 \\
-    \displaystyle
-    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
-		 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}\cdot
-    \Y{i-1,j+1}{m- \Kron_{a'b',s_is_j}}{a',b'} &
-   	 \text{Elif }S_{i}=j \land S_{i+1}=j-1\\
-		 \displaystyle
-		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
-		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-  		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
-		 \cdot\Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
-     \cdot\Z{j,k-1}{m'}{b,b'} &
-		 \text{Elif }S_{i}=k \geq j\\
-		 \displaystyle
-		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_ks_i}\le m}}
-		 \sum_{m'=0}^{m-\Kron_{a'b',s_ks_i}}
-   	 e^{\frac{-E_{(k,i),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
-		 \cdot\Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
-     \cdot\Z{k+1,i-1}{m'}{a',b'} &
-		 \text{Elif }-1 < S_{i}=k < i\\
-		 0 & \text{Otherwise}
-  \end{array}\right.
-\label{eq:Y_rec}
-\end{equation}
-The five cases can be broked down as follows.
-\begin{description}
-\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, then the value is the same
-as if we decrease the lower interval bound by $1$ (i.e. $i-1$), and consider all possible
-nucleotides $a'$ at position $i$, correcting the number of mutants
-in function of $\Kron_{a',s_i}$.
-\item[$S_{i}=j$ and $S_{i+1}=j-1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i+1$ is
-paired with $j-11$, we are in the only case were stacked base pairs can occur. We thus add
-the energy of the stacking and of the isostericity of the base pair $(i,j)$. What is left
-to compute is the \emph{outside} value for the interval $[i-1,j+1]$ over all possible nucleotides 
-$a',b'\in B^2$ at positions $i$ and $j$ respectively.
-\item[$S_{i}=k \geq j$:]If nucleotide $i$ is paired with position $k\geq j$, 
-and is not stacked inside, the 
-only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
-We can then consider the outside interval $[i-1,k+1]$ by multiplying it by the the \emph{inside}
-value of the newly included interval (i.e. $[j,k-1]$), for 
-all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $k$ respectively.
-\item[$-1<S_{i}<i$:]As above but if position $i$ is to pairing with a lower value.
-\item[Else:] In all other cases, we are in a derivation of the SCFG that does not correspond to the 
-secondary structure $S$, and we return $0$.
-
-\end{description}
-
-\subsubsection{Combining Inside and Outside values into point-wise mutations probabilities}
-By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
-be described in function of the \emph{inside} term as $\Z{0,n-1}{m}{X,X}$,
- for any nucleotide $X\in B$ or
-in function of the \emph{outside} term, for any unpaired position $k$:
-$$
-	\Z{0,n-1}{m}{X,X}
-	\equiv
-	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
-	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a}
-$$
-
-We are now left to compute the probability that a given position is a given nucleotide.
-We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
-Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations. 
-\begin{align}
-	\mathbb{P}(s_i = x\mid s,\Omega, S,M) &:= \frac{\mathcal{W}(i,x,s,\Omega,S,M)}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}\\ 
-\mathcal{W}(i,x,s,\Omega,S,M)&=
- \left\{
-	\begin{array}{ll}
-			\sum_{m=0}^{M}
-			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
-		&\text{If }S_i = -1\\
-			\displaystyle
-			\sum_{m=0}^{M}
-			\sum_{\substack{b\in B\\\Kron_{xb,s_is_k\leq m}}}
-			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
-     	 e^{\frac{-E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
-			\cdot\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
-			\cdot\Z{i+1,k-1}{m'}{x,b}
-		&\text{If }S_i=k>i\\
-    \displaystyle
-			\sum_{m=0}^{M}
-			\sum_{\substack{b\in B\\\Kron_{bx,s_ks_i\leq m}}}
-			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
-     	 e^{\frac{-E^{\Omega, \beta}_{(k,i),\varnothing\to bx}}{RT}}
-			\cdot\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
-			\cdot\Z{k+1,i-1}{m'}{b,x}
-		&\text{If }S_i=k<i
-	\end{array}\right.\label{eq:combine}
-\end{align}
-
-In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, 
-for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
-\begin{description}
-\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, we are concerned by the weights
-over all sequences which have at position $i$ nucleotide $x$, which is exactly the sum
-of the values of $\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}$, for all $m$ between $0$ and $M$.
-\item[$S_i=k>i$:] Since we need to respect the derivation of the secondary structure $S$, if 
-position $i$ is paired, we must consider the two partition functions. The \emph{outside} of the 
-base pair, and the \emph{inside}, for all possible values for the nucleotide at position $k$, and
-all possible distribution of the mutant positions between the inside and outside of the base pair. We also add the term of isostericity for this specific base pair.
-\item[$S_i=k<i$:] Same as above, but with position $i$ pairing with a lower position.
-\end{description}
-
-\subsection{Complexity considerations}
-Equations~\eqref{eq:Z_rec} and~\eqref{eq:Y_rec} can be computed using dynamic programming.
+%\subsubsection{Outside}	
+%\begin{figure}
+%\resizebox{\textwidth}{!}{\input{FigDPOutside.pgf}}
+%\caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
+%therefore its efficient implementation requires an implementation of the inside computation.}
+%\end{figure}
+%The \emph{Outside} function, $\mathcal Y$, is the partition function considering only the 
+%contributions of subsequences $[0,i]\cup[j,n-1]$ over the mutants of $s$ having exactly $m$ mutations between $[0,i]\cup[j,n-1]$ and whose nucleotide at position $i+1$ is $a$ 
+%(resp. in position $j-1$ it is $b$).
+%We define function $\Y{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
+%\begin{equation}
+%	\Y{-1,j}{m}{X,X}:=
+%		\displaystyle
+%	  \Z{j,n-1}{m}{X,X}
+%\label{eq:Y_in}
+%\end{equation}
+%The recurrence, as shown below, will increase the interval $[i,j]$ by decreasing $i$ when
+%it is not base paired. If it is with a position $k>j$, we increase $j$ to include it.
+% Thus, when we need
+%to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
+%structure between $(j,n-1)$ must be a set of independent stems. Therefore,
+% all the outside energy between $[j,n-1]$ is
+%equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself is as follows.
+%\begin{equation}
+%	\Y{i,j}{m}{a,b} = \left\{
+%  \begin{array}{ll}
+%		\displaystyle
+%    \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}
+%    \Y{i-1,j}{m- \Kron_{a',s_i}}{a',b} &
+%    \text{Elif }S_i=-1 \\
+%    \displaystyle
+%    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
+%		 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}\cdot
+%    \Y{i-1,j+1}{m- \Kron_{a'b',s_is_j}}{a',b'} &
+%   	 \text{Elif }S_{i}=j \land S_{i+1}=j-1\\
+%		 \displaystyle
+%		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
+%		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
+%  		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
+%		 \cdot\Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
+%     \cdot\Z{j,k-1}{m'}{b,b'} &
+%		 \text{Elif }S_{i}=k \geq j\\
+%		 \displaystyle
+%		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_ks_i}\le m}}
+%		 \sum_{m'=0}^{m-\Kron_{a'b',s_ks_i}}
+%   	 e^{\frac{-E_{(k,i),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
+%		 \cdot\Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
+%     \cdot\Z{k+1,i-1}{m'}{a',b'} &
+%		 \text{Elif }-1 < S_{i}=k < i\\
+%		 0 & \text{Otherwise}
+%  \end{array}\right.
+%\label{eq:Y_rec}
+%\end{equation}
+%The five cases can be broked down as follows.
+%\begin{description}
+%\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, then the value is the same
+%as if we decrease the lower interval bound by $1$ (i.e. $i-1$), and consider all possible
+%nucleotides $a'$ at position $i$, correcting the number of mutants
+%in function of $\Kron_{a',s_i}$.
+%\item[$S_{i}=j$ and $S_{i+1}=j-1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i+1$ is
+%paired with $j-11$, we are in the only case were stacked base pairs can occur. We thus add
+%the energy of the stacking and of the isostericity of the base pair $(i,j)$. What is left
+%to compute is the \emph{outside} value for the interval $[i-1,j+1]$ over all possible nucleotides 
+%$a',b'\in B^2$ at positions $i$ and $j$ respectively.
+%\item[$S_{i}=k \geq j$:]If nucleotide $i$ is paired with position $k\geq j$, 
+%and is not stacked inside, the 
+%only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
+%We can then consider the outside interval $[i-1,k+1]$ by multiplying it by the the \emph{inside}
+%value of the newly included interval (i.e. $[j,k-1]$), for 
+%all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $k$ respectively.
+%\item[$-1<S_{i}<i$:]As above but if position $i$ is to pairing with a lower value.
+%\item[Else:] In all other cases, we are in a derivation of the SCFG that does not correspond to the 
+%secondary structure $S$, and we return $0$.
+%
+%
+%\end{description}
+%
+%\subsubsection{Combining Inside and Outside values into point-wise mutations probabilities}
+%By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
+%be described in function of the \emph{inside} term as $\Z{0,n-1}{m}{X,X}$,
+% for any nucleotide $X\in B$ or
+%in function of the \emph{outside} term, for any unpaired position $k$:
+%$$
+%	\Z{0,n-1}{m}{X,X}
+%	\equiv
+%	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
+%	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a}
+%$$
+%
+%We are now left to compute the probability that a given position is a given nucleotide.
+%We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
+%Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations. 
+%\begin{align}
+%	\mathbb{P}(s_i = x\mid s,\Omega, S,M) &:= \frac{\mathcal{W}(i,x,s,\Omega,S,M)}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}\\ 
+%\mathcal{W}(i,x,s,\Omega,S,M)&=
+% \left\{
+%	\begin{array}{ll}
+%			\sum_{m=0}^{M}
+%			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
+%		&\text{If }S_i = -1\\
+%			\displaystyle
+%			\sum_{m=0}^{M}
+%			\sum_{\substack{b\in B\\\Kron_{xb,s_is_k\leq m}}}
+%			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
+%     	 e^{\frac{-E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
+%			\cdot\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
+%			\cdot\Z{i+1,k-1}{m'}{x,b}
+%		&\text{If }S_i=k>i\\
+%    \displaystyle
+%			\sum_{m=0}^{M}
+%			\sum_{\substack{b\in B\\\Kron_{bx,s_ks_i\leq m}}}
+%			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
+%     	 e^{\frac{-E^{\Omega, \beta}_{(k,i),\varnothing\to bx}}{RT}}
+%			\cdot\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
+%			\cdot\Z{k+1,i-1}{m'}{b,x}
+%		&\text{If }S_i=k<i
+%	\end{array}\right.\label{eq:combine}
+%\end{align}
+%
+%In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, 
+%for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
+%\begin{description}
+%\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, we are concerned by the weights
+%over all sequences which have at position $i$ nucleotide $x$, which is exactly the sum
+%of the values of $\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}$, for all $m$ between $0$ and $M$.
+%\item[$S_i=k>i$:] Since we need to respect the derivation of the secondary structure $S$, if 
+%position $i$ is paired, we must consider the two partition functions. The \emph{outside} of the 
+%base pair, and the \emph{inside}, for all possible values for the nucleotide at position $k$, and
+%all possible distribution of the mutant positions between the inside and outside of the base pair. We also add the term of isostericity for this specific base pair.
+%\item[$S_i=k<i$:] Same as above, but with position $i$ pairing with a lower position.
+%\end{description}
+%
+%\subsection{Complexity considerations}
+%Equations~\eqref{eq:Z_rec} and~\eqref{eq:Y_rec} can be computed using dynamic programming.
 Namely, the $\mathcal{Z}^{*}_{*}$ and $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and interval lengths, memorizing the results as they become available to ensure constant-time access during later stages of the computation.
@@
 Computing any given term therefore requires $\Theta(m)$ operations.
-
-In principle, $\Theta(m\cdot n^2)$ terms, identified by $(m,i,j)$ triplets, should be computed.
-However, a close inspection of the recurrences reveals that the computation can be safely restricted to a subset of intervals $(i,j)$.
-For instance, the inside algorithm only requires computing intervals $[i,j]$ that do not break any base-pair, and whose next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
+%
+%In principle, $\Theta(m\cdot n^2)$ terms, identified by $(m,i,j)$ triplets, should be computed.
+%However, a close inspection of the recurrences reveals that the computation can be safely restricted to a subset of intervals $(i,j)$.
+%For instance, the inside algorithm only requires computing intervals $[i,j]$ that do not break any base-pair, and whose next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
 Similar constraints hold for the outside computation, resulting in a drastic limitation of the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored.

### 2013-01-05 ISMB/methods_ISMB.tex
algo

@@
 \newcommand{\ISO}{\text{ISO}}
+\newcommand{\Prob}{\text{P}}
+
 We introduce a probabilistic model, aiming to design RNA sequences with a 
@@
 We use a Boltzmann weighted distribution based on a pseudo-energy function 
-$\PE{\cdot}$ taking into accounts stacked canonical base pairs free-energy.
+$\ES(\cdot)$ taking into accounts stacked canonical base pairs free-energy.
 
@@
 %\subsubsection{Energy contribution}
-%The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
+\subsection{Definitions}
+\subsubsection{Profile}
+At every position, the nucleotide distribution is the same. 
+It is first initialized equiprobably. A bisection scheme is then applied until the sampled sequences $\Cb+\Gb$ content is as desired. The probability
+of a given nucleotide $a\in\B$ will be given by $\Prob(a)$ at every 
+position.
+
+\subsubsection{Energetic model}
+
+The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
 Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
-%\begin{align*}
-%  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
-%\end{align*}
-%where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
+
+\begin{align*}
+  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
+\end{align*}
+
+where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
 not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$).
 This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).
+
 %The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed 
@@
 %
-%\subsubsection{Definitions}
 %Let $B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
@@
 %Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
+
 %\begin{equation}
@@
 \begin{equation}
-	\Z{i,j}{m}{a,b}:=\left\{
+	\Z{i,j}{a,b}:=\left\{
   \begin{array}{ll}
   		\displaystyle
-      \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}  
-      \Z{i+1,j}{m-\Kron_{a',s_i}}{a',b} & \text{If }S_{i}=-1\\
+      \sum_{a'\in \B}  
+      \Prob(a')
+      \cdot\Z{i+1,j}{a',b} & \text{If }S_{i}=-1\\
       \displaystyle
-      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
-			 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}
-			 \cdot \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
+      \sum_{a',b'\in \B^2}
+			 \Prob(a')
+			 \cdot\Prob(b')
+			 \cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
+			 \cdot \Z{i+1,j-1}{a',b'}&
 			 \text{Elif }S_i=j \land S_{i-1}=j+1\\
 			 \displaystyle
-      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
-      \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-   		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
-      \cdot\Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
-      \cdot\Z{k+1,j}{m'}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
+      \sum_{a',b'\in \B^2}
+      \Prob(a')
+			\cdot\Prob(b')
+			\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
+      \cdot\Z{i+1,k-1}{a',b'}
+      \cdot\Z{k+1,j}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
       0 &\text{Otherwise}
@@
 \item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
-any sequence consists in a, possibly mutated, nucleotide $a'$ at position  $i$, 
-followed by a sequence over $[i+1,j]$ having either $m-\Kron_{a',s_i}$, accounting for a possible mutation 
-at position $i$, and having flanking nucleotides $a'$ and $b$.
-\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two, possibly mutated, nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$ having distance $m-\Kron_{a'b',s_is_j}$ (to avoid exceeding the targeted distance $m$).
+any sequence consists in a nucleotide $a'$ at position  $i$, 
+followed by a sequence over $[i+1,j]$ and having flanking nucleotides $a'$ and $b$.
+\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$.
 Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
@@
 \item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
-only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
+only term contributing directly to the energy is the nucleotide distribution $\Prob(\cdot)$. 
 Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$.
-Since the number of mutations sum to $m$ over the whole sequence must , then a parameter $m'$ is introduced to distribute the remaining mutations between the two sequences.
 \item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
 \end{description}
+
+\subsection{Stochastic backtrack}
+
+\begin{algorithm}
+	\SetAlgoLined
+	\lIf{$i>j$} \Return{$\varepsilon$};\tcp*[f]{Empty sequence}\\
+	rand $\leftarrow$ Random$\left(\Z{i,j}{a,b}\right)$\;
+	k $\leftarrow S_i$\;
+	\If(\tcp*[f]{Unpaired}){$k = -1$}{
+		\For{$a'\in\B$}{
+			rand $\leftarrow$ rand $- \Prob(a')\cdot \Z{i+1,j}{a',b}$\;
+			\lIf{$\text{rand}<0$}\Return{$a'\cdot\text{Backtrack}\left((i+1,j),(a',b),S\right)$}\;
+		}
+	}
+	\Else{
+		\If(\tcp*[f]{Stack}){$k=j \land S_{i-1}=j+1$}{
+			\For{$(a',b')\in\B\times\B$}{
+				rand $\leftarrow$ rand $-
+			 	\Prob(a')
+			 	\cdot\Prob(b')
+			 	\cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
+			 	\cdot \Z{i+1,j-1}{a',b'}	$\;
+				\If{$\text{rand}<0$}{
+					\Return{$a'\cdot\text{Backtrack}\left((i+1,j-1),(a',b'),S\right)\cdot b'$}\;		
+				}
+			}
+		}
+		\Else(\tcp*[f]{Paired not stacked}){
+			\For{$(a',b')\in\B\times\B$}{
+				rand $\leftarrow$ rand $-
+				\Prob(a')
+				\cdot\Prob(b')
+				\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
+  	    \cdot\Z{i+1,k-1}{a',b'}
+    	  \cdot\Z{k+1,j}{b',b}$\;	
+ 				\If{$\text{rand}<0$}{
+					\Return{$a'
+					\cdot\text{Backtrack}\left((i+1,k-1),(a',b'),S\right)
+					\cdot b'
+					\cdot\text{Backtrack}\left((k+1,j),(b',b),S\right)
+					$}\;	
+				}
+			}
+		}
+	}
+\caption{Backtrack$\left((i,j),(a,b),S\right)$\label{alg:back}}
+\end{algorithm}
 

### 2013-01-06 ISMB/introduction_ISMB.tex
fixed ISMB headers

@@
-%!TEX root = main_RECOMB.tex
+%!TEX root = main_ISMB.tex
 \section{Introduction}
 \label{sec:introduction}
+
+At the core of the emerging field of synthetic biology resides our capacity to design and re-engineer molecules with target functions.
+RNA molecules are well tailored for such applications.
+The ease to synthesize them (they are directly transcribed from DNA) and the broad diversity of catalytic and regulation functions they can perform enable to integrate \textit{de-novo} logic  circuits within living cells \cite{Rodrigo:2012fk} or re-program existing regulation mechanisms \cite{Chang:2012uq}.
+Future advances and applications of these techniques in gene-therapy studies will strongly rely on efficient computational methods to design and re-engineer RNA molecules.
+
+Most of RNA functions are, at least partially, encoded by the three-dimensional molecular structures, which are themselves primarily determined by the secondary structures.
+The development of efficient algorithms for designing RNA sequences with pre-defined secondary structures is thus a milestone to enter the synthetic biology era.
+\RNAinverse pioneered RNA secondary structure design algorithms.
+It has been developed and distributed with the Vienna RNA package \cite{Hofacker:1994}.
+However, only posterior experimental studies revealed the potential and practical impact of these techniques.
+Thereby, during the last 6 years many improvements and variants of \RNAinverse have been proposed.
+Conceptually, almost all existing programs follow the same approach.
+First a seed sequence is selected, then a local search strategy is used to mutate the seed and find in its vicinity a sequence with desired folding properties.
+Using this strategy, \INFORNA \cite{Busch:2006uq}, \RNASSD \cite{Aguirre-Hernandez:2007kx} and \NUPACK \cite{Zadeh:2011fk} significantly improved the performance of the RNA secondary structure design algorithms.
+More recent research studies aimed to include more constraints in the selection criteria.\RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \cite{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \cite{Lyngso:2012vn}.
+
+We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \cite{Levin:2012kx}.
+Instead a local search approach, we propose to apply a global sampling strategy of the mutational landscape using our \RNAmutants algorithm \cite{Waldispuhl2008}.
+This methodology offers promising performance but  suffers from a prohibitive running time complexity.
+Following our work, Garcia-Martin and co-workers proposed with \RNAiFOLD an alternate methodology using constraint programming techniques to prune the mutational landscape.
+The latter also suffers from prohibitive running times, but it is worth noting that it proposes a seed-less approach to the RNA secondary structure design problem.
+
+In this paper we introduce \RNApyro, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}.
+\RNApyro addresses previous limitations of \RNAensign and offers new functionalities.
+First, while our previous program had a running time complexity of $\mathcal{O}(n^4)$, \RNApyro has now a linear-time and space complexity that allows him to demonstrate similar speeds to any local search algorithm.
+Next, \RNApyro is \textit{seed-less}.
+Unlike \RNAensign, it does not require a seed sequence to start its search.
+Finally, \RNApyro implements a novel algorithm using weighted sampling techniques \cite{DBLP:journals/corr/abs-1002-0046} that enables us to control, for the first time, \textit{explicitly} the \texttt{C+G}-content of the solution.
+This functionality is essential because wild sequences within living organisms often present medium or low \texttt{C+G}-content.
+Presumably to offer better transcription rates and/or structural plasticity.
+Previous programs do not allow to control this parameter and tend to output sequences with high \texttt{C+G}-contents.
+
+We demonstrate the performance of our algorithms on a set of real RNA structures with sizes ranging from *** to *** nucleotides.
+To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse.
+Remarkably, our glocal methodology overcome both local and global approaches.

### 2013-01-07 ISMB/introduction_ISMB.tex
Method... interrupted

@@
 We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \cite{Levin:2012kx}.
-Instead a local search approach, we propose to apply a global sampling strategy of the mutational landscape using our \RNAmutants algorithm \cite{Waldispuhl2008}.
+Instead of a local search approach, we propose to apply a global sampling strategy of the mutational landscape using our \RNAmutants algorithm \cite{Waldispuhl2008}.
 This methodology offers promising performance but  suffers from a prohibitive running time complexity.

### 2013-01-07 ISMB/methods_ISMB.tex
s
Method... interrupted

@@
 
-\newcommand{\PE}[1]{E(#1)}
-\newcommand{\EI}{\text{EI}}
-\newcommand{\ES}{\text{ES}}
-\newcommand{\ISO}{\text{ISO}}
-\newcommand{\Prob}{\text{P}}
+We introduce a probabilistic model for the design of RNA sequences with a specific $\Gb+\Cb$ content and folding into a predefined secondary structure.
+For the sake of simplicity, we choose to base this proof-of-concept implementation on a simplified free-energy function $\ES(\cdot)$, which only considers the contributions of 
+stacked canonical base-pairs.
 
-We introduce a probabilistic model, aiming to design RNA sequences with a 
-specific $\Gb+\Cb$ content, folding into a predefined 2D conformation.
-We use a Boltzmann weighted distribution based on a pseudo-energy function 
-$\ES(\cdot)$ taking into accounts stacked canonical base pairs free-energy.
+We show how a modification of the dynamic programming scheme introduced in the previously introduced \RNAmutants allow for the sampling of good and diverse design candidates in linear time and space complexities.
 
@@
 \subsection{Definitions}
-\subsubsection{Profile}
-At every position, the nucleotide distribution is the same. 
-It is first initialized equiprobably. A bisection scheme is then applied until the sampled sequences $\Cb+\Gb$ content is as desired. The probability
-of a given nucleotide $a\in\B$ will be given by $\Prob(a)$ at every 
-position.
 
-\subsubsection{Energetic model}
+%\subsubsection{Profile}
+%At every position, the nucleotide distribution is the same. 
+%It is first initialized equiprobably. A bisection scheme is then applied until the sampled sequences $\Cb+\Gb$ content is as desired. The probability
+%of a given nucleotide $a\in\B$ will be given by $\Prob(a)$ at every 
+%position.
 
-The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
+\subsubsection{Secondary structure}
+Todo...
+
+\subsubsection{Energy model}
+The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}).
 Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
-
 \begin{align*}
@@
 \end{align*}
-
 where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
 not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$).
-This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).
+This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = \infty$).
 
-%The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed 
-%(e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
-%\subsubsection{Isostericity contribution}
-%The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
-Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the contribution of the isostericity to the pseudo-energy as
-%\begin{align*}
-%  \ES(s,S,\Omega) &= \sum_{\substack{(i,j)\in S\\ \text{pairs}}}\EI^{\Omega}_{(i,j),s_i s_j}, & \text{where}&& 	\EI^{\Omega}_{(i,j),ab}:=
-%	\frac{
-%		\sum_{s'\in\Omega}
-%			\text{\ISO}((s_i',s_j'),(a,b))}
-%%-		\left(			\text{\ISO}((s_i',s_j'),(s_i,s_j))				\right)	
-%{		
-%		|\Omega|
-%	}
-%\end{align*}
-%is the average isostericity of a base-pair in the candidate sequence, compared with the reference alignment.
-%The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
-Isostericity scores range between $0$ and $9.7$, with $0$ being assigned to a perfect isostericity, and a penalty of $10$ is used for missing entries.
-%The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
-%
-%\subsection{Mutational profile of sequences}
-%
-%
-%Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
-%We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $m$ mutations from $s$
-%%  under the SCFG derivation $S$ 
-%(formally $\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
-%We define a variant of the
-% \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
-% desired probability,  the two  functions
-%$\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$. 
-%
-%The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec}, is analogous to the \emph{inside} algorithm.
-%It is the partition function, i.e. the sum of Boltzmann factors, over all sequences within $[i,j]$, 
-%knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$), within 
-%$m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
-% computes the \emph{outside} algorithm, i.e.  
-%the partition function over sequences within $m$ mutations of $s$, restricted to two intervals $[0,i]\cup[j,n-1]$, and 
-%knowing  that position $i+1$ is composed of $a$ (resp.
-$j-1$ is $b$).
-A suitable combination of these terms, given in Equation~\eqref{eq:combine}, gives the total weight, and in turn the probability, of seeing a specific base at a given position.
-%
-%%Drawing a parallel with stochastic context-free grammars (SCFG), for which the inside-outside algorithm was introduced, the set of sequences can be seen as the language of words having length $n$, generated from a stochastic context-free grammar.
-%%These rules would limit the number of mutations (i.e. )
-%% $S$ can be considered as constraining weighing on the shape of parse trees the derivations in a classic 
-%% of the {SCFG} generating  all  secondary structures of length  $|s|$. 
-%
-%
-%
-%
-%
-%Let $B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
-%Given $s\in B^n$ an RNA sequence, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
-%length $n$, and $S$ a secondary structure without pseudoknots. 
-%Formally, if $(i,j)$ and $(k,l)$ are base pairs in $S$, there is no overlapping extremities
-% $\{i,j\}\cap \{k,l\}=\varnothing$ and either the intersection is empty 
-% ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
-% $[i,j]\subset[k,l])$. 
-%
-%Let us then remind a Hamming distance function $\delta: B^*\times B^* \to \mathbb{N}^+$, which takes two sequences $s'$ and $s''$ as input, $|s'|=|s''|$, and returns the number of differing positions.
-%Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
+\subsection{\GCContent weighted pseudo-Boltzmann ensemble and distribution}
 
-%\begin{equation}
-%  E_{(i,j),ab \to a'b'}^{\Omega,\beta}  = \alpha \cdot\ES^\beta_{a b \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
-%\end{equation}
-%
-%
-%\subsubsection{Inside}
-%\begin{figure}
-%\resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
-%\caption{Principle of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
-%can be decomposed as a sequence over $[i+1,j]$ preceded by a, possibly mutated, base at $i$ 
-%(Unpaired case), a sequence over $[i+1,j-1]$ surrounded by some base-pair (Stacking-pair case), 
-%or as two sequences over $[i+1,k-1]$ and $[k+1,j]$, completed by some base-pair (General base-pairing case). 
-%In each case, one has to investigate any possible ways to distribute mutations between the different sequences 
-%and locally instanciated bases.}
-%\end{figure}
-%
-%The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function, i.e.
-the sum of Boltzmann factors, over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively).
-%Such terms can be defined by recurrence, for which the following initial conditions holds:
-%
-%\begin{equation}
-%	\forall i \in [0,n-1]:\, \Z{i+1,i}{m}{a,b}=\left\{
-%	\begin{array}{ll}
-%		1 &\text{If } m = 0\\
-%		0 &\text{Otherwise.}
-%	\end{array}\right.
-%\label{eq:Z_in}
-%\end{equation}
-%In other words, the set of sequences at distance $m$ of the empty sequence is either empty if $m>0$, or restricted to the empty sequence, having energy $0$, if $m=0$.
-Since the energetic terms only depend on base pairs, they are not involved in the initial conditions.
+In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, let us introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences.
+Let $\Target$ be a targeted secondary structure and $\gc(s)$ be the number of occurrences of \Gb and \Cb in an RNA sequence $s$, then let us define the pseudo-Boltzmann factor $\B_{x}(s,S)$ of a sequence $s$ such that
+\begin{equation}
+\B_{x, \Target}(s) = e^{\frac{-\ES(s,\Target)}{RT}}\cdot x^{\gc(s)}
+\label{def:genBoltz}
+\end{equation}
+where $R$ is the Boltzmann constant and $T$ the temperature in Kelvin.
 
-The main recursion is composed of four terms:
+Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$ gives the pseudo-partition function $\mathcal{Z}_{\Target,x}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that
+\begin{align}\mathcal{Z}_{x,\Target} &= \sum_{\substack{|s|=|\Target|}}\B_{x,\Target}(s)& \text{and}&&
+\Prob_{x,\Target}(s) &= \frac{\B_{x,\Target}(s)}{\mathcal{Z}_{x,\Target}}.\label{def:distribution}\end{align}
+
+It was established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the \GCContent distribution observed in sequences sampled with respect to the above distribution converges to a normal law of expectation in $\mu\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma\cdot\sqrt{n}\cdot(1+o(1))$.
+Furthermore, it can easily be shown that the expected \GCContent is a continuous and strictly increasing monotonic  function of $x$, whose limits are $0$ when $x=0$ and $n$ when $x\to +\infty$.
+
+Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent.
+Such a value can be determined using any numerical recipe, and a simple bisection is used in our implementation.
+Furthermore, the concentration of the distribution, as asserted by the limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp.
+$\Theta(1)$, i.e.
+constant) for the expected number of attempts before generating a sequence having the exact targeted \GCContent  (resp.
+within $[gc-\kappa,gc+\kappa]$, where $\kappa\in\Theta(1/\sqrt n)$).
+
+\subsection{Sampling the pseudo-Boltzmann ensemble}
+
+Let us now describe a linear-time algorithm to sample sequences at random in the pseudo-Boltzmann distribution.
+This algorithm follows the general principles of the recursive approach to random generation~\cite{Wilf1977}, pioneered in the context of RNA by the \SFold algorithm~\cite{Ding2003}.
+The algorithm starts by precomputing the partition function restricted to each valid interval, and then performs a series of stochastic backtracks, using precomputed values to decide on the probability of each alternative.
+
+\subsubsection{Precomputation}
+Let us describe a dynamic programming algorithm to compute $\Z{i,j}{a,b}$, the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively.
+
+Since the only sequence over an empty interval is the empty sequence, having energy $0$, one has
 \begin{equation}
+	\forall i \in [0,n-1]:\, \Z{i+1,i}{a,b}=1.
+	\label{eq:Z_in}
+\end{equation}
+
+The recursion consists in four terms, depending on the base-pairing status and context of $(i,j)$:
+\begin{equation*}
 	\Z{i,j}{a,b}:=\left\{
@@
       \sum_{a'\in \B}  
-      \Prob(a')
-      \cdot\Z{i+1,j}{a',b} & \text{If }S_{i}=-1\\
+      x^{\gc(a')}
+      \cdot\Z{i+1,j}{a',b} &\begin{array}{@{}l@{}}\text{If }\Target_{i}=-1\\ \text{\relsize{-1}Position $i$ unpaired,}\end{array}\\
       \displaystyle
       \sum_{a',b'\in \B^2}
-			 \Prob(a')
-			 \cdot\Prob(b')
+			 x^{\gc(a'.b')}
 			 \cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
 			 \cdot \Z{i+1,j-1}{a',b'}&
-			 \text{Elif }S_i=j \land S_{i-1}=j+1\\
+\begin{array}{@{}l@{}}\text{Elif }\Target_i=j \land \Target_{i-1}=j+1\\ \text{\relsize{-1}$(i,j)$ base-paired, stacked onto $(i$-$1,j$+$1)$,}\end{array}
+			 \\
 			 \displaystyle
       \sum_{a',b'\in \B^2}
-      \Prob(a')
-			\cdot\Prob(b')
+      x^{\gc(a'.b')}
 			\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
       \cdot\Z{i+1,k-1}{a',b'}
-      \cdot\Z{k+1,j}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
+      \cdot\Z{k+1,j}{b',b} & 
+\begin{array}{@{}l@{}}\text{Elif }\Target_i=k \land i < k \leq j\\
+\text{\relsize{-1} Junction or bulge $\Rightarrow$ split, no stacking.}
+\end{array}
+\\
       0 &\text{Otherwise}
@@
 \label{eq:Z_rec}
-\end{equation}
-The cases can be broken down as follows:
-\begin{description}
-\item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
-any sequence consists in a nucleotide $a'$ at position  $i$, 
-followed by a sequence over $[i+1,j]$ and having flanking nucleotides $a'$ and $b$.
-\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$.
-Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
-then a stacking energy contribution is added. 
-\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
-only term contributing directly to the energy is the nucleotide distribution $\Prob(\cdot)$. 
-Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$.
-\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
-\end{description}
+\end{equation*}
+It is worth mentioning that, despite the appearances, this recurrence can be computed in linear $\Theta(|\Target|)$ time and space.
+Indeed, one can show that position $j$ is entirely determined by $i$ within subsequent recursive calls.
 
-\subsection{Stochastic backtrack}
+More precisely, let us consider the following predicate: {\em Let $(i',j')$ be the base-pair such that $i'<i<j'$ while minimizing $i-i'$, then TODO!!!!
 
-\begin{algorithm}
+$j$ is the position preceding the end of the closest base-pair nesting position $i$ if such a base-pair exists, or the last position in the sequence otherwise.} Obviously, the predicate holds for $[0,n-1]$.
+Moreover, if $i$ is unpaired, then $i+1$ and $i$
+
+%The cases can be broken down as follows:
+%\begin{description}
+%\item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
+%any sequence consists in a nucleotide $a'$ at position  $i$, 
+%followed by a sequence over $[i+1,j]$ and having flanking nucleotides $a'$ and $b$.
+%\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$.
+%Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
+%then a stacking energy contribution is added. 
+%\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
+%only term contributing directly to the energy is the nucleotide distribution $\Prob(\cdot)$. 
+%Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$.
+%\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
+%\end{description}
+
+\subsubsection{Stochastic backtrack}
+Once the partition functions have been computed, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, as described in Algorithm~\ref{alg:back}.
+\begin{algorithm}[t]
 	\SetAlgoLined
-	\lIf{$i>j$} \Return{$\varepsilon$};\tcp*[f]{Empty sequence}\\
-	rand $\leftarrow$ Random$\left(\Z{i,j}{a,b}\right)$\;
+\SetKwFunction{Backtrack}{Backtrack}
+\SetKwFunction{Random}{Random}
+	rand $\leftarrow$ \Random$\left(\Z{i,j}{a,b}\right)$\;
 	k $\leftarrow S_i$\;
@@
 		\For{$a'\in\B$}{
-			rand $\leftarrow$ rand $- \Prob(a')\cdot \Z{i+1,j}{a',b}$\;
-			\lIf{$\text{rand}<0$}\Return{$a'\cdot\text{Backtrack}\left((i+1,j),(a',b),S\right)$}\;
+			rand $\leftarrow$ rand $- x^{\gc(a')}\cdot \Z{i+1,j}{a',b}$\;
+			\lIf{$\text{rand}<0$}\Return{$a'.\Backtrack\left((i+1,j),(a',b),S\right)$}\;
 		}
@@
 	\Else{
-		\If(\tcp*[f]{Stack}){$k=j \land S_{i-1}=j+1$}{
+		\If(\tcp*[f]{Stacking pairs}){$k=j \land S_{i-1}=j+1$}{
 			\For{$(a',b')\in\B\times\B$}{
 				rand $\leftarrow$ rand $-
-			 	\Prob(a')
-			 	\cdot\Prob(b')
+			 	x^{\gc(a'.b')}
 			 	\cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
@@
 				\If{$\text{rand}<0$}{
-					\Return{$a'\cdot\text{Backtrack}\left((i+1,j-1),(a',b'),S\right)\cdot b'$}\;		
+					\Return{$a'.\Backtrack\left((i+1,j-1),(a',b'),S\right). b'$}\;		
 				}
@@
 		}
-		\Else(\tcp*[f]{Paired not stacked}){
+		\Else{
+			\If(\tcp*[f]{Paired, not stacked}){$\Target_i=k \land i < k \leq j$}{
 				\For{$(a',b')\in\B\times\B$}{
 					rand $\leftarrow$ rand $-	
-				\Prob(a')
-				\cdot\Prob(b')
+       x^{\gc(a'.b')}
 					\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
@@
 						\Return{$a'
-					\cdot\text{Backtrack}\left((i+1,k-1),(a',b'),S\right)
-					\cdot b'
-					\cdot\text{Backtrack}\left((k+1,j),(b',b),S\right)
+						.\Backtrack\left((i+1,k-1),(a',b'),S\right)
+						.b'
+						.\Backtrack\left((k+1,j),(b',b),S\right)
 						$}\;	
@@
 		 }
+    \Else {\Return{$\varepsilon$};	\tcp*[f]{Empty sequence}}
 		}
-\caption{Backtrack$\left((i,j),(a,b),S\right)$\label{alg:back}}
+	}
+\caption{\protect\Backtrack$\left((i,j),(a,b),S\right)$\label{alg:back}}
 \end{algorithm}
+
+\subsection{Overall complexity}
 

### 2013-01-07 ISMB/results_ISMB.tex
fixed ISMB headers
results + figs

@@
-%!TEX root = main_RECOMB.tex
+%!TEX root = main_ISMB.tex
 \section{Results}
 \label{sec:results}
+
+\subsection{Implementation}
+\label{sec:implementation}
+Our software, \texttt{Incarnation}, was implemented in Python2.7. We used
+\emph{RNAinverse} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.]
+
+TODO HARDWARE
+
+ We used as parameter $\beta=15$, the penalty for an invalid base pair.
+
+We present in Fig.~\ref{fig:time} the average times to generate one sequence
+with the required \texttt{G+C} content. As expected, the time grows linearly
+in function of the length of the structures.  A highly interesting feature
+is the huge decrease in time of computation when there is $15\%$ free 
+nucleotides inside the loops of the secondary structure.
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[scale=0.45]{Figures/time_rnastrand_clustered_rnainverse_100samples_fix}
+	\caption{Average time in seconds to generate one sequence. We explicitly show 
+	the time spent by \texttt{Incarnation} (full line) and \texttt{RNAinverse} (dashed) for various \texttt{G+C} contents. The first plot is in function
+	of the length of the structures. The second is in function of the stack
+	density (i.e. $2\cdot\#stacks/length$) and the latter in function of 
+	the proportion of free nucleotides (i.e. not base paired) inside loops.}
+	\label{fig:time}	
+\end{figure}
+
+\subsection{Dataset}
+To evaluate the quality of our method, we selected a subset of $50$ secondary
+structures from the \textit{RNA STRNAD} database~\cite{andronescu2008rna}.
+Those are known secondary structure from a variety of organisms with 
+ lengths between $20$ and $100$ nucleotides.
+
+ TODO MORE INFO RNASTRAND???
+
+ To ease the visualization of the data, we clustered together the strucures
+ in function of their length, stacks density and proportion of free 
+ nucleotides in loops. We present in Fig.~\ref{fig:bins} the distribution
+ of the structures in the different bins.
+
+ \begin{figure}[ht!]
+ 	\centering
+	\includegraphics[scale=0.45]{Figures/bins_distribution.png}
+	\caption{Number of secondary structures per bins.}
+	\label{fig:bins}
+ \end{figure}
+
+\subsection{Design}
+ To benchmark our method, we first started by sampling a minimum
+ $100$ sequence per structures. After correcting the free nucleotides with
+ \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
+
+A first criterion is which percentage of sequences have as MFE the target
+secondary structures. We are also interested by the number of structures
+for which we can design at least one sequence with the desired MFE. We
+present in Fig.~\ref{fig:mfe_struct_solved} those results. We observe
+that length is not a good predictor for the quality of the results. Instead,
+a high number of free nucleotides in the structure seems to be good measures of its design hardness. 
+On top of that, with a \texttt{G+C} content of $50\%$ and more, at least
+one solutions was found for almost all structures. The difficulty of 
+designing sequences for targets with a high number of free nucleotides 
+ in loops appears clearly in the last subfigure. Even with a high \texttt{G+C} content a few structures were not solved.
+
+\begin{figure}[ht!]	
+	\centering
+	\includegraphics[scale=0.45]{Figures/mfe_struct_solved}
+	\caption{Quality of the results. The first row shows the percentage
+	of sampled sequences folding into the target. The second shows the 	
+	proportion	of structures for which at least one correct sequence was 
+	sampled.}
+	\label{fig:mfe_struct_solved}	
+\end{figure}
+
+To evaluate the global quality of \texttt{Incarnation} sequences, we show
+in Fig.~\ref{fig:ss_sens} the ratio of well predicted base pairs in the
+MFE structure of our sampled sequences. We can observe that, in all cases,
+the hardest sequence to design are those with an extremely low \texttt{G+C}
+content. As anticipated, those are the sequences with the weakest bonds.
+We notice that the most accurate sequences yield a \texttt{G+C} content
+of $70\pm 10\%$.
+
+As discussed in Sec.~\ref{sec:implementation}, we notice a highly decreased
+computational time needed to generate the sequences with $15\%$ free 
+nucleotides in the loops. We remark that those structures also yield 
+a much lower structural sensitivity.
+
+\begin{figure}[ht!]
+ 	\centering
+	\includegraphics[scale=0.45]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
+	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }
+	\label{fig:ss_sens}	
+\end{figure}
+
+\subsection{Samples properties}
+
+In this section, we are only interested  in the samples folding into the 
+targetMFE.
+
+A desirable feature in sequence design, is to produce samples with a high
+diversity. We show in Fig.~\ref{fig:nb_sols_entropy} the number of correct
+solutions as their average entropy and base pair entropy, since \texttt{Incarnation}  only influences the distribution of nucleotides inside stacked base pairs.
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[scale=0.45]{Figures/nb_sols_entropy.png}
+	\caption{Number of solutions generated and their average entropy. 
+	The last row presents only the entropy of the nucleotides inside base 
+	pairs.}
+	\label{fig:nb_sols_entropy}
+\end{figure}
+
+An other critical properties in RNA sequence design is the frequency of
+the MFE. Presented in Fig.~\ref{fig:freq}.
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[scale=0.45]{Figures/freq.png}
+	\caption{Average frequency of the MFE.}
+	\label{fig:freq}
+\end{figure}

### 2013-01-07 ISMB/methods_ISMB.tex
Method v2

@@
 
-We show how a modification of the dynamic programming scheme introduced in the previously introduced \RNAmutants allow for the sampling of good and diverse design candidates in linear time and space complexities.
+We show how a modification of the dynamic programming scheme introduced in the previously contributed\RNAmutants allows for the sampling of good and diverse design candidates, in linear time and space complexities.
 
@@
 
-It was established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the \GCContent distribution observed in sequences sampled with respect to the above distribution converges to a normal law of expectation in $\mu\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma\cdot\sqrt{n}\cdot(1+o(1))$.
-Furthermore, it can easily be shown that the expected \GCContent is a continuous and strictly increasing monotonic  function of $x$, whose limits are $0$ when $x=0$ and $n$ when $x\to +\infty$.
-
-Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent.
-Such a value can be determined using any numerical recipe, and a simple bisection is used in our implementation.
-Furthermore, the concentration of the distribution, as asserted by the limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp.
-$\Theta(1)$, i.e.
-constant) for the expected number of attempts before generating a sequence having the exact targeted \GCContent  (resp.
-within $[gc-\kappa,gc+\kappa]$, where $\kappa\in\Theta(1/\sqrt n)$).
-
 \subsection{Sampling the pseudo-Boltzmann ensemble}
@@
 This algorithm follows the general principles of the recursive approach to random generation~\cite{Wilf1977}, pioneered in the context of RNA by the \SFold algorithm~\cite{Ding2003}.
-The algorithm starts by precomputing the partition function restricted to each valid interval, and then performs a series of stochastic backtracks, using precomputed values to decide on the probability of each alternative.
+The algorithm starts by precomputing the partition function restricted to each valid interval, and then performs a series of recursive stochastic backtracks, using precomputed values to decide on the probability of each alternative.
 
-\subsubsection{Precomputation}
-Let us describe a dynamic programming algorithm to compute $\Z{i,j}{a,b}$, the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively.
-
+\subsubsection{Precomputing the pseudo-partition function}\label{sec:pf}
+Firstly, a dynamic programming algorithm computes $\mathcal{Z}\substack{(i,j)\\ [a,b]}$ the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively.
 Since the only sequence over an empty interval is the empty sequence, having energy $0$, one has
@@
 The recursion consists in four terms, depending on the base-pairing status and context of $(i,j)$:
-\begin{equation*}
+\begin{equation}
 	\Z{i,j}{a,b}:=\left\{
@@
 			 \cdot \Z{i+1,j-1}{a',b'}&
-\begin{array}{@{}l@{}}\text{Elif }\Target_i=j \land \Target_{i-1}=j+1\\ \text{\relsize{-1}$(i,j)$ base-paired, stacked onto $(i$-$1,j$+$1)$,}\end{array}
+\begin{array}{@{}l@{}}\text{Elif }\Target_i=j \land \Target_{i-1}=j+1\\ \text{\relsize{-1}$(i,j)$ stacking pair onto $(i$-$1,j$+$1)$,}\end{array}
 			 \\
@@
 \begin{array}{@{}l@{}}\text{Elif }\Target_i=k \land i < k \leq j\\
-\text{\relsize{-1} Junction or bulge $\Rightarrow$ split, no stacking.}
+\text{\relsize{-1}Junction/bulge $\Rightarrow$ split, no stack,}
 \end{array}
 \\
-      0 &\text{Otherwise}
+      0 &\text{Otherwise.}
 	\end{array}\right.
 \label{eq:Z_rec}
-\end{equation*}
-It is worth mentioning that, despite the appearances, this recurrence can be computed in linear $\Theta(|\Target|)$ time and space.
-Indeed, one can show that position $j$ is entirely determined by $i$ within subsequent recursive calls.
+\end{equation}
 
-More precisely, let us consider the following predicate: {\em Let $(i',j')$ be the base-pair such that $i'<i<j'$ while minimizing $i-i'$, then TODO!!!!
+Despite its apparent quadratic complexity, the recurrence described in Subsection~\ref{sec:pf} can be computed in linear $\Theta(|\Target|)$ time and space, owing to the fact that, on any recursive call, the position $j$ is entirely determined by $i$.
 
-$j$ is the position preceding the end of the closest base-pair nesting position $i$ if such a base-pair exists, or the last position in the sequence otherwise.} Obviously, the predicate holds for $[0,n-1]$.
-Moreover, if $i$ is unpaired, then $i+1$ and $i$
+\begin{theorem}Starting from the complete region $[0,n-1]$, Equation~\eqref{eq:Z_rec} can be computed in $\Theta(|\Target|)$ time and space complexity.
+\end{theorem}
+\begin{proof}
 
-%The cases can be broken down as follows:
-%\begin{description}
-%\item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
-%any sequence consists in a nucleotide $a'$ at position  $i$, 
-%followed by a sequence over $[i+1,j]$ and having flanking nucleotides $a'$ and $b$.
-%\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$.
-%Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
-%then a stacking energy contribution is added. 
-%\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
-%only term contributing directly to the energy is the nucleotide distribution $\Prob(\cdot)$. 
-%Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$.
-%\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
-%\end{description}
+Consider the {\bf covering base-pair} $\CNBP{i}$ of a position $i$, defined as the base-pair $(i',j')$ such that $i'<i<j'$ and which minimizes $i-i'$.
+For the sake of simplicity, and without loss of generality, let us assume that there exists a virtual base-pair $(-1,n)$ in $\Target$, so that any position in $[0,n-1]$ has a covering base-pair.
+Then one easily shows by induction that, starting from the complete interval $[0,n-1]$, the recursive calls only evaluate $\mathcal{Z}$ over intervals $[i,j]$ such that $j=j'-1$, $\CNBP{i}=[i',j']$.
+
+\begin{cframed}This property clearly holds initially for the complete interval $[0,n-1]$, owing to the virtual base-pair $(-1,n)$.
+Furthermore, if it holds for a given $(i,j)$, $\CNBP{i}=[i',j+1]$, then:
+\begin{itemize} 
+  \item If $i$ is unpaired (Case 1), then $\CNBP{i+1}=\CNBP{i}=[i',j+1]$, and the property is satisfied by the subsequent call over $[i+1,j]$. 
+  \item If $i$ is paired with $j$ and stacking onto $(i-1,j+1)$ (Case 2), then one has $\CNBP{i+1}=[i,j]$ and the  subsequent call over $[i+1,j-1]$ satisfies the property.
+  \item If $i$ is paired with some $k$ (Case 3), then $\CNBP{i+1}=[i,k]$ and $\CNBP{k+1}=[i',j+1]$.
+Therefore the subsequent calls, respectively over $[i+1,k-1]$ and $[k+1,j]$, both satisfy the property.
+\end{itemize}
+The case of an empty sequence does not cause a recursive call and must not be verified.
+\end{cframed}
+Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(|\Target|)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges and only contribute to a constant in the complexity.
+\end{proof}
 
 \subsubsection{Stochastic backtrack}
-Once the partition functions have been computed, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, as described in Algorithm~\ref{alg:back}.
+Once the pseudo-partition functions have been computed, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation.
+One or several recursive calls over regions that are left unassigned are then performed, as described by Algorithm~\ref{alg:back}.
+
+Since, during each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position, and that the number of executions of the loops are bounded by a constant, then the complexity of the algorithm grows linearly on the targeted length.
 \begin{algorithm}[t]
+\DontPrintSemicolon
 	\SetAlgoLined
@@
 \SetKwFunction{Random}{Random}
-	rand $\leftarrow$ \Random$\left(\Z{i,j}{a,b}\right)$\;
+	rand $\leftarrow$ \Random$\left(\Z{i,j}{a,b}\right)$\tcp*[r]{Draw random number in $[0,\Z{i,j}{a,b}[$}
+ \lIf(\tcp*[f]{Empty sequence}){$j<i$}{\Return{$\varepsilon$}}
+	\Else{
   k $\leftarrow S_i$\;
-	\If(\tcp*[f]{Unpaired}){$k = -1$}{
+	\If(\tcp*[f]{Unpaired}){$(k = -1)$}{
 		\For{$a'\in\B$}{
 			rand $\leftarrow$ rand $- x^{\gc(a')}\cdot \Z{i+1,j}{a',b}$\;
-			\lIf{$\text{rand}<0$}\Return{$a'.\Backtrack\left((i+1,j),(a',b),S\right)$}\;
+			\lIf{$\text{rand}<0$}\Return{$a'.\Backtrack\left(i+1,j,a',b,\Target\right)$}\;
 		}
 	}
-	\Else{
-		\If(\tcp*[f]{Stacking pairs}){$k=j \land S_{i-1}=j+1$}{
+	\lElse{
+		\If(\tcp*[f]{Stacking pairs}){$(k=j) \land (S_{i-1}=j+1)$}{
 			\For{$(a',b')\in\B\times\B$}{
@@
 			 	\cdot \Z{i+1,j-1}{a',b'}	$\;
-				\If{$\text{rand}<0$}{
-					\Return{$a'.\Backtrack\left((i+1,j-1),(a',b'),S\right). b'$}\;		
+				\lIf{$\text{rand}<0$}{
+					\Return{$a'.\Backtrack\left(i+1,j-1,a',b',\Target\right). b'$}\;		
 				}
@@
 		}
-		\Else{
-			\If(\tcp*[f]{Paired, not stacked}){$\Target_i=k \land i < k \leq j$}{
+		\lElse{
+			\If(\tcp*[f]{Paired, not stacked}){$(\Target_i=k) \land (i < k \leq j)$}{
 				\For{$(a',b')\in\B\times\B$}{
@@
 						\Return{$a'
-						.\Backtrack\left((i+1,k-1),(a',b'),S\right)
+						.\Backtrack\left(i+1,k-1,a',b',\Target\right)
 						.b'
-						.\Backtrack\left((k+1,j),(b',b),S\right)
+						.\Backtrack\left(k+1,j,b',b,\Target\right)
 						$}\;	
@@
 		 }
-    \Else {\Return{$\varepsilon$};	\tcp*[f]{Empty sequence}}
 		}
 	}
-\caption{\protect\Backtrack$\left((i,j),(a,b),S\right)$\label{alg:back}}
+ }
+\caption{\protect\Backtrack$\left(i,j,a,b,\Target\right)$\label{alg:back}}
 \end{algorithm}
 
-\subsection{Overall complexity}
+\subsubsection{Rejecting unsuitable candidates}
+
+Note that, although the parameter $x$ is introduced to match the expected \GCContent with the targeted one, nothing prevents Algorithm~\ref{alg:back} to generate sequences of arbitrary \GCContent.
+Therefore, we use a rejection-based approach~\cite{Bodini2010} previously used by the authors in a similar context~\cite{Waldispuhl2011}.
+
+It was established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the \GCContent distribution observed in sequences sampled with respect to the above distribution converges asymptotically towards a normal law having expectation in $\mu\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma\cdot\sqrt{n}\cdot(1+o(1))$.
+Furthermore, it can easily be shown that the expected \GCContent is a continuous and strictly increasing monotonic  function of $x$, whose limits are $0$ when $x=0$ and $n$ when $x\to +\infty$.
+
+Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent.
+Such a value can be determined using any numerical recipe, and a simple bisection procedure, further described in earlier work~\cite{Waldispuhl2011}, is used in our implementation.
+Furthermore, the concentration of the distribution, as asserted by the limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp.
+$\Theta(1)$, i.e.
+constant) for the expected number of attempts before generating a sequence having the exact targeted \GCContent  (resp.
+within $[gc-\kappa,gc+\kappa]$, where $\kappa\in\Theta(1/\sqrt n)$).
+
+\subsection{Redesigning unpaired regions: A local/global hybrid approach}
 

### 2013-01-07 ISMB/results_ISMB.tex
figs

@@
 Our software, \texttt{Incarnation}, was implemented in Python2.7. We used
-\emph{RNAinverse} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.]
-
-TODO HARDWARE
-
+\emph{RNAinverse} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
+All time benchmarks were run on a one AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.
  We used as parameter $\beta=15$, the penalty for an invalid base pair.
@@
 secondary structures. We are also interested by the number of structures
-for which we can design at least one sequence with the desired MFE. We
-present in Fig.~\ref{fig:mfe_struct_solved} those results. We observe
+for which we can design at least one sequence with the desired MFE.
+We present in 
+ Fig.~\ref{fig:mfe_struct_solved_noinverse} these results when only
+ \texttt{Incarnation} is applied, and 
+in Fig.~\ref{fig:mfe_struct_solved} after processing by \RNAinverse. We observe
 that length is not a good predictor for the quality of the results. Instead,
@@
 	\centering
+	\includegraphics[scale=0.45]{Figures/mfe_struct_solve_nornainverse.png}
+	\caption{Quality of \texttt{Incarnation} results. The first row shows the percentage
+	of sampled sequences folding into the target. The second shows the 	
+	proportion	of structures for which at least one correct sequence was 
+	sampled.}
+	\label{fig:mfe_struct_solved_noinverse}	
+\end{figure}
+
+\begin{figure}[ht!]	
+	\centering
 	\includegraphics[scale=0.45]{Figures/mfe_struct_solved}
-	\caption{Quality of the results. The first row shows the percentage
+	\caption{Quality of \texttt{Incarnation+RNAinverse} results. The first row shows the percentage
 	of sampled sequences folding into the target. The second shows the 	
@@
  	\centering
-	\includegraphics[scale=0.45]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
+	\includegraphics[width=1\textwidth,height=0.16\textheight]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
 	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }

### 2013-01-07 ISMB/methods_ISMB.tex
Further into methods section

@@
 
-%\subsubsection{Profile}
-%At every position, the nucleotide distribution is the same. 
-%It is first initialized equiprobably. A bisection scheme is then applied until the sampled sequences $\Cb+\Gb$ content is as desired. The probability
-%of a given nucleotide $a\in\B$ will be given by $\Prob(a)$ at every 
-%position.
+A targeted secondary structure $\Target$ is abstracted as a non-crossing arc-annotated sequence,  where 
+$\Target_i$ stands for the base-pairing position of position $i$ in $\Target$ if any (and, reciprocally, $\Target_{\Target_i}=i$), or $-1$ otherwise. 
+Let us denote by $\gc(s)$ the number of occurrences of \Gb and \Cb in an RNA sequence $s$.
 
-\subsubsection{Secondary structure}
-Todo...
-
-\subsubsection{Energy model}
+\subsubsection{Simplified energy model}
 The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}).
@@
 
-\subsection{\GCContent weighted pseudo-Boltzmann ensemble and distribution}
+\subsubsection{\GCContent weighted pseudo-Boltzmann ensemble and distribution}
 
 In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, let us introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences.
-Let $\Target$ be a targeted secondary structure and $\gc(s)$ be the number of occurrences of \Gb and \Cb in an RNA sequence $s$, then let us define the pseudo-Boltzmann factor $\B_{x}(s,S)$ of a sequence $s$ such that
+Let $\Target$ be a targeted secondary structure, then let us define the pseudo-Boltzmann factor $\B_{x}(s,S)$ of a sequence $s$ such that
 \begin{equation}
@@
 
-Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$ gives the pseudo-partition function $\mathcal{Z}_{\Target,x}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that
+Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$ gives the pseudo-partition function $\mathcal{Z}_{x,\Target}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that
 \begin{align}\mathcal{Z}_{x,\Target} &= \sum_{\substack{|s|=|\Target|}}\B_{x,\Target}(s)& \text{and}&&
@@
 \end{cframed}
-Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(|\Target|)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges and only contribute to a constant in the complexity.
+Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(|\Target|)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges and only contribute to a constant factor in the complexity.
 \end{proof}
@@
 \subsubsection{Stochastic backtrack}
-Once the pseudo-partition functions have been computed, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation.
-One or several recursive calls over regions that are left unassigned are then performed, as described by Algorithm~\ref{alg:back}.
-
-Since, during each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position, and that the number of executions of the loops are bounded by a constant, then the complexity of the algorithm grows linearly on the targeted length.
+Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation.
+One or several recursive calls over regions that are left unassigned are then performed, as described by Algorithm~\ref{alg:back}, and illustrated by Figure~\ref{fig:stochastic}.
 \begin{algorithm}[t]
@@
 
+Since, during each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position, and that the number of executions of the loops are bounded by a constant, then the complexity of the algorithm grows linearly on the targeted length.
+
+\begin{figure}
+\resizebox{\textwidth}{!}{\input{FigStochasticBacktrack.pgf}}
+\caption{Three main cases in the stochastic backtrack procedure over an interval $[i,j]$: Either position $i$ is left unpaired (top), a base-pair $(i,j)$ is formed, stacking onto $(i+1,j-1)$ (middle), or $i$ is paired somewhere else, defining two regions on which subsequent recursive calls are needed (bottom).
+Positions indicated in red are assigned during at this stage of the backtrack.\label{fig:stochastic}}
+\end{figure}
+
 \subsubsection{Rejecting unsuitable candidates}
@@
 \subsection{Redesigning unpaired regions: A local/global hybrid approach}
+Due to the simplification in our energy model, unpaired regions are not subject design constraints other than the \GCContent, leading to modest similarities of refolded design candidates with the targeted structure.
+To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions, specifying a constraint mask to prevent stacking base-pairs from being modified.
 

### 2013-01-08 ISMB/results_ISMB.tex
graph
res / supp data

@@
 We present in Fig.~\ref{fig:time} the average times to generate one sequence
-with the required \texttt{G+C} content. As expected, the time grows linearly
+with the required \texttt{C+G} content. As expected, the time grows linearly
 in function of the length of the structures.  A highly interesting feature
@@
 	\caption{Average time in seconds to generate one sequence. We explicitly show 
-	the time spent by \texttt{Incarnation} (full line) and \texttt{RNAinverse} (dashed) for various \texttt{G+C} contents. The first plot is in function
+	the time spent by \texttt{Incarnation} (full line) and \texttt{RNAinverse} (dashed) for various \texttt{C+G} contents. The first plot is in function
 	of the length of the structures. The second is in function of the stack
@@
 for which we can design at least one sequence with the desired MFE.
-We present in 
- Fig.~\ref{fig:mfe_struct_solved_noinverse} these results when only
- \texttt{Incarnation} is applied, and 
-in Fig.~\ref{fig:mfe_struct_solved} after processing by \RNAinverse. We observe
+We show the results when using only \texttt{Incarnation} in 
+ Fig.~\ref{fig:mfe_struct_solved_noinverse}. The high rate of failure
+is due to the fact that no selection criterions are applied to
+free nucleotides. A local strategy is thus needed.
+After processing \texttt{Incarnation} sequences with \RNAinverse to 
+locally optimize the free nucleotides, we obtain the results 
+in Fig.~\ref{fig:mfe_struct_solved}. We observe
 that length is not a good predictor for the quality of the results. Instead,
-a high number of free nucleotides in the structure seems to be good measures of its design hardness. 
-On top of that, with a \texttt{G+C} content of $50\%$ and more, at least
+a high number of free nucleotides in the structure seems to be a 
+good measure of its design hardness. 
+On top of that, with a \texttt{C+G} content of $50\%$ and more, at least
 one solutions was found for almost all structures. The difficulty of 
 designing sequences for targets with a high number of free nucleotides 
- in loops appears clearly in the last subfigure. Even with a high \texttt{G+C} content a few structures were not solved.
+ in loops appears clearly in the last subfigure. Even with a high \texttt{C+G} content a our method was not able to generate solutions for a few
+ structures.
 
@@
 MFE structure of our sampled sequences. We can observe that, in all cases,
-the hardest sequence to design are those with an extremely low \texttt{G+C}
+the hardest sequence to design are those with an extremely low \texttt{C+G}
 content. As anticipated, those are the sequences with the weakest bonds.
-We notice that the most accurate sequences yield a \texttt{G+C} content
+We notice that the most accurate sequences yield a \texttt{C+G} content
 of $70\pm 10\%$.
@@
  	\centering
-	\includegraphics[width=1\textwidth,height=0.16\textheight]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
+	\includegraphics[scale=0.45]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
 	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }
@@
 diversity. We show in Fig.~\ref{fig:nb_sols_entropy} the number of correct
-solutions as their average entropy and base pair entropy, since \texttt{Incarnation}  only influences the distribution of nucleotides inside stacked base pairs.
+solutions as their average entropy and base pair entropy, since 
+\texttt{Incarnation}  only influences the distribution of nucleotides inside 
+stacked base pairs. The most constraining factor for generating valid
+ solutions seems to be  low \texttt{C+G} content and the percentage of free nucleotides. Our method is able to generate
+
+Also, a critical properties in RNA sequence design is 
+the frequency of the MFE. 
+The sequence should be in its target conformation long enough to
+perform the desired action. Presented in Fig.~\ref{fig:freq} we see that
+there is a slow decline of the frequency with the increase in size. Yet,
+for an average \texttt{C+G} content, the frequency stays over $10\%$ even
+at size of $100$ nucleotides.
 
@@
 
-An other critical properties in RNA sequence design is the frequency of
-the MFE. Presented in Fig.~\ref{fig:freq}.
-
 \begin{figure}[ht!]

### 2013-01-08 ISMB/results_ISMB.tex
Some typos + figs

--- 2013-01-07 ISMB/results_ISMB.tex
@@
 	\centering
-	\includegraphics[scale=0.45]{Figures/time_rnastrand_clustered_rnainverse_100samples_fix}
+	\includegraphics[width=\textwidth]{Figures/time_rnastrand_clustered_rnainverse_100samples_fix}
 	\caption{Average time in seconds to generate one sequence. We explicitly show 
@@
 To evaluate the quality of our method, we selected a subset of $50$ secondary
-structures from the \textit{RNA STRNAD} database~\cite{andronescu2008rna}.
+structures from the \textit{RNA STRAND} database~\cite{andronescu2008rna}.
 Those are known secondary structure from a variety of organisms with 
@@
  	\centering
-	\includegraphics[scale=0.45]{Figures/bins_distribution.png}
+	\includegraphics[width=\textwidth]{Figures/bins_distribution.png}
 	\caption{Number of secondary structures per bins.}
@@
 	\centering
-	\includegraphics[scale=0.45]{Figures/mfe_struct_solve_nornainverse.png}
+	\includegraphics[width=\textwidth]{Figures/mfe_struct_solve_nornainverse.png}
 	\caption{Quality of \texttt{Incarnation} results. The first row shows the percentage
@@
 	\centering
-	\includegraphics[scale=0.45]{Figures/mfe_struct_solved}
+	\includegraphics[width=\textwidth]{Figures/mfe_struct_solved}
 	\caption{Quality of \texttt{Incarnation+RNAinverse} results. The first row shows the percentage
@@
  	\centering
-	\includegraphics[width=1\textwidth,height=0.16\textheight]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
+	\includegraphics[width=\textwidth]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
+	%\includegraphics[width=1\textwidth,height=0.16\textheight]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
 	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }
@@
 	\centering
-	\includegraphics[scale=0.45]{Figures/nb_sols_entropy.png}
+	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy.png}
 	\caption{Number of solutions generated and their average entropy. 
@@
 	\centering
-	\includegraphics[scale=0.45]{Figures/freq.png}
+	\includegraphics[width=\textwidth]{Figures/freq.png}
 	\caption{Average frequency of the MFE.}

### 2013-01-08 ISMB/introduction_ISMB.tex
updated ref bodini

@@
 Unlike \RNAensign, it does not require a seed sequence to start its search.
-Finally, \RNApyro implements a novel algorithm using weighted sampling techniques \cite{DBLP:journals/corr/abs-1002-0046} that enables us to control, for the first time, \textit{explicitly} the \texttt{C+G}-content of the solution.
+Finally, \RNApyro implements a novel algorithm using weighted sampling techniques \cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \texttt{C+G}-content of the solution.
 This functionality is essential because wild sequences within living organisms often present medium or low \texttt{C+G}-content.

### 2013-01-08 ISMB/results_ISMB.tex
supp

--- 2013-01-08 ISMB/results_ISMB.tex
@@
 
-We present in Fig.~\ref{fig:time} the average times to generate one sequence
-with the required \texttt{C+G} content. As expected, the time grows linearly
-in function of the length of the structures.  A highly interesting feature
+We present the average times spent by \texttt{Incarnation} and \RNAinverse
+ to generate one sequence
+with the required \texttt{C+G} content in Fig.~\ref{fig:time}. As expected, the time grows linearly
+in function of the length of the structures for \texttt{Incarnation}.  A highly interesting feature
 is the huge decrease in time of computation when there is $15\%$ free 
@@
 \subsection{Dataset}
-To evaluate the quality of our method, we selected a subset of $50$ secondary
+To evaluate the quality of our method, we used secondary
 structures from the \textit{RNA STRNAD} database~\cite{andronescu2008rna}.
-Those are known secondary structure from a variety of organisms with 
- lengths between $20$ and $100$ nucleotides.
-
- TODO MORE INFO RNASTRAND???
+Those are known secondary structure from a variety of organisms.
+We select the same subset of $50$ structures as in~\cite{Levin:2012kx}, 
+which have lengths between $20$ and $100$ nucleotides.
 

### 2013-01-08 ISMB/methods_ISMB.tex

@@
 
-\begin{cframed}This property clearly holds initially for the complete interval $[0,n-1]$, owing to the virtual base-pair $(-1,n)$.
-Furthermore, if it holds for a given $(i,j)$, $\CNBP{i}=[i',j+1]$, then:
+\CFramed{This property clearly holds initially for the complete interval $[0,n-1]$, owing to the virtual base-pair $(-1,n)$. Furthermore, if it holds for a given $(i,j)$, $\CNBP{i}=[i',j+1]$, then:
 \begin{itemize} 
@@
 The case of an empty sequence does not cause a recursive call and must not be verified.
-\end{cframed}
+}
 Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(|\Target|)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges and only contribute to a constant factor in the complexity.
@@
 \subsection{Redesigning unpaired regions: A local/global hybrid approach}
-Due to the simplification in our energy model, unpaired regions are not subject design constraints other than the \GCContent, leading to modest similarities of refolded design candidates with the targeted structure.
+Due to the simplification in our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest similarities of refolded design candidates with the targeted structure.
 To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions, specifying a constraint mask to prevent stacking base-pairs from being modified.
+{\em todo: discuss evolution of \GCContent}
 

### 2013-01-08 ISMB/supplementary_ISMB.tex
supp

%!TEX root = main_ISMB.tex
\section{Supplementary data}
\subsection{RNAinverse}
Using the same dataset of $50$ structures, we generated $100$ samples
per structure with \RNAinverse. They yield for most parameters
a high entropy. Although, the $\texttt{C+G}$ content variation is minimal
through all samples. We show the results when using only \RNAinverse, 
only \texttt{Incarnation}, and both.

\begin{figure}[ht!]
	\centering
	\includegraphics[scale=0.4]{Figures/RNAinverse_data_100.png}
	\caption{Entropy and \texttt{C+G} content. The first
		column is in function of the length, the second of the stacks
		density and the last of the percentage of free nucleotides in loops.
		The blue line represents the results and standard deviation when using 
		only \RNAinverse with 100samples.  The dashed lines are the results
		when using only \texttt{Incarnation} and the dashed line are results
		of \texttt{Incarnation} post-processed by \RNAinverse.
		}
	\label{fig:rnainverse}
\end{figure}

### 2013-01-08 ISMB/introduction_ISMB.tex
update abstract

@@
 
-In this paper we introduce \RNApyro, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}.
-\RNApyro addresses previous limitations of \RNAensign and offers new functionalities.
-First, while our previous program had a running time complexity of $\mathcal{O}(n^4)$, \RNApyro has now a linear-time and space complexity that allows him to demonstrate similar speeds to any local search algorithm.
-Next, \RNApyro is \textit{seed-less}.
+In this paper we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}.
+\ourprog addresses previous limitations of \RNAensign and offers new functionalities.
+First, while our previous program had a running time complexity of $\mathcal{O}(n^4)$, \ourprog has now a linear-time and space complexity that allows him to demonstrate similar speeds to any local search algorithm.
+Next, \ourprog is \textit{seed-less}.
 Unlike \RNAensign, it does not require a seed sequence to start its search.
-Finally, \RNApyro implements a novel algorithm using weighted sampling techniques \cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \texttt{C+G}-content of the solution.
-This functionality is essential because wild sequences within living organisms often present medium or low \texttt{C+G}-content.
+Finally, \ourprog implements a novel algorithm using weighted sampling techniques \cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \gc~content of the solution.
+This functionality is essential because wild sequences within living organisms often present medium or low \gc~content.
 Presumably to offer better transcription rates and/or structural plasticity.
-Previous programs do not allow to control this parameter and tend to output sequences with high \texttt{C+G}-contents.
+Previous programs do not allow to control this parameter and tend to output sequences with high \gc~contents.
 
-We demonstrate the performance of our algorithms on a set of real RNA structures with sizes ranging from *** to *** nucleotides.
+We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \textsf{RNA STRAND} database \cite{andronescu2008rna}.
 To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse.

### 2013-01-08 ISMB/main_ISMB.tex
algo
Method... interrupted
Further into methods section
res / supp data
Some typos + figs
update abstract

@@
 \usepackage{xspace}
+\usepackage{tcolorbox}
 \usepackage{relsize}
@@
 
-\newcommand{\ourprog}{\texttt{RNA-MoIP}\xspace}
+\usepackage[noend,ruled,vlined]{algorithm2e}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{corollary}[theorem]{Corollary}
+
+\newcommand{\ourprog}{\texttt{Incarnation}\xspace}
 
@@
 \newcommand{\RNApyro}{\texttt{RNApyro}\xspace}
+\newcommand{\RNAinverse}{\texttt{RNAinverse}\xspace}
+\newcommand{\RNASSD}{\texttt{RNA-SSD}\xspace}
+\newcommand{\INFORNA}{\texttt{INFO-RNA}\xspace}
+\newcommand{\NUPACK}{\texttt{NUPACK}\xspace}
+\newcommand{\RNAiFOLD}{\texttt{RNAiFOLD}\xspace}
+\newcommand{\frankenstein}{\texttt{Frnakenstein}\xspace}
+\newcommand{\RNAexinv}{\texttt{RNAexinv}\xspace}
+\newcommand{\rnaDesign}{\texttt{rnaDesign}\xspace}
+\newcommand{\RNAensign}{\texttt{RNA-ensign}\xspace}
+\newcommand{\GCContent}{\Gb\Cb-content\xspace}
+\newcommand{\SFold}{\texttt{SFold}\xspace}
 
@@
 %\newcommand{\citep}{\cite}
-\newcommand{\Z}[3]{\mathcal{Z}_{\substack{(#1)\\ [#3]}}^{#2}}
+\newcommand{\Z}[2]{\mathcal{Z}{\substack{(#1)\\ [#2]}}}
 \newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
@@
 
-\newcommand{\Ab}{{\sf{A}}}
-\newcommand{\Cb}{{\sf{C}}}
-\newcommand{\Gb}{{\sf{G}}}
-\newcommand{\Ub}{{\sf{U}}}
+\newcommand{\Ab}{{\sf{A}}\xspace}
+\newcommand{\Cb}{{\sf{C}}\xspace}
+\newcommand{\Gb}{{\sf{G}}\xspace}
+\newcommand{\Ub}{{\sf{U}}\xspace}
 
-\title{}
-\author{}
-\date{}
+\newcommand{\gc}{{\#\text{{\sf gc}}}}
+\newcommand{\PE}[1]{E(#1)}
+\newcommand{\EI}{\text{EI}}
+\newcommand{\ES}{\text{{\sf Es}}}
+\newcommand{\ISO}{\text{ISO}}
+\newcommand{\Prob}{\text{P}}
+\newcommand{\Target}{\text{\sf S}}
+\newcommand{\CNBP}[1]{\gamma(#1)}
+
+\newcommand{\CFramed}[2][gray]{\begin{tcolorbox}[colframe=#1,title=To be included only if space allows\ldots,colback=white]#1\end{tcolorbox}}
+\renewcommand{\CFramed}[2][gray]{}
+
+\title{A glocal approach to designing RNA sequences with target secondary structure and nucleotide distribution}
+\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
+\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, ƒcole Polytechnique, Palaiseau, France.\\Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}
 
@@
 \begin{abstract}
+\noindent
+\textbf{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies.
+Most of the current software use similar local search strategies (i.e.
+a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \gc~content in their sequences.
+However, the latter is an important criteria for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.
+We recently introduced a novel global sampling approach for the RNA secondary structure design problem.
 
 \noindent
-\textbf{Key words:} RNA, mutations, secondary structure
+\textbf{Results:}
+In this paper, we introduce \ourprog
+
+\noindent
+\textbf{Key words:} RNA, secondary structure, design, weighted sampling, GC-content.
 \end{abstract}
@@
 
+\newpage
+
+\input{supplementary_ISMB.tex}
+
 \end{document}

### 2013-01-08 ISMB/supplementary_ISMB.tex
sup / fig

@@
 per structure with \RNAinverse. They yield for most parameters
-a high entropy. Although, the $\texttt{C+G}$ content variation is minimal
-through all samples. We show the results when using only \RNAinverse, 
-only \texttt{Incarnation}, and both.
+a high entropy. We present for every different concentration of \texttt{C+G}
+the average sample entropy and the average base pairs entropy.
 
@@
 	\includegraphics[scale=0.4]{Figures/RNAinverse_data_100.png}
-	\caption{Entropy and \texttt{C+G} content. The first
-		column is in function of the length, the second of the stacks
-		density and the last of the percentage of free nucleotides in loops.
-		The blue line represents the results and standard deviation when using 
-		only \RNAinverse with 100samples.  The dashed lines are the results
-		when using only \texttt{Incarnation} and the dashed line are results
-		of \texttt{Incarnation} post-processed by \RNAinverse.
-		}
+	\caption{Entropy and \texttt{C+G} content.}
 	\label{fig:rnainverse}

### 2013-01-08 ISMB/main_ISMB.tex
completed title/abstract/intro

@@
 \author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
-\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, ƒcole Polytechnique, Palaiseau, France.\\Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}
+\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, ƒcole Polytechnique, Palaiseau, France.\\ \small $^*$Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}
 
@@
 a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \gc~content in their sequences.
-However, the latter is an important criteria for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.
-We recently introduced a novel global sampling approach for the RNA secondary structure design problem.
-
+However, the latter is an important criteria for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.\\
 \noindent
 \textbf{Results:}
-In this paper, we introduce \ourprog
-
+In this paper, we introduce \ourprog, a novel algorithm to design RNA sequences folding into target secondary structures with a predefined nucleotide distribution.
+\ourprog uses a global sampling approach and weighted sampling techniques.
+We show that our approach is fast (i.e.
+running time comparable or better than local search methods), seed-less (we remove the bias of the seed in local search heuristics), and successfully generates high-quality sequences (i.e.
+thermodynamically stable) for any \gc~content.
+To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies.
+Remarkably, our glocal methodology overcomes both local and global approaches.\\
 \noindent

### 2013-01-08 ISMB/results_ISMB.tex
typo & supp data

@@
 
-We present the average times spent by \texttt{Incarnation} and \RNAinverse
- to generate one sequence
+We present the average times spent by \texttt{Incarnation} and \RNAinverse to generate one sequence
 with the required \texttt{C+G} content in Fig.~\ref{fig:time}. As expected, the time grows linearly
@@
 To evaluate the quality of our method, we used secondary
-structures from the \textit{RNA STRNAD} database~\cite{andronescu2008rna}.
+structures from the \textit{RNA STRAND} database~\cite{andronescu2008rna}.
 Those are known secondary structure from a variety of organisms.

### 2013-01-08 ISMB/supplementary_ISMB.tex
typo & supp data

@@
 \section{Supplementary data}
-\subsection{RNAinverse}
+\subsection{Benchmark \ourprog vs \RNAinverse}
 Using the same dataset of $50$ structures, we generated $100$ samples
@@
 \begin{figure}[ht!]
-	\centering
+	%\centering
+	\hspace{-5em}
 	\includegraphics[scale=0.4]{Figures/RNAinverse_data_100.png}
@@
 \end{figure}
+
+\subsection{Impact of local search on the \gc~content of \ourprog output}
+
+These data demonstrate that the local search heuristic used to design nucleotides in loop regions has a very limited impact on the \gc content.
+For each class of \gc~content, we show in Table~\ref{table:impact_on_gc}, the observed \gc~content of the sequence sampled by \ourprog, and then we report the observed \gc~content after the \RNAinverse step (as defined in Section \ref{subsec:glocal_method}).
+Our results show that the \gc~content is relatively well conserved.
+
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+target \gc & \ourprog & \ourprog + \RNAinverse \\
+\hline
+10 & 0.15 & 0.21\\
+30 & 0.30 & 0.33\\
+50 & 0.48 & 0.49\\
+70 & 0.71 & 0.69\\
+90 & 0.83 & 0.78\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Observed \gc~content of solutions returned by \ourprog (2nd column) and after the application of the local search heuristic (3rd column).}
+\label{table:impact_on_gc}
+\end{table}

### 2013-01-08 ISMB/supplementary_ISMB.tex
sup

--- 2013-01-08 ISMB/supplementary_ISMB.tex
@@
 per structure with \RNAinverse. They yield for most parameters
-a high entropy. We present for every different concentration of \texttt{C+G}
+a high entropy. For all structures that have been solved 
+by the three methods, only \RNAinverse, only \texttt{Incarnation} and
+\texttt{Incarnation} followed by \RNAinverse,
+we present for every different concentration of \texttt{C+G}
 the average sample entropy and the average base pairs entropy.
@@
 	\includegraphics[scale=0.4]{Figures/RNAinverse_data_100.png}
-	\caption{Entropy and \texttt{C+G} content.}
+	\caption{Entropy and \texttt{C+G} content for structures solved by
+	the 3 methods.}
 	\label{fig:rnainverse}
 \end{figure}
+
+%The same analysis but only for structures solved by \texttt{Incarnation}
+%and \texttt{Incarnation} followed by \RNAinverse, is presented in 
+%Fig.~\ref{fig:inc_rnainv}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics{}
+%	\caption{Entropy and \texttt{C+G} content for structures solved by
+%	the \texttt{Incarnation} and \texttt{Incarnation}$+$\RNAinverse.}
+%	\label{fig:inc_rnainv}
+%\end{figure}

### 2013-01-08 ISMB/methods_ISMB.tex

@@
 \subsection{Redesigning unpaired regions: A local/global hybrid approach}
+\label{subsec:glocal_method}
 Due to the simplification in our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest similarities of refolded design candidates with the targeted structure.

### 2013-01-08 ISMB/supplementary_ISMB.tex

--- 2013-01-08 ISMB/supplementary_ISMB.tex
@@
 	\includegraphics[scale=0.4]{Figures/RNAinverse_data_100.png}
-	\caption{Entropy and \texttt{C+G} content.}
+	\caption{Entropy and \texttt{C+G} content for structures solved by
+	the 3 methods.}
 	\label{fig:rnainverse}
@@
 
+%The same analysis but only for structures solved by \texttt{Incarnation}
+%and \texttt{Incarnation} followed by \RNAinverse, is presented in 
+%Fig.~\ref{fig:inc_rnainv}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics{}
+%	\caption{Entropy and \texttt{C+G} content for structures solved by
+%	the \texttt{Incarnation} and \texttt{Incarnation}$+$\RNAinverse.}
+%	\label{fig:inc_rnainv}
+%\end{figure}
 \subsection{Impact of local search on the \gc~content of \ourprog output}

### 2013-01-08 ISMB/introduction_ISMB.tex
completed title/abstract/intro
typos

@@
 Thereby, during the last 6 years many improvements and variants of \RNAinverse have been proposed.
-Conceptually, almost all existing programs follow the same approach.
-First a seed sequence is selected, then a local search strategy is used to mutate the seed and find in its vicinity a sequence with desired folding properties.
-Using this strategy, \INFORNA \cite{Busch:2006uq}, \RNASSD \cite{Aguirre-Hernandez:2007kx} and \NUPACK \cite{Zadeh:2011fk} significantly improved the performance of the RNA secondary structure design algorithms.
+Conceptually, almost all of existing algorithms follow the same approach.
+First a seed sequence is selected, then a local search strategy is used to mutate the seed and find, in its vicinity, a sequence with desired folding properties.
+Using this strategy, \INFORNA \cite{Busch:2006uq}, \RNASSD \cite{Aguirre-Hernandez:2007kx} and \NUPACK \cite{Zadeh:2011fk} significantly improved the performance of RNA secondary structure design algorithms.
 More recent research studies aimed to include more constraints in the selection criteria.\RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \cite{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \cite{Lyngso:2012vn}.
@@
 We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \cite{Levin:2012kx}.
-Instead of a local search approach, we propose to apply a global sampling strategy of the mutational landscape using our \RNAmutants algorithm \cite{Waldispuhl2008}.
-This methodology offers promising performance but  suffers from a prohibitive running time complexity.
-Following our work, Garcia-Martin and co-workers proposed with \RNAiFOLD an alternate methodology using constraint programming techniques to prune the mutational landscape.
-The latter also suffers from prohibitive running times, but it is worth noting that it proposes a seed-less approach to the RNA secondary structure design problem.
+Instead of a local search approach, we proposed a global sampling strategy of the mutational landscape based on the \RNAmutants algorithm~\cite{Waldispuhl2008}.
+This methodology offered promising performances, but suffered from  prohibitive runtime and memory consumption.
+Following our work, Garcia-Martin and co-workers proposed \RNAiFOLD, an alternate methodology that uses constraint programming techniques to prune the mutational landscape.
+While also suffering from prohibitive running times, it is worth noting that this latter algorithm also proposes a seed-less approach to the RNA secondary structure design problem.
 
-In this paper we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}.
+In this paper, we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}.
 \ourprog addresses previous limitations of \RNAensign and offers new functionalities.
-First, while our previous program had a running time complexity of $\mathcal{O}(n^4)$, \ourprog has now a linear-time and space complexity that allows him to demonstrate similar speeds to any local search algorithm.
+First, while our previous program had a running time complexity of $\mathcal{O}(n^5)$, \ourprog now runs in linear-time and space complexity, allowing it to demonstrate similar speeds as any local search algorithm.
 Next, \ourprog is \textit{seed-less}.
-Unlike \RNAensign, it does not require a seed sequence to start its search.
-Finally, \ourprog implements a novel algorithm using weighted sampling techniques \cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \gc~content of the solution.
-This functionality is essential because wild sequences within living organisms often present medium or low \gc~content.
-Presumably to offer better transcription rates and/or structural plasticity.
-Previous programs do not allow to control this parameter and tend to output sequences with high \gc~contents.
+Unlike \RNAensign, it does not require a seed sequence to initiate its search.
+Finally, \ourprog implements a novel algorithm using weighted sampling techniques~\cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution.
+This functionality is essential because wild sequences within living organisms often present medium or low \GCContent, presumably to offer better transcription rates and/or structural plasticity.
+Previous programs do not allow to control this parameter and tend to output sequences with high \GCContent{}s.
 
-We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \textsf{RNA STRAND} database \cite{andronescu2008rna}.
+We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \RNASTRAND database \cite{andronescu2008rna}.
 To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse.
-Remarkably, our glocal methodology overcome both local and global approaches.
+Remarkably, our glocal methodology overcomes both local and global approaches.

### 2013-01-08 ISMB/main_ISMB.tex
typo & supp data
typos

@@
 \newcommand{\contextfold}{\texttt{contextfold}\xspace}
+
+\newcommand{\RNASTRAND}{{\sf RNA STRAND}\xspace}
+
 %\newcommand{\citep}{\cite}
@@
 
-\title{A glocal approach to designing RNA sequences with target secondary structure and nucleotide distribution}
+\title{A weighted sampling algorithm to designing RNA sequences with target secondary structure and nucleotide distribution}
 \author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
-\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, ƒcole Polytechnique, Palaiseau, France.\\ \small $^*$Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}
+\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.\\ \small $^*$Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}
 

### 2013-01-08 ISMB/methods_ISMB.tex
typos

--- 2013-01-08 ISMB/methods_ISMB.tex
@@
 stacked canonical base-pairs.
-
-We show how a modification of the dynamic programming scheme introduced in the previously contributed\RNAmutants allows for the sampling of good and diverse design candidates, in linear time and space complexities.
+We show how a modification of the dynamic programming scheme used in the \RNAmutants allows for the sampling of good and diverse design candidates, in linear time and space complexities.
 
@@
 \end{align*}
-where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
+where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble pairs (i.e.
 not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$).
@@
 For the sake of simplicity, and without loss of generality, let us assume that there exists a virtual base-pair $(-1,n)$ in $\Target$, so that any position in $[0,n-1]$ has a covering base-pair.
-Then one easily shows by induction that, starting from the complete interval $[0,n-1]$, the recursive calls only evaluate $\mathcal{Z}$ over intervals $[i,j]$ such that $j=j'-1$, $\CNBP{i}=[i',j']$.
-
+Then, it is easily shown by induction that, starting from the complete interval $[0,n-1]$, the recursive calls only evaluate $\mathcal{Z}$ over intervals $[i,j]$ such that $j=j'-1$, $\CNBP{i}=[i',j']$.
 \CFramed{This property clearly holds initially for the complete interval $[0,n-1]$, owing to the virtual base-pair $(-1,n)$. Furthermore, if it holds for a given $(i,j)$, $\CNBP{i}=[i',j+1]$, then:
@@
 
-Since, during each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position, and that the number of executions of the loops are bounded by a constant, then the complexity of the algorithm grows linearly on the targeted length.
+Since, on each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position, and that the number of executions of the loops are bounded by a constant, then the complexity of the algorithm grows linearly on the targeted length.
 
@@
 Such a value can be determined using any numerical recipe, and a simple bisection procedure, further described in earlier work~\cite{Waldispuhl2011}, is used in our implementation.
-Furthermore, the concentration of the distribution, as asserted by the limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp.
+Furthermore, the concentration of the distribution, asserted by its limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp.
 $\Theta(1)$, i.e.
@@
 
-\subsection{Redesigning unpaired regions: A local/global hybrid approach}
-Due to the simplification in our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest similarities of refolded design candidates with the targeted structure.
-To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions, specifying a constraint mask to prevent stacking base-pairs from being modified.
+\subsection{Postprocessing unpaired regions: A local/global hybrid approach}
+\label{subsec:glocal_method}
+Due to our simplified our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest similarities of refolded design candidates with the targeted structure.
+To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions, specifying a constraint mask to prevent stacking base-pairs from being modified and, if necessary, reetablishing stacking-pair content afterward.
 {\em todo: discuss evolution of \GCContent}

### 2013-01-08 ISMB/results_ISMB.tex
add section RNAinverse
typos

@@
 \label{sec:implementation}
-Our software, \texttt{Incarnation}, was implemented in Python2.7. We used
-\emph{RNAinverse} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
+Our software, \texttt{Incarnation}, was implemented in {\tt Python~2.7}. We used
+\RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
 All time benchmarks were run on a one AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.
@@
 We present the average times spent by \texttt{Incarnation} and \RNAinverse to generate one sequence
-with the required \texttt{C+G} content in Fig.~\ref{fig:time}. As expected, the time grows linearly
+with the required \texttt{C+G} content in Figure~\ref{fig:time}. As expected, the time grows linearly
 in function of the length of the structures for \texttt{Incarnation}.  A highly interesting feature
@@
 To evaluate the quality of our method, we used secondary
-structures from the \textit{RNA STRAND} database~\cite{andronescu2008rna}.
+structures from the \RNASTRAND database~\cite{andronescu2008rna}.
 Those are known secondary structure from a variety of organisms.
@@
  in function of their length, stacks density and proportion of free 
- nucleotides in loops. We present in Fig.~\ref{fig:bins} the distribution
+ nucleotides in loops. We present in Figure~\ref{fig:bins} the distribution
  of the structures in the different bins.
@@
  To benchmark our method, we first started by sampling a minimum
- $100$ sequence per structures. After correcting the free nucleotides with
+ of $100$ sequence per structure. After correcting the free nucleotides with
  \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
@@
 We show the results when using only \texttt{Incarnation} in 
- Fig.~\ref{fig:mfe_struct_solved_noinverse}. The high rate of failure
-is due to the fact that no selection criterions are applied to
-free nucleotides. A local strategy is thus needed.
+ Figure~\ref{fig:mfe_struct_solved_noinverse}. As mentioned earlier, the high rate of failure
+is due to the fact that no selection criterion is applied to
+unpaired nucleotides. A local strategy is thus needed.
 After processing \texttt{Incarnation} sequences with \RNAinverse to 
-locally optimize the free nucleotides, we obtain the results 
-in Fig.~\ref{fig:mfe_struct_solved}. We observe
+locally optimize the free nucleotides, we obtained the results 
+shown in Figure~\ref{fig:mfe_struct_solved}. We observed
 that length is not a good predictor for the quality of the results. Instead,
@@
 To evaluate the global quality of \texttt{Incarnation} sequences, we show
-in Fig.~\ref{fig:ss_sens} the ratio of well predicted base pairs in the
+in Figure~\ref{fig:ss_sens} the ratio of well predicted base pairs in the
 MFE structure of our sampled sequences. We can observe that, in all cases,
@@
 
-In this section, we are only interested  in the samples folding into the 
-targetMFE.
+In this section, we further analyze the generated sequences that fold into the 
+target structure.
 
 A desirable feature in sequence design, is to produce samples with a high
-diversity. We show in Fig.~\ref{fig:nb_sols_entropy} the number of correct
+diversity. We show in Figure~\ref{fig:nb_sols_entropy} the number of correct
 solutions as their average entropy and base pair entropy, since 
@@
 The sequence should be in its target conformation long enough to
-perform the desired action. Presented in Fig.~\ref{fig:freq} we see that
+perform the desired action. Presented in Figure~\ref{fig:freq} we see that
 there is a slow decline of the frequency with the increase in size. Yet,
@@
 \end{figure}
+
+\subsection{Global sampling, Local search and Glocal approach}
+
+\label{fig:rnainverse}

### 2013-01-08 ISMB/results_ISMB.tex
add section of comparison with local

@@
 
-\subsection{Global sampling, Local search and Glocal approach}
+\subsection{Global sampling vs Local search vs Glocal approach}
+Here, we are interested in estimating the impact of the design methodology on the performances.
+More precisely, we aim to what are the benefits of a global sampling approach (\ourprog), versus a glocal procedure (\ourprog + \RNAinverse) or the classical local search methodology (\RNAinverse).
+Thus, we compare our method to a pure local search procedure.
+In particular, here we compare to \RNAinverse as it is the program we used to do the local optimization (see Section \ref{subsec:glocal_method}).
+Our results (See Fig.
+\ref{fig:rnainverse} in Supplementary data) show that \ourprog overcome other strategies.
+In particular, the base pair entropy as well as diversity of sequences are favorable to \ourprog.
+This advantage is even more striking for low \gc~content regimes.
+
+To complete this benchmark, we added the results obtained with the glocal procedure (\ourprog + \RNAinverse).
+We note that the results are slightly lower than those of \ourprog alone.
+However, we must stress that the \ourprog data consider only the sample of \ourprog already satisfying the MFE criterion.
+In fact \RNAinverse enable to ``correct'' a lot of sample sampled by \ourprog but without the good folding properties.
+Hence, our conclusion here is that the glocal approach \ourprog + \RNAinverse enable us to conserve the same level of performance but to drastically improve the success rate of our methodology.
 

### 2013-01-08 ISMB/methods_ISMB.tex
fix typos

@@
 
-We introduce a probabilistic model for the design of RNA sequences with a specific $\Gb+\Cb$ content and folding into a predefined secondary structure.
+We introduce a probabilistic model for the design of RNA sequences with a specific \GCContent and folding into a predefined secondary structure.
 For the sake of simplicity, we choose to base this proof-of-concept implementation on a simplified free-energy function $\ES(\cdot)$, which only considers the contributions of 
@@
 \caption{Three main cases in the stochastic backtrack procedure over an interval $[i,j]$: Either position $i$ is left unpaired (top), a base-pair $(i,j)$ is formed, stacking onto $(i+1,j-1)$ (middle), or $i$ is paired somewhere else, defining two regions on which subsequent recursive calls are needed (bottom).
-Positions indicated in red are assigned during at this stage of the backtrack.\label{fig:stochastic}}
+Positions indicated in red are assigned during this stage of the backtrack.\label{fig:stochastic}}
 \end{figure}

### 2013-01-08 ISMB/results_ISMB.tex
fix typos

--- 2013-01-08 ISMB/results_ISMB.tex
@@
 We present the average times spent by \texttt{Incarnation} and \RNAinverse to generate one sequence
-with the required \texttt{C+G} content in Figure~\ref{fig:time}. As expected, the time grows linearly
+with the required \GCContent in Figure~\ref{fig:time}. As expected, the time grows linearly
 in function of the length of the structures for \texttt{Incarnation}.  A highly interesting feature
@@
 	\caption{Average time in seconds to generate one sequence. We explicitly show 
-	the time spent by \texttt{Incarnation} (full line) and \texttt{RNAinverse} (dashed) for various \texttt{C+G} contents. The first plot is in function
+	the time spent by \texttt{Incarnation} (full line) and \texttt{RNAinverse} (dashed) for various \GCContent. The first plot is in function
 	of the length of the structures. The second is in function of the stack
@@
 good measure of its design hardness. 
-On top of that, with a \texttt{C+G} content of $50\%$ and more, at least
-one solutions was found for almost all structures. The difficulty of 
+ The difficulty of 
 designing sequences for targets with a high number of free nucleotides 
- in loops appears clearly in the last subfigure. Even with a high \texttt{C+G} content a our method was not able to generate solutions for a few
- structures.
+ in loops appears clearly in the last column of Figure~\ref{fig:mfe_struct_solved}.
+Even with a \GCContent of $50\%$ and more, when at least
+one solutions was found for almost all structures, most of the samples 
+do not fold into the target. A few structures remained unsolved under
+all \GCContent.
 
@@
 MFE structure of our sampled sequences. We can observe that, in all cases,
-the hardest sequence to design are those with an extremely low \texttt{C+G}
-content. As anticipated, those are the sequences with the weakest bonds.
-We notice that the most accurate sequences yield a \texttt{C+G} content
+the hardest sequence to design are those with an extremely low \GCContent. As anticipated, those are the sequences with the weakest bonds.
+We notice that the most accurate sequences yield a \GCContent
 of $70\pm 10\%$.
@@
 stacked base pairs. The most constraining factor for generating valid
- solutions seems to be  low \texttt{C+G} content and the percentage of free nucleotides. Our method is able to generate
+ solutions seems to be  low \GCContent and the percentage of free nucleotides. Our method is able to generate
 
@@
 there is a slow decline of the frequency with the increase in size. Yet,
-for an average \texttt{C+G} content, the frequency stays over $10\%$ even
+for an average \GCContent, the frequency stays over $10\%$ even
 at size of $100$ nucleotides.

### 2013-01-08 ISMB/conclusion_ISMB.tex
fixed ISMB headers
conclusion v1

@@
-%!TEX root = main_RECOMB.tex
+%!TEX root = main_ISMB.tex
 \section{Conclusion}
+
 \label{sec:conclusion}
+
+In this article, we described a novel algorithm, \ourprog, for the RNA design problem, i.e. the design of an RNA sequence adopting a predefined secondary structure as minimal free-energy fold.
+Implementing a global sampling approach, it optimizes affinity towards the target secondary structure, while granting the user full control over the \GCContent of the resulting sequences.
+This extended control does not necessarily induce additional computational demands, and we showed the linear complexity of both the preprocessing stage and the generation of candidate sequences for the design, allowing for the design of larger and more complex secondary structures in a matter of minutes on a single processor.
+We evaluated the method on a benchmark composed of target secondary structures extracted from the \RNASTRAND database.
+We observed good overall success rate, with the notable exception of very low targeted \GCContent ($10\%$), and a good entropy within designed candidates.
+Finally, we implemented an hybrid approach, using the \RNAinverse software as a post-processing step for unpaired regions.
+This approach greatly increased the success rate of the method, allowing for the design of highly diverse candidates for almost all of the structures in our benchmark, while largely preserving the targeted \GCContent.

### 2013-01-08 ISMB/methods_ISMB.tex
conclusion v1

--- 2013-01-08 ISMB/methods_ISMB.tex
@@
 \label{subsec:glocal_method}
-Due to our simplified our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest similarities of refolded design candidates with the targeted structure.
-To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions, specifying a constraint mask to prevent stacking base-pairs from being modified and, if necessary, reetablishing stacking-pair content afterward.
-{\em todo: discuss evolution of \GCContent}
+Due to our simplified our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest probabilities for refolded design candidates to match the targeted structure.
+To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions.
+We specified a constraint mask to prevent stacking base-pairs from being modified and, whenever necessary, reetablished the content content a posteriori.
 

### 2013-01-08 ISMB/results_ISMB.tex
conclusion v1

--- 2013-01-08 ISMB/results_ISMB.tex
@@
 \label{sec:implementation}
-Our software, \texttt{Incarnation}, was implemented in {\tt Python~2.7}. We used
+Our software, \ourprog, was implemented in {\tt Python~2.7}. We used
 \RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
@@
 
-We present the average times spent by \texttt{Incarnation} and \RNAinverse to generate one sequence
+We present the average runtimes observed for \texttt{Incarnation} and \RNAinverse to generate one sequence
 with the required \texttt{C+G} content in Figure~\ref{fig:time}. As expected, the time grows linearly
@@
 	\caption{Average time in seconds to generate one sequence. We explicitly show 
-	the time spent by \texttt{Incarnation} (full line) and \texttt{RNAinverse} (dashed) for various \texttt{C+G} contents. The first plot is in function
+	the time spent by \ourprog (full line) and \RNAinverse (dashed) for various \texttt{C+G} contents. The first plot is in function
 	of the length of the structures. The second is in function of the stack
@@
 for which we can design at least one sequence with the desired MFE.
-We show the results when using only \texttt{Incarnation} in 
+We show the results when using only \ourprog in 
  Figure~\ref{fig:mfe_struct_solved_noinverse}. As mentioned earlier, the high rate of failure
@@
 unpaired nucleotides. A local strategy is thus needed.
-After processing \texttt{Incarnation} sequences with \RNAinverse to 
+After processing \ourprog sequences with \RNAinverse to 
 locally optimize the free nucleotides, we obtained the results 
 shown in Figure~\ref{fig:mfe_struct_solved}. We observed
-that length is not a good predictor for the quality of the results. Instead,
-a high number of free nucleotides in the structure seems to be a 
-good measure of its design hardness. 
+that length is, in general, not a good predictor for the quality of the results. Instead,
+a high number of free nucleotides in the structure seems to make the design difficult. 
 On top of that, with a \texttt{C+G} content of $50\%$ and more, at least
@@
 	\includegraphics[width=\textwidth]{Figures/mfe_struct_solved}
-	\caption{Quality of \texttt{Incarnation+RNAinverse} results. The first row shows the percentage
+	\caption{Quality of \ourprog+\RNAinverse results. The first row shows the percentage
 	of sampled sequences folding into the target. The second shows the 	
@@
 
-To evaluate the global quality of \texttt{Incarnation} sequences, we show
+To evaluate the global quality of \ourprog sequences, we show
 in Figure~\ref{fig:ss_sens} the ratio of well predicted base pairs in the
@@
 	\includegraphics[scale=0.45]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
-	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }
+	\caption{Structural sensitivity (i.e. $\#$well predicted base/$\#$base pairs in target) of the sampled sequences MFE. }
 	\label{fig:ss_sens}	
@@
 
-A desirable feature in sequence design, is to produce samples with a high
-diversity. We show in Figure~\ref{fig:nb_sols_entropy} the number of correct
+A desirable feature in sequence design, is to produce  highly
+diverse samples. Figure~\ref{fig:nb_sols_entropy} shows the number of correct
 solutions as their average entropy and base pair entropy, since 
-\texttt{Incarnation}  only influences the distribution of nucleotides inside 
+\ourprog  only influences the distribution of nucleotides inside 
 stacked base pairs. The most constraining factor for generating valid
@@
 
-\subsection{Global sampling, Local search and Glocal approach}
+%\subsection{Global sampling, Local search and Glocal approach}
 

### 2013-01-08 ISMB/introduction_ISMB.tex

@@
 Using this strategy, \INFORNA \cite{Busch:2006uq}, \RNASSD \cite{Aguirre-Hernandez:2007kx} and \NUPACK \cite{Zadeh:2011fk} significantly improved the performance of RNA secondary structure design algorithms.
-More recent research studies aimed to include more constraints in the selection criteria.\RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \cite{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \cite{Lyngso:2012vn}.
+More recent research studies aimed to include more constraints in the selection criteria.
+\RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \cite{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \cite{Lyngso:2012vn}.
 

### 2013-01-08 ISMB/supplementary_ISMB.tex
typos;

--- 2013-01-08 ISMB/supplementary_ISMB.tex
@@
 \texttt{Incarnation} followed by \RNAinverse,
-we present for every different concentration of \texttt{C+G}
+we present for every different concentration of \GCContent
 the average sample entropy and the average base pairs entropy.
@@
 \begin{figure}[ht!]
-	\centering
+	%\centering
+	\hspace{-5em}
 	\includegraphics[scale=0.4]{Figures/RNAinverse_data_100.png}
-	\caption{Entropy and \texttt{C+G} content for structures solved by
+	\caption{Entropy and \GCContent  for structures solved by
 	the 3 methods.}
@@
 %\end{figure}
+\subsection{Impact of local search on the \GCContent of \ourprog output}
+
+These data demonstrate that the local search heuristic used to design nucleotides in loop regions has a very limited impact on the \GCContent.
+For each class of \GCContent, we show in Table~\ref{table:impact_on_gc}, the observed \GCContent of the sequence sampled by \ourprog, and then we report the observed \GCContent after the \RNAinverse step (as defined in Section \ref{subsec:glocal_method}).
+Our results show that the \GCContent is relatively well conserved.
+
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+target \gc & \ourprog & \ourprog + \RNAinverse \\
+\hline
+10 & 0.15 & 0.21\\
+30 & 0.30 & 0.33\\
+50 & 0.48 & 0.49\\
+70 & 0.71 & 0.69\\
+90 & 0.83 & 0.78\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Observed \GCContent of solutions returned by \ourprog (2nd column) and after the application of the local search heuristic (3rd column).}
+\label{table:impact_on_gc}
+\end{table}

### 2013-01-08 ISMB/main_ISMB.tex
notation

@@
 
-\newcommand{\ourprog}{\texttt{Incarnation}\xspace}
+\newcommand{\ourprog}{\texttt{IncaRNAtion}\xspace}
 
@@
 Most of the current software use similar local search strategies (i.e.
-a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \gc~content in their sequences.
+a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences.
 However, the latter is an important criteria for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.\\
@@
 running time comparable or better than local search methods), seed-less (we remove the bias of the seed in local search heuristics), and successfully generates high-quality sequences (i.e.
-thermodynamically stable) for any \gc~content.
+thermodynamically stable) for any \GCContent.
 To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies.

### 2013-01-08 ISMB/results_ISMB.tex
typos;
notation

@@
 
-We present the average times spent by \texttt{Incarnation} and \RNAinverse to generate one sequence
+We present the average times spent by \ourprog and \RNAinverse to generate one sequence
 with the required \GCContent in Figure~\ref{fig:time}. As expected, the time grows linearly
-in function of the length of the structures for \texttt{Incarnation}.  A highly interesting feature
+in function of the length of the structures for \ourprog.  A highly interesting feature
 is the huge decrease in time of computation when there is $15\%$ free 
@@
 	\caption{Average time in seconds to generate one sequence. We explicitly show 
-	the time spent by \texttt{Incarnation} (full line) and \texttt{RNAinverse} (dashed) for various \GCContent. The first plot is in function
+	the time spent by \ourprog (full line) and \texttt{RNAinverse} (dashed) for various \GCContent. The first plot is in function
 	of the length of the structures. The second is in function of the stack
@@
 unpaired nucleotides. A local strategy is thus needed.
-After processing \texttt{Incarnation} sequences with \RNAinverse to 
+After processing \ourprog sequences with \RNAinverse to 
 locally optimize the free nucleotides, we obtained the results 
@@
 	\includegraphics[width=\textwidth]{Figures/mfe_struct_solve_nornainverse.png}
-	\caption{Quality of \texttt{Incarnation} results. The first row shows the percentage
+	\caption{Quality of \ourprog results. The first row shows the percentage
 	of sampled sequences folding into the target. The second shows the 	
@@
 solutions as their average entropy and base pair entropy, since 
-\texttt{Incarnation}  only influences the distribution of nucleotides inside 
+\ourprog  only influences the distribution of nucleotides inside 
 stacked base pairs. The most constraining factor for generating valid
@@
 In particular, the base pair entropy as well as diversity of sequences are favorable to \ourprog.
-This advantage is even more striking for low \gc~content regimes.
+This advantage is even more striking for low \GCContent regimes.
 

### 2013-01-08 ISMB/supplementary_ISMB.tex
tab rnainv

@@
 per structure with \RNAinverse. They yield for most parameters
-a high entropy. For all structures that have been solved 
+a high entropy.
+Table.~\ref{tab:nb_rnainv} contains the number of sample with a given \GCContent generated by \RNAinverse. The vast majority resides around $50\%-70\%$, showing one of its major deficiencies. 
+ For all structures that have been solved 
 by the three methods, only \RNAinverse, only \texttt{Incarnation} and
@@
 we present for every different concentration of \GCContent
-the average sample entropy and the average base pairs entropy.
+the average sample entropy and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
+
+\begin{table}[h!]
+	\begin{center}
+		\begin{tabular}{|c|c|}
+		\hline
+		target \GCContent & $\#$\RNAinverse samples\\
+		10\% & 11\\
+			30\%& 437\\
+		50\%& 3362\\
+	70\% &1155\\
+	90\%& 35\\\hline
+		\end{tabular}
+	\end{center}
+	\label{tab:nb_rnainv}
+\end{table}
 
@@
 \hline
-target \gc & \ourprog & \ourprog + \RNAinverse \\
+target \GCContent & \ourprog & \ourprog + \RNAinverse \\
 \hline
-10 & 0.15 & 0.21\\
-30 & 0.30 & 0.33\\
-50 & 0.48 & 0.49\\
-70 & 0.71 & 0.69\\
-90 & 0.83 & 0.78\\
+10\% & 0.15 & 0.21\\
+30\% & 0.30 & 0.33\\
+50\% & 0.48 & 0.49\\
+70\% & 0.71 & 0.69\\
+90\% & 0.83 & 0.78\\
 \hline

### 2013-01-08 ISMB/conclusion_ISMB.tex
conclusion v2

@@
 Implementing a global sampling approach, it optimizes affinity towards the target secondary structure, while granting the user full control over the \GCContent of the resulting sequences.
-This extended control does not necessarily induce additional computational demands, and we showed the linear complexity of both the preprocessing stage and the generation of candidate sequences for the design, allowing for the design of larger and more complex secondary structures in a matter of minutes on a single processor.
+This extended control does not necessarily induce additional computational demands, and we showed the linear complexity of both the preprocessing stage and the generation of candidate sequences for the design, allowing for the design of larger and more complex secondary structures in a matter of minutes on a single processor (e.g.
+28 minutes were required to sample candidate sequences for a 1500nts-long 16s rRNA).
 We evaluated the method on a benchmark composed of target secondary structures extracted from the \RNASTRAND database.
-We observed good overall success rate, with the notable exception of very low targeted \GCContent ($10\%$), and a good entropy within designed candidates.
+We observed good overall success rate, with the notable exception of very low targeted \GCContent ($10\%$), and a good to excellent entropy within designed candidates.
 Finally, we implemented an hybrid approach, using the \RNAinverse software as a post-processing step for unpaired regions.
 This approach greatly increased the success rate of the method, allowing for the design of highly diverse candidates for almost all of the structures in our benchmark, while largely preserving the targeted \GCContent.
+
+In the future, we would like to complement this study by further investigating the potential of hybrid local/global -- or {\em glocal} -- approaches.
+A global sampling approach would capture the positive aspects of design, optimizing affinity towards a given structure while allowing the specification of expressive systems of constraints.
+Designed sequences would serve as a seed for a restricted local approach which, by breaking unwanted symmetries, would perform the negative part of the design, 
+while ideally maintaining obedience to the constraints.
+Another perspective of this work is the incorporation of the full Turner energy model, which should in principle yield better designs for unpaired regions.

### 2013-01-08 ISMB/main_ISMB.tex
conclusion v2

@@
 
-\title{A weighted sampling algorithm to designing RNA sequences with target secondary structure and nucleotide distribution}
+\title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution}
 \author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
@@
 \textbf{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies.
-Most of the current software use similar local search strategies (i.e.
-a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences.
+Most existing algorithms use local search strategies (i.e.
+a random seed is progressively adapted to acquire the desired folding properties) and, more importantly, do not allow the user to control explicitly the nucleotide distribution such as the \gc~content in their sequences.
 However, the latter is an important criteria for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.\\
@@
 running time comparable or better than local search methods), seed-less (we remove the bias of the seed in local search heuristics), and successfully generates high-quality sequences (i.e.
-thermodynamically stable) for any \GCContent.
+thermodynamically stable) for any \gc~content.
 To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies.
-Remarkably, our glocal methodology overcomes both local and global approaches.\\
+Remarkably, our glocal methodology outperforms both local and global approaches.\\
 \noindent

### 2013-01-08 ISMB/results_ISMB.tex
conclusion v2

--- 2013-01-08 ISMB/results_ISMB.tex
@@
 	density (i.e. $2\cdot\#stacks/length$) and the latter in function of 
-	the proportion of free nucleotides (i.e. not base paired) inside loops.}
+	the proportion of free nucleotides (i.e. unpaired bases) within loops.}
 	\label{fig:time}	
@@
 We select the same subset of $50$ structures as in~\cite{Levin:2012kx}, 
-which have lengths between $20$ and $100$ nucleotides.
+whose length ranges between $20$ and $100$ nucleotides.
 
- To ease the visualization of the data, we clustered together the strucures
- in function of their length, stacks density and proportion of free 
- nucleotides in loops. We present in Figure~\ref{fig:bins} the distribution
- of the structures in the different bins.
+ To ease the visualization of the results, we clustered together strucures
+ having similar length, stacks density and proportion of free-nucleotides in loops. 
+  We present in Figure~\ref{fig:bins} the distribution of structures in each category.
 
@@
 	\includegraphics[width=\textwidth]{Figures/bins_distribution.png}
-	\caption{Number of secondary structures per bins.}
+	\caption{Number of secondary structures per bin, according to our three clustering criteria.}
 	\label{fig:bins}
@@
 \subsection{Design}
- To benchmark our method, we first started by sampling a minimum
- of $100$ sequence per structure. After correcting the free nucleotides with
+ To benchmark our method, we first started by sampling about $100$ sequences per structure. After correcting the free nucleotides with
  \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
@@
 shown in Figure~\ref{fig:mfe_struct_solved}. We observed
-that length is not a good predictor for the quality of the results. Instead,
-a high number of free nucleotides in the structure seems to be a 
+that length is, in general, not a good predictor for the hardness of designing a structure. 
+Instead, a high number of free nucleotides in the structure seems to be a 
 good measure of its design hardness. 
@@
 A desirable feature in sequence design, is to produce samples with a high
-diversity. We show in Figure~\ref{fig:nb_sols_entropy} the number of correct
+diversity. Figure~\ref{fig:nb_sols_entropy} shows the number of correct
 solutions as their average entropy and base pair entropy, since 
@@
 Here, we are interested in estimating the impact of the design methodology on the performances.
-More precisely, we aim to what are the benefits of a global sampling approach (\ourprog), versus a glocal procedure (\ourprog + \RNAinverse) or the classical local search methodology (\RNAinverse).
-Thus, we compare our method to a pure local search procedure.
-In particular, here we compare to \RNAinverse as it is the program we used to do the local optimization (see Section \ref{subsec:glocal_method}).
+More precisely, we aim to determine the benefits of a global sampling approach (\ourprog), compared to both a glocal procedure (\ourprog + \RNAinverse) and a classical local search methodology (\RNAinverse).
+Thus, we compare our method to a pure local search procedure (see Section \ref{subsec:glocal_method}).
 Our results (See Fig.
-\ref{fig:rnainverse} in Supplementary data) show that \ourprog overcome other strategies.
+\ref{fig:rnainverse} in Supplementary data) show that \ourprog outperforms other strategies.
 In particular, the base pair entropy as well as diversity of sequences are favorable to \ourprog.
-This advantage is even more striking for low \gc~content regimes.
+This advantage is even more striking for low targeted \GCContent{}s regimes.
 

### 2013-01-08 ISMB/supplementary_ISMB.tex
last

@@
 we present for every different concentration of \GCContent
-the average sample entropy and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
+the average sample entropy. and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
 
@@
 \begin{figure}[ht!]
-	%\centering
-	\hspace{-5em}
+	\centering
+	%\hspace{-5em}
 	\includegraphics[scale=0.4]{Figures/RNAinverse_data_100.png}

### 2013-01-08 ISMB/results_ISMB.tex
f

--- 2013-01-08 ISMB/results_ISMB.tex
@@
 stacked base pairs. The most constraining factor for generating valid
- solutions seems to be  low \GCContent and the percentage of free nucleotides. Our method is able to generate
+ solutions seems to be  low \GCContent and the percentage of free nucleotides.
+
+%Our method is able to generate
 
@@
 	\includegraphics[width=\textwidth]{Figures/freq.png}
-	\caption{Average frequency of the MFE.}
+	\caption{Average Boltzmann probability of the MFE.}
 	\label{fig:freq}
@@
 \subsection{Global sampling vs Local search vs Glocal approach}
+
 Here, we are interested in estimating the impact of the design methodology on the performances.
-More precisely, we aim to what are the benefits of a global sampling approach (\ourprog), versus a glocal procedure (\ourprog + \RNAinverse) or the classical local search methodology (\RNAinverse).
-Thus, we compare our method to a pure local search procedure.
-In particular, here we compare to \RNAinverse as it is the program we used to do the local optimization (see Section \ref{subsec:glocal_method}).
+More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNAinverse).
+
 Our results (See Fig.
-\ref{fig:rnainverse} in Supplementary data) show that \ourprog overcome other strategies.
+\ref{fig:rnainverse} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone.
+Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces more entropic sequences than \ourprog for the central \GCContent regime ($30\%-70\%$).
+
+It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Table~\ref{tab:nb_rnainv} in Supplementary Material), thereby requiring a large number of runs before producing a suitable \GCContent sequence.
+Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent.
+
+% \ourprog outperforms other strategies for non-centered .
 In particular, the base pair entropy as well as diversity of sequences are favorable to \ourprog.
-This advantage is even more striking for low \GCContent regimes.
-
-To complete this benchmark, we added the results obtained with the glocal procedure (\ourprog + \RNAinverse).
+This advantage is even more striking for low targeted \GCContent{}s regimes.
+%To complete this benchmark, we added the results obtained with the glocal procedure (\ourprog + \RNAinverse).
 We note that the results are slightly lower than those of \ourprog alone.
@@
 In fact \RNAinverse enable to ``correct'' a lot of sample sampled by \ourprog but without the good folding properties.
-Hence, our conclusion here is that the glocal approach \ourprog + \RNAinverse enable us to conserve the same level of performance but to drastically improve the success rate of our methodology.
+Hence, our conclusion here is that the glocal approach enable us to conserve the same level of performance but to drastically improve the success rate of our methodology.
 

### 2013-01-08 ISMB/main_ISMB.tex
last

--- 2013-01-08 ISMB/main_ISMB.tex
@@
 To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies.
-Remarkably, our glocal methodology overcomes both local and global approaches.\\
+Remarkably, our glocal methodology overcomes both local and global approaches  for sampling sequences with a specific GC content and target structure.\\
 \noindent

### 2013-01-09 Recomb/introduction_RECOMB.tex
recomb typos

@@
 Nonetheless, these high-throughput techniques typically have high error rates that make their applications to metagenomics (a.k.a. environmental
-genomics) studies challenging. For instance, pyrosequencing as implemented by Roche's 454 produces may have a error rate raising up to 10\%.
+genomics) studies challenging. For instance, pyrosequencing as implemented by Roche's 454 produces may have an error rate raising up to 10\%.
 Because there is no cloning step, resequencing to increase accuracy is not possible and it is therefore vital to disentangle noise from true sequence
 diversity in this type of data \cite{Quince:2009uq}. Errors can be significantly reduced  when large multiple sequence alignments with close homologs
-are available, but in studies of new or not well known organisms, such information is rather sparse. In particular, it is common to is not enough  similarity to 
+are available, but in studies of new or not well known organisms, such information is rather sparse. In particular, it is common that there is not enough  similarity to 
 differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
@@
 
-In this paper, we introduce \RNApyro, a novel algorithm to that enable us to calculate precisely mutational probabilities in RNA sequences with a
+In this paper, we introduce \RNApyro, a novel algorithm that enables us to calculate precisely mutational probabilities in RNA sequences with a
 conserved consensus secondary structure. We show how our techniques can exploit the structural information embedded in physics-based energy models, 

### 2013-01-09 Recomb/methods_RECOMB.tex
recomb typos

--- 2012-10-07 Recomb/methods_RECOMB.tex
@@
  ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
- $[i,j]\subset[k,l])$. Let $R$ be the Boltzmann constant, $T$ the temperature in Kelvin and
-  the function $\delta$ such that: 
- $\forall a,a' \in B, \delta_{a,a'}:=\left\{\begin{array}{ll}
-															1 & \text{If } a\equiv a'\\
-															0 & \text{Otherwise.}
-														\end{array}\right.$
+ $[i,j]\subset[k,l])$.
 
-Finally, we denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
+Let us then remind the Hamming distance function $\delta: B^*\times B^* \to \mathbb{N}^+$, which takes two sequences $s'$ and $s''$ as input, $|s'|=|s''|$, and returns the number of differing positions.
+Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
 \begin{equation}

### 2013-01-09 Recomb/main_RECOMB.tex
format llncs

@@
-\documentclass[11pt]{article}
-\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
-\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
+\documentclass{llncs}
+%\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
+%\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
 %\geometry{landscape}                % Activate for for rotated page geometry
@@
 \usepackage{amsmath}
-\usepackage{amsthm} %pour les newtheorem
+%\usepackage{amsthm} %pour les newtheorem
 \usepackage{epstopdf}
@@
 %\usepackage{subfig} %Subfigures
-\usepackage{subcaption} %Subfigures
+%\usepackage{subcaption} %Subfigures
 
@@
 
+\newcommand{\Ab}{{\sf{A}}}
+\newcommand{\Cb}{{\sf{C}}}
+\newcommand{\Gb}{{\sf{G}}}
+\newcommand{\Ub}{{\sf{U}}}
+
 %\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
-\title{A linear inside-outside algorithm for correcting sequencing errors in structured RNA}
-\author{Vladimir Reinharz, Yann Ponty$^{*}$, J\'er\^{o}me Waldisp\"{u}hl$^*$}
+\title{A Linear Inside-Outside Algorithm for Correcting Sequencing Errors in Structured RNA}
+\author{Vladimir Reinharz\inst{1}, Yann Ponty\inst{2}$^{*}$ \and J\'er\^{o}me Waldisp\"{u}hl\inst{1}$^*$}
 \date{}
+\institute{School of Computer Science, McGill University, Montreal, Canada.
+	\and  Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.
+	 \\\email{jeromew@cs.mcgill.ca}, \email{yann.ponty@lix.polytechnique.fr}}
 

### 2013-01-09 Recomb/methods_RECOMB.tex
format llncs

@@
 
-\subsection{Probabilistic model}
+\subsection{Probabilistic Model}
 Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
@@
 
-\subsubsection{Energy contribution}
+\subsubsection{Energy Contribution}
 The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
@@
 (e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
-\subsubsection{Isostericity contribution}
+\subsubsection{Isostericity Contribution}
 The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
@@
 
-\subsection{Mutational profile of sequences}
+\subsection{Mutational Profile of Sequences}
 
@@
 
-\subsubsection{Combining Inside and Outside values into point-wise mutations probabilities}
+\subsubsection{Combining Inside and Outside Values into Point-Wise Mutations Probabilities}
 By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
@@
 
-\subsection{Complexity considerations}
+\subsection{Complexity Considerations}
 Equations~\eqref{eq:Z_rec} and~\eqref{eq:Y_rec} can be computed using dynamic programming.

### 2013-01-09 Recomb/results_RECOMB.tex
Revamped part of method section One more hunt for typos
format llncs

@@
 The software was implemented in Python2.7 using the \textit{mpmath}~\cite{mpmath} library
-for  arbitrary floating point precision. The source code is freely available at \verb+https://github.com/McGill-CSB/RNApyro+.
+for  arbitrary floating point precision. The source code is freely available at: \\\hfill\verb+https://github.com/McGill-CSB/RNApyro+.
 
@@
  efficiency and scalability, we present in Table~\ref{tab:time} typical runtimes required to compute the probabilities for  every nucleotide at every positions for a vast set of parameters.
-For those test
+For those tests,
  both the sequences and the target secondary structure were generated at random.
 
-\begin{table}[b]
+\begin{table}[t]
 \begin{center}
@@
 
-\subsection{Error correction in 5s rRNA}
+\subsection{Error Correction in 5s rRNA}
 
@@
 \begin{figure}
-\begin{subfigure}[b]{0.3\textwidth}
-\includegraphics[width=1.2\textwidth]{figures/ROC_6.pdf}
-\caption{6 mutations}
-\label{fig:ROC6mut}
-\end{subfigure}
-\hfill
-\begin{subfigure}[b]{0.3\textwidth}
-\includegraphics[width=1.2\textwidth]{figures/ROC_12.pdf}
-\caption{12 mutations}
-\label{fig:ROC12mut}
-\end{subfigure}
-\hfill
-\begin{subfigure}[b]{0.3\textwidth}
-\includegraphics[width=1.2\textwidth]{figures/ROC_24.pdf}
-\caption{24 mutations}
-\label{fig:ROC24mut}
-\end{subfigure}
+\begin{center}
+	\includegraphics[width=\textwidth]{subfigs_perform.png}
+\end{center}
+%\begin{subfigure}[b]{0.3\textwidth}
+%\includegraphics[width=1.2\textwidth]{figures/ROC_6.pdf}
+%\caption{6 mutations}
+%\label{fig:ROC6mut}
+%\end{subfigure}
+%\hfill
+%\begin{subfigure}[b]{0.3\textwidth}
+%\includegraphics[width=1.2\textwidth]{figures/ROC_12.pdf}
+%\caption{12 mutations}
+%\label{fig:ROC12mut}
+%\end{subfigure}
+%\hfill
+%\begin{subfigure}[b]{0.3\textwidth}
+%\includegraphics[width=1.2\textwidth]{figures/ROC_24.pdf}
+%\caption{24 mutations}
+%\label{fig:ROC24mut}
+%\end{subfigure}
 \caption{Performance of error-correction. Subfigures accuracy with under-estimated error rate (6 mutations), exact estimates (12 mutations) and over estimates 
@@
 mutations considered. Indeed, good AUC values are achieved with low estimates of the number of errors in the sequence (c.f. 50\% of the errors  in
-Fig.~\ref{fig:ROC6mut}) and with large  values (c.f.~200\% of the errors  in Fig.~\ref{fig:ROC24mut}). It is worth noting that scoring scheme with larger weight on
+Fig.~\ref{fig:ROCall}(a)) and with large  values (c.f.~200\% of the errors  in Fig.~\ref{fig:ROCall}(c)). It is worth noting that scoring scheme with larger weight on
 the isostericity scores (low $\alpha$ values) seem more robust to under- and over-estimate of the number of errors.

### 2013-01-09 Recomb/results_RECOMB.tex
typo

@@
 The software was implemented in Python2.7 using the \textit{mpmath}~\cite{mpmath} library
-for  arbitrary floating point precision. The source code is freely available at: \\\hfill\verb+https://github.com/McGill-CSB/RNApyro+.
+for  arbitrary floating point precision. The source code is freely available at: \\\verb+https://github.com/McGill-CSB/RNApyro+.
 
@@
  of mutations. $\alpha$ is
-set at $0.5$,  $\beta$ to $1.5$ and $|\Omega|=44$.}
+set at $0.5$,  $\beta$ to $15$ and $|\Omega|=44$.}
 \label{tab:time}
@@
 mutations. Our sequences have $119$ nucleotides, thus the number of mutations corresponds to an error-rate of 10\%. We repeated this operation 
-$10$ times. The value of $\beta$ has been set to $1.5$ (larger values gave similar results). To estimate the impact of the distribution of the weights
+$10$ times. The value of $\beta$ has been set to $15$ (larger values gave similar results). To estimate the impact of the distribution of the weights
 between the energy term and isostericity  score, we used 4 different values of $\alpha = {0, 0.5, 0.8, 1.0}$. Similarly, we also investigated the impact of 

### 2013-01-09 Recomb/introduction_RECOMB.tex
shrunk figures ... could do better but got to go

@@
 are available, but in studies of new or not well known organisms, such information is rather sparse. In particular, it is common that there is not enough  similarity to 
-differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
+differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often leading to artificially inflated diversity estimates~\cite{Kunin2010}.
 A few techniques have been developed to remedy to this problem~\cite{Quinlan2008,Medvedev2011} but they do not take

### 2013-01-09 Recomb/methods_RECOMB.tex
shrinking equations
shrunk figures ... could do better but got to go

@@
 
-\subsubsection{Energy Contribution}
+\subsubsection{Energy Contribution.}
 The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
@@
 where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e.
-not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$).
+neither $\Gb\Ub$, $\Ub\Gb$, $\Cb\Gb$, $\Gb\Cb$, $\Ab\Ub$ nor $\Ub\Ab$).
 This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).
@@
 (e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
-\subsubsection{Isostericity Contribution}
+\subsubsection{Isostericity Contribution.}
 The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
@@
 The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
-Isostericity scores range between $0$ and $9.7$, with $0$ being assigned to a perfect isostericity, and a penalty of $10$ is used for missing entries.
+Isostericity scores range between $0$ and $9.7$, $0$ corresponding to a perfect isostericity, and a penalty of $10$ is used for missing entries.
 The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
@@
 
-\subsubsection{Definitions}
+\subsubsection{Definitions.}
 Let $B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
@@
 
-\subsubsection{Inside}
-\begin{figure}
+\subsubsection{Inside computation.}
+\begin{figure}[t]
+\begin{minipage}{0.60\textwidth}
 \resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
+\end{minipage}
+\begin{minipage}{0.38\textwidth}
 \caption{Principle of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
-can be decomposed as a sequence over $[i+1,j]$ preceded by a, possibly mutated, base at $i$ 
-(Unpaired case), a sequence over $[i+1,j-1]$ surrounded by some base-pair (Stacking-pair case), 
-or as two sequences over $[i+1,k-1]$ and $[k+1,j]$, completed by some base-pair (General base-pairing case). 
+can be decomposed as a sequence preceded by a, possibly mutated, base at $i$ 
+(Unpaired case), a sequence surrounded by some base-pair (Stacking-pair case), 
+or as two sequences separated by some base-pair (General base-pairing case). 
 In each case, one has to investigate any possible ways to distribute mutations between the different sequences 
-and locally instanciated bases.}
+and locally instantiated bases.}
+\end{minipage}
 \end{figure}
@@
 
-The main recursion is composed of four terms:
+The main recursion is composed of four terms:\\
+
 \begin{equation}
-	\Z{i,j}{m}{a,b}:=\left\{
+\label{eq:Z_rec}
+\resizebox{\textwidth}{!}{%
+	$\Z{i,j}{m}{a,b}:=\left\{
   \begin{array}{ll}
@@
       0 &\text{Otherwise}
-	\end{array}\right.
-\label{eq:Z_rec}
+	\end{array}\right.$}
 \end{equation}
+\
 The cases can be broken down as follows:
@@
 
-\subsubsection{Outside}	
-\begin{figure}
+\subsubsection{Outside computation.}	
+\begin{figure}[t]
+\begin{minipage}{0.75\textwidth}
 \resizebox{\textwidth}{!}{\input{FigDPOutside.pgf}}
+\end{minipage}
+\begin{minipage}{0.24\textwidth}
 \caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
-therefore its efficient implementation requires an implementation of the inside computation.}
+therefore its efficient implementation requires a precomputation of the inside contributions.}
+\end{minipage}
 \end{figure}
@@
  all the outside energy between $[j,n-1]$ is
-equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself is as follows.
+equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself unfolds as follows.
 \begin{equation}
-	\Y{i,j}{m}{a,b} = \left\{
+\resizebox{\textwidth}{!}{%
+	$\Y{i,j}{m}{a,b} = \left\{
   \begin{array}{ll}
@@
 		 0 & \text{Otherwise}
-  \end{array}\right.
-\label{eq:Y_rec}
+  \end{array}\right.$
+}\label{eq:Y_rec}
 \end{equation}
@@
 
-\subsubsection{Combining Inside and Outside Values into Point-Wise Mutations Probabilities}
+\subsubsection{Combining Inside and Outside Values into Point-Wise Mutations Probabilities.}
 By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
@@
 Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations. 
-\begin{align}
-	\mathbb{P}(s_i = x\mid s,\Omega, S,M) &:= \frac{\mathcal{W}(i,x,s,\Omega,S,M)}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}\\ 
-\mathcal{W}(i,x,s,\Omega,S,M)&=
- \left\{
+\begin{equation}
+	\mathbb{P}(s_i = x\;|\; M) := \frac{\mathcal{W}^M_{\substack{i, [x]}}}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}
+\end{equation}
+where $\mathcal{W}{\substack{\ast\\ \ast}}$ is defined by:
+\begin{equation}
+\resizebox{\textwidth}{!}{%
+$ \mathcal{W}^M_{\substack{i, [x]}} =  \left\{
 	\begin{array}{ll}
@@
 		&\text{If }S_i=k<i
-	\end{array}\right.\label{eq:combine}
-\end{align}
+	\end{array}\right.
+$}\label{eq:combine}
+\end{equation}
 

### 2013-01-09 Recomb/results_RECOMB.tex
shrunk figures ... could do better but got to go

@@
 The software was implemented in Python2.7 using the \textit{mpmath}~\cite{mpmath} library
-for  arbitrary floating point precision. The source code is freely available at: \\\verb+https://github.com/McGill-CSB/RNApyro+.
+for  arbitrary floating point precision. The source code is freely available at:
 
-The time benchmarks were done on a MacMini 2010, 2.3GHz dual-core Intel Core i5, 8GB of RAM.
-Since applications of \RNApyro implies a need for
+{\centering \url{https://github.com/McGill-CSB/RNApyro}\\}
+
+The time benchmarks were performed on a MacMini 2010, 2.3GHz dual-core Intel Core i5, 8GB of RAM.
+Since typical use-cases of \RNApyro require
  efficiency and scalability, we present in Table~\ref{tab:time} typical runtimes required to compute the probabilities for  every nucleotide at every positions for a vast set of parameters.
@@
 \begin{table}[t]
-\begin{center}
+\begin{tabular}{cp{27em}}
 \begin{tabular}{lccc}
-$|s|$&\multicolumn{3}{c}{Number of mutations}\\\cline{2-4}
+$|s|$&\multicolumn{3}{c}{\#mutations}\\\cline{2-4}
 		 			  & 6   &  12  & 24\\\cline{2-4}
@@
 500         & 5400s&       &  21003s    \\
-\end{tabular}
-\end{center}
+\end{tabular}&
 \caption{Time to compute all probabilities. The first column indicates the length and  the column indexes indicate the number
@@
 set at $0.5$,  $\beta$ to $15$ and $|\Omega|=44$.}
+\end{tabular}
 \label{tab:time}
@@
 
-We designed our benchmark using a leave-one-out strategy. We randomly picked one sequence from our data set and performed $12$ random
-mutations. Our sequences have $119$ nucleotides, thus the number of mutations corresponds to an error-rate of 10\%. We repeated this operation 
-$10$ times. The value of $\beta$ has been set to $15$ (larger values gave similar results). To estimate the impact of the distribution of the weights
-between the energy term and isostericity  score, we used 4 different values of $\alpha = {0, 0.5, 0.8, 1.0}$. Similarly, we also investigated the impact of 
-an under- and over- estimate of the number of errors, and use values equal to 50\% (6 mutations) and 200\% (24 mutations) of the exact number of
-errors (i.e. 12).
+We designed our benchmark using a leave-one-out strategy. We randomly picked a single sequence from our data set and performed $12$ random
+mutations, corresponding to an error-rate of 10\%. We repeated this operation $10$ times. The value of $\beta$ was set to $15$ (larger values gave similar results). 
+To estimate the impact on the distribution of the relative weights of energy and isostericity, we used 4 different values of $\alpha = {0, 0.5, 0.8, 1.0}$. 
+Similarly, we also investigated the impact of an under- and over- estimate of the number of errors, by setting the presumed number of errors to 50\% (6 mutations) and 200\% (24 mutations) of their exact number (i.e.
+$12$).
 
@@
 by \RNApyro. More specifically, we fixed a threshold $\lambda \in [0,1]$, and predicted an error at position $i$ in sequence $\omega$ if and only if the
-probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceeds this threshold. To correct the errors we used the set of nucleotides with a probability
-higher than this threshold is  $C(i) = \{ n \; | \;  n \in \{ A,C,G,U \} \mbox{ and } P(i,n) > \lambda \mbox{ and }  n \neq \omega[i] \}$, where $\omega[i]$ is
-the nucleotide at position $i$ in the input sequence. We note that for the lowest thresholds multiple nucleotides can be available in $C(i)$ to correct
-the sequence. Here, we remind that our aim is to estimate the potential of error-correction of \RNApyro and not to develop an error-correction pipe-line, which  
-will be the subject of further studies. Finally, we progressively varied $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
-under the curve (AUC). We report our results in Figure~\ref{fig:ROCall}.
+probability $P(i,x)$ of a nucleotide $x \in \{ A,C,G,U \}$ exceeds this threshold. To correct the errors we used the set of nucleotides having probability
+greated than $\lambda$, that is  $$C(i) = \{ x \; | \;  x \in \{ A,C,G,U \} \mbox{ and } P(i,x) > \lambda \mbox{ and }  n \neq \omega[i] \},$$ where $\omega[i]$ is
+the nucleotide at position $i$ in the input sequence. We note that, for lower thresholds, multiple nucleotides may be available in $C(i)$ to correct
+the sequence. Here, we remind that our aim is to estimate the potential of error-correction of \RNApyro, and not to develop a full-fledged error-correction pipe-line, which  
+shall be the subject of further studies. Finally, we progressively varied $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
+under the curve (AUC). Our results are reported in Figure~\ref{fig:ROCall}.
 
@@
 %\end{subfigure}
-\caption{Performance of error-correction. Subfigures accuracy with under-estimated error rate (6 mutations), exact estimates (12 mutations) and over estimates 
+\caption{Performance of error-correction. Subfigures show accuracy with under-estimated error rates (6 mutations), exact estimates (12 mutations) and over estimates 
 (24 mutations). We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values 
@@
 
-Our data demonstrate that our algorithm exhibits interesting performance for error-correction applications. First, the AUC values (up to $0.86$) indicate that a
+Our data demonstrates that our algorithm exhibits interesting potential for error-correction applications. First, the AUC values (up to $0.86$) indicate that a
 signal has been successfully extracted. This result has been achieved with errors in loop regions (i.e. without base pairing information) and thus suggests
 that correction rates in structured regions (i.e. base paired regions) could be even higher. Next, the optimal values of $\alpha$ tend to be close to $0.0$. This 
-finding suggests that at this point the information issued from the stacking energy is currently modest. However, specific examples showed improved performance
-using this energy term. Further studies must be conducted to understand how to make the best use it. Finally, our algorithm seems robust to the number of
-mutations considered. Indeed, good AUC values are achieved with low estimates of the number of errors in the sequence (c.f. 50\% of the errors  in
-Fig.~\ref{fig:ROCall}(a)) and with large  values (c.f.~200\% of the errors  in Fig.~\ref{fig:ROCall}(c)). It is worth noting that scoring scheme with larger weight on
-the isostericity scores (low $\alpha$ values) seem more robust to under- and over-estimate of the number of errors.
+finding suggests that, at this point, the information issued from the consideration of stacking energies is currently modest. However, specific examples showed improved performance
+using this energy term. Further studies must be conducted to understand how to make the best use of it. Finally, our algorithm seems robust to the number of
+presumed mutations. Indeed, good AUC values are achieved even with conservative estimates for the number of errors (c.f.~50\% of the errors, leading to 
+Fig.~\ref{fig:ROCall}(a)), as well as with large  values (cf~200\% of the errors  in Fig.~\ref{fig:ROCall}(c)). It is worth noting that scoring schemes giving a larger weight on
+the isostericity scores (i.e. for low $\alpha$ values) seem more robust to under- and over-estimating the number of errors.

### 2013-01-16 Recomb/methods_RECOMB.tex
Page limit

@@
 Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
-then any sequence $s$ is assigned a probability proportional to its Boltzmann factor
+then any sequence $s$ has probability proportional to its Boltzmann factor
 \begin{align*}
@@
 The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
-Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the contribution of the isostericity to the pseudo-energy as
+Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the isostericity contribution to the pseudo-energy as
 \begin{align*}
@@
 We define a variant of the
- \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
- desired probability,  the two  functions
-$\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.
+ \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to compute these
+ probability. 
+%from the two  functions $\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.
 
@@
 \end{equation}
-
 \subsubsection{Inside computation.}
@@
 \begin{minipage}{0.38\textwidth}
-\caption{Principle of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
-can be decomposed as a sequence preceded by a, possibly mutated, base at $i$ 
+\caption{Principle of the inside computation (partition function). Any sequence with mutations  
+can be decomposed as a sequence preceded by a, possibly mutated, base 
 (Unpaired case), a sequence surrounded by some base-pair (Stacking-pair case), 
-or as two sequences separated by some base-pair (General base-pairing case). 
-In each case, one has to investigate any possible ways to distribute mutations between the different sequences 
-and locally instantiated bases.}
+or as two sequences segregated by some base-pair (General base-pairing case), and mutations must be distributed between sub-sequences 
+and instantiated bases.}
 \end{minipage}
+\SpaceCheating
 \end{figure}
@@
 The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function, i.e.
-the sum of Boltzmann factors, over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively).
+the sum of Boltzmann factors over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively).
 Such terms can be defined by recurrence, for which the following initial conditions holds:
@@
 Since the energetic terms only depend on base pairs, they are not involved in the initial conditions.
-
-The main recursion is composed of four terms:\\
+The main recursion itself is composed of four terms:\\
 
@@
 \label{eq:Z_rec}
-\resizebox{\textwidth}{!}{%
+\resizebox{.93\textwidth}{!}{%
 	$\Z{i,j}{m}{a,b}:=\left\{
@@
 \begin{description}
-\item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
-any sequence consists in a, possibly mutated, nucleotide $a'$ at position  $i$, 
-followed by a sequence over $[i+1,j]$ having either $m-\Kron_{a',s_i}$, accounting for a possible mutation 
-at position $i$, and having flanking nucleotides $a'$ and $b$.
-\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two, possibly mutated, nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$ having distance $m-\Kron_{a'b',s_is_j}$ (to avoid exceeding the targeted distance $m$).
+\item[$S_{i}=-1$:] If nucleotide at position $i$ is unpaired, then 
+any sequence is a concatenation of a, possibly mutated, nucleotide $a'$ at position  $i$, 
+followed by a sequence over $[i+1,j]$ having $m-\Kron_{a',s_i}$ mutations (accounting for a possible mutation 
+at position $i$), and having flanking nucleotides $a'$ and $b$.
+\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two, possibly mutated, nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$ having distance $m-\Kron_{a'b',s_is_j}$.
 Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
@@
 \begin{figure}[t]
-\begin{minipage}{0.75\textwidth}
+\begin{minipage}{0.79\textwidth}
 \resizebox{\textwidth}{!}{\input{FigDPOutside.pgf}}
 \end{minipage}
-\begin{minipage}{0.24\textwidth}
+\begin{minipage}{0.20\textwidth}
 \caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
@@
 \end{minipage}
+\SpaceCheating
 \end{figure}
@@
 (resp. in position $j-1$ it is $b$).
-We define function $\Y{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
+The resulting terms $\Y{i,j}{m}{a,b}$ can be computed by recurrence, using as initial conditions:
 \begin{equation}
@@
 		\displaystyle
-	  \Z{j,n-1}{m}{X,X}
+	  \Z{j,n-1}{m}{X,X}.
 \label{eq:Y_in}
 \end{equation}
-The recurrence, as shown below, will increase the interval $[i,j]$ by decreasing $i$ when
-it is not base paired. If it is with a position $k>j$, we increase $j$ to include it.
- Thus, when we need
-to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
-structure between $(j,n-1)$ must be a set of independent stems. Therefore,
- all the outside energy between $[j,n-1]$ is
-equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself unfolds as follows.
+The recurrence below extends the interval $[i,j]$, by including $i-1$ when
+position $i$ not base paired, or extended in both directions if $i$ is paired with a position $k>j$.
+% Thus, when we need
+%to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
+%structure between $(j,n-1)$ must be a set of independent stems. Therefore,
+% all the outside energy between $[j,n-1]$ is
+%equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. 
+The recursion itself unfolds as follows:
 \begin{equation}
@@
 and is not stacked inside, the 
-only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
-We can then consider the outside interval $[i-1,k+1]$ by multiplying it by the the \emph{inside}
+only term contributing directly to the energy is the isostericity of base pair $(i,k)$. 
+Therefore, we consider the outside interval $[i-1,k+1]$, multiplying it by the \emph{inside}
 value of the newly included interval (i.e. $[j,k-1]$), for 
 all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $k$ respectively.
-\item[$-1<S_{i}<i$:]As above but if position $i$ is to pairing with a lower value.
-\item[Else:] In all other cases, we are in a derivation of the SCFG that does not correspond to the 
+\item[$-1<S_{i}<i$:]As above but if position $i$ is paired with a lower value.
+\item[Else:] Any other derivation of the SCFG does not correspond to the 
 secondary structure $S$, and we return $0$.
@@
 \subsubsection{Combining Inside and Outside Values into Point-Wise Mutations Probabilities.}
-By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
-be described in function of the \emph{inside} term as $\Z{0,n-1}{m}{X,X}$,
- for any nucleotide $X\in B$ or
-in function of the \emph{outside} term, for any unpaired position $k$:
+By construction, the partition function over all sequences at exactly $m$ mutations of a reference sequence $s$ can 
+be either described from the \emph{inside} contribution $\Z{0,n-1}{m}{X,X}$ of the whole sequence,
+$\forall X\in B$, or from \emph{outside} terms as:
 $$
@@
 	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
-	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a}
+	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a},\forall k \text{	unpaired.}
 $$
@@
 We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
-Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations. 
+Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations, one defines
 \begin{equation}

### 2013-01-16 Recomb/results_RECOMB.tex
Page limit

@@
 \begin{table}[t]
-\begin{tabular}{cp{27em}}
-\begin{tabular}{lccc}
-$|s|$&\multicolumn{3}{c}{\#mutations}\\\cline{2-4}
+\begin{minipage}{.4\textwidth}
+\begin{tabular}{rccc}
+Length &\multicolumn{3}{c}{\#mutations}\\\cline{2-4}
 		 			  & 6   &  12  & 24\\\cline{2-4}
@@
 500         & 5400s&       &  21003s    \\
-\end{tabular}&
-\caption{Time to compute all probabilities. The first column indicates the length and  the column indexes indicate the number
+\end{tabular}
+\end{minipage}
+\begin{minipage}{.59\textwidth}
+\caption{Time required by the computation of probabilities. First column indicates the length and  the column indexes indicate the number
  of mutations. $\alpha$ is
 set at $0.5$,  $\beta$ to $15$ and $|\Omega|=44$.}
-\end{tabular}
 \label{tab:time}
+\end{minipage}
+\SpaceCheating
 \end{table}
@@
 probability $P(i,x)$ of a nucleotide $x \in \{ A,C,G,U \}$ exceeds this threshold. To correct the errors we used the set of nucleotides having probability
-greated than $\lambda$, that is  $$C(i) = \{ x \; | \;  x \in \{ A,C,G,U \} \mbox{ and } P(i,x) > \lambda \mbox{ and }  n \neq \omega[i] \},$$ where $\omega[i]$ is
+greated than $\lambda$, that is  $C(i) = \{ x \; | \;  x \in \{ A,C,G,U \} \mbox{ and } P(i,x) > \lambda \mbox{ and }  n \neq \omega[i] \},$ where $\omega[i]$ is
 the nucleotide at position $i$ in the input sequence. We note that, for lower thresholds, multiple nucleotides may be available in $C(i)$ to correct
@@
 \begin{figure}
-\begin{center}
-	\includegraphics[width=\textwidth]{subfigs_perform.png}
-\end{center}
+\centering
+	\includegraphics[width=\textwidth]{subfigs_perform.png}\\
+
 %\begin{subfigure}[b]{0.3\textwidth}
@@
 ranging of $\alpha=\{0,0.5,0.8,1.0\}$. The AUC is indicated in the legend of the figures. Each individual ROC curve represent the average performance over the 10 experiments.}
-\label{fig:ROCall}
+\label{fig:ROCall}\SpaceCheating
 \end{figure}

### 2013-05-17 JCB/introduction_JCB.tex
Initial commit JCB extended version

--- 2013-01-09 Recomb/introduction_RECOMB.tex
@@
-%!TEX root = main_RECOMB.tex
+%!TEX root = main_RNAPyro_JCB.tex
 \section{Introduction}
@@
 
-Ribonucleic acids (RNAs) are found in every living organisms and exhibit a broad range of functions, from catalyzing
-chemical reactions as the RNase P or the group II introns, hybridizing  messenger RNA to regulate gene expression,
+Ribonucleic acids (RNAs) are found in every living organism, and exhibit a broad range of functions, ranging from catalyzing
+chemical reactions, as the RNase P or the group II introns, hybridizing  messenger RNA to regulate gene expressions,
 to ribosomal RNA (rRNA) synthesizing proteins. Those functions  require specific structures,  encoded in their nucleotide

### 2013-05-17 JCB/methods_JCB.tex
Initial commit JCB extended version

--- 2013-01-16 Recomb/methods_RECOMB.tex
@@
-%!TEX root = main_RECOMB.tex
+%!TEX root = main_RNAPyro_JCB.tex
 \section{Methods}
@@
 the sum of Boltzmann factors over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively).
+\TODOYann{Reformat inside Figure as in Incarnation.}
 Such terms can be defined by recurrence, for which the following initial conditions holds:
@@
 \end{equation}
-\
+\TODOYann{Rephrase in term of tree, as in Incarnation??}
 The cases can be broken down as follows:
@@
 \end{figure}
+
 The \emph{Outside} function, $\mathcal Y$, is the partition function considering only the 
@@
 (resp. in position $j-1$ it is $b$).
+\TODOYann{Reformat outside Figure as in Incarnation.}
 The resulting terms $\Y{i,j}{m}{a,b}$ can be computed by recurrence, using as initial conditions:
@@
 \end{equation}
+\TODOYann{Rephrase in term of tree, as in Incarnation??}
 The five cases can be broked down as follows.
@@
 
+\TODOYann{}
+\TODOVlad{Add illustration from RECOMB presentation}
 In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, 

### 2013-05-17 JCB/introduction_JCB.tex
Added a few 'todo' notes

@@
 
+\TODOTous{Relire le papier et ajouter references apparues entre temps (ex: en preparant la presentation).}
 Ribonucleic acids (RNAs) are found in every living organism, and exhibit a broad range of functions, ranging from catalyzing

### 2013-05-17 JCB/methods_JCB.tex
Added a few 'todo' notes

@@
 length $n$, and $S$ a secondary structure without pseudoknots. 
+\TODOYann{Move above, define secondary structure and add fancy VARNA illustration}
+
 Formally, if $(i,j)$ and $(k,l)$ are base pairs in $S$, there is no overlapping extremities
@@
 
-\TODOYann{}
-\TODOVlad{Add illustration from RECOMB presentation}
+\TODOYann{Rephrase as tree constructs}
+\TODOVlad{Add illustration of Inside/Outside combination (adapted from RECOMB presentation??)}
 In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, 

### 2013-05-18 JCB/methods_JCB.tex
Started restructuring figures

@@
 %!TEX root = main_RNAPyro_JCB.tex
+
 \section{Methods}
@@
 
-\newcommand{\PE}[1]{E(#1)}
-\newcommand{\EI}{\text{EI}}
-\newcommand{\ES}{\text{ES}}
-\newcommand{\ISO}{\text{ISO}}
 We introduce a probabilistic model, which aims at capturing both the stability of the folded RNA and its ability to adopt a predefined 3D conformation.
@@
 Given $s\in B^n$ an RNA sequence, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
-length $n$, and $S$ a secondary structure without pseudoknots. 
+length $n$, and $\Struct$ a secondary structure without pseudoknots. 
 \TODOYann{Move above, define secondary structure and add fancy VARNA illustration}
@@
 \begin{figure}[t]
-\begin{minipage}{0.60\textwidth}
 \resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
-\end{minipage}
-\begin{minipage}{0.38\textwidth}
 \caption{Principle of the inside computation (partition function). Any sequence with mutations  
@@
 and instantiated bases.}
-\end{minipage}
-\SpaceCheating
 \end{figure}
@@
 	$\Z{i,j}{m}{a,b}:=\left\{
-  \begin{array}{ll}
+  \begin{array}{l}
   		\displaystyle
       \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}  
-      \Z{i+1,j}{m-\Kron_{a',s_i}}{a',b} & \text{If }S_{i}=-1\\
+      \Z{i+1,j}{m-\Kron_{a',s_i}}{a',b} \\
+       \text{[If }S_{i}=-1\text{]}\\
       \displaystyle
@@
 			 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}
-			 \cdot \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
-			 \text{Elif }S_i=j \land S_{i-1}=j+1\\
+			 \cdot \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}\\
+			 \text{[Else if }S_i=j \land S_{i-1}=j+1\text{]}\\
 			 \displaystyle
@@
       \cdot\Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
-      \cdot\Z{k+1,j}{m'}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
-      0 &\text{Otherwise}
+      \cdot\Z{k+1,j}{m'}{b',b} \\
+       \text{[Elif }S_i=k \land i < k \leq j\text{]}\\
+      0 \\\text{[Otherwise]}
 	\end{array}\right.$}

### 2013-05-18 scripts/art_main.py
update art

@@ -1,0 +1,134 @@
+import os
+import sys
+import shlex
+from subprocess import call
+from tempfile import mkdtemp, NamedTemporaryFile as NTF 
+from shutil import rmtree
+from pprint import pprint
+from itertools import combinations
+
+
+def get_raw_data(seq,l_reads,fold):
+    dir_temp = mkdtemp(dir='.')
+    cwd = os.getcwd()
+    os.chdir(dir_temp)
+
+    temp_file = NTF(dir='.',delete=False,suffix='.fasta')
+    temp_file.file.write('>seq\n%s' % seq)
+    temp_file.file.close()
+
+    try:
+        cmd = 'art_illumina -i %s -l %s -f %s -o out' % (
+                temp_file.name,
+                l_reads,
+                fold)
+        cmd  = shlex.split(cmd)
+        call(cmd)
+    except:
+        os.chdir(cwd)
+        rmtree(dir_temp)
+        print "Error with command:"
+        print '\t', cmd
+        sys.exit(1)
+
+    raw_reads = [x.strip() for x in open('out.fq')]
+    raw_align = [x.strip() for x in open('out.aln')]
+
+    os.chdir(cwd)
+    rmtree(dir_temp)
+
+    return raw_reads, raw_align
+
+def reverse_sens(seq):
+    dic = {'A':'U','U':'A','C':'G','G':'C'}
+    return ''.join(dic[x] for x in reversed(seq))
+
+def clean_align(raw_align,seq):
+    l = len(seq)
+    d_clean = {}
+
+    for i,x in enumerate(raw_align):
+        if x.startswith('>'):
+            _, name, start, orientation = x.split()
+            start = int(start)
+            d_clean[name] = {}
+
+            if orientation == '+':
+                d_clean[name]['ref'] = raw_align[i+1].replace('T','U')
+                d_clean[name]['seq'] = raw_align[i+2].replace('T','U')
+                d_clean[name]['start'] = start
+            else:
+                d_clean[name]['ref'] = reverse_sens(raw_align[i+1].replace('T','U'))
+                d_clean[name]['seq'] = reverse_sens(raw_align[i+2].replace('T','U'))
+                d_clean[name]['start'] = l - start - len(d_clean[name]['seq'])
+
+    return d_clean
+
+def check_align(wt,ref,seq,start):
+    wt = wt[start:start+len(ref)]
+    if wt != ref:
+        return False
[…]

### 2013-05-19 scripts/art_main.py
art pyrosequencing max

@@ -43,8 +43,8 @@ def reverse_sens(seq):
     dic = {'A':'U','U':'A','C':'G','G':'C'}
     return ''.join(dic[x] for x in reversed(seq))
 
-def clean_align(raw_align,seq):
-    l = len(seq)
+def clean_align(raw_align,l_seq):
+    l = l_seq
     d_clean = {}
 
     for i,x in enumerate(raw_align):
@@ -79,22 +79,22 @@ def merge_reads(d_clean_align,d_nt_probs):
     first_pos = d_clean_align[l_names_sorted[0]]['start']
     seq = d_clean_align[l_names_sorted[0]]['seq']
     d_seq_probs = {'start':first_pos,
-                   'probs':[{x:[d_nt_probs[l_names_sorted[0]][i]] for i,x in enumerate(seq)}]
+                   'probs':{i+first_pos:{x:[d_nt_probs[l_names_sorted[0]][i]]} for i,x in enumerate(seq)}
                   }
-
+    
     for name in l_names_sorted[1:]:
         seq = d_clean_align[name]['seq']
         start = d_clean_align[name]['start']
 
         for i,x in enumerate(seq):
             i = i + start
-            if i < len(d_seq_probs['probs']):
+            if i in d_seq_probs['probs']:
                 if x in d_seq_probs['probs'][i]:
                     d_seq_probs['probs'][i][x].append(d_nt_probs[name][i-start])
                 else:
                     d_seq_probs['probs'][i][x] = [d_nt_probs[name][i-start]]
             else:
-                d_seq_probs['probs'].append({x:[d_nt_probs[name][i-start]]})
+                d_seq_probs['probs'][i] = {x:[d_nt_probs[name][i-start]]}
 
 
     return d_seq_probs
@@ -108,12 +108,58 @@ def get_nt_probs(raw_reads):
 
     return d_nt_probs
 
+def get_max_vote(d_merge_probs):
+    seq = []
+    for pos in sorted(d_merge_probs['probs'].keys()):
+        t = tuple((x,len(d_merge_probs['probs'][pos][x])) 
+                  for x in d_merge_probs['probs'][pos])
+        nt = sorted(t, key=lambda z:z[1])[-1][0] 
+        if pos > len(seq):
+            seq.extend(['-']*(pos-len(seq)))
+        elif pos < len(seq):
+             print "problem get_max_vote order probs"
+             sys.exit(1)
+        seq.append(nt)
[…]

### 2013-05-20 JCB/methods_JCB.tex
New inside/outside figures
Started adapting text

@@
 \subsubsection{Inside computation.}
-\begin{figure}[t]
-\resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
+\begin{figure}[t]\centering
+\includegraphics[scale=\ScaleDP]{FigDPInsideWrapper}
 \caption{Principle of the inside computation (partition function). Any sequence with mutations  
@@
 
-The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function, i.e.
-the sum of Boltzmann factors over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively).
+The \emph{Inside} function $\Z{\Struct}{m}{x,y}$ is simply the partition function, i.e.
+the sum of Boltzmann factors over all sequences for a substructure $\Struct$, featuring $m$ mutations/errors compared to $s$, and having flanking nucleotides $x$ and $y$.
 \TODOYann{Reformat inside Figure as in Incarnation.}
-Such terms can be defined by recurrence, for which the following initial conditions holds:
-
+Such terms can be computed by recurrence, whose initial values are:
 \begin{equation}
-	\forall i \in [0,n-1]:\, \Z{i+1,i}{m}{a,b}=\left\{
+	\forall x,y\in B\times B, m\in [0,M]:\, \Z{\varepsilon}{m}{x,y}=\left\{
 	\begin{array}{ll}
@@
 \end{equation}
-In other words, the set of sequences at distance $m$ of the empty sequence is either empty if $m>0$, or restricted to the empty sequence, having energy $0$, if $m=0$.
+In other words, either there is no sequence at distance $m>0$ of the empty sequence, or the only allowed sequence is the empty sequence ($m=0$), having energy $0$.
 Since the energetic terms only depend on base pairs, they are not involved in the initial conditions.
@@
 \label{eq:Z_rec}
-\resizebox{.93\textwidth}{!}{%
-	$\Z{i,j}{m}{a,b}:=\left\{
-  \begin{array}{l}
+	\Z{i,j}{m}{a,b}:=\left\{
+  \begin{array}{ll}
   		\displaystyle
       \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}  
-      \Z{i+1,j}{m-\Kron_{a',s_i}}{a',b} \\
+      \Z{i+1,j}{m-\Kron_{a',s_i}}{a',b} &
        \text{[If }S_{i}=-1\text{]}\\
@@
 			 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}
-			 \cdot \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}\\
+			 \cdot \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
 			 \text{[Else if }S_i=j \land S_{i-1}=j+1\text{]}\\
@@
       \cdot\Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
-      \cdot\Z{k+1,j}{m'}{b',b} \\
+      \cdot\Z{k+1,j}{m'}{b',b} &
        \text{[Elif }S_i=k \land i < k \leq j\text{]}\\
       0 \\\text{[Otherwise]}
-	\end{array}\right.$}
+	\end{array}\right.
 \end{equation}
@@
 \subsubsection{Outside computation.}	
-\begin{figure}[t]
-\begin{minipage}{0.79\textwidth}
-\resizebox{\textwidth}{!}{\input{FigDPOutside.pgf}}
-\end{minipage}
-\begin{minipage}{0.20\textwidth}
+\begin{figure}[t]\centering
+\includegraphics[scale=\ScaleDP]{FigDPOutsideWrapper}
 \caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
 therefore its efficient implementation requires a precomputation of the inside contributions.}
-\end{minipage}
-\SpaceCheating
 \end{figure}

### 2013-05-22 JCB/methods_JCB.tex
Almost there on outside...

@@
 
-We introduce a probabilistic model, which aims at capturing both the stability of the folded RNA and its ability to adopt a predefined 3D conformation.
-To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment.
+We introduce a probabilistic model, which aims at capturing both the stability of the folded RNA and its affinity towards a predefined 3D conformation.
+To that purpose, a Boltzmann weighted distribution is assumed, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment.
 In this model, the probability that the nucleotide at a given position needs to be mutated (i.e.
-corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990}.
+corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990} in time which scales linearly with the sequence length.
+
+\subsubsection{Definitions.}
+Let $\B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
+Given an RNA sequence $s\in \B^n$, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
+length $n$.
+$\Struct$ is a secondary structure without pseudoknots, denoted by a dot-parenthesis string (well-parenthesized expression with dots).
+In any such expression, matched parentheses induce and unambiguous set of corresponding positions, associated with base-pairing positions, mediated by hydrogen bonds.
+It follows that if $(i,j)$ and $(k,l)$ are base pairs in $S$, there is no overlapping extremities  $\{i,j\}\cap \{k,l\}=\varnothing$ and either the intersection is empty
+ ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
+ $[i,j]\subset[k,l])$.
+Let us finally denote by $\delta: \B^*\times \B^* \to \mathbb{N}^+$ the Hamming distance, i.e.
+the number of differing positions between two sequences $s'$ and $s''$ such that $|s'|=|s''|$.
+\TODOYann{Move above, define secondary structure and add fancy VARNA illustration}
 
@@
 \begin{align*}
-  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
+  {B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
 \end{align*}
@@
 \subsubsection{Energy Contribution.}
-The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
+The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
 Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
@@
 This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).
-The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed 
-(e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
+The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed.
+For instance, multiloops do not consider base-specific contributions, and therefore their consideration would constitute a criterion for preferring a sequence over another.
+Furthermore, it greatly eases the design and implementation of dynamic-programming equations.
 \subsubsection{Isostericity Contribution.}
@@
 Isostericity scores range between $0$ and $9.7$, $0$ corresponding to a perfect isostericity, and a penalty of $10$ is used for missing entries.
-The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
+The isostericity contribution exponentially favors sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
+\TODOYann{Avoid duplicate reference to Stombaugh et al}
 
-\subsection{Mutational Profile of Sequences}
+\subsubsection{Combining contributions.}
+Let us remark that any of the individual contributions can be associated to (a subset of) the base-pairs occurring in the structure, possibly complemented, in the case of stacking pairs, with the knowledge of flanking base-pairing nucleotide.
+Therefore, let us denote by $\EBP{i,j}{xy}{a'b'}$ the local contribution of a base-pair $(i,j)$ of nucleotides $(a',b')$, surrounded by a stacking pair $(x,y)$ (or $\varnothing$ otherwise), to the pseudo-energy:
+\begin{equation}
+  \EBP{i,j}{xy}{a'b'}  = \alpha \cdot\ES^\beta_{xy \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
+\end{equation}
+
+\subsection{Computing the Mutational Profile of Sequences}
 
 Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
-We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $m$ mutations from $s$
+We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $M$ mutations from $s$
 %  under the SCFG derivation $S$ 
-(formally $\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
+(formally $\mathbb{P}(s_i = x\mid s,\Omega, \Struct,M)$). 
 We define a variant of the
- \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to compute these
- probability. 
+ \emph{Inside-Outside algorithm}~\cite{Lari1990} to compute this
+ probability, based on two sets of dynamic programming equations. 
 %from the two  functions $\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.
 
-The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec}, is analogous to the \emph{inside} algorithm.
-It is the partition function, i.e. the sum of Boltzmann factors, over all sequences within $[i,j]$, 
+The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C}, is analogous to an \emph{inside} computation.
+Considering an interval of the sequence/secondary structure, it consists in the accumulated contributions of any possible sequences that have suitable Hamming distance within the interval.
+It is therefore similar to a partition function, i.e.
+the sum of Boltzmann factors over all sequences within $[i,j]$,
 knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$), within 
 $m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
- computes the \emph{outside} algorithm, i.e.  
-the partition function over sequences within $m$ mutations of $s$, restricted to two intervals $[0,i]\cup[j,n-1]$, and 
-knowing  that position $i+1$ is composed of $a$ (resp.
-$j-1$ is $b$).
-A suitable combination of these terms, given in Equation~\eqref{eq:combine}, gives the total weight, and in turn the probability, of seeing a specific base at a given position.
+ computes the \emph{outside} algorithm,   
+the partition function over sequences within $m$ mutations of $s$ outside the interval (restricted to two intervals $[0,i]\cup[j,n-1]$), 
+knowing  that flanking inner positions $(i+1,j-1)$ contain nucleotides $a$ and $b$ respectively.
+A suitable multiplicative combination of these terms, computed as shown in Equation~\eqref{eq:combine}, gives the total weight of every possible sequences that support a given base-pair (or unpaired position) and, in turn, the probability of seeing a specific base at a given position.
 
@@
 
-\subsubsection{Definitions.}
-Let $B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
-Given $s\in B^n$ an RNA sequence, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
-length $n$, and $\Struct$ a secondary structure without pseudoknots. 
-\TODOYann{Move above, define secondary structure and add fancy VARNA illustration}
-
-Formally, if $(i,j)$ and $(k,l)$ are base pairs in $S$, there is no overlapping extremities
- $\{i,j\}\cap \{k,l\}=\varnothing$ and either the intersection is empty 
- ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
- $[i,j]\subset[k,l])$.
-
-Let us then remind the Hamming distance function $\delta: B^*\times B^* \to \mathbb{N}^+$, which takes two sequences $s'$ and $s''$ as input, $|s'|=|s''|$, and returns the number of differing positions.
-Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
-\begin{equation}
-  E_{(i,j),ab \to a'b'}^{\Omega,\beta}  = \alpha \cdot\ES^\beta_{a b \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
-\end{equation}
 \subsubsection{Inside computation.}
@@
 \includegraphics[scale=\ScaleDP]{FigDPInsideWrapper}
-\caption{Principle of the inside computation (partition function). Any sequence with mutations  
+\caption{Principle of the inside computation (partition function). Any sequence (mutated)  
 can be decomposed as a sequence preceded by a, possibly mutated, base 
 (Unpaired case), a sequence surrounded by some base-pair (Stacking-pair case), 
-or as two sequences segregated by some base-pair (General base-pairing case), and mutations must be distributed between sub-sequences 
-and instantiated bases.}
+or as two sequences segregated by some base-pair (General base-pairing case). In this latter case, mutations must be distributed between sub-sequences.\label{fig:inside}}
 \end{figure}
@@
 The \emph{Inside} function $\Z{\Struct}{m}{x,y}$ is simply the partition function, i.e.
-the sum of Boltzmann factors over all sequences for a substructure $\Struct$, featuring $m$ mutations/errors compared to $s$, and having flanking nucleotides $x$ and $y$.
-\TODOYann{Reformat inside Figure as in Incarnation.}
-Such terms can be computed by recurrence, whose initial values are:
+the sum of Boltzmann factors over all sequences for a substructure $\Struct$ (implicitly attached to an interval $[i,j]$ of the sequence), featuring $m$ mutations/errors compared to $s$, and having flanking nucleotides $x$ and $y$.
+Such terms can be computed by recurrence, using the following equation for the initial case:
 \begin{equation}
-	\forall x,y\in B\times B, m\in [0,M]:\, \Z{\varepsilon}{m}{x,y}=\left\{
+	\forall x,y\in B\times \B, m\in [0,M]:\, \Z{\varepsilon}{m}{x,y}=\left\{
 	\begin{array}{ll}
-		1 &\text{If } m = 0\\
-		0 &\text{Otherwise.}
+		1 &\text{ If } m = 0\\
+		0 &\text{ Otherwise.}
 	\end{array}\right.
@@
 In other words, either there is no sequence at distance $m>0$ of the empty sequence, or the only allowed sequence is the empty sequence ($m=0$), having energy $0$.
-Since the energetic terms only depend on base pairs, they are not involved in the initial conditions.
-The main recursion itself is composed of four terms:\\
+Since the contributions only depend on base pairs, they do not appear in the initial conditions.
 
+The main recursion considers a general structure $\Struct$, flanked by two nucleotides (outside the region of interest) $x$ and $y$, respectively on the $5'$ and $3'$ end of the sequence.
+As illustrated by Figure~\ref{fig:inside}, it is computed, for each subinterval $[i,j]$, by considering one of the three following cases, dependent on the base-pairing status and context of the leftmost position in the sequence/structure:
+\begin{itemize}
+\item {\bf Case 1: Unpaired leftmost position.} If the first position  is unpaired in the structure, then $\Struct$ can be further decomposed, using a dot-parenthesis notation, as $\Struct = \ub \Struct'$.
+Let $a\in B$ be the nucleotide found at the leftmost position in the initial structure, then one has:
 \begin{equation}
-\label{eq:Z_rec}
-	\Z{i,j}{m}{a,b}:=\left\{
-  \begin{array}{ll}
-  		\displaystyle
-      \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}  
-      \Z{i+1,j}{m-\Kron_{a',s_i}}{a',b} &
-       \text{[If }S_{i}=-1\text{]}\\
-      \displaystyle
-      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
-			 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}
-			 \cdot \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
-			 \text{[Else if }S_i=j \land S_{i-1}=j+1\text{]}\\
-			 \displaystyle
-      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
-      \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-   		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
-      \cdot\Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
-      \cdot\Z{k+1,j}{m'}{b',b} &
-       \text{[Elif }S_i=k \land i < k \leq j\text{]}\\
-      0 \\\text{[Otherwise]}
-	\end{array}\right.
+\label{eq:Z_rec_A}
+	\Z{\Struct}{m}{x,y} =
+      \sum_{\substack{a'\in \B,\\ \Kron_{a,a'}\le m}}  
+      \Z{\Struct'}{m-\Kron_{a,a'}}{a',y}.
 \end{equation}
-\TODOYann{Rephrase in term of tree, as in Incarnation??}
-The cases can be broken down as follows:
-\begin{description}
-\item[$S_{i}=-1$:] If nucleotide at position $i$ is unpaired, then 
-any sequence is a concatenation of a, possibly mutated, nucleotide $a'$ at position  $i$, 
-followed by a sequence over $[i+1,j]$ having $m-\Kron_{a',s_i}$ mutations (accounting for a possible mutation 
-at position $i$), and having flanking nucleotides $a'$ and $b$.
-\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two, possibly mutated, nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$ having distance $m-\Kron_{a'b',s_is_j}$.
-Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
-then a stacking energy contribution is added. 
-\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
-only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
-Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$.
-Since the number of mutations sum to $m$ over the whole sequence must , then a parameter $m'$ is introduced to distribute the remaining mutations between the two sequences.
-\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
-\end{description}
+Indeed, any suitable sequence is a concatenation of a, possibly mutated, nucleotide $a'$ at the first position, followed by a sequence over the remaining interval, having $m-\Kron_{a,a'}$ mutations (accounting for a possible mutation at first position), and having flanking nucleotides $a'$ and $y$.
+\item {\bf Case 2: Paired ends, stacking onto another base-pair.} If both ends of the considered interval form a base-pair ($\Struct = \op\Struct'\cp$), stacking onto another base pair just outside whole region, then the isosteric contribution of the base-pair must be supplemented with a specific "stacking-pairs" bonus.
+Let $a$ and $b$ be the nucleotides found on both ends of the interval (positions $i$ and $j$), then one has
+\begin{equation}
+\label{eq:Z_rec_B}
+	\Z{\Struct}{m}{x,y} =
+      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{ab,a'b'}\le m}}
+			 e^{\frac{-\EBP{i,j}{xy}{a'b'}}{RT}}
+			 \cdot \Z{S'}{m-\Kron_{ab,a'b'}}{a',b'}.
+\end{equation}
+Any sequence generated here consists of two, possibly mutated, nucleotides $a'$ and $b'$, flanking a sequence over the remaining portion.
+In order for the total distance to sum to $m$, this portion must feature $m-\Kron_{ab,a'b'}$ additional point wise mutations.
+
+\item {\bf Case 3: Paired leftmost position, but no stacking pairs.} In this case, the structure is split into two parts by the base pair ($\Struct = \op \Struct'\cp\Struct''$).
+Let us denote by $k$ the partner of position $i$, and by $a$, $b$ and $c$ the bases found at positions $i$, $k$ and $j$ respectively, then one has:
+\begin{equation}
+\label{eq:Z_rec_C}
+	\Z{\Struct}{m}{x,y}=\sum_{\substack{a',b'\in \B^2,\\ \Kron_{ab, a'b'}\le m}}
+      \sum_{m'=0}^{m-\Kron_{ab,a'b'}}
+   		 e^{\frac{-\EBP{i,k}{\varnothing}{a'b'}}{RT}}
+      \cdot\Z{\Struct'}{m-m'-\Kron_{ab,a'b'}}{a',b'}
+      \cdot\Z{\Struct''}{m'}{b',y}.
+\end{equation}
+In other words, if the leftmost position is paired, and the base-pair is not stacking onto another base-pair, then the  only term contributing directly to the energy is the isostericity of the base pair.
+Admissible sequences for $\Struct$ consist of two paired nucleotides $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence for $\Struct'$ (over an interval $[i+1,k-1]$), and followed by a (possibly empty) sequence for $\Struct''$ (over $[k+1,j]$).
+Since the total number of mutations sums to $m$, a parameter $m'$ is introduced to distribute the remaining mutations between the two sequences.
+\end{itemize}
 
@@
 
-The \emph{Outside} function, $\mathcal Y$, is the partition function considering only the 
-contributions of subsequences $[0,i]\cup[j,n-1]$ over the mutants of $s$ having exactly $m$ mutations between $[0,i]\cup[j,n-1]$ and whose nucleotide at position $i+1$ is $a$ 
-(resp. in position $j-1$ it is $b$).
+The \emph{Outside} function, $\Y{\Struct}{m}{x,y}$ is the partition function considering only the 
+contributions of subsequences excluding a given structure/interval $\Struct$, occupying the open interval $]i,j[$ in the sequence, at Hamming distance exactly $m$ to the initial sequence $s$, and assuming that nucleotides $x$ and $y$ were previously chosen for $i+1$ and $j-1$, the outermost portions of the excluded structure.
 \TODOYann{Reformat outside Figure as in Incarnation.}
-The resulting terms $\Y{i,j}{m}{a,b}$ can be computed by recurrence, using as initial conditions:
+The associated terms $\Y{\Struct}{m}{x,y}$ can then be computed recursively, considering as the base case any prefix $\Struct$ of the complete structure:
 \begin{equation}
-	\Y{-1,j}{m}{X,X}:=
-		\displaystyle
-	  \Z{j,n-1}{m}{X,X}.
+	\forall x,y\in B\times \B, \forall \Struct'\text{ s.t. }\Struct=\Struct'.\Struct'', m\in [0,M]:\, \Y{\Struct^\ast}{m}{x,y}=\Z{\Struct''}{m}{y,X}
 \label{eq:Y_in}
 \end{equation}
-The recurrence below extends the interval $[i,j]$, by including $i-1$ when
-position $i$ not base paired, or extended in both directions if $i$ is paired with a position $k>j$.
-% Thus, when we need
-%to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
-%structure between $(j,n-1)$ must be a set of independent stems. Therefore,
-% all the outside energy between $[j,n-1]$ is
-%equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. 
-The recursion itself unfolds as follows:
+where $X\in\B$ can be any nucleotide, and will not affect further computations.
+
+The main recurrence below extends the excluded structure leftward, and considers four cases depending on its base-pairing status and directionality:
+\begin{itemize}
+\item {\bf Case 1: Unpaired position.}
 \begin{equation}
-\resizebox{\textwidth}{!}{%
-	$\Y{i,j}{m}{a,b} = \left\{
-  \begin{array}{ll}
-		\displaystyle
-    \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}
-    \Y{i-1,j}{m- \Kron_{a',s_i}}{a',b} &
-    \text{Elif }S_i=-1 \\
-    \displaystyle
-    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
-		 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}\cdot
-    \Y{i-1,j+1}{m- \Kron_{a'b',s_is_j}}{a',b'} &
-   	 \text{Elif }S_{i}=j \land S_{i+1}=j-1\\
-		 \displaystyle
-		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
+	\Y{\Struct}{m}{x,y} = \sum_{\substack{a'\in \B,\\ \Kron_{a,a'}\le m}}
+    \Y{\ub\Struct}{m- \Kron_{a,a'}}{a',y}
+\label{eq:Y_rec_A}
+\end{equation}
+\item {\bf Case 2: Stacking base-pair.}
+\begin{equation}
+	\Y{\Struct}{m}{x,y} = 
+    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{ab,a'b'}\le m}}
+		 e^{\frac{-\EBP{i,j}{xy}{a'b'}}{RT}}\cdot
+    \Y{\op\Struct\cp}{m- \Kron_{ab,a'b'}}{a',b'} 
+\label{eq:Y_rec}
+\end{equation}
+\item {\bf Case 3: Next position paired rightward.}
+\begin{equation}
+	\Y{i,j}{m}{a,b} = \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
 		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
@@
 		 \cdot\Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
-     \cdot\Z{j,k-1}{m'}{b,b'} &
-		 \text{Elif }S_{i}=k \geq j\\
-		 \displaystyle
+     \cdot\Z{j,k-1}{m'}{b,b'} 
+\label{eq:Y_rec}
+\end{equation}
+\item {\bf Case 4: Next position paired leftward.}
+\begin{equation}
+	\Y{i,j}{m}{a,b} = 
 		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_ks_i}\le m}}
@@
 		 \cdot\Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
-     \cdot\Z{k+1,i-1}{m'}{a',b'} &
-		 \text{Elif }-1 < S_{i}=k < i\\
-		 0 & \text{Otherwise}
-  \end{array}\right.$
-}\label{eq:Y_rec}
+     \cdot\Z{k+1,i-1}{m'}{a',b'}
+\label{eq:Y_rec}
 \end{equation}
+\end{itemize}
 \TODOYann{Rephrase in term of tree, as in Incarnation??}

### 2013-05-22 JCB/methods_JCB.tex
Finished outside

@@
 the sum of Boltzmann factors over all sequences for a substructure $\Struct$ (implicitly attached to an interval $[i,j]$ of the sequence), featuring $m$ mutations/errors compared to $s$, and having flanking nucleotides $x$ and $y$.
-Such terms can be computed by recurrence, using the following equation for the initial case:
+Such terms can be computed recursively, using the following equation for the initial case:
 \begin{equation}
@@
 \subsubsection{Outside computation.}
-\begin{figure}[t]\centering
-\includegraphics[scale=\ScaleDP]{FigDPOutsideWrapper}
-\caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
-therefore its efficient implementation requires a precomputation of the inside contributions.}
-\end{figure}
 
@@
 \TODOYann{Reformat outside Figure as in Incarnation.}
-The associated terms $\Y{\Struct}{m}{x,y}$ can then be computed recursively, considering as the base case any prefix $\Struct$ of the complete structure:
+The associated terms $\Y{\Struct}{m}{x,y}$ can then be computed recursively, initially considering the case of any prefix $\Struct'$ of the complete structure:
 \begin{equation}
-	\forall x,y\in B\times \B, \forall \Struct'\text{ s.t. }\Struct=\Struct'.\Struct'', m\in [0,M]:\, \Y{\Struct^\ast}{m}{x,y}=\Z{\Struct''}{m}{y,X}
+	\forall x,y\in \B\times \B, \forall \Struct'\text{ s.t. }\Struct=\Struct'.\Struct'', m\in [0,M]:\, \Y{\Struct'}{m}{x,y}=\Z{\Struct''}{m}{y,z}
 \label{eq:Y_in}
 \end{equation}
-where $X\in\B$ can be any nucleotide, and will not affect further computations.
+where $z\in\B$ can be any nucleotide, and provably does not affect further computations. 
+In other words, the sequences explored by an outside computation, excluding a prefix of $\Struct$, are exactly the sequences generated on the corresponded suffix.
+This set of sequence is also the inside term on the suffix structure.
+It is also worth pointing out that $\Y{\Struct}{m}{x,y}= \Z{\varepsilon}{m}{y,X}=\mathbbm{1}_{m=0}$.
 
-The main recurrence below extends the excluded structure leftward, and considers four cases depending on its base-pairing status and directionality:
+\begin{figure}[t]\centering
+\includegraphics[scale=\ScaleDP]{FigDPOutsideWrapper}
+\caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
+therefore its efficient implementation requires the precomputation and storage of the inside contributions.}
+\end{figure}
+In general, the main recurrence below works by extending the excluded structure $\Struct$ (covering the $]i,j[$ interval) in the leftward direction, and considers four cases depending on the base-pairing status and directionality of the next considered position:
 \begin{itemize}
-\item {\bf Case 1: Unpaired position.}
+\item {\bf Case 1: Unpaired position.} When the innermost leftward base is unpaired, any base $a\in\B$ may be chosen for this position, and the rest of the sequence is generated recursively.
+The number of remaining mutations may be decreased by the choice of $a$, leading to the following recurrence:
 \begin{equation}
@@
 \end{equation}
-\item {\bf Case 2: Stacking base-pair.}
+\item {\bf Case 2: Stacking base-pair.} If both ends of the excluded structure are paired together, and are nested within another base-pair in the remaining structure, then an additional contribution for stacking has to be considered.
+The outside terms are then computed by simulating the choice of two base-pairing nucleotides, and by recurring on the remaining portion, as follows:
 \begin{equation}
@@
     \Y{\op\Struct\cp}{m- \Kron_{ab,a'b'}}{a',b'} 
-\label{eq:Y_rec}
+\label{eq:Y_rec_B}
 \end{equation}
-\item {\bf Case 3: Next position paired rightward.}
+\item {\bf Case 3: Next position paired rightward, in the absence of stacking pair.} In this case, the innermost leftward position $i$ is paired to the right at some position $k$ . 
+Let us assume that its partner resides outside the excluded structure $\Struct$ (This assumption is provably without loss of generality, and directly follows from the fact that $\Struct$ is well-parenthesized).
+The structure within the base-pair may be described by the expression $\op\Struct \Struct'\cp$, where $\Struct'$ may possibly be empty (except if $\Struct$ is enclosed within corresponding brackets, to avoid stacking pairs).
+Therefore, any sequence considered by the outside computation consists in three independent parts: two nucleotides for the paired positions, a sequence for the region excluding $\Struct''$, and a sequence for $\Struct'$.
+It follows that:
 \begin{equation}
-	\Y{i,j}{m}{a,b} = \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
-		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
-  		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
-		 \cdot\Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
-     \cdot\Z{j,k-1}{m'}{b,b'} 
-\label{eq:Y_rec}
+	\Y{\Struct}{m}{x,y} = \sum_{\substack{a'b'\in \B^2,\\ \Kron_{ab,a'b'}\le m}}
+		 \sum_{m'=0}^{m-\Kron_{ab,a'b'}}
+  		 e^{\frac{-\EBP{i,k}{\varnothing}{a'b'}}{RT}}
+		 \cdot\Y{\op\Struct \Struct'\cp}{m- \Kron_{ab,a'b'} - m'}{a',b'}
+     \cdot\Z{\Struct'}{m'}{y,b'} .
+\label{eq:Y_rec_C}
 \end{equation}
-\item {\bf Case 4: Next position paired leftward.}
+\item {\bf Case 4: Next position paired leftward.} If the innermost, leftward, position $i$ is paired to some position $k<i$, delimiting a substructure $\Struct'$, then any sequence considered by the outside computation consists in three parts: two nucleotides $a$ and $b$, a sequence for $\Struct'$, and a sequence for the region excluding $\op\Struct'\cp\Struct$.
+Consequently, one has:
 \begin{equation}
-	\Y{i,j}{m}{a,b} = 
-		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_ks_i}\le m}}
-		 \sum_{m'=0}^{m-\Kron_{a'b',s_ks_i}}
-   	 e^{\frac{-E_{(k,i),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
-		 \cdot\Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
-     \cdot\Z{k+1,i-1}{m'}{a',b'}
-\label{eq:Y_rec}
+	\Y{i,j}{m}{x,y} = 
+		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{ab,a'b'}\le m}}
+		 \sum_{m'=0}^{m-\Kron_{ab,a'b'}}
+   	 e^{\frac{-\EBP{k,i}{\varnothing}{a'b'}}{RT}}
+		 \cdot\Y{\op\Struct'\cp\Struct}{m- \Kron_{ab,a'b'} - m'}{a',b}
+     \cdot\Z{\Struct'}{m'}{a',b'}
+\label{eq:Y_rec_D}
 \end{equation}
 \end{itemize}
-\TODOYann{Rephrase in term of tree, as in Incarnation??}
-The five cases can be broked down as follows.
-\begin{description}
-\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, then the value is the same
-as if we decrease the lower interval bound by $1$ (i.e. $i-1$), and consider all possible
-nucleotides $a'$ at position $i$, correcting the number of mutants
-in function of $\Kron_{a',s_i}$.
-\item[$S_{i}=j$ and $S_{i+1}=j-1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i+1$ is
-paired with $j-11$, we are in the only case were stacked base pairs can occur. We thus add
-the energy of the stacking and of the isostericity of the base pair $(i,j)$. What is left
-to compute is the \emph{outside} value for the interval $[i-1,j+1]$ over all possible nucleotides 
-$a',b'\in B^2$ at positions $i$ and $j$ respectively.
-\item[$S_{i}=k \geq j$:]If nucleotide $i$ is paired with position $k\geq j$, 
-and is not stacked inside, the 
-only term contributing directly to the energy is the isostericity of base pair $(i,k)$. 
-Therefore, we consider the outside interval $[i-1,k+1]$, multiplying it by the \emph{inside}
-value of the newly included interval (i.e. $[j,k-1]$), for 
-all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $k$ respectively.
-\item[$-1<S_{i}<i$:]As above but if position $i$ is paired with a lower value.
-\item[Else:] Any other derivation of the SCFG does not correspond to the 
-secondary structure $S$, and we return $0$.
 
-\end{description}
+\subsubsection{Combining Inside and Outside Computations into Point-Wise Mutations Probabilities.}
 
-\subsubsection{Combining Inside and Outside Values into Point-Wise Mutations Probabilities.}
-By construction, the partition function over all sequences at exactly $m$ mutations of a reference sequence $s$ can 
-be either described from the \emph{inside} contribution $\Z{0,n-1}{m}{X,X}$ of the whole sequence,
-$\forall X\in B$, or from \emph{outside} terms as:
-$$
-	\Z{0,n-1}{m}{X,X}
-	\equiv
-	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
-	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a},\forall k \text{	unpaired.}
-$$
+We are now left to compute the probability that a a given nucleotide $a\in\B$ is found at a given position $i$.
+This quantity can also be expressed as the ratio of $\mathcal{W}^M_{\substack{i, [a]}}$, the total Boltzmann weight 
+of the set of sequences featuring the nucleotide, and $\Z{\varepsilon}{\le M}{X,X}$ the total weight of sequences having at most $M$ mutations:
+\begin{equation}
+	\mathbb{P}(s_i = a\;|\; M) := \frac{\mathcal{W}^M_{\substack{i, [a]}}}{\sum_{m=0}^{M}\Z{\varepsilon}{m}{X,X}}\label{eq:normalize}
+\end{equation}
 
-We are now left to compute the probability that a given position is a given nucleotide.
-We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
-Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations, one defines
-\begin{equation}
-	\mathbb{P}(s_i = x\;|\; M) := \frac{\mathcal{W}^M_{\substack{i, [x]}}}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}
-\end{equation}
+To that purpose, we leverage the \emph{Inside-Outside} construction, and obtain the three following cases:
+\begin{itemize}
+\item 
 where $\mathcal{W}{\substack{\ast\\ \ast}}$ is defined by:
@@
 \end{equation}
+\end{itemize}
 
@@
 
+By construction, the partition function over all sequences at exactly $m$ mutations of a reference sequence $s$ can 
+be either described from the \emph{inside} contribution $\Z{\Struct}{m}{X,X}$ of the whole sequence,
+$\forall X\in B$, or from a subset of the \emph{outside} terms, such that:
+$$
+	\Z{\Struct}{m}{X,X}
+	\equiv
+	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
+	\Y{\ub}{m-\Kron_{a,s[k]}}{a,a},\forall k \text{	unpaired.}
+$$
+
 \subsection{Complexity Considerations}
-Equations~\eqref{eq:Z_rec} and~\eqref{eq:Y_rec} can be computed using dynamic programming.
+Equations~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C} and~\eqref{eq:Y_rec_A}-\eqref{eq:Y_rec_D} can be computed using dynamic programming.
 Namely, the $\mathcal{Z}^{*}_{*}$ and $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and interval lengths, memorizing the results as they become available to ensure constant-time access during later stages of the computation.

### 2013-05-23 scripts/art_main.py
more script16s

@@ -108,7 +108,21 @@ def get_nt_probs(raw_reads):
 
     return d_nt_probs
 
-def get_max_vote(d_merge_probs):
+def list_ss_bp(ss):
+    d_lbp = {}
+    d_rbp = {}
+    left = []
+    for i,x in enumerate(ss):
+        if x == '(':
+            left.append(i)
+        elif x == ')':
+            l = left.pop()
+            r = i
+            d_lbp[l] = r
+            d_rbp[r] = l
+    return d_lbp,d_rbp
+
+def get_max_vote(d_merge_probs,ss):
     seq = []
     for pos in sorted(d_merge_probs['probs'].keys()):
         t = tuple((x,len(d_merge_probs['probs'][pos][x])) 
@@ -122,7 +136,8 @@ def get_max_vote(d_merge_probs):
         seq.append(nt)
     return ''.join(seq)
 
-def main(seq,l_reads,fold):
+
+def main(seq,l_reads,fold,ss):
     seq = seq.upper().replace('U','T')
     if any(x not in ('ACGT') for x in seq):
         print help()
@@ -133,7 +148,7 @@ def main(seq,l_reads,fold):
     d_clean_align = clean_align(raw_align,l_seq)
     d_nt_probs = get_nt_probs(raw_reads)
     d_merge_probs = merge_reads(d_clean_align,d_nt_probs)
-    max_vote_seq = get_max_vote(d_merge_probs)
+    max_vote_seq = get_max_vote(d_merge_probs,ss)
     print seq
     print max_vote_seq
 
@@ -165,6 +180,7 @@ if __name__ == '__main__':
     opts = sys.argv
     
     seq = ''
+    ss = ''
     l_reads = 0
     fold = 0
 
@@ -174,12 +190,14 @@ if __name__ == '__main__':
         elif x == '-s':
             seq = opts[i+1].upper().replace('U','T')
         elif x == '-f':
-            fold = opts[i+1]
+            fold = int(opts[i+1])
+        elif x == '-ss':
+            ss = opts[i+1]
 
-    if not seq or not l_reads or not fold:
+    if not seq or not l_reads or not fold or not ss:
         print seq, l_reads,fold
         help()
         sys.exit(1)
 
-    main(seq,l_reads,fold)
+    main(seq,l_reads,fold,ss)
     

### 2013-05-23 src/RNAPyro.py
Change

@@ -382,13 +382,23 @@ def energy((a,b),(a2,b2),alpha):
   E = STACKING_ENERGY[a,a2,b2,b]
   return  math.exp(-(alpha*E)/(BOLTZMANN*T))
 
+def isGapChar(c):
+  return c=="." or c=="-"
+
 def isostericity(seq,ref_seq,(i,j),(a,b), alpha):
   if not ref_seq:
     return 1.
   #isostericity of going from original base pair to (a,b)
-  iso_mut = sum(ISO[(ref[i],ref[j]),(a,b)] for ref in ref_seq)
-  iso_start = sum(ISO[(ref[i],ref[j]),(seq[i],seq[j])] for ref in ref_seq)
-  iso = mpf(iso_mut-iso_start)/len(ref_seq)
+  iso_delta = 0.
+  nb_correct = 0
+  for ref in ref_seq:
+    if not (isGapChar(ref[i]) or isGapChar(ref[j])):
+      iso_delta += ISO[(ref[i],ref[j]),(a,b)]-ISO[(ref[i],ref[j]),(seq[i],seq[j])]
+      nb_correct += 1
+  if nb_correct>0:
+    iso = mpf(iso_delta)/nb_correct
+  else:
+    iso = mpf(0.)
   return  math.exp(-((1-alpha)*iso)/(BOLTZMANN*T))
 
 @memoize

### 2013-05-23 scripts/pyroc.py
ROC curve script

@@ -1,0 +1,315 @@
+#!/usr/bin/env python
+# encoding: utf-8
+"""
+PyRoc.py
+
+Created by Marcel Caraciolo on 2009-11-16.
+Copyright (c) 2009 Federal University of Pernambuco. All rights reserved.
+
+Python Module for calculating the area under the receive operating characteristic curve, given a dataset.
+
+0.1  - First Release
+
+"""
+
+import random
+
+try:
+	import pylab
+except:
+	print "error:\tcan't import pylab module, you must install the module:\n"
+	print "\tmatplotlib to plot charts!'\n"
+
+
+def random_mixture_model(pos_mu=.6,pos_sigma=.1,neg_mu=.4,neg_sigma=.1,size=200):
+	pos = [(1,random.gauss(pos_mu,pos_sigma),) for x in xrange(size/2)]
+	neg = [(0,random.gauss(neg_mu,neg_sigma),) for x in xrange(size/2)]
+	return pos+neg
+
+
+def plot_multiple_rocs_separate(rocList,title='', labels = None, equal_aspect = True):
+	""" Plot multiples ROC curves as separate at the same painting area. """
+	pylab.clf()
+	pylab.title(title)
+	for ix, r in enumerate(rocList):
+		ax = pylab.subplot(4,4,ix+1)
+		pylab.ylim((0,1))
+		pylab.xlim((0,1))
+		ax.set_yticklabels([])
+		ax.set_xticklabels([])
+		if equal_aspect:
+			cax = pylab.gca()
+			cax.set_aspect('equal')
+		
+		if not labels:
+			labels = ['' for x in rocList]
+		
+		pylab.text(0.2,0.1,labels[ix],fontsize=8)
+		pylab.plot([x[0] for x in r.derived_points],[y[1] for y in r.derived_points], 'r-',linewidth=2)
+	
+	pylab.show()
+	
+
+
+def _remove_duplicate_styles(rocList):
+ 	""" Checks for duplicate linestyles and replaces duplicates with a random one."""
+	pref_styles = ['cx-','mx-','yx-','gx-','bx-','rx-']
+	points = 'ov^>+xd'
+	colors = 'bgrcmy'
+	lines = ['-','-.',':']
+	
+	rand_ls = []
+	
+	for r in rocList:
+		if r.linestyle not in rand_ls:
+			rand_ls.append(r.linestyle)
+		else:
+			while True:
+				if len(pref_styles) > 0:
[…]

### 2013-05-23 scripts/art_main.py
benchmarks 16

@@ -1,6 +1,7 @@
 import os
 import sys
 import shlex
+from random import choice
 from subprocess import call
 from tempfile import mkdtemp, NamedTemporaryFile as NTF 
 from shutil import rmtree
@@ -137,7 +138,21 @@ def get_max_vote(d_merge_probs,ss):
     return ''.join(seq)
 
 
-def main(seq,l_reads,fold,ss):
+def get_most_reads_nt(d):
+    listed_keys = sorted(d.keys(), key=lambda x:len(d[x]))
+    return listed_keys[-1]
+
+
+def get_most_reads_seq(d_merge_probs, seq):
+    new_seq = []
+    for i,x in enumerate(seq):
+        if i in d_merge_probs['probs']:
+            new_seq.append(get_most_reads_nt(d_merge_probs['probs'][i]))
+        else:
+            new_seq.append(choice('ACGU'))
+    return ''.join(new_seq)
+
+def main(seq,l_reads,fold):
     seq = seq.upper().replace('U','T')
     if any(x not in ('ACGT') for x in seq):
         print help()
@@ -148,9 +163,8 @@ def main(seq,l_reads,fold,ss):
     d_clean_align = clean_align(raw_align,l_seq)
     d_nt_probs = get_nt_probs(raw_reads)
     d_merge_probs = merge_reads(d_clean_align,d_nt_probs)
-    max_vote_seq = get_max_vote(d_merge_probs,ss)
-    print seq
-    print max_vote_seq
+    new_seq = get_most_reads_seq(d_merge_probs, seq)
+    print new_seq
 
 
 def help():
@@ -180,7 +194,6 @@ if __name__ == '__main__':
     opts = sys.argv
     
     seq = ''
-    ss = ''
     l_reads = 0
     fold = 0
 
@@ -191,13 +204,11 @@ if __name__ == '__main__':
             seq = opts[i+1].upper().replace('U','T')
         elif x == '-f':
             fold = int(opts[i+1])
-        elif x == '-ss':
-            ss = opts[i+1]
 
-    if not seq or not l_reads or not fold or not ss:
+    if not seq or not l_reads or not fold:
         print seq, l_reads,fold
         help()
         sys.exit(1)
 
-    main(seq,l_reads,fold,ss)
+    main(seq,l_reads,fold)
     

### 2013-05-23 src/RNAPyro.py

@@ -24,10 +24,11 @@
 #(INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS#
 #SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE                  #
 ###############################################################################
-fimport os,sys
+import os,sys
 import itertools
 import random
 import math
+import time
 
 def MPMATH_MISSING():
   print """The module `mpmath` was not found. This might impeed the
@@ -692,7 +693,7 @@ def parse_fasta(file_name):
       line = line.strip()
       if not line:
         continue
-      if all(x in 'AUGC' for x in line):
+      if all(x in 'AUGC.' for x in line):
         seq.append(line)
         continue
       if all(x in '(.)' for x in line):
@@ -828,8 +829,10 @@ if __name__ == "__main__":
 
   n = len(seq)
   if not f_no_profile:
+    tic = time.time()
     results = all_probabilities(seq,ref_seq,struct,m,alpha)
     display_all_probabilities(results)
+    print time.time() - tic
   else:
     forward(seq,ref_seq,struct,(0,n-1),('X', 'X'),m,alpha)
 

### 2013-05-23 JCB/methods_JCB.tex

@@
 Let us remark that any of the individual contributions can be associated to (a subset of) the base-pairs occurring in the structure, possibly complemented, in the case of stacking pairs, with the knowledge of flanking base-pairing nucleotide.
-Therefore, let us denote by $\EBP{i,j}{xy}{a'b'}$ the local contribution of a base-pair $(i,j)$ of nucleotides $(a',b')$, surrounded by a stacking pair $(x,y)$ (or $\varnothing$ otherwise), to the pseudo-energy:
+Dropping the implicit dependency on $\Omega$ and $\beta$, let us denote by $\EBP{i,j}{xy}{a'b'}$ the local contribution of a base-pair $(i,j)$ of nucleotides $(a',b')$, surrounded by a stacking pair $(x,y)$ (or $\varnothing$ otherwise), to the pseudo-energy:
 \begin{equation}
@@
 knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$), within 
-$m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
+$m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec_A}-\eqref{eq:Y_rec_D},
  computes the \emph{outside} algorithm,   
@@
 knowing  that flanking inner positions $(i+1,j-1)$ contain nucleotides $a$ and $b$ respectively.
-A suitable multiplicative combination of these terms, computed as shown in Equation~\eqref{eq:combine}, gives the total weight of every possible sequences that support a given base-pair (or unpaired position) and, in turn, the probability of seeing a specific base at a given position.
+A suitable combination of these terms, computed as shown in Equations~\eqref{eq:combine_A}-\eqref{eq:combine_C}, gives the total weight of every possible sequences that support a given base-pair (or unpaired position) and, in turn, the probability of seeing a specific base at a given position.
 
@@
 \caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
-therefore its efficient implementation requires the precomputation and storage of the inside contributions.}
+therefore its efficient implementation requires the precomputation and storage of the inside contributions.\label{fig:outside}}
 \end{figure}
-In general, the main recurrence below works by extending the excluded structure $\Struct$ (covering the $]i,j[$ interval) in the leftward direction, and considers four cases depending on the base-pairing status and directionality of the next considered position:
+In general, the main recurrence below works by extending the excluded structure $\Struct$ (covering the $]i,j[$ interval) in the leftward direction.
+As shown in Figure~\ref{fig:outside}, four cases must be considered, depending on the base-pairing status and directionality of the next considered position:
 \begin{itemize}
@@
 \item {\bf Case 2: Stacking base-pair.} If both ends of the excluded structure are paired together, and are nested within another base-pair in the remaining structure, then an additional contribution for stacking has to be considered.
-The outside terms are then computed by simulating the choice of two base-pairing nucleotides, and by recurring on the remaining portion, as follows:
+The outside terms are then computed by simulating any pair of base-pairing nucleotides for $i,j$, and by proceeding recursively on the remaining portion, as follows:
 \begin{equation}
@@
 \begin{equation}
-	\Y{i,j}{m}{x,y} = 
+	\Y{\Struct}{m}{x,y} = 
 		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{ab,a'b'}\le m}}
@@
 \begin{equation}
-	\mathbb{P}(s_i = a\;|\; M) := \frac{\mathcal{W}^M_{\substack{i, [a]}}}{\sum_{m=0}^{M}\Z{\varepsilon}{m}{X,X}}\label{eq:normalize}
+	\mathbb{P}(s_i = a\;|\; M) := \frac{\mathcal{W}^M_{\substack{i, [a]}}}{\Z{\varepsilon}{\le M}{X,X'}} = \frac{\mathcal{W}^M_{\substack{i, [a]}}}{\sum_{m=0}^{M}\Z{\varepsilon}{m}{X,X'}}\label{eq:normalize}
 \end{equation}
+where $X,X'\in\B$ may be any nucleotides (no impact on energy/weights).
 
-To that purpose, we leverage the \emph{Inside-Outside} construction, and obtain the three following cases:
+To that purpose, one leverages the \emph{Inside-Outside} construction.
+Namely, while computing the total contribution of sequences featuring a base $a'\in\B$ at position $i$, one must consider the three following situations:
 \begin{itemize}
-\item 
-where $\mathcal{W}{\substack{\ast\\ \ast}}$ is defined by:
+\item {\bf Case 1: Unpaired position.} In this case, the supporting sequences are simply those excluding the structure $\ub$ at position $i$, where a base $c\in\B$ was formerly found.
+Summing over any possible number of mutations for the \emph{outside} region, one obtains:
 \begin{equation}
-\resizebox{\textwidth}{!}{%
-$ \mathcal{W}^M_{\substack{i, [x]}} =  \left\{
-	\begin{array}{ll}
+ \mathcal{W}^M_{\substack{i, [a]}} =  \sum_{m=\Kron_{a,c}}^{M}
+			\Y{\ub}{m-\Kron_{a,c}}{a,a}.
+\label{eq:combine_A}
+\end{equation}
+\item {\bf Case 2: Position is paired rightward.} Here, position $i$ is paired to some position $k>i$, whose content impacts the base-pair contribution.
+Any sequence having $a$ at position $i$ can be decomposed as a base-pair, nesting a sequence for the structure $\Struct'$ on the interval $[i+1,k-1]$, and an outside sequence, excluding the structure $\op\Struct'\cp$.
+Consequently, let $c$ and $d$ be the original nucleotides at positions $i$ and $k$, one has:
+\begin{equation}
+ \mathcal{W}^M_{\substack{i, [a]}} =  
 			\sum_{m=0}^{M}
-			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
-		&\text{If }S_i = -1\\
-			\displaystyle
+			\sum_{\substack{b\in \B\\\Kron_{ab,cd}\leq m}}
+			\sum_{m'=0}^{m-\Kron_{ab,cd}}
+     	 e^{\frac{-\EBP{i,k}{\varnothing}{ab}}{RT}}
+			\cdot\Y{\op\Struct'\cp}{m-\Kron_{ab,cd}}{a,b}
+			\cdot\Z{\Struct'}{m'}{a,b}
+\label{eq:combine_B}
+\end{equation}
+\item {\bf Case 3: Position is paired leftward.} This case is symmetrical to the previous one, with the exception that $k<i$.
+Consequently, let $c$ and $d$ be the original nucleotides at positions $i$ and $k$, one has:
+\begin{equation}
+ \mathcal{W}^M_{\substack{i, [a]}} =  
 			\sum_{m=0}^{M}
-			\sum_{\substack{b\in B\\\Kron_{xb,s_is_k\leq m}}}
-			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
-     	 e^{\frac{-E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
-			\cdot\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
-			\cdot\Z{i+1,k-1}{m'}{x,b}
-		&\text{If }S_i=k>i\\
-    \displaystyle
-			\sum_{m=0}^{M}
-			\sum_{\substack{b\in B\\\Kron_{bx,s_ks_i\leq m}}}
-			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
-     	 e^{\frac{-E^{\Omega, \beta}_{(k,i),\varnothing\to bx}}{RT}}
-			\cdot\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
-			\cdot\Z{k+1,i-1}{m'}{b,x}
-		&\text{If }S_i=k<i
-	\end{array}\right.
-$}\label{eq:combine}
+			\sum_{\substack{b\in \B\\\Kron_{ba,cd}\leq m}}
+			\sum_{m'=0}^{m-\Kron_{ba,cd}}
+     	 e^{\frac{-\EBP{k,i}{\varnothing}{ba}}{RT}}
+			\cdot\Y{\op\Struct'\cp}{m-\Kron_{ba,cd}}{b,a}
+			\cdot\Z{\Struct'}{m'}{b,a}
+\label{eq:combine_C}
 \end{equation}
@@
 
-\TODOYann{Rephrase as tree constructs}
-\TODOVlad{Add illustration of Inside/Outside combination (adapted from RECOMB presentation??)}
-In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, 
-for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
-\begin{description}
-\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, we are concerned by the weights
-over all sequences which have at position $i$ nucleotide $x$, which is exactly the sum
-of the values of $\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}$, for all $m$ between $0$ and $M$.
-\item[$S_i=k>i$:] Since we need to respect the derivation of the secondary structure $S$, if 
-position $i$ is paired, we must consider the two partition functions. The \emph{outside} of the 
-base pair, and the \emph{inside}, for all possible values for the nucleotide at position $k$, and
-all possible distribution of the mutant positions between the inside and outside of the base pair. We also add the term of isostericity for this specific base pair.
-\item[$S_i=k<i$:] Same as above, but with position $i$ pairing with a lower position.
-\end{description}
-
-By construction, the partition function over all sequences at exactly $m$ mutations of a reference sequence $s$ can 
-be either described from the \emph{inside} contribution $\Z{\Struct}{m}{X,X}$ of the whole sequence,
-$\forall X\in B$, or from a subset of the \emph{outside} terms, such that:
-$$
-	\Z{\Struct}{m}{X,X}
-	\equiv
-	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
-	\Y{\ub}{m-\Kron_{a,s[k]}}{a,a},\forall k \text{	unpaired.}
-$$
-
 \subsection{Complexity Considerations}
-Equations~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C} and~\eqref{eq:Y_rec_A}-\eqref{eq:Y_rec_D} can be computed using dynamic programming.
-Namely, the $\mathcal{Z}^{*}_{*}$ and $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and interval lengths, memorizing the results as they become available to ensure constant-time access during later stages of the computation.
-Furthermore, energy terms $E(\cdot)$ can be accessed in constant time thanks to a simple precomputation (not described)  of the isostericity contributions in $\Theta(n\cdot|\Omega|)$.
-Computing any given term therefore requires $\Theta(m)$ operations.
+Clearly, Equations~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C} and~\eqref{eq:Y_rec_A}-\eqref{eq:Y_rec_D} can be computed using dynamic programming.
+Namely, the $\mathcal{Z}^{*}_{*}$ and $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and structure lengths, storing the results as they become available to ensure constant-time access during later stages of the computation.
+Furthermore, energy terms $E(\cdot)$ may be accessed in constant time after a simple precomputation of the isostericity contributions in $\Theta(n\cdot|\Omega|)$.
+Computing any given term therefore requires $\Theta(m)$ operations due to the explicit distribution of the number of mutations.
 

### 2013-05-23 JCB/methods_JCB.tex
fd

@@
 \begin{equation}
-	\forall x,y\in B\times \B, m\in [0,M]:\, \Z{\varepsilon}{m}{x,y}=\left\{
+	\forall x,y\in \B\times \B, m\in [0,M]:\, \Z{\varepsilon}{m}{x,y}=\left\{
 	\begin{array}{ll}
@@
 contributions of subsequences excluding a given structure/interval $\Struct$, occupying the open interval $]i,j[$ in the sequence, at Hamming distance exactly $m$ to the initial sequence $s$, and assuming that nucleotides $x$ and $y$ were previously chosen for $i+1$ and $j-1$, the outermost portions of the excluded structure.
-\TODOYann{Reformat outside Figure as in Incarnation.}
 The associated terms $\Y{\Struct}{m}{x,y}$ can then be computed recursively, initially considering the case of any prefix $\Struct'$ of the complete structure:
@@
 \end{equation}
-\item {\bf Case 2: Stacking base-pair.} If both ends of the excluded structure are paired together, and are nested within another base-pair in the remaining structure, then an additional contribution for stacking has to be considered.
+\item {\bf Case 2: Stacking base-pair.} If both ends of the excluded structure are paired together, and are nested within another base-pair in the remaining structure, then an additional contribution, stemming from a stacking pairs energy, has to be considered.
 The outside terms are then computed by simulating any pair of base-pairing nucleotides for $i,j$, and by proceeding recursively on the remaining portion, as follows:

### 2013-05-23 16S/extract_gapless.py
bench

@@ -1,0 +1,199 @@
+import os
+import sys
+import cPickle
+import art_main
+from shutil import rmtree
+from tempfile import NamedTemporaryFile as NTF, mkdtemp
+from subprocess import check_output
+from parse_stockholm import get_data, l_basepairs
+from pprint import pprint
+from random import choice
+from numpy import mean
+
+DATA = 'RF00177_seed.stockholm.txt'
+NAME = 'RF00177'
+IUPACBASES = {
+  'A':['A'],
+  'C':['C'],
+  'G':['G'],
+  'U':['U'],
+  'N':['A','C','G','U'],
+  'R':['A','G'],
+  'Y':['C','U'],
+  'S':['G','C'],
+  'W':['A','U'],
+  'K':['G','U'],
+  'M':['A','C'],
+  'B':['C','G','U'],
+  'D':['A','G','U'],
+  'H':['A','C','U'],
+  'V':['A','C','G']
+  }
+
+def ss_from_l_bp(l_bp,s):
+    l = set(x[0] for x in l_bp)
+    r = set(x[1] for x in l_bp)
+    ss = []
+    for i in range(len(s)):
+        if s[i] == '.':
+            continue
+        if i in l:
+            ss.append('(')
+        elif i in r:
+            ss.append(')')
+        else:
+            ss.append('.')
+    return ''.join(ss)
+
+def remove_seq_gap_ss(l_s,ss):
+    l_ss = l_basepairs(ss)
+    l_s_1 = []
+    for s in l_s:
+        for x,y in l_ss:
+            if s[x] == '.' or s[y] == '.':
+                break
+        else:
+            l_s_1.append(s)
+
+    return l_s_1
+
+def cluster_by_seq_same_gaps(l_s):
+    l_c = []
+    s_done = set()
+    for i,s in enumerate(l_s):
+        if s in s_done:
+            continue
+        l_c.append([s])
+        s_done.add(s)
+        for s2 in l_s[i+1:]:
+            if (all(s2[j] == '.' for j,nt in enumerate(s) 
+                     if nt == '.') and 
+                    all(s[j] == '.' for j,nt in enumerate(s2)
+                     if nt == '.')):
+                l_c[-1].append(s2)
+                s_done.add(s2)
+    return l_c
+
+def rand_iupac(x):
+    return choice(IUPACBASES[x])
+
+def cluster_remove_ref_gaps(l_s,ss):
+    l_ss = l_basepairs(ss)
[…]

### 2013-05-23 scripts/art_main.py

@@ -1,214 +1,0 @@
-import os
-import sys
-import shlex
-from random import choice
-from subprocess import call
-from tempfile import mkdtemp, NamedTemporaryFile as NTF 
-from shutil import rmtree
-from pprint import pprint
-from itertools import combinations
-
-
-def get_raw_data(seq,l_reads,fold):
-    dir_temp = mkdtemp(dir='.')
-    cwd = os.getcwd()
-    os.chdir(dir_temp)
-
-    temp_file = NTF(dir='.',delete=False,suffix='.fasta')
-    temp_file.file.write('>seq\n%s' % seq)
-    temp_file.file.close()
-
-    try:
-        cmd = 'art_illumina -i %s -l %s -f %s -o out' % (
-                temp_file.name,
-                l_reads,
-                fold)
-        cmd  = shlex.split(cmd)
-        call(cmd)
-    except:
-        os.chdir(cwd)
-        rmtree(dir_temp)
-        print "Error with command:"
-        print '\t', cmd
-        sys.exit(1)
-
-    raw_reads = [x.strip() for x in open('out.fq')]
-    raw_align = [x.strip() for x in open('out.aln')]
-
-    os.chdir(cwd)
-    rmtree(dir_temp)
-
-    return raw_reads, raw_align
-
-def reverse_sens(seq):
-    dic = {'A':'U','U':'A','C':'G','G':'C'}
-    return ''.join(dic[x] for x in reversed(seq))
-
-def clean_align(raw_align,l_seq):
-    l = l_seq
-    d_clean = {}
-
-    for i,x in enumerate(raw_align):
-        if x.startswith('>'):
-            _, name, start, orientation = x.split()
-            start = int(start)
-            d_clean[name] = {}
-
-            if orientation == '+':
-                d_clean[name]['ref'] = raw_align[i+1].replace('T','U')
-                d_clean[name]['seq'] = raw_align[i+2].replace('T','U')
-                d_clean[name]['start'] = start
-            else:
-                d_clean[name]['ref'] = reverse_sens(raw_align[i+1].replace('T','U'))
-                d_clean[name]['seq'] = reverse_sens(raw_align[i+2].replace('T','U'))
-                d_clean[name]['start'] = l - start - len(d_clean[name]['seq'])
-
-    return d_clean
-
-def check_align(wt,ref,seq,start):
-    wt = wt[start:start+len(ref)]
-    if wt != ref:
-        return False
-    elif '.' in ref+seq:
[…]

### 2013-05-23 JCB/acknowledgments_JCB.tex

%!TEX root = main_RNAPyro_JCB.tex
\section{Acknowledgments}
\label{sec:acknowledgments}
The authors would like to thank Rob Knight for his suggestions and comments.
This work was funded by the French Agence Nationale de la Recherche (ANR) through the {\sc Magnum} {\tt ANR 2010 BLAN 0204} 
project (to YP), the
FQRNT team grant 232983 (to VR and JW) and NSERC Discovery grant 219671 (to JW).

### 2013-05-23 JCB/conclusion_JCB.tex
Initial commit JCB extended version

--- 2012-10-08 Recomb/conclusion_RECOMB.tex
@@
-%!TEX root = main_RECOMB.tex
+%!TEX root = main_RNAPyro_JCB.tex
 \section{Conclusion}
@@
 
+\TODOJerome{Etoffer la conclusion}
 In this article we presented a new and efficient way of exploring the mutational landscape of an RNA under structural constraints,
@@
 isostericity seems to hold a higher accuracy on the estimation of errors.
-This indicates that an extractable signal is contained in the isostericity.
+This indicates that an exploitable signal is captured by the isostericity.
 Importantly, the implementation is fast enough for practical applications.

### 2013-05-23 JCB/main_RNAPyro_JCB.tex
Initial commit JCB extended version
Started restructuring figures
New inside/outside figures
Almost there on outside...
Finished outside

--- 2013-01-09 Recomb/main_RECOMB.tex
@@
 \documentclass{llncs}
-%\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
+\usepackage[top=1.5in, bottom=1.5in,right=1.5in,left=1.5in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
 %\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
@@
 %\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
+%\usepackage{jcompbio}
 \usepackage{graphicx}
@@
 \usepackage{relsize}
+\usepackage{bbm}
 \usepackage[numbers]{natbib}
-\usepackage[colorinlistoftodos]{todonotes}
+\usepackage[colorinlistoftodos,textsize=small]{todonotes}
+
+\makeatletter
+\renewcommand\bibsection%
+{
+  \section*{\refname
+    \@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}
+}
+\makeatother
 
@@
 \usetikzlibrary{snakes}
+\usetikzlibrary{shadows}
 \pgfdeclarelayer{background}
@@
 
-\newcommand{\ourprog}{\texttt{RNA-MoIP}\xspace}
-
-\usepackage[applemac]{inputenc} %for the encoding
-
-\newcommand{\RNAmutants}{\texttt{RNAmutants}\xspace}
-\newcommand{\RNApyro}{\texttt{RNApyro}\xspace}
-
-\newcommand{\red}[1]{{\color{red}#1}}
-\newcommand{\farna}{\texttt{FARNA}\xspace}
-\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
-\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
-\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
-\newcommand{\nast}{\texttt{NAST}\xspace}
-\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
-\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
-\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
-\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
-\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
-\newcommand{\contrafold}{\texttt{contrafold}\xspace}
-\newcommand{\unafold}{\texttt{unafold}\xspace}
-\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
-\newcommand{\assemble}{\texttt{assemble}\xspace}
-\newcommand{\fred}{\texttt{FR3D}\xspace}
-\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
-\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
-\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
-\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
-\newcommand{\contextfold}{\texttt{contextfold}\xspace}
-%\newcommand{\citep}{\cite}
-\newcommand{\Z}[3]{\mathcal{Z}_{\substack{(#1)\\ [#3]}}^{#2}}
-\newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
-\newcommand{\B}{\mathcal{B}}
-\newcommand{\Kron}{\delta}
-\newcommand{\ub}{\bullet}
-
-\newcommand{\Ab}{{\sf{A}}}
-\newcommand{\Cb}{{\sf{C}}}
-\newcommand{\Gb}{{\sf{G}}}
-\newcommand{\Ub}{{\sf{U}}}
+\input{macros.tex}
 
 %\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
-\title{A Linear Inside-Outside Algorithm for Correcting Sequencing Errors in Structured RNA}
+\title{Using structural and evolutionary information to correct RNA pyrosequencing errors in non-coding RNAs}
 \author{Vladimir Reinharz\inst{1}, Yann Ponty\inst{2}$^{*}$ \and J\'er\^{o}me Waldisp\"{u}hl\inst{1}$^*$}
@@
 \begin{document}
+
+\ShowTODO{\setcounter{tocdepth}{1}
+\listoftodos} 
 \maketitle
@@
 
-\newpage
+%\newpage
 
-\input{introduction_RECOMB.tex}
+\input{introduction_JCB.tex}
 
-\input{methods_RECOMB.tex}
+\input{methods_JCB.tex}
 
-\input{results_RECOMB.tex}
+\input{results_JCB.tex}
 
-\input{conclusion_RECOMB.tex}
+\input{conclusion_JCB.tex}
 
-\input{acknowledgments_RECOMB.tex}
+\input{acknowledgments_JCB.tex}
+%\newpage
 
-\newpage
-
-\bibliographystyle{plainnat}
+\bibliographystyle{jcompbiol}
 \bibliography{RNApyro}

### 2013-05-23 JCB/results_JCB.tex
Initial commit JCB extended version
Added a few 'todo' notes

--- 2013-01-16 Recomb/results_RECOMB.tex
@@
-%!TEX root = main_RECOMB.tex
+%!TEX root = main_RNAPyro_JCB.tex
 \section{Results}
@@
  both the sequences and the target secondary structure were generated at random.
+\TODOVlad{Add more extensive time benchmark, as a plot}
 
@@
 
-\subsection{Error Correction in 5s rRNA}
+\subsection{Error Correction in 5s rRNAs}
 
@@
 by \RNApyro. More specifically, we fixed a threshold $\lambda \in [0,1]$, and predicted an error at position $i$ in sequence $\omega$ if and only if the
-probability $P(i,x)$ of a nucleotide $x \in \{ A,C,G,U \}$ exceeds this threshold. To correct the errors we used the set of nucleotides having probability
-greated than $\lambda$, that is  $C(i) = \{ x \; | \;  x \in \{ A,C,G,U \} \mbox{ and } P(i,x) > \lambda \mbox{ and }  n \neq \omega[i] \},$ where $\omega[i]$ is
+probability $P(i,x)$ of a nucleotide $x$ exceeds this threshold. To correct the errors we used the set of nucleotides having probability
+greater than $\lambda$, that is  
+$$C(i) = \{ x \; | \;  x \in \{ \Ab,\Cb,\Gb,\Ub \}, P(i,x) > \lambda \mbox{ and }  n \neq \omega[i] \},$$
+ where $\omega[i]$ is
 the nucleotide at position $i$ in the input sequence. We note that, for lower thresholds, multiple nucleotides may be available in $C(i)$ to correct

### 2013-05-24 JCB/methods_JCB.tex
Stated problem more explicitly

@@
 \subsubsection{Isostericity Contribution.}
-The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
+The concept of isostericity score is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
 Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the isostericity contribution to the pseudo-energy as
@@
 Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
-We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $M$ mutations from $s$
-%  under the SCFG derivation $S$ 
-(formally $\mathbb{P}(s_i = x\mid s,\Omega, \Struct,M)$). 
-We define a variant of the
- \emph{Inside-Outside algorithm}~\cite{Lari1990} to compute this
- probability, based on two sets of dynamic programming equations. 
+We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $M$ mutations from $s$.
+Formally, let $\mathcal{D}_{s,M}=\{s'\;|\;\delta(s,s')\le M\}$ be the set of admissible sequences, one aims at computing the probabilities
+\begin{equation}
+\mathbb{P}(s_i = x\mid s,\Omega, \Struct,M) = \frac{\sum_{\substack{s'\in\mathcal{D}_{s,M}\\\text{s.t. }s'_{i}=x}}B(s')}{\sum_{\substack{s''\in\mathcal{D}_{s,M}}}B(s'')}
+\end{equation}
+
+Clearly, the number of sequences in $\mathcal{D}_{s,M}$ grows exponentially with the sequence length, therefore one cannot realistically rely on an exhaustive enumeration to compute the mutational profile.
+To work around this issue, we propose a linear-time variant of the
+ \emph{Inside-Outside algorithm}~\cite{Lari1990} to compute this probability, based on two sets of dynamic programming equations. 
 %from the two  functions $\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$. 
-
 The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C}, is analogous to an \emph{inside} computation.
@@
 \subsection{Complexity Considerations}
-Clearly, Equations~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C} and~\eqref{eq:Y_rec_A}-\eqref{eq:Y_rec_D} can be computed using dynamic programming.
+Using dynamic programming, Equations~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C} and~\eqref{eq:Y_rec_A}-\eqref{eq:Y_rec_D} can be computed in linear time and memory.
 Namely, the $\mathcal{Z}^{*}_{*}$ and $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and structure lengths, storing the results as they become available to ensure constant-time access during later stages of the computation.
@@
 For instance, the inside algorithm only requires computing intervals $[i,j]$ that do not break any base-pair, and whose next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
-Similar constraints hold for the outside computation, resulting in a drastic limitation of the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored.
-Consequently the overall complexity of the algorithm is $\Theta(n\cdot(|\Omega|+m^2))$ arithmetic operations and $\Theta(n\cdot(|\Omega|+m))$ memory.
+Similar constraints hold for the outside computation,  following from the linearity of the outside recurrence (i.e.
+the computation of the outside term for a given excluded structure only relies on the computation of another, strictly larger, structure).
+
+These properties drastically limit the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored.
+Consequently the overall complexity of the algorithm is $\Theta(n\cdot(|\Omega|+M^2))$ arithmetic operations and $\Theta(n\cdot(|\Omega|+M))$ memory.

### 2013-05-24 JCB/methods_JCB.tex
Typos

@@
 %from the two  functions $\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.
-The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C}, is analogous to an \emph{inside} computation.
-Considering an interval of the sequence/secondary structure, it consists in the accumulated contributions of any possible sequences that have suitable Hamming distance within the interval.
+
+The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C}, is analogous to an \emph{inside} computation: Considering a given substructure of the input structure, it computes the accumulated contributions of any possible sequences that have suitable Hamming distance within the interval.
 It is therefore similar to a partition function, i.e.
@@
 knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$), within 
-$m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec_A}-\eqref{eq:Y_rec_D},
+$m$ mutations of $s$.
+
+The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec_A}-\eqref{eq:Y_rec_D},
  computes the \emph{outside} algorithm,   

### 2013-05-24 16S/extract_gapless.py
first bench

@@ -1,15 +1,17 @@
 import os
 import sys
 import cPickle
-import art_main
+#import art_main
 from shutil import rmtree
 from tempfile import NamedTemporaryFile as NTF, mkdtemp
 from subprocess import check_output
-from parse_stockholm import get_data, l_basepairs
+from multiprocessing import Pool
+#from parse_stockholm import get_data, l_basepairs
 from pprint import pprint
-from random import choice
+from random import choice, shuffle
 from numpy import mean
 
+PROCS = 18
 DATA = 'RF00177_seed.stockholm.txt'
 NAME = 'RF00177'
 IUPACBASES = {
@@ -29,6 +31,8 @@ IUPACBASES = {
   'H':['A','C','U'],
   'V':['A','C','G']
   }
+
+d_s_trim = {}
 
 def ss_from_l_bp(l_bp,s):
     l = set(x[0] for x in l_bp)
@@ -105,15 +109,15 @@ def get_RNAPyro_out(seq,ref_seq,struct,m,alpha):
     tmp_in.close()
 
     
-    cmd = ['python','/Users/vreinharz/Projects/RNApyro/src/RNAPyro.py',
-           '-f', tmp_in.name, '-m',str(m),'-a',str(alpha)]
+    cmd = ['python','/scratch/Vreinh/Applications/RNApyro/src/RNAPyro.py',
+           '-f', tmp_in.name, '-m',str(m),'-a',str(alpha), '-p', '15']
     print ' '.join(cmd)
 
     try:
         out = check_output(cmd)
     except:
         rmtree(tmp_dir)
-        print "Error calling RNAPyro"
+        print "Error calling RNAPyro", cmd
         sys.exit(1)
 
     rmtree(tmp_dir)
@@ -142,6 +146,74 @@ def do_benchmarks(d_s_trim):
         alpha = 0.5
         prob, time = get_RNAPyro_out(seq,ref_seq,struct,m,alpha)
 
+def l_basepairs(ss):
+    l_bp = []
+    l = []
+    for i, x in enumerate(ss):
+        if x == '(':
+            l.append(i)
+        elif x == ')':
+            l_bp.append((l.pop(), i))
+    return l_bp
+
+def correct_ss(ss, seq):
+
+    l_bp = l_basepairs(ss)
+    new_l_bp = []
+    for l,r in l_bp:
+        if (seq[l],seq[r]) in (('A', 'U'), ('U', 'A'),
[…]

### 2013-05-24 benchmark/bench_weight.py

@@ -3,12 +3,25 @@ import os
 import subprocess
 import cPickle
 
+def  fix_bad_substitution(struct,l_rnas,data_2):
+  new_data_set = []
+  for i,rna in enumerate(l_rnas):
+    rnainv = data_2[i]
+    new_rna = []
+    for j,x in enumerate(rna):
+      if struct[j] == '.':
+        new_rna.append(rnainv[j])
+      else:
+        new_rna.append(x)
+    new_data_set.append(''.join(new_rna))
+  return new_data_set
+
 def benchmark(list_sec_struc,weights,nb,output_cPickle=None):
   """f
   Will output a dictionary with keys (sec_struct,gc_content_target)
   and value a tuple (res,data_dump)
     res = list of sequences with +- 10% of GC target
-    data_dump = list alterning 'C' weight in profile,
+    data_dump = list updating 'C' weight in profile,
                 list of GC contents for sequences in given round
 
   If a 'output_cPickle' arguement is given, will save the dic in it
@@ -39,6 +52,36 @@ def benchmark(list_sec_struc,weights,nb,output_cPickle=None):
   os.remove(tmp_struct_file)
   return d_all_bench
 
+def format_rna(struct,rna):
+  new_rna = []
+  for i,x in enumerate(rna):
+    if not struct[i] == '.':
+      new_rna.append(x.lower())
+    else:
+      new_rna.append(x)
+  return ''.join(new_rna)
+
+def rnainverse(struct,l_rnas):
+  l_rnas = l_rnas.split()
+  tmp_rnainverse = 'tmp_file_rnainverse.tmp'
+  with open(tmp_rnainverse,'w') as f:
+    for r in l_rnas[:-1]:
+      r = format_rna(struct,r)
+      f.write('%s\n%s\n' % (struct,r)) 
+    f.write('%s\n%s' % (struct,l_rnas[-1])) 
+  to_do = 'RNAinverse < %s' % tmp_rnainverse
+  data = subprocess.Popen(to_do,shell=True,stdout=subprocess.PIPE)
+  data = data.communicate()[0]
+  data = [x for x in data.split('\n') if x]
+  data_2 = []
+  for x in data:
+    y = x.split()[0]
+    if all(z in 'ACGU' for z in y.upper()):
+      data_2.append(y.upper())
+  os.remove(tmp_rnainverse)
+  fix_bad_substitution(struct,l_rnas,data_2)
[…]

### 2013-05-24 JCB/conclusion_JCB.tex

@@
 In this article we presented a new and efficient way of exploring the mutational landscape of an RNA under structural constraints,
-and apply our techniques to identify and fix sequencing errors. In addition, we introduce a new
-scoring scheme combining the nearest-neighbour energy model to new isostericity matrices in order to account for geometrical
-discrepancies occurring during base pair replacements.
-The algorithm runs in  $\Theta(n\cdot(|\Omega|+m^2))$ time and $\Theta(n\cdot(|\Omega|+m))$ memory, where $n$ is the length of the RNA,
-$m$ the number of mutations and $\Omega$ the size of the multiple sequence alignment.
+and apply our techniques to identify and fix sequencing errors. In addition, we introduced a new scoring scheme to measure the
+likelihood of sequencing errors that combines the classical nearest-neighbour energy model parameters \cite{Turner2010} to the recently introduced
+isostericity matrices \cite{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring during base pair replacements.
+Importantly, our algorithm runs in  $\Theta(n\cdot(|\Omega|+m^2))$ time and $\Theta(n\cdot(|\Omega|+m))$ memory, where $n$ is the length of the RNA,
+$m$ the number of mutations and $\Omega$ the size of the multiple sequence alignment. This achievement enables us to envision applications to high-throughput sequencing pipe-lines.
 
-By combining  into \texttt{RNApyro} these two approaches,  the 
+By combining these two approaches into \texttt{RNApyro} ,  the 
 mutational landscape exploration and the pseudo energy model,

### 2013-05-24 scripts/createROC.py
Computed Delta, still a bug as needs to compare

@@ -4,6 +4,7 @@
 
 from pyroc import *
 import pylab
+import os
 
 def loadMutationalProfiles(path):
 	fileHandler = open(path,'r')
@@ -12,6 +13,11 @@ def loadMutationalProfiles(path):
 	struct = lines[1]
 	nbMut = int(lines[2])
 	profiles = eval(lines[3])
+	for alpha in profiles:
+		profile = profiles[alpha]
+		for i in range(len(profile)):
+			for j in range(len(profile[i])):
+				profile[i][j] = float(profile[i][j])
 	fileHandler.close()
 	return seq,struct,nbMut,profiles
 
@@ -53,46 +59,65 @@ def classifyProfile(profile,seq,struct,mode):
 				if b.lower() == bestB.lower():
 					result.append((cl,prob))
 				else:
-					result.append((cl,2.0))
-	print result
+					result.append((cl,0.0))
+	#print result
 	return result
+
+def evaluatePrediction(profile,seq,nbMut):
+	ok = 0.
+	for i in range(len(profile)):
+		bestB,probB = "",0.
+		for j in range(len(profile[i])):
+			prob = profile[i][j]
+			b = BASES[j]
+			if prob > probB:
+				bestB,probB = b,prob
+		if bestB==seq[i]:
+			ok += 1
+	return ok-(len(seq)-nbMut)
 
 def computePlotROCs(path,mode):
 	seq,struc,nbMut,profiles = loadMutationalProfiles(path)
 	rocs = []
 	lbls = []
 	pylab.rc('text', usetex=True)
-	for alpha in profiles:
+	alphas = profiles.keys()
+	alphas.sort()
+	for alpha in alphas:
 		c = classifyProfile(profiles[alpha],seq,struc,mode)
+		delta = evaluatePrediction(profiles[alpha],seq,nbMut)
 		roc = ROCData(c)
 		rocs.append(roc)
 		lbls.append("$\\alpha=%.1f,$ {\\sf AUC} $=%.1f\\%%$"%(alpha,(100.*roc.auc(0))))
-		print "Alpha: %s, ROC AUC: %s" % (alpha,str(roc.auc(0)))
+		print "File:%s, Mode:%s, Alpha:%s, ROC AUC:%.2f, Delta:%s" % (path.split(os.sep)[-1],mode,alpha,(100.*roc.auc(0)),delta)
 	plt = plot_multiple_roc(rocs,"", lbls)
 	return plt
 
 if __name__ == '__main__':
[…]

### 2013-05-24 scripts/jcb_graphs.py
auto graphs

@@ -1,0 +1,30 @@
+import os
+import sys
+from subprocess import call
+
+def main(m='a'):
+    with open(os.path.join('..','16S','key_to_Bench.txt')) as f:
+        key = eval(f.readline())
+    d_key = {x[0]:x[1:] for x in key}
+
+    for f_name in os.listdir(os.path.join('..','16S','Bench')):
+        id = int(f_name.split('.')[0])
+        fold = d_key[id][-1]
+        try:
+            os.mkdir('Figs_Bench')
+        except OSError:
+            pass
+        cmd = ['python',
+               'createROC.py',
+               '-f', '%s' % os.path.join('..','16S','Bench',f_name),
+               '-o', '%s' % os.path.join('Figs_Bench','%sid_%sfold.pdf' % (id, fold)),
+               '-m', '%s' % m]
+        print cmd
+        call(cmd)
+
+if __name__ == '__main__':
+    m = 'a'
+    for i, opt in enumerate(sys.argv):
+        if opt == '-m':
+            m = sys.argv[i + 1]
+    main(m)

### 2013-05-24 src/RNAPyro.py

@@ -804,7 +804,7 @@ def help(file_name=False,
       -a <float>    The weight given to match the MSE or the secondary
                     structure, between [0,1]. A weight of 1 will only
                     take into account the secondary structure, 0 only
-                    the MSE
+                    the MSA
     Optional:
       -p <float>    The penality for a non canonical base pair (default
                     sys.maxint)

### 2013-05-24 JCB/conclusion_JCB.tex

@@
 isostericity matrices \cite{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring during base pair replacements.
-Importantly, our algorithm runs in  $\Theta(n\cdot(|\Omega|+m^2))$ time and $\Theta(n\cdot(|\Omega|+m))$ memory, where $n$ is the length of the RNA,
-$m$ the number of mutations and $\Omega$ the size of the multiple sequence alignment. This achievement enables us to envision applications to high-throughput sequencing pipe-lines.
+Importantly, our algorithm runs in  $\Theta(n\cdot(|\Omega|+M^2))$ time and $\Theta(n\cdot(|\Omega|+M))$ memory, where $n$ is the length of the RNA,
+$M$ the number of mutations and $\Omega$ the size of the multiple sequence alignment. This achievement enables us to envision applications to high-throughput sequencing pipe-lines.
 

### 2013-05-24 JCB/introduction_JCB.tex

@@
 
-Here, we expand the range of algorithmic techniques previously introduced with \RNAmutants~\cite{Waldispuhl2008}.
+Here, we expand the range of algorithmic techniques previously introduced with the \RNAmutants software~\cite{Waldispuhl2008,Waldispuhl2011}.
 Instead of exploring the full conformational landscape and sample mutants, we develop an inside-outside algorithm that enables us
 to explore the complete mutational landscape with a \emph{fixed} secondary structure and to calculate exactly mutational probability values. In addition
-to a gain into the numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot m^2)$ for the
-original version of  \RNAmutants to $\mathcal{O}(n \cdot m^2)$ for \RNApyro, where $n$ is the size of the sequence and $m$ the number of mutations).
+to a gain into the numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot M^2)$ for the
+original version of  \RNAmutants to $\mathcal{O}(n \cdot M^2)$ for \RNApyro, where $n$ is the size of the sequence and $M$ the maximal number of mutations).
 
@@
 
+\TODOTous{Mentionner technologie Illumina 50-75 nts, et modele d'erreur/mutations.}
 We illustrate and benchmark our techniques for point-wise error corrections on the 5S ribosomal RNA. We choose the latter since it has been extensively

### 2013-05-25 scripts/jcb_graphs.py
iso

@@ -1,15 +1,23 @@
 import os
 import sys
+import cPickle
 from subprocess import call
 
 def main(m='a'):
     with open(os.path.join('..','16S','key_to_Bench.txt')) as f:
         key = eval(f.readline())
     d_key = {x[0]:x[1:] for x in key}
+    with open(os.path.join('..','16S','d_clusters_trim.cPickle')) as f:
+        data = cPickle.load(f)
+
 
     for f_name in os.listdir(os.path.join('..','16S','Bench')):
         id = int(f_name.split('.')[0])
+        seq = d_key[id][0]
         fold = d_key[id][-1]
+        mut = data[seq]['mut'][fold]
+        with open(os.path.join('Figs_Bench','%sid_%sfold.mut' % (id, fold)), 'w') as f:
+            f.write(mut)
         try:
             os.mkdir('Figs_Bench')
         except OSError:

### 2013-05-25 src/RNAPyro_iso.py

@@ -370,6 +370,27 @@ class memoize(dict):
   def resetCache(self):
       self.clear()
 
+class memoize_iso(dict):
+  """Generically memoizes a function results."""
+  fun = None
+  
+  def __init__(self, f):
+      self.fun = f
+  
+  def __call__(self,*args):
+      nargs = [args[0]]
+      nargs.extend(args[2:])
+      nargs = tuple(nargs)
+      if nargs in self:
+          return self[nargs]
+      else:
+          val = mpf(self.fun(*args))
+          self[nargs] = val
+          return val
+  def resetCache(self):
+      self.clear()
+
+
 def delta(seq,i,c):
   if i<0 or i>=len(seq):
     return 0
@@ -386,6 +407,7 @@ def energy((a,b),(a2,b2),alpha):
 def isGapChar(c):
   return c=="." or c=="-"
 
+@memoize_iso
 def isostericity(seq,ref_seq,(i,j),(a,b), alpha):
   if not ref_seq:
     return 1.

### 2013-05-25 scripts/jcb_graphs.py
iso

@@ -11,7 +11,7 @@ def main(m='a'):
         data = cPickle.load(f)
 
 
-    for f_name in os.listdir(os.path.join('..','16S','Bench')):
+    for f_name in os.listdir(os.path.join('..','16S','Bench_iso')):
         id = int(f_name.split('.')[0])
         seq = d_key[id][0]
         fold = d_key[id][-1]
@@ -24,7 +24,7 @@ def main(m='a'):
             pass
         cmd = ['python',
                'createROC.py',
-               '-f', '%s' % os.path.join('..','16S','Bench',f_name),
+               '-f', '%s' % os.path.join('..','16S','Bench_iso',f_name),
                '-o', '%s' % os.path.join('Figs_Bench','%sid_%sfold.pdf' % (id, fold)),
                '-m', '%s' % m]
         print cmd

### 2013-05-25 scripts/jcb_graphs.py
graphs

@@ -25,7 +25,7 @@ def main(m='a'):
         cmd = ['python',
                'createROC.py',
                '-f', '%s' % os.path.join('..','16S','Bench_iso',f_name),
-               '-o', '%s' % os.path.join('Figs_Bench','%sid_%sfold.pdf' % (id, fold)),
+               '-o', '%s' % os.path.join('Figs_Bench','%sid_%sfold_%s.pdf' % (id, fold,m)),
                '-m', '%s' % m]
         print cmd
         call(cmd)

### 2013-05-25 JCB/methods_JCB.tex
modifs figs

@@
 
-To that purpose, one leverages the \emph{Inside-Outside} construction.
+\begin{figure}[t]\centering
+\includegraphics[scale=\ScaleDP]{FigDPCombineWrapper}
+\caption{Combining the inside and outside contributions to compute the total Boltzmann weight of all sequences having a given base (case 1) or base-pair (cases 2 and 3) at a given position.\label{fig:combine}}
+\end{figure}
+
+To that purpose, one leverages the \emph{Inside-Outside} construction, as  illustrated by Figure~\ref{fig:combine}.
 Namely, while computing the total contribution of sequences featuring a base $a'\in\B$ at position $i$, one must consider the three following situations:

### 2013-05-26 JCB/acknowledgments_JCB.tex
updated conclusion (missing statement of results)

@@
 This work was funded by the French Agence Nationale de la Recherche (ANR) through the {\sc Magnum} {\tt ANR 2010 BLAN 0204} 
-project (to YP), the
-FQRNT team grant 232983 (to VR and JW) and NSERC Discovery grant 219671 (to JW).
+project (to YP), the FQRNT team grant 232983 and NSERC Discovery grant 219671 (to JW).

### 2013-05-26 JCB/conclusion_JCB.tex

@@
 isostericity matrices \cite{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring during base pair replacements.
+
+We combined our algorithm for exploring the mutational neighbourhood of an input sequence with known secondary structure to
+this new pseudo energy model, and create a tool to predict point-wise sequencing errors in structured RNAs.
 Importantly, our algorithm runs in  $\Theta(n\cdot(|\Omega|+M^2))$ time and $\Theta(n\cdot(|\Omega|+M))$ memory, where $n$ is the length of the RNA,
@@
 
-By combining these two approaches into \texttt{RNApyro} ,  the 
-mutational landscape exploration and the pseudo energy model,
- we created a tool predicting the positions
- yielding point-wise sequencing error and correcting them.
-We validated our model with the 5s rRNA,
-as presented in Sec.~\ref{sec:results}.
-We observed that the models
-with larger weights on the
-isostericity seems to hold a higher accuracy on the estimation of errors.
-This indicates that an exploitable signal is captured by the isostericity.
-Importantly, the implementation is fast enough for practical applications.
+We validated our model on the 5s rRNA and 16s rRNA (See Sec.~\ref{sec:results}) and showed that our technology enables us to recover 
+mutational errors with [INSERT PERFS]. Interestingly, we observed that using the isostericity matrices alone yields higher performance than with the 
+nearest-neighbour energy model alone. This findings supports our hypothesis that isostericity matrices provide a valuable source of informations that
+can be efficiently used for RNA sequence and structure analysis.
 
-We must recall that our approach is restricted to
- the correction of point-wise error in structured regions (i.e. base paired nucleotides).
- Nonetheless it should supplement well existing tools, by using previously discarded
-information holding, as shown, a strong signal.
+Our techniques are designed to correcting point-wise errors in structured regions (i.e. base paired nucleotides). Nonetheless, our software 
+\texttt{RNApyro} can be easily combined with other methodologies previously developed for correcting other types of sequencing errors such as indels in 
+unstructured regions or repeats \cite{Quinlan2008,Quince:2009uq}.
+In further works, we also intend to include in our model errors stemming from insertions or deletions.
+It is indeed theoretically possible to considers these scenario within our dynamic programming scheme \cite{Waldispuhl:2002fk} with minor impacts of the complexity.
 
-Further research, given the potential of error-correction of \texttt{RNApyro}, 
-will evaluate its impact over large datasets with different existing
-  NGS error-correction pipe-line.
+Finally, we hope that integrating our software to current sequencing pipe-lines used in metagenomics studies will permit to improve the estimate of microbial diversity.

### 2013-05-26 scripts/jcb_graphs.py
all mut files

@@ -9,15 +9,15 @@ def main(m='a'):
     d_key = {x[0]:x[1:] for x in key}
     with open(os.path.join('..','16S','d_clusters_trim.cPickle')) as f:
         data = cPickle.load(f)
-
+    
 
     for f_name in os.listdir(os.path.join('..','16S','Bench_iso')):
+        if f_name.endswith('mut'):
+            continue
         id = int(f_name.split('.')[0])
         seq = d_key[id][0]
         fold = d_key[id][-1]
         mut = data[seq]['mut'][fold]
-        with open(os.path.join('Figs_Bench','%sid_%sfold.mut' % (id, fold)), 'w') as f:
-            f.write(mut)
         try:
             os.mkdir('Figs_Bench')
         except OSError:

### 2013-05-26 scripts/jcb_graphs.py
batch processing

@@ -3,7 +3,7 @@ import sys
 import cPickle
 from subprocess import call
 
-def main(m='a'):
+def ind(m='a'):
     with open(os.path.join('..','16S','key_to_Bench.txt')) as f:
         key = eval(f.readline())
     d_key = {x[0]:x[1:] for x in key}
@@ -30,9 +30,50 @@ def main(m='a'):
         print cmd
         call(cmd)
 
+def batch(m='a'):
+    with open(os.path.join('..','16S','key_to_Bench.txt')) as f:
+        key = eval(f.readline())
+    d_key = {x[0]:x[1:] for x in key}
+    with open(os.path.join('..','16S','d_clusters_trim.cPickle')) as f:
+        data = cPickle.load(f)
+   
+    l_f_name = os.listdir(os.path.join('..','16S','Bench_iso'))
+    l_id = set([x for x in l_f_name if not x.endswith('mut')])
+    d_fold = {}
+    for id in l_id:
+        fold = d_key[int(id.split('.')[0])][-1]
+        if fold not in d_fold:
+            d_fold[fold] = [id]
+        else:
+            d_fold[fold].append(id)
+
+    for fold in d_fold:
+        to_do = d_fold[fold]
+
+        cmd = ['python',
+               'createROC.py',
+               '-f']
+        for x in to_do:
+            cmd.append(os.path.join('..','16S','Bench_iso',x))
+
+        cmd.extend(['-o', '%s' % os.path.join('Figs_Bench','%sfold.pdf' % (fold)),
+               '-m', '%s' % m])
+        print cmd
+        continue
+        call(cmd)
+
+
 if __name__ == '__main__':
     m = 'a'
+    batch_flag = False
     for i, opt in enumerate(sys.argv):
         if opt == '-m':
             m = sys.argv[i + 1]
-    main(m)
+        elif opt == '-b':
+            batch_flag = True
+
+                  
+    if batch_flag:
+        batch(m)
+    else:
+        ind(m)

### 2013-05-27 JCB/conclusion_JCB.tex
updated conclusion

@@
 and apply our techniques to identify and fix sequencing errors. In addition, we introduced a new scoring scheme to measure the
-likelihood of sequencing errors that combines the classical nearest-neighbour energy model parameters \cite{Turner2010} to the recently introduced
-isostericity matrices \cite{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring during base pair replacements.
+likelihood of sequencing errors that combines the classical nearest-neighbour energy model parameters \cite{Turner2010} to the
+recently introduced isostericity matrices \cite{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring
+during base pair replacements.
 
 We combined our algorithm for exploring the mutational neighbourhood of an input sequence with known secondary structure to
-this new pseudo energy model, and create a tool to predict point-wise sequencing errors in structured RNAs.
-Importantly, our algorithm runs in  $\Theta(n\cdot(|\Omega|+M^2))$ time and $\Theta(n\cdot(|\Omega|+M))$ memory, where $n$ is the length of the RNA,
-$M$ the number of mutations and $\Omega$ the size of the multiple sequence alignment. This achievement enables us to envision applications to high-throughput sequencing pipe-lines.
+this new pseudo energy model, and create a tool to predict point-wise sequencing errors in structured RNAs. Importantly, our
+algorithm runs in  $\Theta(n\cdot(|\Omega|+M^2))$ time and $\Theta(n\cdot(|\Omega|+M))$ memory, where $n$ is the length of
+the RNA, $M$ the number of mutations and $\Omega$ the size of the multiple sequence alignment. This achievement enables
+us to envision applications to high-throughput sequencing pipe-lines.
 
-We validated our model on the 5s rRNA and 16s rRNA (See Sec.~\ref{sec:results}) and showed that our technology enables us to recover 
-mutational errors with [INSERT PERFS]. Interestingly, we observed that using the isostericity matrices alone yields higher performance than with the 
-nearest-neighbour energy model alone. This findings supports our hypothesis that isostericity matrices provide a valuable source of informations that
-can be efficiently used for RNA sequence and structure analysis.
+We validated our model on the 5s rRNA and 16s rRNA (See Sec.~\ref{sec:results}) and showed that our technology enables us
+to recover mutational errors with high accuracy (Area under the ROC curve between $0.95$ and $0.99$ when the mutations are
+located in in base paired regions). Interestingly, we observed that using the isostericity matrices alone yields higher performance
+than with the nearest-neighbour energy model alone. This finding supports our hypothesis that isostericity matrices provide a
+valuable source of informations that can be efficiently used for RNA sequence and structure analysis. Nonetheless, we also found
+that the nearest-neighbour model seems to provide a valuable signal when we under-estimate of the number of errors in the input
+sequences.
 
-Our techniques are designed to correcting point-wise errors in structured regions (i.e. base paired nucleotides). Nonetheless, our software 
-\texttt{RNApyro} can be easily combined with other methodologies previously developed for correcting other types of sequencing errors such as indels in 
-unstructured regions or repeats \cite{Quinlan2008,Quince:2009uq}.
-In further works, we also intend to include in our model errors stemming from insertions or deletions.
-It is indeed theoretically possible to considers these scenario within our dynamic programming scheme \cite{Waldispuhl:2002fk} with minor impacts of the complexity.
+Our techniques are designed to correcting point-wise errors in structured regions (i.e. base paired nucleotides). Nonetheless,
+our software \texttt{RNApyro} can be easily combined with other methodologies previously developed for correcting other types
+of sequencing errors such as indels in unstructured regions or repeats \cite{Quinlan2008,Quince:2009uq}. In further works, we
+also intend to include in our model errors stemming from insertions or deletions. It is indeed theoretically possible to considers
+these scenario within our dynamic programming scheme \cite{Waldispuhl:2002fk} with minor impacts of the complexity.
 
-Finally, we hope that integrating our software to current sequencing pipe-lines used in metagenomics studies will permit to improve the estimate of microbial diversity.
+Finally, we hope that integrating our software to current sequencing pipe-lines used in metagenomics studies will permit to
+improve the estimate of microbial diversity.

### 2013-05-27 scripts/timeBenchmark.py
Preliminary time benchmark

@@ -1,0 +1,79 @@
+import os,sys,os.path;
+from  math import log;
+import random
+from subprocess import call
+import time
+
+def parseSecStr(secstr):
+  p = []
+  result = []
+  for i in range(len(secstr)):
+    c = secstr[i]
+    if c =="(":
+      p.append(i)
+    elif c==")":
+      j = p.pop()
+      result.append((j,i))
+  return result
+      
+BASES = ["A","C","G","U"]
+BASE_PAIRS = [("A","U"),
+  ("U","A"),
+  ("U","G"),
+  ("G","C"),
+  ("G","U"),
+  ("C","G")]
+
+def genMSA(secstr,numseqs):
+  result = []
+  bps = parseSecStr(secstr)
+  for i in range(numseqs):
+    seq = [random.choice(BASES) for j in secstr]
+    for (x,y) in bps:
+      a,b = random.choice(BASE_PAIRS)
+      seq[x]=a
+      seq[y]=b
+    result.append("".join(seq))
+  return result
+
+def genRNAPyroInputFiles(basepath,fromlength=50,tolength=500,lengthincr=25,numseqs=25,alignmentlength=50):
+  for i in range(fromlength,tolength,25):
+    sys.stderr.write("*%s"%(i))
+    os.system("grgfreqs --draw RNALoopsOut.ggd %s %s > out.txt 2> dummy"%(i,numseqs))
+    j = 1
+    for l in open("out.txt"):
+      data = l[:-1].replace("H","(").replace("a","(").replace("b",")").replace("m",".").replace("u",".").replace("M",".").split()
+      secstr = "".join(data)
+      seqs = genMSA(secstr,alignmentlength)
+      fname  = "%s-length%s-num%s.dat"%(basepath,i,j)
+      print fname
+      outfile = open(fname,"w")
+      for s in seqs:
+        outfile.write(s+"\n")
+      outfile.write("".join(data)+"\n")
+      outfile.close()    
+      j += 1
+
+
+
+if __name__=="__main__":
+#  if not os.path.exists("TimeBench"):
+#    os.mkdir("TimeBench")
+#  genRNAPyroInputFiles(os.path.join("TimeBench","MSA"),50,500,25,50,50)
+  exec_name = os.path.join('..','src','RNAPyro_iso.py')
+  nbmuts = 10
+  alpha = .5
+  outfile = open("TimeBench.dat","w")
+  for f_name in os.listdir(os.path.join("TimeBench")):
[…]

### 2013-05-27 src/RNAPyro.py

@@ -781,7 +781,7 @@ def help(file_name=False,
 
   print """Required:
       -f <file_path>  Path to the file containing the reference
-                      sequence, the MSE (optional) and the secondary
+                      sequence, the MSA (optional) and the secondary
                       structure. The FIRST sequence in the file
                       will be used as reference
       -m <int>        The numbers of mutants required

### 2013-05-28 JCB/ResponseToReviewers-RNAPyro-JCB-2013.tex
Some benchmark + method

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt,hyperref,draft]{article} % use larger type; default would be 10pt

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{xcolor} % to change the page dimensions
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{amsmath} % for much better looking tables
\usepackage{xspace}
\usepackage{url}
\usepackage{amssymb} % for much better looking tables
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{ntheorem} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
\usepackage{framed}
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

\newcommand{\Answer}[1]{\noindent\textsf{\textbf{Response: }}{\sf#1}\\}
\newcommand{\Comment}[1]{\noindent\textsf{\textbf{Comment: }}{\it#1}\\[.5em]}

%%% END Article customizations

\input{macros}

%%% The "real" document content comes below...

\title{Using structural and evolutionary information to detect and correct pyrosequencing errors in non-coding RNAs}
\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
\date{$^1$~School of Computer Science, McGill University, Montreal, Canada, $^2$~Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.\\ \small $^*$~Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}

%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

In general, the reviewers did not raise very specific points, but were mostly concerned by the applicability of the method in a high-throughput context. In this extended version, we demonstrated the scalability of the method by running processing longer sequences (16S rRNAs) obtained from random reads (generated by the {\tt ART} pyrosequencing simulator), a setting which is closer to real use-cases in metagenomics. Furthermore, our python implementation  is still more time-consuming than would be ideal in a high-throughput context. However, we propose a pipeline which first maps short reads to the reference alignment, followed by an alignment/correction of the consensus sequence against the reference multiple sequence alignment. In this setting, {\tt RNAPyo} is only executed once for each rRNA, as opposed to once for every read, and the typical runtime of our software does not induce such a strong overhead. Additionally, we investigated the error-correction potential as a function of the coverage (depth) of the sequencing.

\section*{Detailed responses}

\Comment{(Reviewer \#1) While the paper is algorithmic interesting, and clearly publishable in principle, I'm a bit concerned with the practical application of RNApyro. If many reads are available for the same sequence, the tool is obviously not necessary since purely sequence-based error detection should be applicable. In the scenario envisioned here, however, the model makes a very strong assumption, name a completely known consensus structure to which the query sequence must fit without indels. While plausible if the consensus is derived from closely related species (in which case point mutations that disrupt the structure are easily visible in most cases) one would expect also structural variations as well.\\
(Reviewer \#2) However, assuming that it works as expected, I understand that a successful "error correction" involves replacing an unlikely secondary structure base pair by a more likely one. Couldn't this become a problem for outlier RNA sequences that do not exactly fold according to the canonical model used in RFAM? If this is the case, then the program would correct certain sequences/species better than others, which is not quite a desirable property in error correction. 
}
\Answer{We agree with the reviewer that the error-correction potential of our method may be challenged by sequences whose folding is substantially different from the consensus structure. However, assuming some level of structural conservation within RNA families that share a given function is generally held as reasonable, and constitutes the main paradigm in RNA comparative bioinformatics, e.g. underlying the RFAM database and homologous sequences retrieval. This paradigm is, of course, challenged in the case of structure-disrupting mutations (which are increasingly identified and linked with RNA-based genetic conditions), but should remain reasonably consensual in the context of metagenomics, where structural outliers (often pathological, or even lethal) are perhaps of lesser interest. \\
Furthermore, the consensus structure needs not be completely known. In this extended version, we considered the potential for error-correction of {\tt RNAPyro}, restricted paired regions, based on the dual rationale that 1) our energy model does not (yet!) capture unpaired regions, and 2) unpaired region may be paired but not sufficiently conserved throughout the evolution (or simply misaligned). The excellent areas under the ROC curves observed in this case suggests that a partial consensus may not too strongly impede the quality of our error corrections. \\
Finally, regarding a more general structural diversity in families of structural RNAs (e.g. only preserving a domain, a opposed to a general architecture), we believe such cases would be revealed by the level of detail of the consensus structure and, by disallowing error-corrections within unpaired regions, our method would not overly erroneous and inhomogeneous. }

\Comment{(Reviewer \#1) Also, how realistic is it that one would trust a single 454 read with 6 or 12 expected errors ? In practice, on would at least want to have a prior of the quality of the nucleotide call included.}
\Answer{We addressed this comment by estimating the number of errors from the coverage. We simply fitted the number of errors empirically observed on sequences assembled from Illumina simulated short reads, as a function of the coverage and sequence length.}

\Comment{(Reviewer \#2) Another issue is that of computing time. Although speed may be improved relative to prior versions of the algorithm, the times presented in Table I (is it for a single sequence?) do not appear to be compatible with NGS applications.}
\Answer{We improved the speed of our implementation by performing further preprocessing, and show in the extended version a more complete time benchmark. As noted earlier in this response, we run {\tt RNApyro} on a consensus sequence obtained by mapping sequences onto the multiple sequence alignment. Consequently, one does not need to run the software once for every read, but rather once every (Length$\times$Coverage$/$Read-Length)-th read, and the improve runtime should now realistically allow for embedding {\tt RNAPyro} within a high-throughput pipeline.
}

\Comment{(Reviewer \#3) The problem setting is unique, but applicability may be limited to known (annotated) ncRNAs. The algorithm may be more naturally applied to optimization by computational evolution.}
\Answer{As suggested by the reviewer, we leveraged the inside computation into a 	statistical sampling algorithm addressing the inverse folding (aka RNA design) problem (manuscript accepted, scheduled for an oral presentation at the ISMB'13 conference~\cite{Reinharz2013}).}

\Comment{(Reviewer \#2) Typos: Please check sentences...}
\Answer{Fixed}

\bibliographystyle{amsplain}
\bibliography{RNApyro}
\end{document}

### 2013-05-28 JCB/results_JCB.tex
Some benchmark + method

@@
 {\centering \url{https://github.com/McGill-CSB/RNApyro}\\}
+\TODOJerome{Did we set up an ad hoc website?}
 
@@
 Since typical use-cases of \RNApyro require
- efficiency and scalability, we present in Table~\ref{tab:time} typical runtimes required to compute the probabilities for  every nucleotide at every positions for a vast set of parameters.
+ efficiency and scalability, we present in TableFigure~\ref{fig:time} typical runtimes required to compute the probabilities for  every nucleotide at every positions for a vast set of parameters.
 For those tests,
- both the sequences and the target secondary structure were generated at random.
+ both the multiple sequence alignment and the reference sequence were generated uniformly at random, based on a realistic random secondary structure, generated as described in Levin~\emph{et al}~\cite{Levin:2012kx}.
 \TODOVlad{Add more extensive time benchmark, as a plot}
 
-\begin{table}[t]
-\begin{minipage}{.4\textwidth}
-\begin{tabular}{rccc}
-Length &\multicolumn{3}{c}{\#mutations}\\\cline{2-4}
-		 			  & 6   &  12  & 24\\\cline{2-4}
-100  				& 35s  & 238s & 1023s\\
-300  			& 135s & 594s &2460s\\\cline{2-4}
-		 						& 25   &    &	50		\\\cline{2-4}
-500         & 5400s&       &  21003s    \\
-\end{tabular}
-\end{minipage}
-\begin{minipage}{.59\textwidth}
-\caption{Time required by the computation of probabilities. First column indicates the length and  the column indexes indicate the number
- of mutations. $\alpha$ is
-set at $0.5$,  $\beta$ to $15$ and $|\Omega|=44$.}
-\label{tab:time}
-\end{minipage}
-\SpaceCheating
-\end{table}
+\begin{figure}[t]
+{\centering \includegraphics[width=.5\linewidth]{figures/TimeBenchmark}\\}
+
+\caption{Typical runtimes required by the computation of mutational profiles 
+averaged on $50$ random sequences for each length ranging from $100$ and $400$ nts, while allowing for a maximal number mutations equal to $M=10$. 
+For each sequence, a random multiple sequence alignment was generated, consisting of $51$ aligned sequences, each compatible with a randomly generated consensus secondary structure. 
+}
+\label{fig:time}
+\end{figure}
 
@@
 
-Our data demonstrates that our algorithm exhibits interesting potential for error-correction applications. First, the AUC values (up to $0.86$) indicate that a
+Our data demonstrates that our algorithm shows interesting potential for error-correction applications. First, the AUC values (up to $0.86$) indicate that a
 signal has been successfully extracted. This result has been achieved with errors in loop regions (i.e. without base pairing information) and thus suggests
@@
 the isostericity scores (i.e. for low $\alpha$ values) seem more robust to under- and over-estimating the number of errors.
+
+\subsection{Error Correction in 16s rRNAs}
+
+\TODOTous{Describe and present results????}

### 2013-05-28 JCB/ResponseToReviewers-RNAPyro-JCB-2013.tex
Typos

@@
 Finally, regarding a more general structural diversity in families of structural RNAs (e.g.
-only preserving a domain, a opposed to a general architecture), we believe such cases would be revealed by the level of detail of the consensus structure and, by disallowing error-corrections within unpaired regions, our method would not overly erroneous and inhomogeneous.
+only preserving a domain, a opposed to a general architecture), we believe such cases would be revealed by the level of detail of the consensus structure and, by disallowing error-corrections within unpaired regions, our method should not be overly erroneous and inhomogeneous (at the obvious cost of missing corrections within unpaired regions).
 }

### 2013-05-28 JCB/results_JCB.tex
comment time & URL
method

--- 2013-05-23 JCB/results_JCB.tex
@@
 
-The time benchmarks were performed on a MacMini 2010, 2.3GHz dual-core Intel Core i5, 8GB of RAM.
+The time benchmarks were performed on a MacMini 2010, 2.3GHz dual-core Intel Core i5, 8GB of RAM for the Sec.~\ref{sec:5S} and Sec.~\ref{sec:16S} on AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.20 cores and 175 GB ram.
 Since typical use-cases of \RNApyro require
@@
 \subsection{Error Correction in 5s rRNAs}
-
+\label{sec:5S}
 To illustrate the potential of our algorithm, we applied our techniques to identify and correct point-wise errors in RNA sequences
@@
 the isostericity scores (i.e. for low $\alpha$ values) seem more robust to under- and over-estimating the number of errors.
+
+\subsection{Error Correction in 16s rRNAs}
+\label{sec:16S}
+
+To complete our benchmark we also looked at the Bacterial small subunit ribosomal RNA. 
+We gathered the seed multiple sequence alignment (MSA) composed of 93 sequences from the Rfam Database 11.0 (Rfam id: RF00177)~\cite{gardner2011rfam}.
+Those sequences are between 1461 and 1568 nucleotides long and yield a similarity of $69\%$.
+
+As in Sec.~\ref{sec:5S} we used a leave-one-out strategy. Since few structures contain all gaps 
+at the same positions, after selecting a sequence of interest,  we used the following strategy.
+ It was applied to each sequence in the alignment. Starting with the full alignment, all positions corresponding to gaps in the sequence of interest were removed, for all sequences in the alignment.
+
+To simulate sequencing error in the sequence of interest,
+ we use the the next-generation sequencing read simulator 
+ART~\cite{huang2012art}. The Illumina technology main error mode is base substitution. We used
+this model since those
+are the errors that \RNApyro detects. As parameters, the reads reads generated
+were $75\text{bps}$ long for an average fold coverage of $5$, $10$ and $15$. 
+ They yield respectively an average error rate of $2.4\%$, $0.9\%$ and $0.6\%$.
+
+ As in Sec.~\ref{sec:5S} we used a value of $\beta=15$ and evaluated $\alpha$ at $0, 0.5, 0.8, 1.0$. 
+ Additionally since the alignment still contains gap, we made sure to evaluate the isostericity term

### 2013-05-28 JCB/methods_JCB.tex
Typo

@@
 
-In principle, $\Theta(m\cdot n^2)$ terms, identified by $(m,i,j)$ triplets, should be computed.
+In principle, $\Theta(M\cdot n^2)$ terms, identified by different $(\Struct,m)$ triplets, should be computed.
 However, a close inspection of the recurrences reveals that the computation can be safely restricted to a subset of intervals $(i,j)$.
 For instance, the inside algorithm only requires computing intervals $[i,j]$ that do not break any base-pair, and whose next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$.
-Similar constraints hold for the outside computation,  following from the linearity of the outside recurrence (i.e.
+A similar property holds for the outside computation,  following from the linearity of the outside recurrence (i.e.
 the computation of the outside term for a given excluded structure only relies on the computation of another, strictly larger, structure).

### 2013-05-28 JCB/results_JCB.tex

--- 2013-05-28 JCB/results_JCB.tex
@@
 
-{\centering \url{https://github.com/McGill-CSB/RNApyro}\\}
-\TODOJerome{Did we set up an ad hoc website?}
+%{\centering \url{https://github.com/McGill-CSB/RNApyro}\\}
+%\TODOJerome{Did we set up an ad hoc website?}
 
+{\centering \url{http://csb.cs.mcgill.ca/RNApyro}\\}
 The time benchmarks were performed on a MacMini 2010, 2.3GHz dual-core Intel Core i5, 8GB of RAM.
-Since typical use-cases of \RNApyro require
- efficiency and scalability, we present in TableFigure~\ref{fig:time} typical runtimes required to compute the probabilities for  every nucleotide at every positions for a vast set of parameters.
-For those tests,
- both the multiple sequence alignment and the reference sequence were generated uniformly at random, based on a realistic random secondary structure, generated as described in Levin~\emph{et al}~\cite{Levin:2012kx}.
+%on AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.20 cores and 175 GB ram.
+Since typical use-cases of \RNApyro require efficiency and scalability, we present in TableFigure~\ref{fig:time}
+typical runtimes required to compute the probabilities for  every nucleotide at every positions for a vast set of parameters.
+For those tests, both the multiple sequence alignment and the reference sequence were generated uniformly at random,
+based on a realistic random secondary structure, generated as described in Levin~\emph{et al}~\cite{Levin:2012kx}.
 \TODOVlad{Add more extensive time benchmark, as a plot}
+\TODOVlad{Is time benchmark still on Mac mini or mcbbigram?}
 
@@
 \subsection{Error Correction in 16s rRNAs}
+\label{sec:16S}
+
+To complete our benchmark we also looked at the Bacterial small subunit ribosomal RNA. 
+We gathered the seed multiple sequence alignment (MSA) composed of 93 sequences from the Rfam Database 11.0 (Rfam id: RF00177)~\cite{gardner2011rfam}.
+Those sequences are between 1461 and 1568 nucleotides long and yield a similarity of $69\%$.
+
+As in Sec.~\ref{sec:5S} we used a leave-one-out strategy. Since few structures contain all gaps 
+at the same positions, after selecting a sequence of interest,  we used the following strategy.
+ It was applied to each sequence in the alignment. Starting with the full alignment, all positions corresponding to gaps in the sequence of interest were removed, for all sequences in the alignment.
+
+To simulate sequencing error in the sequence of interest,
+ we use the the next-generation sequencing read simulator 
+ART~\cite{huang2012art}. The Illumina technology main error mode is base substitution. We used
+this model since those
+are the errors that \RNApyro detects. As parameters, the reads reads generated
+were $75\text{bps}$ long for an average fold coverage of $5$, $10$ and $15$. 
+ They yield respectively an average error rate of $2.4\%$, $0.9\%$ and $0.6\%$.
+
+ As in Sec.~\ref{sec:5S} we used a value of $\beta=15$ and evaluated $\alpha$ at $0, 0.5, 0.8, 1.0$. 
+ Additionally since the alignment still contains gap, we made sure to evaluate the isostericity term
+
+\subsection{Error Correction in 16s rRNAs}
 

### 2013-05-28 JCB/results_JCB.tex
res

@@
 {\centering \url{http://csb.cs.mcgill.ca/RNApyro}\\}
-The time benchmarks were performed on a MacMini 2010, 2.3GHz dual-core Intel Core i5, 8GB of RAM.
-%on AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.20 cores and 175 GB ram.
+The time benchmarks were performed on an AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.20 cores and 175 GB ram.
 Since typical use-cases of \RNApyro require efficiency and scalability, we present in TableFigure~\ref{fig:time}
@@
 \TODOVlad{Add more extensive time benchmark, as a plot}
-\TODOVlad{Is time benchmark still on Mac mini or mcbbigram?}
+%\TODOVlad{Is time benchmark still on Mac mini or mcbbigram?}
 
@@
 \label{sec:16S}
+\TODOTous{Describe and present results????}
 
@@
  Additionally since the alignment still contains gap, we made sure to evaluate the isostericity term
-
-\subsection{Error Correction in 16s rRNAs}
-
-\TODOTous{Describe and present results????}

### 2013-05-28 JCB/results_JCB.tex
res 16s

@@
 this model since those
-are the errors that \RNApyro detects. As parameters, the reads reads generated
+are the errors that \RNApyro detects. As parameters, the reads  generated
 were $75\text{bps}$ long for an average fold coverage of $5$, $10$ and $15$. 
@@
 
- As in Sec.~\ref{sec:5S} we used a value of $\beta=15$ and evaluated $\alpha$ at $0, 0.5, 0.8, 1.0$. 
- Additionally since the alignment still contains gap, we made sure to evaluate the isostericity term
+We used a value of $\beta=15$ and evaluated $\alpha$ at $0, 0.5, 0.8, 1.0$. 
+ Additionally since the alignment still contains gap, we made sure to evaluate the isostericity 
+ term only
+ with sequences in the MSA where both positions in the base pair are not gaps. 
+ Due to the low number of errors, we set the presumed as twice the average error rate made by ART, i.e. $4.8\%$, $1.8\%$ and $1.2\%$ for fold coverages of $5$, $10$ and $15$ respectively.
+
+To evaluate our method, we simulated with a random sequences from the MSA every time,
+ $12$ times the $5-$fold, $10$ times the $10-$fold and $14$ times the $15-$fold coverage. And we 
+ evaluated the ROC curve as in Sec.~\ref{sec:5S}.

### 2013-05-28 JCB/results_JCB.tex
added graphs results

@@
  evaluated the ROC curve as in Sec.~\ref{sec:5S}.
+
+ \begin{figure}
+\centering
+\begin{subfigure}{.33\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{figures/5fold-a}
+  \caption{$5-$fold coverage}
+\end{subfigure}%
+\begin{subfigure}{.33\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{figures/10fold-a}
+    \caption{$10-$fold coverage}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{figures/15fold-a}
+    \caption{$15-$fold coverage}
+\end{subfigure}
+\caption{Performance of error-correction over all positions. Subfigures show accuracy when
+ART fold parameter is set to 5, 10 and $15-$fold coverage.
+We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values ranging of $\alpha = \{0, 0.5, 0.8, 1.0\}$.
+The AUC is indicated in the legend of the figures.
+Each individual ROC curve represent the average performance over at least 10 experiments.}
+
+\label{fig:16s_all}
+\end{figure}
+
+ \begin{figure}
+\centering
+\begin{subfigure}{.33\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{figures/5fold-b}
+  \caption{$5-$fold coverage}
+\end{subfigure}%
+\begin{subfigure}{.33\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{figures/10fold-b}
+    \caption{$10-$fold coverage}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{figures/15fold-b}
+    \caption{$15-$fold coverage}
+\end{subfigure}
+\caption{Performance of error-correction over positions inside a base pair. Subfigures show accuracy when
+ART fold parameter is set to 5, 10 and $15-$fold coverage.
+We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values ranging of $\alpha = \{0, 0.5, 0.8, 1.0\}$.
+The AUC is indicated in the legend of the figures.
+Each individual ROC curve represent the average performance over at least 10 experiments.}
+\label{fig:16s_bp}
+\end{figure}

### 2013-05-28 scripts/timeBenchmark.py
Progressive benchmark

@@ -63,17 +63,33 @@ if __name__=="__main__":
   exec_name = os.path.join('..','src','RNAPyro_iso.py')
   nbmuts = 10
   alpha = .5
-  outfile = open("TimeBench.dat","w")
+
+  dataPoints = {}
   for f_name in os.listdir(os.path.join("TimeBench")):
-    outfile.write("%s\t"%(f_name))
-    cmd = ['python',
-                   exec_name,
-                   '-f', '%s' % os.path.join("TimeBench",f_name),
-                   '-m', '%s' % nbmuts,
-                   '-a', '%s' % alpha
-                   ]
-    tic = time.time()
-    call(cmd)
-    tac = time.time() - tic
-    print "%s\t%s"%(f_name,tac)
-    outfile.write("%s\n"%(tac))
+    args = f_name.split("-")
+    length = int(args[1][len("length"):])
+    if (length not in dataPoints):
+      dataPoints[length] = []
+    dataPoints[length].append(f_name)
+  maxSize = max([len(x) for x in dataPoints.values()])
+
+  outfile = open("TimeBenchTst.dat","w")
+  lengths = dataPoints.keys()
+  lengths.sort()
+  for i in range(maxSize):
+    for length in lengths:
+      if len(dataPoints[length])>i:
+        f_name = dataPoints[length][i]
+        outfile.write("%s\t"%(f_name))
+        cmd = ['python',
+                       exec_name,
+                       '-f', '%s' % os.path.join("TimeBench",f_name),
+                       '-m', '%s' % nbmuts,
+                       '-a', '%s' % alpha
+                       ]
+        tic = time.time()
+        call(cmd)
+        tac = time.time() - tic
+        print "%s\t%s"%(f_name,tac)
+        outfile.write("%s\n"%(tac))
+        outfile.flush()

### 2013-05-28 JCB/results_JCB.tex
results

@@
  $12$ times the $5-$fold, $10$ times the $10-$fold and $14$ times the $15-$fold coverage. And we 
- evaluated the ROC curve as in Sec.~\ref{sec:5S}.
+ evaluated the ROC curve mostly as in Sec.~\ref{sec:5S}. In Fig.~\ref{fig:16s_all} they are computed
+ over all positions, while in Fig.~\ref{fig:16s_bp} only over positions inside a base pair.
 
@@
 \end{figure}
+
+Our data demonstrates that even on long sequences our algorithm shows interesting potential for 
+error-correction applications. First, the AUC values (up to 0.81) when looking at all positions, 
+in Fig.~\ref{fig:16s_all},  indicate that a signal has been successfully extracted. When restraining
+to structured regions (i.e. base paired regions) we obtain AUC values up to 0.998.  Since 
+contributions to the energy only comes from structured regions, this was an expected result.
+
+An interesting feature is that best results are almost always when $\alpha=0$, i.e. all the contribution
+comes from the isostericity. A notable exception is when the number of mutation is underestimated, 
+in that case best performances were observed for non-nil values of $\alpha$.

### 2013-05-28 JCB/results_JCB.tex
Rewrite

--- 2013-05-28 JCB/results_JCB.tex
@@
 
-To complete our benchmark we also looked at the Bacterial small subunit ribosomal RNA. 
-We gathered the seed multiple sequence alignment (MSA) composed of 93 sequences from the Rfam Database 11.0 (Rfam id: RF00177)~\cite{gardner2011rfam}.
-Those sequences are between 1461 and 1568 nucleotides long and yield a similarity of $69\%$.
+To complete our benchmark, we considered the small subunit ribosomal RNA, a molecule wich is of particular interest  in metagenomics an phylogenetics.
+We gathered the seed sequences of the bacterial multiple sequence alignment (MSA) retrieved from the RFAM Database 11.0 (Rfam id: RF00177)~\cite{gardner2011rfam}.
+This alignment is composed of 93 sequences, ranging between 1461 and 1568 nucleotides long, and yield an average pairwise sequence similarity of $69\%$.
 

### 2013-05-28 JCB/results_JCB.tex
art description

--- 2013-05-28 JCB/results_JCB.tex
@@
  we use the the next-generation sequencing read simulator 
-ART~\cite{huang2012art}. The Illumina technology main error mode is base substitution. We used
+\ART~\cite{huang2012art}. The Illumina technology main error mode is base substitution. We used
 this model since those
@@
 were $75\text{bps}$ long for an average fold coverage of $5$, $10$ and $15$. 
- They yield respectively an average error rate of $2.4\%$, $0.9\%$ and $0.6\%$.
+ They yield respectively an average error rate of $2.4\%$, $0.9\%$ and $0.6\%$. \ART reads
+ are automatically mapped to the sequence of interest. For every position covered by reads, 
+ we selected the nucleotide appearing more frequently. At un-covered position, we picked a random
+ nucleotide.
 
@@
  with sequences in the MSA where both positions in the base pair are not gaps. 
- Due to the low number of errors, we set the presumed as twice the average error rate made by ART, i.e. $4.8\%$, $1.8\%$ and $1.2\%$ for fold coverages of $5$, $10$ and $15$ respectively.
+ Due to the low number of errors, we set the presumed as twice the average error rate made by \ART, i.e. $4.8\%$, $1.8\%$ and $1.2\%$ for fold coverages of $5$, $10$ and $15$ respectively. \ART reads
 
@@
 \end{subfigure}
-\caption{Performance of error-correction over positions inside a base pair. Subfigures show accuracy when
+\caption{Performance of error-correction over structured positions (i.e. inside a base pair). Subfigures show accuracy when
 ART fold parameter is set to 5, 10 and $15-$fold coverage.

### 2013-05-28 scripts/timeBenchmark.py
bench time yamaska

@@ -73,7 +73,7 @@ if __name__=="__main__":
     dataPoints[length].append(f_name)
   maxSize = max([len(x) for x in dataPoints.values()])
 
-  outfile = open("TimeBenchTst.dat","w")
+  outfile = open("TimeBench_yamaska.dat","w")
   lengths = dataPoints.keys()
   lengths.sort()
   for i in range(maxSize):

### 2013-05-28 JCB/results_JCB.tex
Rephrased 16s

--- 2013-05-28 JCB/results_JCB.tex
@@
 
-\subsection{Error Correction in 5s rRNAs}
+\subsection{Homogenous Error Correction in 5s rRNAs}
 \label{sec:5S}
@@
 
-\subsection{Error Correction in 16s rRNAs}
+\subsection{Correcting Illumina sequencing errors in 16s rRNAs}
 \label{sec:16S}
@@
 
-To complete our benchmark, we considered the small subunit ribosomal RNA, a molecule wich is of particular interest  in metagenomics an phylogenetics.
+To complete our benchmark, we turned to the small subunit ribosomal RNA in bacteriae, a molecule which is of particular interest  in metagenomics and phylogenetics.
+Our aim was to get as close as possible to a pyrosequencing context, in which reads are produced non-uniformly by an Illumina sequencer, impacting the distribution of errors in the sequence.
+We chose this setting both because of the popularity of Illumina sequencing in metagenomics, and since the underlying sequencing technique only considers base substitutions (no insertions), the only type of errors currently detected by \RNApyro.
+To that purpose, we used simulated Illumina reads, mapped the reads back to a reference alignment, and run \RNApyro on a consensus sequence derived from the mapped reads, estimating the maximal amount of mutations from both the length of the sequence and the sequencing depth.
+
 We gathered the seed sequences of the bacterial multiple sequence alignment (MSA) retrieved from the RFAM Database 11.0 (Rfam id: RF00177)~\cite{gardner2011rfam}.
-This alignment is composed of 93 sequences, ranging between 1461 and 1568 nucleotides long, and yield an average pairwise sequence similarity of $69\%$.
+This alignment is composed of 93 sequences, whose length varies between 1461 and 1568 nucleotides, and has an average pairwise sequence identity of $69\%$.
+We used the pseudoknot-free version of the consensus secondary structure.
+A secondary structure for a specific reference sequence was obtained by simply mapping the structure back to sequence, i.e.
+by removing any base-pair having at least one partner involved in a gapped position.
+For similar reasons, we locally excluded from our calculation of the isostericity contribution for a given base-pair, described in Section~\ref{sec:model}, any sequence featuring at least a gap on the corresponding positions.
 
-As in Sec.~\ref{sec:5S} we used a leave-one-out strategy. Since few structures contain all gaps 
-at the same positions, after selecting a sequence of interest,  we used the following strategy.
- It was applied to each sequence in the alignment. Starting with the full alignment, all positions corresponding to gaps in the sequence of interest were removed, for all sequences in the alignment.
+To simulate sequencing errors, we used the next-generation sequencing read simulator 
+{\tt ART}~\cite{huang2012art}.
+The \emph{Illumina technology} setting was chosen as the main error mode, generating reads of $75~\text{bps}$, featuring mostly base substitution errors.
+Reads were mapped back to the original sequence, and a consensus sequence was determined, from the sequencing output, by a simple majority vote in the case of multiple coverage.
+Uncovered regions were simply generated at random.
+The average  rate of errors observed in the final consensus sequence was empirically estimated to represent $2.4\%$, $0.9\%$ and $0.6\%$ of the reference sequence, for prescribed sequencing coverages of $5$, $10$ and $15$ fold respectively.
 
-To simulate sequencing error in the sequence of interest,
- we use the the next-generation sequencing read simulator 
-ART~\cite{huang2012art}. The Illumina technology main error mode is base substitution. We used
-this model since those
-are the errors that \RNApyro detects. As parameters, the reads  generated
-were $75\text{bps}$ long for an average fold coverage of $5$, $10$ and $15$. 
- They yield respectively an average error rate of $2.4\%$, $0.9\%$ and $0.6\%$.
-
-We used a value of $\beta=15$ and evaluated $\alpha$ at $0, 0.5, 0.8, 1.0$. 
- Additionally since the alignment still contains gap, we made sure to evaluate the isostericity 
- term only
- with sequences in the MSA where both positions in the base pair are not gaps. 
- Due to the low number of errors, we set the presumed as twice the average error rate made by ART, i.e. $4.8\%$, $1.8\%$ and $1.2\%$ for fold coverages of $5$, $10$ and $15$ respectively.
-
-To evaluate our method, we simulated with a random sequences from the MSA every time,
- $12$ times the $5-$fold, $10$ times the $10-$fold and $14$ times the $15-$fold coverage. And we 
- evaluated the ROC curve mostly as in Sec.~\ref{sec:5S}. In Fig.~\ref{fig:16s_all} they are computed
- over all positions, while in Fig.~\ref{fig:16s_bp} only over positions inside a base pair.
+As in Section~\ref{sec:5S}, we evaluated the predictive power of \RNApyro-computed mutational profiles using a leave-one-out strategy.
+We picked a sequence at random from the MSA, sequenced/mutated it as above, and run \RNApyro to establish its mutational profile.
+In this execution, we used values of $\beta=15$ and $\alpha\in [0, 0.5, 0.8, 1.0]$, and set the presumed number of mutations to twice the average error rate made by {\tt ART}, i.e.
+$4.8\%$, $1.8\%$ and $1.2\%$ for fold coverages of $5$, $10$ and $15$ respectively.
+We repeated the whole procedure $12$/$10$/$14$ times for the $5$/$10$/$15-$fold coverage. 
+We evaluated the predictive power of the profile by a computing a joint ROC curve for each value of $\alpha$, and each coverage, as described in Section~\ref{sec:5S}.
+Figure~\ref{fig:16s_all} shows the ROC curves, computed over all positions, while Figure~\ref{fig:16s_bp} only focuses on positions that are paired in the consensus secondary structure.
 
@@
 to structured regions (i.e. base paired regions) we obtain AUC values up to 0.998.  Since 
-contributions to the energy only comes from structured regions, this was an expected result.
+contributions to the energy and isostericity only arise from structured regions, this was an expected result.
 
@@
 comes from the isostericity. A notable exception is when the number of mutation is underestimated, 
-in that case best performances were observed for non-nil values of $\alpha$.
+in that case best performances were observed for non-null values of $\alpha$.

### 2013-05-28 JCB/methods_JCB.tex

@@
 
-\subsection{Probabilistic Model}
+\subsection{Probabilistic Model}\label{sec:model}
 Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 

### 2013-05-28 JCB/results_JCB.tex
time minor

@@
 in that case best performances were observed for non-null values of $\alpha$.
+
+Another observation is when we look at all positions as in Fig.~\ref{fig:16s_all}.
+The first $20\%$ to $30\%$ of sensitivity are obtained almost without any errors and correspond to the nucleotides in
+structured regions. The rest of the predictions are done on un-paired nucleotides.

### 2013-05-29 src/RNAPyro.py
updateded rnapiro

@@ -370,6 +370,27 @@ class memoize(dict):
   def resetCache(self):
       self.clear()
 
+class memoize_iso(dict):
+  """Generically memoizes a function results."""
+  fun = None
+  
+  def __init__(self, f):
+      self.fun = f
+  
+  def __call__(self,*args):
+      nargs = [args[0]]
+      nargs.extend(args[2:])
+      nargs = tuple(nargs)
+      if nargs in self:
+          return self[nargs]
+      else:
+          val = mpf(self.fun(*args))
+          self[nargs] = val
+          return val
+  def resetCache(self):
+      self.clear()
+
+
 def delta(seq,i,c):
   if i<0 or i>=len(seq):
     return 0
@@ -386,6 +407,7 @@ def energy((a,b),(a2,b2),alpha):
 def isGapChar(c):
   return c=="." or c=="-"
 
+@memoize_iso
 def isostericity(seq,ref_seq,(i,j),(a,b), alpha):
   if not ref_seq:
     return 1.
@@ -781,7 +803,7 @@ def help(file_name=False,
 
   print """Required:
       -f <file_path>  Path to the file containing the reference
-                      sequence, the MSA (optional) and the secondary
+                      sequence, the MSE (optional) and the secondary
                       structure. The FIRST sequence in the file
                       will be used as reference
       -m <int>        The numbers of mutants required

### 2013-05-29 src/RNAPyro_iso.py

@@ -1,880 +1,0 @@
-###############################################################################
-#Copyright (c) 2013, Vladimir Reinharz, Yann Ponty & Jerome Waldispuhl        #
-#All rights reserved.                                                         #
-#                                                                             #
-#Redistribution and use in source and binary forms, with or without           #
-#modification, are permitted provided that the following conditions are met:  #
-#* Redistributions of source code must retain the above copyright             #
-#notice, this list of conditions and the following disclaimer.                #
-#* Redistributions in binary form must reproduce the above copyright          #
-#notice, this list of conditions and the following disclaimer in the          #
-#documentation and/or other materials provided with the distribution.         #
-#* Neither the name of the <organization> nor the                             #
-#names of its contributors may be used to endorse or promote products         #
-#derived from this software without specific prior written permission.        #
-#                                                                             #
-#THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"  #
-#AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE    #
-#IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE   #
-#ARE DISCLAIMED. IN NO EVENT SHALL <COPYRIGHT HOLDER> BE LIABLE FOR ANY       #
-#DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES   #
-#(INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; #
[…]

### 2013-05-30 JCB/main_RNAPyro_JCB.tex
Typos
added graphs results
art description
Made the paper 'Arxiv-friendly'

@@
 \usepackage{relsize}
-\usepackage{bbm}
+%\usepackage{bbm}
+\usepackage{bbold}
 \usepackage[numbers]{natbib}
 \usepackage[colorinlistoftodos,textsize=small]{todonotes}
+
+\usepackage{caption}
+%\usepackage{subcaption}
 
@@
 
-\usepackage{tikz}
-\usepackage{amsmath}
-\usepackage{ifthen}
-\usepackage{relsize}
-\usetikzlibrary{positioning}
-\usetikzlibrary{matrix}
-\usetikzlibrary{calc}
-\usetikzlibrary{shapes}
-\usetikzlibrary{arrows}
-\usetikzlibrary{fit}
-\usetikzlibrary{decorations}
-\usetikzlibrary{snakes}
-\usetikzlibrary{shadows}
-\pgfdeclarelayer{background}
-\pgfdeclarelayer{backbackground}
-\pgfdeclarelayer{foreground}
-\pgfsetlayers{backbackground,background,main,foreground}
-
 \input{macros.tex}
@@
 %\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
-\title{Using structural and evolutionary information to correct RNA pyrosequencing errors in non-coding RNAs}
+\title{Using structural and evolutionary information to detect and correct pyrosequencing errors in non-coding RNAs}
 \author{Vladimir Reinharz\inst{1}, Yann Ponty\inst{2}$^{*}$ \and J\'er\^{o}me Waldisp\"{u}hl\inst{1}$^*$}
@@
 %              pdftex}
+\usepackage{subcaption}
+\captionsetup{compatibility=false}
+
+\newcommand{\ART}{\texttt{ART}\xspace}
 
@@
 We develop a scoring scheme combining classical stacking base pair energies to novel isostericity scales, and apply our techniques
-to correct point-wise errors in 5s rRNA sequences. Our results suggest that \RNApyro is a promising algorithm to complement existing
+to correct point-wise errors in 5s and 16s rRNA sequences. Our results suggest that \RNApyro is a promising algorithm to complement existing
 tools in the NGS error-correction pipeline.

### 2013-05-31 JCB/introduction_JCB.tex
stuff

--- 2013-05-17 JCB/introduction_JCB.tex
@@
 
+%\TODOTous{Relire le papier et ajouter references apparues entre temps (ex: en preparant la presentation).}
 Ribonucleic acids (RNAs) are found in every living organism, and exhibit a broad range of functions, ranging from catalyzing
@@
 
+%\TODOTous{Mentionner technologie Illumina 50-75 nts, et modele d'erreur/mutations.}
 We illustrate and benchmark our techniques for point-wise error corrections on the 5S ribosomal RNA. We choose the latter since it has been extensively

### 2013-05-31 JCB/main_RNAPyro_JCB.tex
Yet another TODO note
stuff

@@
 \usepackage{relsize}
-%\usepackage{bbm}
-\usepackage{bbold}
+\usepackage{bbm}
 \usepackage[numbers]{natbib}
@@
 
-\ShowTODO{\setcounter{tocdepth}{1}
-\listoftodos} 
+%\ShowTODO{\setcounter{tocdepth}{1}
+%\listoftodos} 
 \maketitle
@@
 We develop a scoring scheme combining classical stacking base pair energies to novel isostericity scales, and apply our techniques
-to correct point-wise errors in 5s and 16s rRNA sequences. Our results suggest that \RNApyro is a promising algorithm to complement existing
+to correct point-wise errors in 5s rRNA sequences. Our results suggest that \RNApyro is a promising algorithm to complement existing
 tools in the NGS error-correction pipeline.

### 2013-05-31 JCB/methods_JCB.tex
Made the paper 'Arxiv-friendly'
stuff

@@
 the number of differing positions between two sequences $s'$ and $s''$ such that $|s'|=|s''|$.
-\TODOYann{Move above, define secondary structure and add fancy VARNA illustration}
+%\TODOYann{Move above, define secondary structure and add fancy VARNA illustration}
 
@@
 The isostericity contribution exponentially favors sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
-\TODOYann{Avoid duplicate reference to Stombaugh et al}
+%\TODOYann{Avoid duplicate reference to Stombaugh et al}
 
@@
 This set of sequence is also the inside term on the suffix structure.
-It is also worth pointing out that $\Y{\Struct}{m}{x,y}= \Z{\varepsilon}{m}{y,X}=\mathbbm{1}_{m=0}$.
+It is also worth pointing out that $\Y{\Struct}{m}{x,y}= \Z{\varepsilon}{m}{y,X}=\mathbb{1}_{m=0}$.
 

### 2013-07-01 scripts/timeBenchmark.py
time bench

@@ -36,7 +36,7 @@ def genMSA(secstr,numseqs):
     result.append("".join(seq))
   return result
 
-def genRNAPyroInputFiles(basepath,fromlength=50,tolength=500,lengthincr=25,numseqs=25,alignmentlength=50):
+def genRNAPyroInputFiles(basepath,fromlength=50,tolength=500,lengthincr=25,numseqs=25,alignmentlength=75):
   for i in range(fromlength,tolength,25):
     sys.stderr.write("*%s"%(i))
     os.system("grgfreqs --draw RNALoopsOut.ggd %s %s > out.txt 2> dummy"%(i,numseqs))
@@ -60,7 +60,7 @@ if __name__=="__main__":
 #  if not os.path.exists("TimeBench"):
 #    os.mkdir("TimeBench")
 #  genRNAPyroInputFiles(os.path.join("TimeBench","MSA"),50,500,25,50,50)
-  exec_name = os.path.join('..','src','RNAPyro_iso.py')
+  exec_name = os.path.join('..','src','RNAPyro.py')
   nbmuts = 10
   alpha = .5
 
@@ -73,7 +73,7 @@ if __name__=="__main__":
     dataPoints[length].append(f_name)
   maxSize = max([len(x) for x in dataPoints.values()])
 
-  outfile = open("TimeBench_yamaska.dat","w")
+  outfile = open("TimeBench_yamaska_75.dat","w")
   lengths = dataPoints.keys()
   lengths.sort()
   for i in range(maxSize):

### 2013-07-01 scripts/jcb_graphs.py
time multi

@@ -95,7 +95,7 @@ def createTimeFigMultiAl():
     dataPoints = {}
     pylab.figure()
     for yamaska_name in ("TimeBench_yamaska.dat",
-                 "TimeBench_yamaska_75.dat",
+                         "TimeBench_yamaska_75.dat",
                          "TimeBench_yamaska_100.dat"):
         for l in open(yamaska_name,"r"):
             data = l[:-1].split()

### 2013-07-01 JCB/results_JCB.tex
stuff
Minor changes/typos

@@
 \end{subfigure}
-\caption{Performance of error-correction over structured positions (i.e. inside a base pair). Subfigures show accuracy when
+\caption{Performance of error-correction restricted to structured positions (i.e. involved in a canonical base pair). Subfigures show accuracy when
 ART fold parameter is set to 5, 10 and $15-$fold coverage.
@@
 comes from the isostericity. A notable exception is when the number of mutation is underestimated, 
-in that case best performances were observed for non-null values of $\alpha$.
+in that case best performances were observed for larger values of $\alpha$.
 
-Another observation is when we look at all positions as in Fig.~\ref{fig:16s_all}.
-The first $20\%$ to $30\%$ of sensitivity are obtained almost without any errors and correspond to the nucleotides in
+A final observation is related to  all positions as in Fig.~\ref{fig:16s_all}. The first $20\%$ to $30\%$ of sensitivity are obtained almost without any errors and correspond to the nucleotides in 
 structured regions. The rest of the predictions are done on un-paired nucleotides.

### 2013-07-01 JCB/conclusion_JCB.tex

@@
 
-\TODOJerome{Etoffer la conclusion}
+%\TODOJerome{Etoffer la conclusion}
 In this article we presented a new and efficient way of exploring the mutational landscape of an RNA under structural constraints,

### 2013-07-01 JCB/results_JCB.tex
update text commnet ROC

--- 2013-05-28 JCB/results_JCB.tex
@@
 based on a realistic random secondary structure, generated as described in Levin~\emph{et al}~\cite{Levin:2012kx}.
-\TODOVlad{Add more extensive time benchmark, as a plot}
+%\TODOVlad{Add more extensive time benchmark, as a plot}
 %\TODOVlad{Is time benchmark still on Mac mini or mcbbigram?}
@@
 \label{sec:16S}
-\TODOTous{Describe and present results????}
+%\TODOTous{Describe and present results????}
 
@@
 $4.8\%$, $1.8\%$ and $1.2\%$ for fold coverages of $5$, $10$ and $15$ respectively.
-We repeated the whole procedure $12$/$10$/$14$ times for the $5$/$10$/$15-$fold coverage. 
+We repeated the whole procedure $30$ times for each ($5$/$10$/$15-$fold) coverage. 
 We evaluated the predictive power of the profile by a computing a joint ROC curve for each value of $\alpha$, and each coverage, as described in Section~\ref{sec:5S}.
@@
 The AUC is indicated in the legend of the figures.
-Each individual ROC curve represent the average performance over at least 10 experiments.}
+Each individual ROC curve represent the average performance over 30 experiments.}
 
@@
 in Fig.~\ref{fig:16s_all},  indicate that a signal has been successfully extracted. When restraining
-to structured regions (i.e. base paired regions) we obtain AUC values up to 0.998.  Since 
+to structured regions (i.e. base paired regions) we obtain AUC values up to 0.988.  Since 
 contributions to the energy and isostericity only arise from structured regions, this was an expected result.
@@
 An interesting feature is that best results are almost always when $\alpha=0$, i.e. all the contribution
-comes from the isostericity. A notable exception is when the number of mutation is underestimated, 
-in that case best performances were observed for non-null values of $\alpha$.
+comes from the isostericity.
+% A notable exception is when the number of mutation is underestimated, 
+%in that case best performances were observed for non-null values of $\alpha$.
 

### 2013-07-04 JCB/conclusion_JCB.tex
minor change conclusion

@@
 
-Our techniques are designed to correcting point-wise errors in structured regions (i.e. base paired nucleotides). Nonetheless,
+Our techniques are designed to correcting point-wise errors in structured regions (i.e. base paired nucleotides). However,
 our software \texttt{RNApyro} can be easily combined with other methodologies previously developed for correcting other types

### 2013-07-05 JCB/results_JCB.tex

@@
 
-Another observation is when we look at all positions as in Fig.~\ref{fig:16s_all}.
-The first $20\%$ to $30\%$ of sensitivity are obtained almost without any errors and correspond to the nucleotides in
+A final observation is related to  all positions as in Fig.~\ref{fig:16s_all}. The first $20\%$ to $30\%$ of sensitivity are obtained almost without any errors and correspond to the nucleotides in 
 structured regions. The rest of the predictions are done on un-paired nucleotides.

### 2014-02-04 JCB/conclusion_JCB.tex
JCBFinal

@@
 and apply our techniques to identify and fix sequencing errors. In addition, we introduced a new scoring scheme to measure the
-likelihood of sequencing errors that combines the classical nearest-neighbour energy model parameters \cite{Turner2010} to the
-recently introduced isostericity matrices \cite{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring
+likelihood of sequencing errors that combines the classical nearest-neighbour energy model parameters \citep{Turner2010} to the
+recently introduced isostericity matrices \citep{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring
 during base pair replacements.
@@
 We validated our model on the 5s rRNA and 16s rRNA (See Sec.~\ref{sec:results}) and showed that our technology enables us
-to recover mutational errors with high accuracy (Area under the ROC curve between $0.95$ and $0.99$ when the mutations are
-located in in base paired regions). Interestingly, we observed that using the isostericity matrices alone yields higher performance
+to recover mutational errors with high accuracy (Area under the ROC curve between $0.95$ and $0.99$ when restricting predicted mutations to base paired regions).
+Interestingly, we observed that using the isostericity matrices alone yields higher performance
 than with the nearest-neighbour energy model alone. This finding supports our hypothesis that isostericity matrices provide a
 valuable source of informations that can be efficiently used for RNA sequence and structure analysis. Nonetheless, we also found
-that the nearest-neighbour model seems to provide a valuable signal when we under-estimate of the number of errors in the input
-sequences.
+that the nearest-neighbour model seems to provide a valuable signal when the number of errors in the input sequences is under-estimated.
 
@@
 our software \texttt{RNApyro} can be easily combined with other methodologies previously developed for correcting other types
-of sequencing errors such as indels in unstructured regions or repeats \cite{Quinlan2008,Quince:2009uq}. In further works, we
+of sequencing errors such as indels in unstructured regions or repeats \citep{Quinlan2008,Quince:2009uq}.
+
+In furture works, we
 also intend to include in our model errors stemming from insertions or deletions. It is indeed theoretically possible to considers
-these scenario within our dynamic programming scheme \cite{Waldispuhl:2002fk} with minor impacts of the complexity.
-
+these scenario within our dynamic programming scheme \citep{Waldispuhl:2002fk} with minor impacts of the complexity.
 Finally, we hope that integrating our software to current sequencing pipe-lines used in metagenomics studies will permit to

### 2014-02-04 JCB/introduction_JCB.tex
JCBFinal

@@
 
-For half a century, biological molecules have been studied as a proxy to understand evolution~\cite{Zuckerkandl1965}, and due
+For half a century, biological molecules have been studied as a proxy to understand evolution~\citep{Zuckerkandl1965}, and due
 to their fundamental functions and remarkably conserved structures, rRNAs have always been a prime candidate for phylogenetic
-studies~\cite{Olsen1986, Olsen1993}. In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007} benefited
+studies~\citep{Olsen1986, Olsen1993}. In recent years, studies as the \emph{Human Microbiome Project}~\citep{Turnbaugh2007} benefited
 of new technologies such as the NGS techniques to sequence as many new organisms as possible and extract an unprecedented flow of new information. 
@@
 Because there is no cloning step, resequencing to increase accuracy is not possible and it is therefore vital to disentangle noise from true sequence
-diversity in this type of data \cite{Quince:2009uq}. Errors can be significantly reduced  when large multiple sequence alignments with close homologs
+diversity in this type of data \citep{Quince:2009uq}. Errors can be significantly reduced  when large multiple sequence alignments with close homologs
 are available, but in studies of new or not well known organisms, such information is rather sparse. In particular, it is common that there is not enough  similarity to 
-differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often leading to artificially inflated diversity estimates~\cite{Kunin2010}.
-A few techniques have been developed to remedy to this problem~\cite{Quinlan2008,Medvedev2011} but they do not take
+differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often leading to artificially inflated diversity estimates~\citep{Kunin2010}.
+A few techniques have been developed to remedy to this problem~\citep{Quinlan2008,Medvedev2011} but they do not take
 into account all the available information. It is therefore essential to develop methods that can exploit any type of signal available to correct errors.
@@
 
-Here, we expand the range of algorithmic techniques previously introduced with the \RNAmutants software~\cite{Waldispuhl2008,Waldispuhl2011}.
+Here, we expand the range of algorithmic techniques previously introduced with the \RNAmutants software~\citep{Waldispuhl2008,Waldispuhl2011}.
 Instead of exploring the full conformational landscape and sample mutants, we develop an inside-outside algorithm that enables us
@@
 
-We design a new scoring scheme combining nearest-neighbor models \cite{Turner2010} to isostericity metrics \cite{Stombaugh2009}.
-Classical approaches use a Boltzmann distribution whose weights are estimated using a nearest-neighbour energy model~\cite{Turner2010}. However, the
-latter only accounts for  canonical and wobble, base pairs. As was shown by Leontis and Westhof~\cite{Leontis2001},
+We design a new scoring scheme combining nearest-neighbor models \citep{Turner2010} to isostericity metrics \citep{Stombaugh2009}.
+Classical approaches use a Boltzmann distribution whose weights are estimated using a nearest-neighbour energy model~\citep{Turner2010}. However, the
+latter only accounts for  canonical and wobble, base pairs. As was shown by Leontis and Westhof~\citep{Leontis2001},
 the diversity of base pairs observed in tertiary structures is much larger, albeit their energetic contribution remains unknown. To quantify geometrical discrepancies, 
-an isostericity distance has been designed \cite{Stombaugh2009}, increasing as two base pairs geometrically differ from each other in space. Therefore, we 
+an isostericity distance has been designed \citep{Stombaugh2009}, increasing as two base pairs geometrically differ from each other in space. Therefore, we 
 incorporate these scores in the Boltzmann weights used by \RNApyro.
@@
 We illustrate and benchmark our techniques for point-wise error corrections on the 5S ribosomal RNA. We choose the latter since it has been extensively
-used for phylogenetic reconstructions~\cite{Hori1987} and its sequence has been recovered for over 712 species (in the Rfam seed alignment with id
+used for phylogenetic reconstructions~\citep{Hori1987} and its sequence has been recovered for over 712 species (in the Rfam seed alignment with id
 \texttt{RF00001}). Using a leave one out strategy, we perform random distributed mutations on a sequence. While our methodology is restricted to the correction of 

### 2014-02-04 JCB/methods_JCB.tex
JCBFinal

@@
 In this model, the probability that the nucleotide at a given position needs to be mutated (i.e.
-corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990} in time which scales linearly with the sequence length.
+corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\citep{Lari1990} in time which scales linearly with the sequence length.
 
@@
 \subsubsection{Energy Contribution.}
-The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}.
+The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using values from the Turner 2004 model retrieved from the NNDB~\citep{Turner2010}.
 Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
@@
 \subsubsection{Isostericity Contribution.}
-The concept of isostericity score is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
+The concept of isostericity score is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\citep{Stombaugh2009}.
 Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the isostericity contribution to the pseudo-energy as
@@
 is the average isostericity of a base-pair in the candidate sequence, compared with the reference alignment.
-The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}.
+The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\citep{Stombaugh2009}.
 Isostericity scores range between $0$ and $9.7$, $0$ corresponding to a perfect isostericity, and a penalty of $10$ is used for missing entries.
@@
 To work around this issue, we propose a linear-time variant of the
- \emph{Inside-Outside algorithm}~\cite{Lari1990} to compute this probability, based on two sets of dynamic programming equations. 
+ \emph{Inside-Outside algorithm}~\citep{Lari1990} to compute this probability, based on two sets of dynamic programming equations. 
 %from the two  functions $\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.
@@
 \begin{figure}[t]\centering
-\includegraphics[scale=\ScaleDP]{FigDPInsideWrapper}
+\includegraphics[scale=\ScaleDP]{Figure1-FigDPInsideWrapper}
 \caption{Principle of the inside computation (partition function). Any sequence (mutated)  
@@
 
-\item {\bf Case 3: Paired leftmost position, but no stacking pairs.} In this case, the structure is split into two parts by the base pair ($\Struct = \op \Struct'\cp\Struct''$).
+\item {\bf Case 3: Paired leftmost position + absence of stacking pairs.} In this case, the structure is split into two parts by the base pair ($\Struct = \op \Struct'\cp\Struct''$).
 Let us denote by $k$ the partner of position $i$, and by $a$, $b$ and $c$ the bases found at positions $i$, $k$ and $j$ respectively, then one has:
@@
 \begin{figure}[t]\centering
-\includegraphics[scale=\ScaleDP]{FigDPOutsideWrapper}
+\includegraphics[scale=\ScaleDP]{Figure2-FigDPOutsideWrapper}
 \caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
@@
 \end{equation}
-\item {\bf Case 3: Next position paired rightward, in the absence of stacking pair.} In this case, the innermost leftward position $i$ is paired to the right at some position $k$ . 
+\item {\bf Case 3: Next position paired rightward + absence of stacking pair.} In this case, the innermost leftward position $i$ is paired to the right at some position $k$ . 
 Let us assume that its partner resides outside the excluded structure $\Struct$ (This assumption is provably without loss of generality, and directly follows from the fact that $\Struct$ is well-parenthesized).
@@
 \begin{figure}[t]\centering
-\includegraphics[scale=\ScaleDP]{FigDPCombineWrapper}
+\includegraphics[scale=\ScaleDP]{Figure3-FigDPCombineWrapper}
 \caption{Combining the inside and outside contributions to compute the total Boltzmann weight of all sequences having a given base (case 1) or base-pair (cases 2 and 3) at a given position.\label{fig:combine}}

### 2014-02-04 JCB/results_JCB.tex
JCBFinal

@@
 \subsection{Implementation}
-The software was implemented in Python2.7 using the \textit{mpmath}~\cite{mpmath} library
+The software was implemented in Python2.7 using the \textit{mpmath}~\citep{mpmath} library
 for  arbitrary floating point precision. The source code is freely available at:
@@
 For those tests, both the multiple sequence alignment and the reference sequence were generated uniformly at random,
-based on a realistic random secondary structure, generated as described in Levin~\emph{et al}~\cite{Levin:2012kx}.
+based on a realistic random secondary structure, generated as described in~\cite{Levin:2012kx}.
 %\TODOVlad{Add more extensive time benchmark, as a plot}
@@
 \begin{figure}[t]
-{\centering \includegraphics[width=.5\linewidth]{figures/TimeBenchmark}\\}
+{\centering \includegraphics[width=.5\linewidth]{Figure4-TimeBenchmark}\\}
 
@@
 biggest MSA  which contains $130$ sequences (out of $712$ in the original Rfam MSA). Then, in order to avoid overfitting, we used \texttt{cd-hit}
-\cite{Li:2006fk} to remove sequences with more than 80\% of sequence similarity. This operation resulted in a data set of $45$ sequences.
+\citep{Li:2006fk} to remove sequences with more than 80\% of sequence similarity. This operation resulted in a data set of $45$ sequences.
 
@@
 \centering
-	\includegraphics[width=\textwidth]{subfigs_perform.png}\\
+	\includegraphics[width=\textwidth]{Figure5-subfigs_perform}\\
 
@@
 
-We gathered the seed sequences of the bacterial multiple sequence alignment (MSA) retrieved from the RFAM Database 11.0 (Rfam id: RF00177)~\cite{gardner2011rfam}.
+We gathered the seed sequences of the bacterial multiple sequence alignment (MSA) retrieved from the RFAM Database 11.0 (Rfam id: RF00177)~\citep{gardner2011rfam}.
 This alignment is composed of 93 sequences, whose length varies between 1461 and 1568 nucleotides, and has an average pairwise sequence identity of $69\%$.
@@
 To simulate sequencing errors, we used the next-generation sequencing read simulator 
-{\tt ART}~\cite{huang2012art}.
+{\tt ART}~\citep{huang2012art}.
 The \emph{Illumina technology} setting was chosen as the main error mode, generating reads of $75~\text{bps}$, featuring mostly base substitution errors.
@@
 
- \begin{figure}
+ \begin{figure}[t]
 \centering
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/5fold-a}
+  \includegraphics[width=\linewidth]{Figure6-5fold-a}
   \caption{$5-$fold coverage}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/10fold-a}
+  \includegraphics[width=\linewidth]{Figure7-10fold-a}
     \caption{$10-$fold coverage}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/15fold-a}
+  \includegraphics[width=\linewidth]{Figure8-15fold-a}
     \caption{$15-$fold coverage}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/5fold-b}
+  \includegraphics[width=\linewidth]{Figure9-5fold-b}
   \caption{$5-$fold coverage}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/10fold-b}
+  \includegraphics[width=\linewidth]{Figure10-10fold-b}
     \caption{$10-$fold coverage}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/15fold-b}
+  \includegraphics[width=\linewidth]{Figure11-15fold-b}
     \caption{$15-$fold coverage}
@@
 
-An interesting feature is that best results are almost always when $\alpha=0$, i.e. all the contribution
+An interesting feature is that best results are almost always obtained when $\alpha=0$, i.e. when all the contribution
 comes from the isostericity.
